### 2017-10-19 docs/manuscript.tex

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{geometry} 
\geometry{a4paper}  
\usepackage[parfill]{parskip} 
\usepackage{graphicx}	
\usepackage{amsmath}
\usepackage{fullpage}
\usepackage{setspace} 
\usepackage{lineno}
\usepackage[none]{hyphenat}

\usepackage[round]{natbib}

\title{High throughput data extraction from primary literature: The metaDigitise R package}
\author{Joel L. Pick ?...? \& Daniel W.A. Noble}
% \date{}

\begin{document}
\doublespacing
\raggedright

\maketitle
Prepared for: Journal of Statistical Software?

\section*{Abstract}

Keywords: meta-analysis, data extraction, R

\clearpage

Meta-analysis is becoming increasingly common in many fields. It relies foremost on data extracted from primary literature. This data if often presented in figures and so needs to be manually extracted. Although there are several existing tools to perform tasks like this, 
these tools are either not made for this specific purpose (i.e. meta-analysis) and so require a large amount of downstream data manipulation or do not allow easy import into statistical software (such as R). Furthermore these tools are generally not designed for a high throughput of images. 
Here, we present an interactive R package designed for large scale data extraction from figures, with meta-analysis in mind. To this end, we provide tools specific to data extraction from common plot types (mean and error plots, boxplots, scatterplots and histograms) and functions that allow for the summary statistics needed for downstream analysis to be calculated (i.e. mean sd and n and r). %d? 
The package

\section{Functionality}
\subsection{Figure Rotation}
-for extraction from older, copied papers. also to flip sideways plots

\subsection{Calibration}

\subsection{Bulk Digitisation}

\section{Plot Types}
\subsection{Mean and error plots}

\subsection{Boxplots}

\subsection{Scatterplots}

\subsection{Histograms}

\section{Summary Data}

\section{Reproducibility}

\end{document}

### 2017-10-19 R/ME_Box_extraction.R
argument descriptions

@@ -1,4 +1,5 @@
 #' @title single_group_extract
+#' @param plot_type Type of plot
 #' @description Takes points user defined points from a single group mean_error plot or boxplot, in a set order, and returns them.
 single_group_extract <- function(plot_type){
 	
@@ -16,6 +17,9 @@ single_group_extract <- function(plot_type){
 }
 
 #' @title groups_extract
+#' @param plot_type The type of plot
+#' @param nGroups number of groups
+#' @param image_width width of image
 #' @description Extraction of data from boxplots of mean_error plots, from multiple groups
 groups_extract <- function(plot_type, nGroups, image_width){
 	nRows <- ifelse(plot_type=="mean_error",2,5)
@@ -46,6 +50,9 @@ groups_extract <- function(plot_type, nGroups, image_width){
 }
 
 #' @title convert_group_data
+#' @param cal_data Calibrated data
+#' @param plot_type The type of plot
+#' @param nGroups number of groups
 #' @description Converts, pre-calibrated points clicked into a meaningful dataframe 
 convert_group_data <- function(cal_data, plot_type, nGroups){
 	nRows <- ifelse(plot_type=="mean_error",2,5)

### 2017-10-19 R/metaDigitise.R
add required package info

@@ -14,8 +14,8 @@ metaDigitise <- function(file, plot_type=NULL, summary_stats=FALSE){
 	new_image <- graph_rotate(image)
 	flush.console()
 
-	image_width <- image_info(new_image)["width"]
-	image_height <- image_info(new_image)["height"]
+	image_width <- magick::image_info(new_image)["width"]
+	image_height <- magick::image_info(new_image)["height"]
 
 	output <- list()
 

### 2017-10-30 R/import_metaDigitise.R
Added a check in to make sure user ends path names with / to signify folder. Also added back in, who the hell knows how it got deleted, teh details object in process_new_digitise function

@@ -6,7 +6,7 @@
 #' @export
 
 import_metaDigitise <- function(dir, summary = TRUE ) {
-	
+	## THIS FUNCTION MAYBE OBSELETE NOW. REPLACED WITH process_new_files as all component functions are embed within.
 	    details <- dir_details(dir)
 	    metaDig <- load_metaDigitise(details$doneCalFiles)
 	 plot_types <- lapply(metaDig, function(x) x$plot_type)

### 2017-10-30 R/metaDigitise.R

@@ -11,6 +11,11 @@
 #' @export
 
 metaDigitise<-function(dir, summary = TRUE){
+		# Check dir has a / at the end.
+	if( (substring(dir, nchar(dir)) == "/") == FALSE){
+		dir <- paste0(dir, "/")
+	}
+
 	cat("Do you want to...\n")
 	
 	Q <- menu(c("Process new images", "Import existing data", "Edit existing data"))
@@ -27,6 +32,7 @@ metaDigitise<-function(dir, summary = TRUE){
 process_new_files <- function(dir, summary = TRUE) {
 
 			       setup_calibration_dir(dir)
+			      details <- dir_details(dir)
 	     done_details <- get_notDone_file_details(dir)
 		   type <- user_options("Are all plot types the same? (diff/same)" , c("diff", "same"))
 	

### 2017-10-30 R/import_metaDigitise.R
Fixed a few important functions

@@ -6,9 +6,9 @@
 #' @export
 
 import_metaDigitise <- function(dir, summary = TRUE ) {
-	## THIS FUNCTION MAYBE OBSELETE NOW. REPLACED WITH process_new_files as all component functions are embed within.
+	
 	    details <- dir_details(dir)
-	    metaDig <- load_metaDigitise(details$doneCalFiles)
+	    metaDig <- load_metaDigitise(details$doneCalFiles, details$name)
 	 plot_types <- lapply(metaDig, function(x) x$plot_type)
 
 	if(summary == TRUE){
@@ -28,11 +28,9 @@ import_metaDigitise <- function(dir, summary = TRUE ) {
 #' @description Loads metaDigitise calibration files from a directory that is partially or fully digitised already
 #' @param doneCalFiles The calibration files that have already been finished taken from directory details
 #' @return Returns a list of metaDigitised objects that have already been completed
-
-load_metaDigitise <- function(doneCalFiles){
-	details <- get_notDone_file_details(dir)
+load_metaDigitise <- function(doneCalFiles, names){
 	metaDig <- lapply(doneCalFiles, function(x) readRDS(x))
-	names(metaDig) <- details$calibrations
+	names(metaDig) <- names
 	return(metaDig)
 }
 
@@ -44,10 +42,10 @@ load_metaDigitise <- function(doneCalFiles){
 
 order_lists <- function(list, plot_types){
 
-	mean_error <- list[match("mean_error", unlist(plot_types), nomatch = TRUE)]
-   scatterplot <- list[match("scatterplot", unlist(plot_types),  nomatch = FALSE)]
-	   boxplot <- list[match("boxplot", unlist(plot_types),  nomatch = FALSE)]
-	      hist <- list[match("histogram", unlist(plot_types),  nomatch = FALSE)]
+	mean_error <- list[which(unlist(plot_types) == "mean_error")]
+   scatterplot <- list[which(unlist(plot_types) == "scatterplot")]
+	   boxplot <- list[which(unlist(plot_types) == "boxplot")]
+	      hist <- list[which(unlist(plot_types) == "histogram")]
 
 	dat_list <- as.list(c(mean_error = mean_error, boxplot=boxplot, hist=hist, scatterplot=scatterplot))
 

### 2017-10-30 R/metaDigitise.R
Working now

@@ -32,12 +32,12 @@ metaDigitise<-function(dir, summary = TRUE){
 process_new_files <- function(dir, summary = TRUE) {
 
 			       setup_calibration_dir(dir)
-			      details <- dir_details(dir)
-	     done_details <- get_notDone_file_details(dir)
+			      done_details <- dir_details(dir)
+	     details <- get_notDone_file_details(dir)
 		   type <- user_options("Are all plot types the same? (diff/same)" , c("diff", "same"))
 	
 	if(length(done_details$calibrations) >= 1){	
-		done_objects <- load_metaDigitise(details$doneCalFiles)
+		done_objects <- load_metaDigitise(done_details$doneCalFiles, done_details$names)
 		done_plot_types <- lapply(done_objects, function(x) x$plot_type)
 	}
 
@@ -58,10 +58,12 @@ process_new_files <- function(dir, summary = TRUE) {
 		plot_type <- c(done_plot_types, complete_plot_types)
 
 	if(summary == TRUE){
+
 		return(do.call(rbind, list(summary(done_objects), summary(data_list))))
+
 	}else{
-			done_figs <- extract_digitised(done_objects, types = plot_types, summary = summary)
-			new_figs <- extract_digitised(data_list, types = plot_types, summary = summary)
+			done_figs <- extract_digitised(done_objects,  summary = summary)
+			new_figs <- extract_digitised(data_list, summary = summary)
 		return(order_lists(c(done_figs, new_figs), plot_types = plot_type))
 	}
 
@@ -91,16 +93,12 @@ specify_type <- function(){
 #' @return The function will return a data frame with the data across all the digitised files 
 #' @export
 
-extract_digitised <- function(list, types = c("same", "diff"), summary = TRUE) {
+extract_digitised <- function(list, summary = TRUE) {
 
 	if(summary == TRUE) {
 		data <- do.call(rbind, lapply (list, function(x) summary(x)))
 		rownames(data) <- 1:nrow(data)
 		return(data)
-	}
-
-	if(types == "same"){
-		return( do.call (rbind, lapply (list, function(x) x$processed_data)))
 	} else {
 		return(lapply (list, function(x) x$processed_data))
 	}

### 2017-10-30 R/import_metaDigitise.R
fixes and working still

@@ -27,7 +27,9 @@ import_metaDigitise <- function(dir, summary = TRUE ) {
 #' @title load_metaDigitise
 #' @description Loads metaDigitise calibration files from a directory that is partially or fully digitised already
 #' @param doneCalFiles The calibration files that have already been finished taken from directory details
+#' @param names The names of the done calibration files
 #' @return Returns a list of metaDigitised objects that have already been completed
+
 load_metaDigitise <- function(doneCalFiles, names){
 	metaDig <- lapply(doneCalFiles, function(x) readRDS(x))
 	names(metaDig) <- names

### 2017-10-30 R/metaDigitise.R

@@ -88,7 +88,6 @@ specify_type <- function(){
 #' @title extract_digitised
 #' @param list A list of objects returned from metaDigitise
 #' @param summary A logical 'TRUE' or 'FALSE' indicating whether metaDigitise should print summary statitics from each figure and group.
-#' @param types Whether the plots are the same or different types
 #' @description Function for extracting the data from a metaDigitise list
 #' @return The function will return a data frame with the data across all the digitised files 
 #' @export
@@ -100,7 +99,9 @@ extract_digitised <- function(list, summary = TRUE) {
 		rownames(data) <- 1:nrow(data)
 		return(data)
 	} else {
-		return(lapply (list, function(x) x$processed_data))
+		tmp <- lapply (list, function(x) x$processed_data)
+		names(tmp) <- unlist(lapply(list, function(x) x$image_file))
+		return(tmp)
 	}
 }
 

### 2017-10-30 R/metaDigitise.R
Sorted a very small bug taht would occur under conditions in which the caldat/ folder was empty

@@ -55,8 +55,12 @@ process_new_files <- function(dir, summary = TRUE) {
 	
 		complete_plot_types <- lapply(data_list, function(x) x$plot_type)
 
-		plot_type <- c(done_plot_types, complete_plot_types)
-
+		if( length(done_plot_types) > 1){
+			plot_type <- c(done_plot_types, complete_plot_types)
+		} else{
+			plot_type <- complete_plot_types
+		}
+		
 	if(summary == TRUE){
 
 		return(do.call(rbind, list(summary(done_objects), summary(data_list))))

### 2017-10-30 R/metaDigitise.R
Opps accidently pshed some unneeded figs

@@ -1,9 +1,12 @@
 #' @title metaDigitise
-#' @description Single or batch processes figures of .png, .jpg, .tiff, .pdf extensions within a set directory, consolidates the data and exports the data for each image and type
+#' @description Single or batch processing of figures with .png, .jpg, .tiff, .pdf extensions within a set directory. metaDigitise consolidates the data and exports the data for each image and image type. It can also summarise the data, provide the raw data (if scatterplots) and automatically imports previously finished data and merges it with newly digitised data. metaDigitise also allows users to check their calibration along with editing previous digitisations.
 #' @param dir the path name to the directory / folder where the files are located
 #' @param summary whether the digitised data should be returned as a summary (TRUE) or as a concatenated list of similar types. 
 #' @details 
-#' metaDigitise can be used on a directory with a whole host of different figure (mean and error, scatter plots, box plots and histograms) and file types (.jpeg, .png, .tiff, .pdf). It will automatically cycle through all files within a directory in order, prompting the user for specific information as they go. It will also write calibration files (also containing processed data), into a special caldat/ folder within the directory. Importantly, as new files are added to a directory that has already been "completed", metaDigitise will recognize these unfinished files and only cycle through the digitisation of these new files. 
+#' metaDigitise can be used on a directory with a whole host of different figure types (mean and error, scatter plots, box plots and histograms) and file types (.jpeg, .png, .tiff, .pdf). It will automatically cycle through all files within a directory in order, prompting the user for specific information as they go. It will automatically write metaDigitise object files (in .RDS format containing processed and calibration data along with directory and file details), into a special caldat/ folder within the directory. Importantly, as new files are added to a directory that has already been "completed", metaDigitise will recognize these unfinished files and only cycle through the digitisation of these new files. 
+#' 
+#' metDigitise is built for reproducibility and ease of operation. This means that, so long as the caldat/ folder along with respective images are maintained, anyone using metaDigitise can simply import existing digitisations, modify them and fix them. 
+#' @author Joel Pick - joel.l.pick@gmail.com; Daniel Noble - daniel.wa.noble@gmail.com
 #' @examples
 #' # data <- metaDigitise(dir = "./example_figs/", summary = TRUE)
 #' # summary(data)
@@ -25,9 +28,10 @@ metaDigitise<-function(dir, summary = TRUE){
 
 
 #' @title process_new_files
[…]

### 2017-10-30 docs/manuscript.tex
some manuscript writing

@@
 
-\title{High throughput data extraction from primary literature: The metaDigitise R package}
-\author{Joel L. Pick ?...? \& Daniel W.A. Noble}
+\title{Flexible, high throughput data extraction from primary literature: The metaDigitise R package}
+\author{? Joel L. Pick ?...? \& Daniel W.A. Noble ?}
 % \date{}
@@
 This data if often presented in figures and so needs to be manually extracted.
-Although there are several existing tools to perform tasks like this,
-these tools are either not made for this specific purpose (i.e.
-meta-analysis) and so require a large amount of downstream data manipulation or do not allow easy import into statistical software (such as R).
+Although there are several existing tools to perform tasks like this, these tools are not made for this specific purpose (i.e.
+meta-analysis).
+Specifically, they do not differentiate between common types of plot in which are use to present data, meaning they require a large amount of downstream data manipulation.
+Many exist in stand-alone software and so do not allow easy import into commonly used statistical software (such as R).
+These softwares often do not allow for different grouping of points to be made, and frequently do not allow metadata (such as variable names, group names, sample sizes) to be entered at the time of data extraction.
 Furthermore these tools are generally not designed for a high throughput of images.
 Here, we present an interactive R package designed for large scale data extraction from figures, with meta-analysis in mind.
-To this end, we provide tools specific to data extraction from common plot types (mean and error plots, boxplots, scatterplots and histograms) and functions that allow for the summary statistics needed for downstream analysis to be calculated (i.e.
-mean sd and n and r).
+To this end, we provide tools specific to data extraction from common plot types (mean and error plots, box plots, scatter plots and histograms, see Figure \ref{fig:plot_type}) and functions that allow for the summary statistics needed for downstream analysis to be calculated (e.g.
+mean, standard deviation and sample size).
 %d?
-The package
+
+\begin{figure}[ht] 
+%\onehalfspacing
+ \includegraphics[width=0.9\textwidth]{fig_plot_type.pdf} 
+ \caption{Plot types that metaDigitise is designed to extract data from: A) mean and error plot, B) box plot, C) scatter plot and D) histogram. }
+\label{fig:plot_type}
+\end{figure}
 
 \section{Functionality}
+There is one main function in the package metaDigitise, which interactively takes the user through the process of extracting figures from graphs.
+
+\subsection{Bulk Digitisation}
+metaDigitise was created with the idea that the user would have multiple images to extract from.
+It therefore operates in the same way whether the user has one or multiple images.
+It assumes that the user has put all the figures needed for extraction in one parent directory.
+It first creates a new directory `caldat' inside this parent directory, in which calibration files are saved (see below).
+
+It then prompts the user to select whether they want to extract from new images, import previously extracted data or edit previously extracted data.
+Extracting from new images takes the user through the steps outlined below, figure by figure, giving the user the option to quit the process after each figure.
+If this process has already been started when the user starts metaDigitise, then it starts where the user left off.
+Similarly if the user adds new images to the parent directory after all other images are extracted (or part way through), the new images will be extracted from.
+After extraction (or upon quitting the function), the extracted data will be returned.
+%% explain summary etc
+
+If the process of data extraction has been finished, or the user wants to reimport summary or processed data ...
+
 \subsection{Figure Rotation}
--for extraction from older, copied papers. also to flip sideways plots
+Figure may have been extracted from old publications, for example from scanned images, and so are not perfectly orientated on the image.
+This will make the calibration of the points in the figure from the image problematic.
+metaDigitise allows users to rotate the image.
+By clicking two points on the x-axis, metaDigitse calculates the angle needed to rotate the image so the x-axis is horizontal, and rotates it.
+(Figure \ref{fig:rotate} )
+
+\begin{figure}[!h] 
+%\onehalfspacing
+ \includegraphics[width=0.9\textwidth]{fig_rotate.pdf} 
+ \caption{ }
+\label{fig:rotate}
+\end{figure}
+
+Furthermore, in some figures mean and error, boxplots or histograms may be presented with horizontal bars.
+metaDigitise assumes that the bars are vertical, but allows the user to flip the image so that the bars are vertical (Figure \ref{fig:flip}).
+
+\begin{figure}[!h] 
+%\onehalfspacing
+ \includegraphics[width=0.9\textwidth]{fig_flip.pdf} 
+ \caption{ }
+\label{fig:flip}
+\end{figure}
 
 \subsection{Calibration}
+metaDigitise requires the user to calibrate the axes in the figure.
+To do this the user is required to click on two known points on the axis in question, and then enter the value of those points in the figure (Figure \ref{fig:calibrate}).
+Using this information, metaDigitise then calculates the value of any clicked points in terms of the figure axes.
+In the case of mean and error plots and box plots, it calibrates only the y axis (assuming the x axis is redundant).
+For scatter plots and histograms both axes are calibrated.
 
-\subsection{Bulk Digitisation}
+\begin{figure}[!h] 
+%\onehalfspacing
+ \includegraphics[width=0.9\textwidth]{fig_calibrate.pdf} 
+ \caption{ }
+\label{fig:calibrate}
+\end{figure}
 
-\section{Plot Types}
-\subsection{Mean and error plots}
+\subsection{Plot Types}
+metaDigitise recognises 4 main types of plot; Mean and error plots, box plots, scatter plots and histograms, shown in Figure \ref{fig:plot_type}.
 
-\subsection{Boxplots}
+\subsubsection{Mean and error plots} 
+metaDigitise prompts the user to enter groups names and allows the user to enter sample sizes, which are used in downstream processing.
+The user is then prompted to click on an error bar followed by the mean.
+Error bars above or below the mean can be clicked - sometimes one is clearer than the other.
+metaDigitise assumes that the error bars are symmetrical.
+Where the user has clicked the error is displayed in a different colour to the mean.
+The user can subsequently add more groups or remove groups.
 
-\subsection{Scatterplots}
+\subsubsection{Box plots}
+ metaDigitise prompts the user to enter groups names and allows the user to enter sample sizes, which are used in downstream processing.
+The user is then prompted to click on the maximum, upper quartile, median, lower quartile and minimum.
+metaDigitise will check that the maximum is greater than the minimum, and return an warning if that is not the case.
+The user can subsequently add more groups or remove groups.
 
-\subsection{Histograms}
+\subsubsection{Scatter plots}
+ metaDigitise prompts the user to enter groups names and then to click on points.
+Points added by mistake can be deleted.
+The user can subsequently add groups, edit groups (add or remove points) or delete groups.
+Different groups are plotted in different colours and shapes, with a legend at the bottom of the figure (Figure \ref{fig:scatter_points}.
+The x and y coordinates are returned.
 
-\section{Summary Data}
+\begin{figure}[!h] 
+%\onehalfspacing
+ \includegraphics[width=0.9\textwidth]{fig_scatter_points.pdf} 
+ \caption{ }
+\label{fig:scatter_points}
+\end{figure}
+
+\subsubsection{Histograms}
+metaDigitise prompts the user to enter groups names and then to click on the top corners of each bar.
+bars can subsequently be deleted.
+MetaDigitise then calculates the midpoint and the frequency for each bar.
+
+\subsection{Summary Data}
+From all plot types, metaDigitise summarises the data form the figure to a mean, standard deviation and sample size, for each identified group.
+These are the summary statistical needed to create all relevant statistics for a meta-analysis.
+In the case of scatter plots, metaDigitise also returns the correlation coefficient between the points within each identified group.
+
+%%% Equations for mean and sd
+
+%% pooled stats: 
+% \begin{align}
+% \mu _{X}&={\frac {\sum _{i}{N_{X_{i}}\mu _{X_{i}}}}{\sum _{i}{N_{X_{i}}}}}
+% \\
+% \sigma _{X}&={\sqrt {{\frac {1}{\sum _{i}{N_{X_{i}}-1}}}\left(\sum _{i}{\left[(N_{X_{i}}-1)\sigma _{X_{i}}^{2}+N_{X_{i}}\mu _{X_{i}}^{2}\right]}-\left[\sum _{i}{N_{X_{i}}}\right]\mu _{X}^{2}\right)}}
+% \end{align}
 
@@
 
+\subsection{Plotting data extraction}
+check extraction
+
+\subsection{Editing data extraction}
+Including addition of N later
+
 \end{document}

### 2017-10-30 docs/figures.R

@@ -1,0 +1,97 @@
+
+set.seed(5)
+library(MASS)
+n=50
+plot_dat <- mvrnorm(n,c(0,0.5),matrix(c(1,0.5,0.5,1),ncol=2))
+x <- plot_dat[,1]
+y <- plot_dat[,2]
+id <- gl(2,n)
+idx <- gl(2,n/2)
+means <- tapply(c(x,y), id, mean)
+ses <- tapply(c(x,y), id, function(x) sd(x)/sqrt(length(x)))
+
+setEPS()
+pdf("/Users/joelpick/Dropbox/0_postdoc/10_metaDigitise/docs/fig_plot_type.pdf", height=9, width=9)
+
+par(mfrow=c(2,2), mar=c(6,5,2,1), cex.lab=1.5)
+plot(means, ylim = c(min(means-ses)-0.1,max(means+ses)+0.1), xlim=c(0.5,2.5), xaxt="n", pch=19, cex=2, ylab="variable", xlab="group")
+mtext("A", 3, adj=0, line=0)
+arrows(c(1,2),means+ses, c(1,2), means-ses, code=3, angle=90, length=0.1)
+axis(1,c(1,2),c(1,2))
+boxplot(c(x,y)~id, ylab="variable", xlab="group", col="grey")
+mtext("B", 3, adj=0, line=0)
+plot(x,y, pch=c(19,1)[idx])
+mtext("C", 3, adj=0, line=0)
+hist(x, main="", col="grey")
+mtext("D", 3, adj=0, line=0)
+
+dev.off()
+
+
+png("/Users/joelpick/Dropbox/0_postdoc/10_metaDigitise/docs/scatter_plot.png")
+plot(x~y, pch=19)
+dev.off()
+
+
+
+setEPS()
+pdf("/Users/joelpick/Dropbox/0_postdoc/10_metaDigitise/docs/fig_rotate.pdf", height=5, width=10)
+
+par(mfrow=c(1,2), oma=c(0,0,3,0), cex.lab=1.5)
+internal_redraw("/Users/joelpick/Dropbox/0_postdoc/10_metaDigitise/docs/scatter_rotate.png")
+mtext("A", side=3, outer = TRUE, adj=0.1, cex=2)
+#user_rotate_graph("/Users/joelpick/Dropbox/0_postdoc/10_metaDigitise/docs/scatter_rotate.png")
+
[…]

### 2017-10-30 R/import_metaDigitise.R
Updates to solve some bugs

@@ -49,7 +49,7 @@ order_lists <- function(list, plot_types){
 	   boxplot <- list[which(unlist(plot_types) == "boxplot")]
 	      hist <- list[which(unlist(plot_types) == "histogram")]
 
-	dat_list <- as.list(c(mean_error = mean_error, boxplot=boxplot, hist=hist, scatterplot=scatterplot))
+	dat_list <- list(mean_error = mean_error, boxplot=boxplot, hist=hist, scatterplot=scatterplot)
 
 return(dat_list)
 

### 2017-10-30 R/metaDigitise.R

@@ -38,7 +38,7 @@ process_new_files <- function(dir, summary = TRUE) {
 			       setup_calibration_dir(dir)
 			      done_details <- dir_details(dir)
 	     details <- get_notDone_file_details(dir)
-		   type <- user_options("Are all plot types the same? (diff/same)" , c("diff", "same"))
+		   type <- user_options("Are all plot types the same? (diff/same)\n" , c("diff", "same"))
 	
 	if(length(done_details$calibrations) >= 1){	
 		done_objects <- load_metaDigitise(done_details$doneCalFiles, done_details$names)
@@ -108,7 +108,7 @@ extract_digitised <- function(list, summary = TRUE) {
 		return(data)
 	} else {
 		tmp <- lapply (list, function(x) x$processed_data)
-		names(tmp) <- unlist(lapply(list, function(x) x$image_file))
+		names(tmp) <- unlist(lapply(list, function(x) filename(x)))
 		return(tmp)
 	}
 }

### 2017-10-31 R/metaDigitise.R
fixed extract_digitise bug

@@ -108,7 +108,7 @@ extract_digitised <- function(list, summary = TRUE) {
 		return(data)
 	} else {
 		tmp <- lapply (list, function(x) x$processed_data)
-		names(tmp) <- unlist(lapply(list, function(x) filename(x)))
+		names(tmp) <- unlist(lapply(list, function(x) filename(x$image_file)))
 		return(tmp)
 	}
 }

### 2017-10-31 R/redraw.R
fixed redraw bug

@@ -123,12 +123,12 @@ internal_redraw <- function(image_file, flip=FALSE, rotate=0, plot_type=NULL, va
 	op <- par(mar=c(3,0,2,0))
 
 	image <- magick::image_read(image_file)
-	new_image <- redraw_rotation(image=image, flip=flip, rotate=rotate)
+	if(rotation) image <- redraw_rotation(image=image, flip=flip, rotate=rotate)
 
-	image_details <- c(width = magick::image_info(new_image)["width"][[1]], height = magick::image_info(new_image)["height"][[1]])
+	image_details <- c(width = magick::image_info(image)["width"][[1]], height = magick::image_info(image)["height"][[1]])
 
 
-	plot(new_image)
+	plot(image)
 	mtext(filename(image_file),3, 1)
 
 	if(!is.null(plot_type)) mtext(plot_type,3, 0)

### 2017-10-31 docs/manuscript.tex
manuscript figures update with iris data

@@
  \includegraphics[width=0.9\textwidth]{fig_plot_type.pdf} 
- \caption{Plot types that metaDigitise is designed to extract data from: A) mean and error plot, B) box plot, C) scatter plot and D) histogram. }
+ \caption{Four plot types that metaDigitise is designed to extract data from: A) mean and error plot, B) box plot, C) scatter plot and D) histogram.
+Data is taken from the iris dataset in R.
+A and B are plotted with the whole dataset, C and D are just the data for the species setosa.}
 \label{fig:plot_type}
@@
 
+\begin{figure}[!h] 
+%\onehalfspacing
+ \includegraphics[width=0.9\textwidth]{fig_all_extract.pdf} 
+ \caption{ }
+\label{fig:all_extract}
+\end{figure}
+
 \subsubsection{Box plots}

### 2017-10-31 docs/figures.R

@@ -1,78 +1,116 @@
 
-set.seed(5)
-library(MASS)
-n=50
-plot_dat <- mvrnorm(n,c(0,0.5),matrix(c(1,0.5,0.5,1),ncol=2))
-x <- plot_dat[,1]
-y <- plot_dat[,2]
-id <- gl(2,n)
-idx <- gl(2,n/2)
-means <- tapply(c(x,y), id, mean)
-ses <- tapply(c(x,y), id, function(x) sd(x)/sqrt(length(x)))
+data(iris)
+head(iris)
+data(airquality)
+head(airquality)
+
+#plot(Temp~Wind,airquality, subset = Month %in% c(5,6), pch=c(19,1)[as.factor(airquality$Month)])
+
+cex.mtext = 2
 
 setEPS()
 pdf("/Users/joelpick/Dropbox/0_postdoc/10_metaDigitise/docs/fig_plot_type.pdf", height=9, width=9)
 
 par(mfrow=c(2,2), mar=c(6,5,2,1), cex.lab=1.5)
-plot(means, ylim = c(min(means-ses)-0.1,max(means+ses)+0.1), xlim=c(0.5,2.5), xaxt="n", pch=19, cex=2, ylab="variable", xlab="group")
-mtext("A", 3, adj=0, line=0)
-arrows(c(1,2),means+ses, c(1,2), means-ses, code=3, angle=90, length=0.1)
-axis(1,c(1,2),c(1,2))
-boxplot(c(x,y)~id, ylab="variable", xlab="group", col="grey")
-mtext("B", 3, adj=0, line=0)
-plot(x,y, pch=c(19,1)[idx])
-mtext("C", 3, adj=0, line=0)
-hist(x, main="", col="grey")
-mtext("D", 3, adj=0, line=0)
+
+means <- with(iris,tapply(Sepal.Length, Species, mean))
+ses <- with(iris,tapply(Sepal.Length, Species, function(x) sd(x)/sqrt(length(x))))*1.96
+
+plot(means, ylim = c(min(means-ses)-0.1,max(means+ses)+0.1), xlim=c(0.5,3.5), xaxt="n", pch=19, cex=2, ylab="Sepal Length", xlab="Species")
+mtext("A", 3, adj=0, line=0, cex=cex.mtext)
+arrows(1:length(means),means+ses, 1:length(means), means-ses, code=3, angle=90, length=0.1)
+axis(1,1:length(means),names(means))
+
[…]

### 2017-10-31 R/redraw.R
testing stuff

@@ -93,7 +93,7 @@ redraw_points <- function(plot_type,raw_data,image_details){
 			bar_data <- subset(raw_data, bar==i)
 			points(y~x,bar_data, pch=19, col="red")
 			lines(y~x, bar_data, lwd=2, col="red")
-			text(mean(bar_data$x),mean(bar_data$y)+legend_pos,bar_data$bar[1], col="red", cex=0.5)
+			text(mean(bar_data$x),mean(bar_data$y)+legend_pos,bar_data$bar[1], col="red", cex=1)
 		}
 	}
 }

### 2017-10-31 docs/manuscript.tex
manuscript editing

@@
 
-\title{Flexible, high throughput data extraction from primary literature: The metaDigitise R package}
+\title{Reproducible, flexible and high throughput data extraction from primary literature: The metaDigitise R package}
 \author{? Joel L. Pick ?...? \& Daniel W.A. Noble ?}
@@
 If this process has already been started when the user starts metaDigitise, then it starts where the user left off.
-Similarly if the user adds new images to the parent directory after all other images are extracted (or part way through), the new images will be extracted from.
+Similarly if the user adds new images to the parent directory after all other images are extracted (or part way through), the new images will be integrated into the extraction procedure.
 After extraction (or upon quitting the function), the extracted data will be returned.
@@
 By clicking two points on the x-axis, metaDigitse calculates the angle needed to rotate the image so the x-axis is horizontal, and rotates it.
-(Figure \ref{fig:rotate} )
+(Figure \ref{fig:rotate}A,B)
 
@@
  \includegraphics[width=0.9\textwidth]{fig_rotate.pdf} 
- \caption{ }
+ \caption{Figure rotation. A) and B) show how non-aligned imaged can be realigned through user defined rotation. C) and D) show show figures can be re-orientated so as to aid data input.}
 \label{fig:rotate}
@@
 Furthermore, in some figures mean and error, boxplots or histograms may be presented with horizontal bars.
-metaDigitise assumes that the bars are vertical, but allows the user to flip the image so that the bars are vertical (Figure \ref{fig:flip}).
-
-\begin{figure}[!h] 
-%\onehalfspacing
- \includegraphics[width=0.9\textwidth]{fig_flip.pdf} 
- \caption{ }
-\label{fig:flip}
-\end{figure}
+metaDigitise assumes that the bars are vertical, but allows the user to flip the image so that the bars are vertical (Figure \ref{fig:rotate}C,D).
 
@@
  \includegraphics[width=0.9\textwidth]{fig_calibrate.pdf} 
- \caption{ }
+ \caption{Axis calibration. The user defines two points on each axis and labels them according to the values shown in the figure.}
 \label{fig:calibrate}
@@
 
+\begin{figure}[!h] 
+%\onehalfspacing
+ \includegraphics[width=0.9\textwidth]{fig_all_extract.pdf} 
+ \caption{Demonstration of data extraction from different plot types}
+\label{fig:all_extract}
+\end{figure}
+
 \subsubsection{Mean and error plots} 
@@
 metaDigitise assumes that the error bars are symmetrical.
-Where the user has clicked the error is displayed in a different colour to the mean.
+Where the user has clicked the error is displayed in a different colour to the mean (Figure \ref{fig:all_extract}A).
 The user can subsequently add more groups or remove groups.
-
-\begin{figure}[!h] 
-%\onehalfspacing
- \includegraphics[width=0.9\textwidth]{fig_all_extract.pdf} 
- \caption{ }
-\label{fig:all_extract}
-\end{figure}
 
@@
 The user can subsequently add groups, edit groups (add or remove points) or delete groups.
-Different groups are plotted in different colours and shapes, with a legend at the bottom of the figure (Figure \ref{fig:scatter_points}.
+Different groups are plotted in different colours and shapes, with a legend at the bottom of the figure (Figure \ref{fig:all_extract}C).
 The x and y coordinates are returned.
-
-\begin{figure}[!h] 
-%\onehalfspacing
- \includegraphics[width=0.9\textwidth]{fig_scatter_points.pdf} 
- \caption{ }
-\label{fig:scatter_points}
-\end{figure}
 
@@
 
+\section{Example}
+%% scatterplot from airquality data
+
+\section{Testing}
+%% simulated data
+%% x people doing y plots
+
 \end{document}

### 2017-10-31 docs/figures.R

@@ -7,6 +7,11 @@ head(airquality)
 #plot(Temp~Wind,airquality, subset = Month %in% c(5,6), pch=c(19,1)[as.factor(airquality$Month)])
 
 cex.mtext = 2
+setosa <- subset(iris, Species == "setosa")
+
+
+
+## first plot
 
 setEPS()
 pdf("/Users/joelpick/Dropbox/0_postdoc/10_metaDigitise/docs/fig_plot_type.pdf", height=9, width=9)
@@ -25,8 +30,6 @@ boxplot(Sepal.Length ~ Species, iris, ylab="Sepal Length", xlab="Species", col="
 mtext("B", 3, adj=0, line=0, cex=cex.mtext)
 
 
-setosa <- subset(iris, Species == "setosa")
-
 set.seed(5)
 plot(Sepal.Length~jitter(Petal.Length),setosa, pch=19, ylab="Sepal Length", xlab="Petal Length")
 mtext("C", 3, adj=0, line=0, cex=cex.mtext)
@@ -35,6 +38,10 @@ hist(setosa$Sepal.Length,, main="", col="grey", xlab="Sepal Length")
 mtext("D", 3, adj=0, line=0, cex=cex.mtext)
 
 dev.off()
+
+
+
+
 
 ### make pngs
 png("/Users/joelpick/Dropbox/0_postdoc/10_metaDigitise/docs/images/iris_mean_error.png")
@@ -60,13 +67,15 @@ hist(setosa$Sepal.Length,, main="", col="grey", xlab="Sepal Length")
 dev.off()
 
 
+png("/Users/joelpick/Dropbox/0_postdoc/10_metaDigitise/docs/images/iris_boxplot_flip.png")
+par(mar=c(6,5,1,1), cex.lab=1.5)
+boxplot(Petal.Width ~ Species, iris, ylab="Petal Width", xlab="Species", col="grey", range=0, horizontal=TRUE)
+dev.off()
+
+
+
 
 dat <- metaDigitise("/Users/joelpick/Dropbox/0_postdoc/10_metaDigitise/docs/images/")
-
-
-
-
-
 
 object_scatter <- readRDS("/Users/joelpick/Dropbox/0_postdoc/10_metaDigitise/docs/images/caldat/iris_scatter_rotate")
[…]

### 2017-10-31 R/metaDigitise.R
Bug fix

@@ -43,6 +43,8 @@ process_new_files <- function(dir, summary = TRUE) {
 	if(length(done_details$calibrations) >= 1){	
 		done_objects <- load_metaDigitise(done_details$doneCalFiles, done_details$names)
 		done_plot_types <- lapply(done_objects, function(x) x$plot_type)
+	} else{
+		done_plot_types = FALSE
 	}
 
 	 plot_types <-  if (type == "diff") {NULL} else { specify_type() }
@@ -59,10 +61,10 @@ process_new_files <- function(dir, summary = TRUE) {
 	
 		complete_plot_types <- lapply(data_list, function(x) x$plot_type)
 
-		if( length(done_plot_types) > 1){
-			plot_type <- c(done_plot_types, complete_plot_types)
-		} else{
+		if( done_plot_types == FALSE){
 			plot_type <- complete_plot_types
+		} else{
+			plot_type <- c(done_plot_types, complete_plot_types)
 		}
 
 	if(summary == TRUE){

### 2017-10-31 R/metaDigitise.R
Trying to fix bug

@@ -2,11 +2,8 @@
 #' @description Single or batch processing of figures with .png, .jpg, .tiff, .pdf extensions within a set directory. metaDigitise consolidates the data and exports the data for each image and image type. It can also summarise the data, provide the raw data (if scatterplots) and automatically imports previously finished data and merges it with newly digitised data. metaDigitise also allows users to check their calibration along with editing previous digitisations.
 #' @param dir the path name to the directory / folder where the files are located
 #' @param summary whether the digitised data should be returned as a summary (TRUE) or as a concatenated list of similar types. 
-#' @details 
-#' metaDigitise can be used on a directory with a whole host of different figure types (mean and error, scatter plots, box plots and histograms) and file types (.jpeg, .png, .tiff, .pdf). It will automatically cycle through all files within a directory in order, prompting the user for specific information as they go. It will automatically write metaDigitise object files (in .RDS format containing processed and calibration data along with directory and file details), into a special caldat/ folder within the directory. Importantly, as new files are added to a directory that has already been "completed", metaDigitise will recognize these unfinished files and only cycle through the digitisation of these new files. 
-#' 
-#' metDigitise is built for reproducibility and ease of operation. This means that, so long as the caldat/ folder along with respective images are maintained, anyone using metaDigitise can simply import existing digitisations, modify them and fix them. 
-#' @author Joel Pick - joel.l.pick@gmail.com; Daniel Noble - daniel.wa.noble@gmail.com
[…]

### 2017-10-31 R/import_metaDigitise.R
Big squishing some more in import digitise AND process new

@@ -8,6 +8,9 @@
 import_metaDigitise <- function(dir, summary = TRUE ) {
 	
 	    details <- dir_details(dir)
+
+	if(length(details$doneCalFiles) == 0) stop("No digitised files to import!")
+
 	    metaDig <- load_metaDigitise(details$doneCalFiles, details$name)
 	 plot_types <- lapply(metaDig, function(x) x$plot_type)
 

### 2017-10-31 R/metaDigitise.R

@@ -41,6 +41,7 @@ process_new_files <- function(dir, summary = TRUE) {
 		done_objects <- load_metaDigitise(done_details$doneCalFiles, done_details$names)
 		done_plot_types <- lapply(done_objects, function(x) x$plot_type)
 	} else{
+		done_objects = 0
 		done_plot_types = 0
 	}
 
@@ -69,9 +70,15 @@ process_new_files <- function(dir, summary = TRUE) {
 		return(do.call(rbind, list(summary(done_objects), summary(data_list))))
 
 	}else{
-			done_figs <- extract_digitised(done_objects,  summary = summary)
-			new_figs <- extract_digitised(data_list, summary = summary)
-		return(order_lists(c(done_figs, new_figs), plot_types = plot_type))
+		if(done_objects == 0){
+				done_figs <- extract_digitised(done_objects,  summary = summary)
+				new_figs <- extract_digitised(data_list, summary = summary)
+			return(order_lists(c(done_figs, new_figs), plot_types = plot_type))
+			
+		} else{
+				new_figs <- extract_digitised(data_list, summary = summary)
+				return(order_lists(new_figs, plot_types = plot_type))
+		}
 	}
 
 }

### 2017-10-31 R/import_metaDigitise.R
Bugs

@@ -52,7 +52,7 @@ order_lists <- function(list, plot_types){
 	   boxplot <- list[which(unlist(plot_types) == "boxplot")]
 	      hist <- list[which(unlist(plot_types) == "histogram")]
 
-	dat_list <- list(mean_error = mean_error, boxplot=boxplot, hist=hist, scatterplot=scatterplot)
+	dat_list <- purrr::compact(list(mean_error = mean_error, boxplot=boxplot, hist=hist, scatterplot=scatterplot))
 
 return(dat_list)
 

### 2017-10-31 R/metaDigitise.R

@@ -41,8 +41,8 @@ process_new_files <- function(dir, summary = TRUE) {
 		done_objects <- load_metaDigitise(done_details$doneCalFiles, done_details$names)
 		done_plot_types <- lapply(done_objects, function(x) x$plot_type)
 	} else{
-		done_objects = 0
-		done_plot_types = 0
+		done_objects = NULL
+		done_plot_types = NULL
 	}
 
 	 plot_types <-  if (type == "diff") {NULL} else { specify_type() }
@@ -59,10 +59,10 @@ process_new_files <- function(dir, summary = TRUE) {
 	
 		complete_plot_types <- lapply(data_list, function(x) x$plot_type)
 
-		if( done_plot_types >= 1){
-			plot_type <- c(done_plot_types, complete_plot_types)
-		} else{
+		if( length(done_plot_types) == 0){
 			plot_type <- complete_plot_types
+		} else{
+			plot_type <- c(done_plot_types, complete_plot_types)
 		}
 
 	if(summary == TRUE){
@@ -70,14 +70,14 @@ process_new_files <- function(dir, summary = TRUE) {
 		return(do.call(rbind, list(summary(done_objects), summary(data_list))))
 
 	}else{
-		if(done_objects == 0){
+		if(length(done_objects) == 0){
+				new_figs <- extract_digitised(data_list, summary = summary)
+				return(order_lists(new_figs, plot_types = plot_type))
+				
+		} else{
 				done_figs <- extract_digitised(done_objects,  summary = summary)
 				new_figs <- extract_digitised(data_list, summary = summary)
 			return(order_lists(c(done_figs, new_figs), plot_types = plot_type))
-			
-		} else{
-				new_figs <- extract_digitised(data_list, summary = summary)
-				return(order_lists(new_figs, plot_types = plot_type))
 		}
 	}
 

### 2017-10-31 R/metaDigitise.R
Some major fixes; hopefully this works now and doesn't breal

@@ -40,9 +40,12 @@ process_new_files <- function(dir, summary = TRUE) {
 	if(length(done_details$calibrations) >= 1){	
 		done_objects <- load_metaDigitise(done_details$doneCalFiles, done_details$names)
 		done_plot_types <- lapply(done_objects, function(x) x$plot_type)
+		names <- lapply(done_objects, function(x) filename(x$image_file))
+		names(done_plot_types) <- names
 	} else{
 		done_objects = NULL
 		done_plot_types = NULL
+		names = NULL
 	}
 
 	 plot_types <-  if (type == "diff") {NULL} else { specify_type() }
@@ -66,14 +69,19 @@ process_new_files <- function(dir, summary = TRUE) {
 		}
 
 	if(summary == TRUE){
-
-		return(do.call(rbind, list(summary(done_objects), summary(data_list))))
+		if(length(done_objects) == 0){
+			return(lapply(data_list, function(x) summary(x)))
+		} else{
+			sum_dat <- do.call(rbind, c(lapply(done_objects, function(x) summary(x)), lapply(data_list, function(x) summary(x))))
+			rownames(sum_dat) <- 1:nrow(sum_dat)
+			return(sum_dat)
+		}
 
 	}else{
 		if(length(done_objects) == 0){
 				new_figs <- extract_digitised(data_list, summary = summary)
-				return(order_lists(new_figs, plot_types = plot_type))
-				
+				return(new_figs)
+
 		} else{
 				done_figs <- extract_digitised(done_objects,  summary = summary)
 				new_figs <- extract_digitised(data_list, summary = summary)

### 2017-10-31 R/H_extract.R
editing group names

@@ -5,16 +5,26 @@
 #' @description Extraction of data from histograms
 histogram_extract <- function(edit=FALSE, raw_data = data.frame(), ...){
 
+	group_id <- readline("\nGroup identifier: ")
+
+	if(edit){ 
+		idQ <- user_options("Change group identifier? (y/n) ",c("y","n"))
+		if(idQ=="y"){
+			group_id <- readline("\nGroup identifier: ")
+			raw_data$id <- group_id
+		}
+	}
+
 	histQ <- if(edit){ "b" }else{ "a" }
 	i <- if(edit){ max(unique(raw_data$bar)) }else{ 0 }
 
-	while(histQ != "c"){
+	while(histQ != "f"){
 		
 		if(histQ=="a"){
 			i=i+1
 			cat("Click on left then right upper corners of bar\n")
 			bar_points <- locator(2, type="o", col="red", lwd=2, pch=19)
-			raw_data <- rbind(raw_data, data.frame(x=bar_points$x,y=bar_points$y, bar=i))
+			raw_data <- rbind(raw_data, data.frame(id=group_id, x=bar_points$x,y=bar_points$y, bar=i))
 		}
 
 		if(histQ=="d"){
@@ -23,7 +33,7 @@ histogram_extract <- function(edit=FALSE, raw_data = data.frame(), ...){
 		}
 
 		internal_redraw(raw_data=raw_data,...)
-		histQ <- readline("Add bar, delete bar or continue? a/d/c ")
+		histQ <- readline("Add bar, Delete bar or Finish plot? a/d/f ")
 
 	}
 	return(raw_data)

### 2017-10-31 R/ME_Box_extraction.R

@@ -6,7 +6,6 @@ single_MB_extract <- function(plot_type){
 	if(plot_type=="mean_error"){
 		cat("\nClick on Error Bar, followed by the Mean\n\n")
 		group_points <- locator(2, type="o", col="red", lwd=2, pch=19)
-		#points(group_points$x[1],group_points$y[1], pch=20, col="yellow")
 	}
 
 	if(plot_type=="boxplot"){
@@ -25,49 +24,25 @@ single_MB_extract <- function(plot_type){
 #' @param raw_data raw data
 #' @param ... further arguments to internal_redraw
 #' @description Extraction of data from boxplots of mean_error plots, from multiple groups
-# MB_extract <- function(plot_type, nGroups,entered_N,...){
-
-# 	nRows <- ifelse(plot_type=="mean_error",2,5)
-# 	raw_data <- as.data.frame(matrix(NA, ncol=4, nrow=nGroups*nRows, dimnames=list(NULL, c("id","x","y","n"))))
-	
-# 	for(i in 1:nGroups) {
-# 		rowStart <- (i-1)*nRows +1
-# 		rows<- rowStart:(rowStart+nRows-1)
-# 		add_removeQ <- "r"
-# 		while(add_removeQ=="r") {
-			
-# 			group_id <- user_unique(paste("\nGroup identifier",i,": "), unique(raw_data$id))
-
-# 			group_N <- if(entered_N){ user_count("\nGroup sample size: ") }else{ NA }
-
-# 			group_points <- single_MB_extract(plot_type)
-			
-# 			raw_data[rows,"id"] <- group_id
-# 			raw_data[rows,"x"] <- group_points$x
-# 			raw_data[rows,"y"] <- group_points$y
-# 			raw_data[rows,"n"] <- group_N
-
-# 			internal_redraw(plot_type=plot_type, raw_data=raw_data, ...)
-
-# 			if(plot_type=="boxplot" & group_points$y[1]<group_points$y[5]) warning("max is smaller than min", call. = FALSE, immediate. = TRUE)
-
-# 			add_removeQ <- user_options("\nReclick or Continue? r/c ", c("c","r"))
-			
-# 			if(add_removeQ=="r") {
-# 				internal_redraw(plot_type=plot_type, raw_data=raw_data[-rows,], ...)
-# 				raw_data[rows,"id"] <- NA
-# 			}		
-# 		}
-# 	}
-# 	return(raw_data)
-# }
 
 
[…]

### 2017-10-31 R/import_metaDigitise.R
Fixed a bug to make sure output is consistent

@@ -11,12 +11,13 @@ import_metaDigitise <- function(dir, summary = TRUE ) {
 
 	if(length(details$doneCalFiles) == 0) stop("No digitised files to import!")
 
-	    metaDig <- load_metaDigitise(details$doneCalFiles, details$name)
+	    metaDig <- load_metaDigitise(details$doneCalFiles, details$calibrations)
 	 plot_types <- lapply(metaDig, function(x) x$plot_type)
 
 	if(summary == TRUE){
 	
 		import <- do.call(rbind,lapply(metaDig, function(x) summary(x)))
+		rownames(import) <- 1:nrow(import)
 	
 	} else{
 		tmp <- lapply(metaDig, function(x) x$processed_data)

### 2017-10-31 R/metaDigitise_functions.R
Updating documentation

@@ -3,6 +3,7 @@
 #' @description Extracts points from a single figure and processes data
 #' @param image_file Image file
 #' @param plot_type Type of plot from "mean_error","boxplot","scatterplot" or"histogram". Function will prompt if not entered by user.
+#' @param cex point size for replotting
 #' @return List of 
 #' @author Joel Pick
 #' @export

### 2017-11-03 R/conversion.R
log capabilities in conversion.R

@@ -2,10 +2,16 @@
 #' @param raw_data The raw data
 #' @param calpoints The calibration points
 #' @param point_vals The point values
+#' @param xlog whether x is logged
+#' @param ylog whether y is logged
 #' @param ... further arguments passed to or from other methods
 #' @description Converts x and y coordinates from original plot coords to actual coords using previous identified coordinates. Modified from digitise package
 
-calibrate <- function(raw_data, calpoints, point_vals,...) {
+calibrate <- function(raw_data, calpoints, point_vals, xlog=FALSE, ylog=FALSE, ...) {
+	
+	if(ylog) point_vals[1:2] <- log(point_vals[1:2])
+	if(xlog) point_vals[3:4] <- log(point_vals[3:4])
+
 	cy <- lm(formula = point_vals[1:2] ~ calpoints$y[1:2])$coeff
  	raw_data$y <- raw_data$y * cy[2] + cy[1]
 
@@ -15,6 +21,10 @@ calibrate <- function(raw_data, calpoints, point_vals,...) {
 	}else{
 		raw_data$x <- raw_data$x 
 	}
+
+	if(ylog) raw_data$y <- exp(raw_data$y)
+	if(xlog) raw_data$x <- exp(raw_data$x)
+
 	return(raw_data)
 }
 

### 2017-11-03 docs/figures.R

@@ -9,12 +9,13 @@ head(airquality)
 cex.mtext = 2
 setosa <- subset(iris, Species == "setosa")
 
+setwd("/Users/joelpick/Dropbox/0_postdoc/10_metaDigitise")
 
 
 ## first plot
 
 setEPS()
-pdf("/Users/joelpick/Dropbox/0_postdoc/10_metaDigitise/docs/fig_plot_type.pdf", height=9, width=9)
+pdf("./docs/fig_plot_type.pdf", height=9, width=9)
 
 par(mfrow=c(2,2), mar=c(6,5,2,1), cex.lab=1.5)
 
@@ -42,59 +43,65 @@ dev.off()
 
 
 
-
 ### make pngs
-png("/Users/joelpick/Dropbox/0_postdoc/10_metaDigitise/docs/images/iris_mean_error.png")
+png("./docs/images/iris_mean_error.png")
 par(mar=c(6,5,1,1), cex.lab=1.5)
 plot(means, ylim = c(min(means-ses)-0.1,max(means+ses)+0.1), xlim=c(0.5,3.5), xaxt="n", pch=19, cex=2, ylab="Sepal Length", xlab="Species")
 arrows(1:length(means),means+ses, 1:length(means), means-ses, code=3, angle=90, length=0.1)
 axis(1,1:length(means),names(means))
 dev.off()
 
-png("/Users/joelpick/Dropbox/0_postdoc/10_metaDigitise/docs/images/iris_scatter.png")
+png("./docs/images/iris_scatter.png")
 par(mar=c(6,5,1,1), cex.lab=1.5)
 plot(Sepal.Length~jitter(Petal.Length),setosa, pch=19, ylab="Sepal Length", xlab="Petal Length")
 dev.off()
 
-png("/Users/joelpick/Dropbox/0_postdoc/10_metaDigitise/docs/images/iris_boxplot.png")
+png("./docs/images/iris_boxplot.png")
 par(mar=c(6,5,1,1), cex.lab=1.5)
 boxplot(Sepal.Length ~ Species, iris, ylab="Sepal Length", xlab="Species", col="grey", range=0)
 dev.off()
 
-png("/Users/joelpick/Dropbox/0_postdoc/10_metaDigitise/docs/iris_histogram.png")
+png("./docs/iris_histogram.png")
 par(mar=c(6,5,1,1), cex.lab=1.5)
 hist(setosa$Sepal.Length,, main="", col="grey", xlab="Sepal Length")
 dev.off()
[…]

### 2017-11-03 R/metaDigitise.R
import one file

@@ -27,7 +27,7 @@ metaDigitise<-function(dir, summary = TRUE){
 	
 	Q <- menu(c("Process new images", "Import existing data", "Edit existing data"))
 	
-	switch(Q, process_new_files(dir, summary = summary), import_metaDigitise(dir, summary = summary), bulk_edit(dir, summary = summary))
+	switch(Q, process_new_files(dir, summary = summary), import_menu(dir, summary = summary), bulk_edit(dir, summary = summary))
 
 }
 

### 2017-11-03 R/graph_rotation.R
scaling points on the plot to be a standard size

@@ -25,13 +25,10 @@ Otherwise chose continue\n
 	flip=FALSE
 	rotate=0
 
-	internal_redraw(image_file, flip = flip, rotate = rotate)
-
 	rotateQ <- "a"
 	while(rotateQ != "c") {
 		if(rotateQ=="f"){
 			flip <- ifelse(!flip, TRUE, FALSE)
-			internal_redraw(image_file, flip = flip,rotate = rotate)
 		}
 		if(rotateQ=="r"){
 			cat("Click left hand then right hand side of x axis\n")
@@ -44,12 +41,10 @@ Otherwise chose continue\n
 			
 			f <- atan2(y.dist, x.dist) * 180/pi
 			rotate <- rotate + f
-			internal_redraw(image_file, flip = flip, rotate = rotate)	
 		}
+		internal_redraw(image_file, flip = flip, rotate = rotate)	
 		rotateQ <- readline("Flip, rotate or continue f/r/c ")
 	}
-
-	internal_redraw(image_file, flip = flip, rotate = rotate)
 
 	out <- list(flip=flip, rotate=rotate)
 

### 2017-11-03 R/redraw.R

@@ -18,17 +18,19 @@ redraw_rotation <- function(image, flip, rotate){
 #' @param calpoints The calibration points
 #' @param point_vals The point values
 #' @param image_details image_details
+#' @param scalar amount to scale size points/text/lines according to plot size
 #' @description plots calibration data on graph
 
-redraw_calibration <- function(plot_type, variable, calpoints,point_vals,image_details){
+redraw_calibration <- function(plot_type, variable, calpoints,point_vals,image_details, scalar){
 	x_shift <- image_details["width"]/30
 	y_shift <- image_details["height"]/30
-	text_cex <- 2
+	text_cex <- 2*scalar
+	line_width <- 2*scalar
 	cal_col <- "blue"
 
-	points(calpoints, pch=3, col=cal_col, lwd=2)
+	points(calpoints, pch=3, col=cal_col, lwd=line_width)
 
-	lines(calpoints[1:2,], col=cal_col, lwd=2)
+	lines(calpoints[1:2,], col=cal_col, lwd=line_width)
 	text(calpoints$x[1:2] - rep(x_shift, 2), calpoints$y[1:2], point_vals[1:2], col=cal_col, cex=text_cex)
 	if(plot_type=="histogram"){
 		text(mean(calpoints$x[3:4]), mean(calpoints$y[3:4]) - y_shift*1.5, variable[1], col=cal_col, cex=text_cex)
@@ -37,7 +39,7 @@ redraw_calibration <- function(plot_type, variable, calpoints,point_vals,image_d
 		}
 
 	if(!plot_type %in% c("mean_error","boxplot")){
-		lines(calpoints[3:4,], col=cal_col, lwd=2)
+		lines(calpoints[3:4,], col=cal_col, lwd=line_width)
 		text(calpoints$x[3:4], calpoints$y[3:4] - rep(y_shift, 2), point_vals[3:4], col=cal_col, cex=text_cex)
 		text(mean(calpoints$x[3:4]), mean(calpoints$y[3:4]) - y_shift*1.5, variable[2], col=cal_col, cex=text_cex)
 	}
@@ -49,33 +51,39 @@ redraw_calibration <- function(plot_type, variable, calpoints,point_vals,image_d
 #' @param plot_type plot_type
 #' @param raw_data The raw data
 #' @param image_details image_details
[…]

### 2017-11-03 docs/figures.R

@@ -79,8 +79,11 @@ boxplot(Petal.Length ~ Species, iris, ylab="Sepal Length", xlab="Species", col="
 dev.off()
 
 
+coords <- cal_coords("boxplot")
+vals <- getVals(coords)
+raw <- locator()
 
-
+calibrate(raw,coords,vals)
 
 dat <- metaDigitise("./docs/images/")
 

### 2017-11-03 tests/testthat/test-misc_func.R

@@ -39,19 +39,18 @@ test_that("Checking user_options..", {
 })
  
 
-user_unique_tester_func <- function(...) {
-	with_mock(
-		readline = function(question) "d",
-		user_unique (...)
-		)
-}
+
 
 test_that("Checking user_unique..", {
 	expect_equal(
- 		user_unique_tester_func("question", c("a","b","c")), 
+ 		with_mock(
+			readline = function(question) "d",
+ 			user_unique("question", c("a","b","c"))
+ 		), 
 		"d",
 		info = "user_unique failed")
 })
+
 
 
 user_numeric_tester_func <- function(..., user_entry) {

### 2017-11-03 tests/testthat/test-point_extraction.R
point_extraction testing

@@ -1,0 +1,29 @@
+context("Check point_extraction...")
+
+
+point_extract_tester_func <- function(...) {
+	with_mock(
+		MB_extract = function(...) 1,
+		group_scatter_extract = function(...) 2,
+		histogram_extract = function(...) 3,
+		point_extraction (...)
+		)
+}
+
+test_that("Checking point_extraction..", {
+	expect_equal(
+ 		point_extract_tester_func(object=list(plot_type="mean_error")), 
+		1,
+		info = "user_options failed")
+
+	expect_equal(
+ 		point_extract_tester_func(object=list(plot_type="scatterplot")), 
+		2,
+		info = "user_options failed")
+
+	expect_equal(
+ 		point_extract_tester_func(object=list(plot_type="histogram")), 
+		3,
+		info = "user_options failed")
+})
+ 

### 2017-11-03 docs/manuscript.tex
MS writing

@@
 \geometry{a4paper}  
-\usepackage[parfill]{parskip} 
+%\usepackage[parfill]{parskip} 
 \usepackage{graphicx}	
@@
 \usepackage[none]{hyphenat}
+\usepackage{listings}
 
@@
 
-\title{Reproducible, flexible and high throughput data extraction from primary literature: The metaDigitise R package}
-\author{? Joel L. Pick ?...? \& Daniel W.A. Noble ?}
+\title{Reproducible, flexible and high throughput data extraction from primary literature for meta-analyses: The metaDigitise R package}
+\author{Joel L. Pick, Shinichi Nakagawa \& Daniel W.A. Noble}
 % \date{}
@@
 \doublespacing
-\raggedright
+%\raggedright
 
 \maketitle
-Prepared for: Journal of Statistical Software?
+Prepared for: Journal of Statistical Software
 
@@
 
-Keywords: meta-analysis, data extraction, R
+Keywords: meta-analysis, data extraction, R, reproducibility, figures, images, summary statistics
 
 \clearpage
+\section*{Introduction}
 
-Meta-analysis is becoming increasingly common in many fields.
-It relies foremost on data extracted from primary literature.
-This data if often presented in figures and so needs to be manually extracted.
+Meta-analysis is becoming increasingly common in many research fields.
+They make use of various summary statistics from studies to generate effect size estimates and their sampling error variance to understand whether particular experimental manipulations, or relationships between variables, show strong overall effects and attempt to explain variation among these effects (Nakagawa et al.
+2017).
+Meta-analysis relies foremost on data extracted from primary literature – more specifically statistics that have been reported in the text or tables of research papers.
+However, necessary summary statistics are often presented in figures, and so, these need to be manually extracted using digitsiing programs to be used in the final analysis.
 Although there are several existing tools to perform tasks like this, these tools are not made for this specific purpose (i.e.
-meta-analysis).
-Specifically, they do not differentiate between common types of plot in which are use to present data, meaning they require a large amount of downstream data manipulation.
-Many exist in stand-alone software and so do not allow easy import into commonly used statistical software (such as R).
-These softwares often do not allow for different grouping of points to be made, and frequently do not allow metadata (such as variable names, group names, sample sizes) to be entered at the time of data extraction.
-Furthermore these tools are generally not designed for a high throughput of images.
-Here, we present an interactive R package designed for large scale data extraction from figures, with meta-analysis in mind.
-To this end, we provide tools specific to data extraction from common plot types (mean and error plots, box plots, scatter plots and histograms, see Figure \ref{fig:plot_type}) and functions that allow for the summary statistics needed for downstream analysis to be calculated (e.g.
-mean, standard deviation and sample size).
-%d?
+a meta-analysis).
+Specifically, they do not differentiate between common plot types that are use to present data, meaning that they require a large amount of downstream data manipulation.
+For example, extracting often only provides the raw data or values from plots.
+This means that these data need to be imported into other programs to calculate correlations (if scatter plots) or errors on means (from mean error plots) need to be back calculated by hand to convert them to a common type (e.g., standard errors to standard deviations).
+Digitising programs often do not allow for different grouping of points to be made, and frequently do not integrate metadata (such as variable names, group names and sample sizes) at the time of data extraction.
+Additionally, existing stand-alone software does not allow easy import into commonly used statistical software (such as R), or try to optimize the research pipelines to facilitate editing and reproducibility.
+These are major issues because extracting from figures can be an incredibly time-consuming process.
+Furthermore, meta-analyses can have a major impact on research fields, and should be expanded upon as more research comes to light.
+Having a tool that facilitates reproducibility in meta-analyses will increase transparency and go a long way to resolving the reproducibility crises we are seeing in many fields.
+
+Here, we present an interactive R package, `metaDigitise`, which is designed for large scale data extraction from figures, specifically catering to the the needs of meta-analysts.
+To this end, we provide tools specific to data extraction from common plot types (mean and error plots, box plots, scatter plots and histograms, see Figure \ref{fig:plot_type}).
+`metaDigitise` operates within the R environment making data extraction, analysis and export more streamlined.
+It also provides users with options to conduct the necessary calculations on raw data immediately after extraction so that comparable summary statistics can be obtained quickly.
+Summaries will condense multiple figures into data frames or lists (depending on the type of figures) and these objects can easily be exported from R, or if using the raw data, analysed in any way the user desires.
+Conveniently, when needing to process many figures at different times `metaDigitise` will only import figures not already completed within a directory.
+This makes it easy to add new figures at anytime.
+`metaDigitise` has also been built for reproducibility in mind.
+It has functions that allow users to redraw their digitisations on figures, correct anything and access the raw calibration data which is written automatically for each figure that is digitised into a special folder within the directory.
+This makes sharing figure digitisation and reproducing the work of others simple and easy, and allows meta-analysts to update meta-analyses more easily.
 
@@
 If the process of data extraction has been finished, or the user wants to reimport summary or processed data ...
+
+\begin{verbatim}
+	install.packages("devtools")
+	devtools::install_github("daniel1noble/metaDigitise")
+	library(metaDigitise)
+\end{verbatim}
 
@@
 
-\section{Reproducibility}
+\section{Directory Structure and Reproducibility}
+
+The `metaDigitise` package is quite flexible.
+Users can extract single figures (if this is all they have) using the `metaDigitise` function with a path name to the directory with the file.
+However, often many figures need extracting from a single paper or set of papers.
+`metaDigitise` will also handle these situations seamlessly by simply cycling through all figures within a directory.
+This is useful because it expedites digitising figures as it prevents users from having to constantly specify the directories and / or paths where files are stored.
+`metaDigitise` essentially will bring up each figure within a folder automatically and allow the user to click and enter the relevant information about a figure as they go.
+This information is then all stored in a data frame or list at the end of the process, saving quite a bit of time.
+Users can stop mid-way through a folder by simply exiting after the last plot they have digitised.
+The data from completed figures will automatically be written to the `caldat/` folder for later use and editing, should the user need to do this.
+
+`metaDigitise` can work on a directory with figures (currently .png, .jpg, .tiff, .pdf images can be used) from many different papers and that are of different types.
+However, users can get creative in how they set up the directories of figures to facilitate extraction.
+For example, one might have 3–4 figures from a single paper that need extracting and the user may want to focus on a single paper at a time while the information about a paper is on hand.
+This could be done by simply setting up a file structure as follows and then using `metaDigitise` with path names (i.e., directory) for each papers folder:
+
+\begin{lstlisting}
+	* Main project directory
+		+ Paper1_P1
+			+ Figure1.png
+			+ Figure2.png
+			+ Figure3.png
+		+ Paper2_P2
+			+ Figure1.png
+			+ Figure2.png
+			+ Figure3.png
+\end{lstlisting}
+
+An alternative directory structure (and probably the most flexible) would be to simply have a set of different figures with an informative and relevant naming scheme to make it easy to identify the paper and figure the data come from.
+This cuts out the need to change directories constantly.
+For example the directory structure could look like:
+
+\begin{lstlisting}
+	* Main project directory
+		+ FiguresToExtract
+			+ P1_Figure1_trait1.png
+			+ P1_Figure2_trait2.png
+			+ P1_Figure3_trait3.png
+			+ P2_Figure1_trait1.png
+			+ P2_Figure2_trait2.png
+			+ P2_Figure3_trait3.png
+\end{lstlisting}
+
+The above directory structure is probably the easiest in combination with a clear and unambiguous naming scheme for each figure.
+Even if only figures from a single paper are digitised, one paper at a time, an overall figure directory will work perfectly because `metaDigitise` will only cycle through incomplete figures, so figures can be added at anytime.
+
+Nonetheless, how users set up their directory is really up to them.
+However, it is important for users to think carefully about reproducibility at this stage.
+Would they like to share the entire project folder with colleagues?
+or would they prefer to simply share the image folder?
+The answers to these questions are important because relative path names are stored in `metaDigitise` objects, meaning that the directory structure (currently at least) needs to be the same for colleagues to re-load previously digitised objects.
+Therefore, the working directory needs to be set to either the "main project directory" or the "FiguresToExtract" directory and relative path names (e.g., dir = "./FiguresToExtract/" if setting the working directory to "Main project directory") used as opposed to absolute paths.
+This facilitates reproducibility should colleagues want to see the digitisations or integrate previously digitised figures with newly updated figures.
+All users need to provide them with is the project or image folder containing the caldat/ folder and the images.
 
 \subsection{Plotting data extraction}
-check extraction
+
+%check extraction
 
 \subsection{Editing data extraction}
-Including addition of N later
+
+%Including addition of N later
 

### 2017-11-03 docs/manuscript.tex
Writings MS

@@
 
-\title{Reproducible, flexible and high throughput data extraction from primary literature for meta-analyses: The metaDigitise R package}
+\title{Reproducible, flexible and high throughput data extraction from primary literature: The metaDigitise R package}
 \author{Joel L. Pick, Shinichi Nakagawa \& Daniel W.A. Noble}
@@
 
-Keywords: meta-analysis, data extraction, R, reproducibility, figures, images, summary statistics
+Keywords: meta-analysis, comparative analysis, data extraction, R, reproducibility, figures, images, summary statistics
 
@@
 
-\section{Functionality}
-There is one main function in the package metaDigitise, which interactively takes the user through the process of extracting figures from graphs.
-
-\subsection{Bulk Digitisation}
-metaDigitise was created with the idea that the user would have multiple images to extract from.
-It therefore operates in the same way whether the user has one or multiple images.
-It assumes that the user has put all the figures needed for extraction in one parent directory.
-It first creates a new directory `caldat' inside this parent directory, in which calibration files are saved (see below).
-
-It then prompts the user to select whether they want to extract from new images, import previously extracted data or edit previously extracted data.
-Extracting from new images takes the user through the steps outlined below, figure by figure, giving the user the option to quit the process after each figure.
-If this process has already been started when the user starts metaDigitise, then it starts where the user left off.
-Similarly if the user adds new images to the parent directory after all other images are extracted (or part way through), the new images will be integrated into the extraction procedure.
-After extraction (or upon quitting the function), the extracted data will be returned.
-%% explain summary etc
-
-If the process of data extraction has been finished, or the user wants to reimport summary or processed data ...
-
-\begin{verbatim}
-	install.packages("devtools")
-	devtools::install_github("daniel1noble/metaDigitise")
-	library(metaDigitise)
-\end{verbatim}
-
-\subsection{Figure Rotation}
-Figure may have been extracted from old publications, for example from scanned images, and so are not perfectly orientated on the image.
-This will make the calibration of the points in the figure from the image problematic.
-metaDigitise allows users to rotate the image.
-By clicking two points on the x-axis, metaDigitse calculates the angle needed to rotate the image so the x-axis is horizontal, and rotates it.
-(Figure \ref{fig:rotate}A,B)
-
-\begin{figure}[!h] 
-%\onehalfspacing
- \includegraphics[width=0.9\textwidth]{fig_rotate.pdf} 
- \caption{Figure rotation. A) and B) show how non-aligned imaged can be realigned through user defined rotation. C) and D) show show figures can be re-orientated so as to aid data input.}
-\label{fig:rotate}
-\end{figure}
-
-Furthermore, in some figures mean and error, boxplots or histograms may be presented with horizontal bars.
-metaDigitise assumes that the bars are vertical, but allows the user to flip the image so that the bars are vertical (Figure \ref{fig:rotate}C,D).
-
-\subsection{Calibration}
-metaDigitise requires the user to calibrate the axes in the figure.
-To do this the user is required to click on two known points on the axis in question, and then enter the value of those points in the figure (Figure \ref{fig:calibrate}).
-Using this information, metaDigitise then calculates the value of any clicked points in terms of the figure axes.
-In the case of mean and error plots and box plots, it calibrates only the y axis (assuming the x axis is redundant).
-For scatter plots and histograms both axes are calibrated.
-
-\begin{figure}[!h] 
-%\onehalfspacing
- \includegraphics[width=0.9\textwidth]{fig_calibrate.pdf} 
- \caption{Axis calibration. The user defines two points on each axis and labels them according to the values shown in the figure.}
-\label{fig:calibrate}
-\end{figure}
-
-\subsection{Plot Types}
-metaDigitise recognises 4 main types of plot; Mean and error plots, box plots, scatter plots and histograms, shown in Figure \ref{fig:plot_type}.
-
-\begin{figure}[!h] 
-%\onehalfspacing
- \includegraphics[width=0.9\textwidth]{fig_all_extract.pdf} 
- \caption{Demonstration of data extraction from different plot types}
-\label{fig:all_extract}
-\end{figure}
-
-\subsubsection{Mean and error plots} 
-metaDigitise prompts the user to enter groups names and allows the user to enter sample sizes, which are used in downstream processing.
-The user is then prompted to click on an error bar followed by the mean.
-Error bars above or below the mean can be clicked - sometimes one is clearer than the other.
-metaDigitise assumes that the error bars are symmetrical.
-Where the user has clicked the error is displayed in a different colour to the mean (Figure \ref{fig:all_extract}A).
-The user can subsequently add more groups or remove groups.
-
-\subsubsection{Box plots}
- metaDigitise prompts the user to enter groups names and allows the user to enter sample sizes, which are used in downstream processing.
-The user is then prompted to click on the maximum, upper quartile, median, lower quartile and minimum.
-metaDigitise will check that the maximum is greater than the minimum, and return an warning if that is not the case.
-The user can subsequently add more groups or remove groups.
-
-\subsubsection{Scatter plots}
- metaDigitise prompts the user to enter groups names and then to click on points.
-Points added by mistake can be deleted.
-The user can subsequently add groups, edit groups (add or remove points) or delete groups.
-Different groups are plotted in different colours and shapes, with a legend at the bottom of the figure (Figure \ref{fig:all_extract}C).
-The x and y coordinates are returned.
-
-\subsubsection{Histograms}
-metaDigitise prompts the user to enter groups names and then to click on the top corners of each bar.
-bars can subsequently be deleted.
-MetaDigitise then calculates the midpoint and the frequency for each bar.
-
-\subsection{Summary Data}
-From all plot types, metaDigitise summarises the data form the figure to a mean, standard deviation and sample size, for each identified group.
-These are the summary statistical needed to create all relevant statistics for a meta-analysis.
-In the case of scatter plots, metaDigitise also returns the correlation coefficient between the points within each identified group.
-
-%%% Equations for mean and sd
-
-%% pooled stats: 
-% \begin{align}
-% \mu _{X}&={\frac {\sum _{i}{N_{X_{i}}\mu _{X_{i}}}}{\sum _{i}{N_{X_{i}}}}}
-% \\
-% \sigma _{X}&={\sqrt {{\frac {1}{\sum _{i}{N_{X_{i}}-1}}}\left(\sum _{i}{\left[(N_{X_{i}}-1)\sigma _{X_{i}}^{2}+N_{X_{i}}\mu _{X_{i}}^{2}\right]}-\left[\sum _{i}{N_{X_{i}}}\right]\mu _{X}^{2}\right)}}
-% \end{align}
-
-\section{Directory Structure and Reproducibility}
+\section*{Directory Structure and Reproducibility}
 
 The `metaDigitise` package is quite flexible.
-Users can extract single figures (if this is all they have) using the `metaDigitise` function with a path name to the directory with the file.
-However, often many figures need extracting from a single paper or set of papers.
-`metaDigitise` will also handle these situations seamlessly by simply cycling through all figures within a directory.
-This is useful because it expedites digitising figures as it prevents users from having to constantly specify the directories and / or paths where files are stored.
-`metaDigitise` essentially will bring up each figure within a folder automatically and allow the user to click and enter the relevant information about a figure as they go.
-This information is then all stored in a data frame or list at the end of the process, saving quite a bit of time.
-Users can stop mid-way through a folder by simply exiting after the last plot they have digitised.
-The data from completed figures will automatically be written to the `caldat/` folder for later use and editing, should the user need to do this.
-
-`metaDigitise` can work on a directory with figures (currently .png, .jpg, .tiff, .pdf images can be used) from many different papers and that are of different types.
+There is one main function in the package, metaDigitise, which interactively takes the user through the process of extracting figures from graphs.
+`metaDigitise` can work on a directory with figures (currently .png, .jpg, .tiff, .pdf images can be used) from many different papers and that are of different types (see below).
 However, users can get creative in how they set up the directories of figures to facilitate extraction.
@@
 The answers to these questions are important because relative path names are stored in `metaDigitise` objects, meaning that the directory structure (currently at least) needs to be the same for colleagues to re-load previously digitised objects.
-Therefore, the working directory needs to be set to either the "main project directory" or the "FiguresToExtract" directory and relative path names (e.g., dir = "./FiguresToExtract/" if setting the working directory to "Main project directory") used as opposed to absolute paths.
-This facilitates reproducibility should colleagues want to see the digitisations or integrate previously digitised figures with newly updated figures.
-All users need to provide them with is the project or image folder containing the caldat/ folder and the images.
+Therefore, the working directory needs to be set to either the "main project directory" or the "FiguresToExtract" directory and relative path names (e.g., users should specify dir = "./FiguresToExtract/" if setting the working directory to "Main project directory") used as opposed to absolute paths.
+This facilitates reproducibility should colleagues want to see the digitisations or integrate previously digitised figures with newly digitised figures.
+All users need to provide them with is the project or image folder.
 
-\subsection{Plotting data extraction}
+\section*{High-throughput Digitisation of Figure(s) within a Directory}
+metaDigitise was created with the idea that the user would likely have multiple images to extract from.
+It therefore operates in the same way whether the user has one or multiple images.
+It assumes that the user has put all the figures needed for extraction in one parent directory (see above).
+Users can extract single figures (if this is all they have) using the `metaDigitise` function with a path name to the directory with the file.
+However, often many figures need extracting from a single paper or set of papers.
+`metaDigitise` will also handle these situations seamlessly by simply cycling through all figures within a directory.
+This is useful because it expedites digitising figures by preventing users from having to constantly change directories and / or open new images.
+It operates by recognizing all the figures in a folder and bringing up each figure automatically allowing the user to click and enter relevant information about a figure as they go.
+All this information is then stored in a data frame or list at the end of the process, saving quite a bit of time.
+Users can stop mid-way through a folder by simply exiting after the last plot they have digitised.
+The data from completed figures will automatically be written to a special `caldat/` folder that is created when first executing the metaDigitise function.
+Digitised data along with metadata for figures are stored in caldat/ for later use and editing (see below)
+
+In order to correctly extract data from figures metaDigitise always requires the user to calibrate the axes in the figure.
+To do this, the user is required to click on two known points on the axis in question, and then enter the value of those points in the figure (Figure \ref{fig:calibrate}).
+Using this information, metaDigitise then calculates the value of any clicked points in terms of the figure axes.
+In the case of mean and error plots and box plots, it calibrates only the y axis (assuming the x axis is redundant).
+For scatter plots and histograms both axes are calibrated.
+
+\begin{figure}[!h] 
+%\onehalfspacing
+ \includegraphics[width=0.9\textwidth]{fig_calibrate.pdf} 
+ \caption{Axis calibration. The user defines two points on each axis and labels them according to the values shown in the figure.}
+\label{fig:calibrate}
+\end{figure}
+
+A particularly powerful and flexible aspect of metaDigitise is its ability to recognize figures that are not already completed and only bring up those figures that have not been digitised in subsequent calls of the function.
+This means that new figures can be added at any time as the project develops and previously digitised figures are simply ignore during processing, but their data re-integrated within the final output after new files are completed.
+
+From all plot types, metaDigitise summarises the data from a figure or set of figures as a mean, standard deviation and sample size, for each identified group within the plot (should multiple groups exist).
+These are the summary statistics needed to create many of the relevant effect sizes and sampling error for a meta-analysis.
+In the case of scatter plots, metaDigitise also returns the correlation coefficient between the points within each identified group.
+
+\subsection{Diverse Plot Types}
+metaDigitise recognises four main types of plot; Mean and error plots, box plots, scatter plots and histograms, shown in Figure \ref{fig:plot_type}.
+Each of these can be processed together and integrated into a single output.
+Alternatively, users can keep like figures together and process them separately.
+
+%%% Equations for mean and sd
+%% pooled stats: 
+% \begin{align}
+% \mu _{X}&={\frac {\sum _{i}{N_{X_{i}}\mu _{X_{i}}}}{\sum _{i}{N_{X_{i}}}}}
+% \\
+% \sigma _{X}&={\sqrt {{\frac {1}{\sum _{i}{N_{X_{i}}-1}}}\left(\sum _{i}{\left[(N_{X_{i}}-1)\sigma _{X_{i}}^{2}+N_{X_{i}}\mu _{X_{i}}^{2}\right]}-\left[\sum _{i}{N_{X_{i}}}\right]\mu _{X}^{2}\right)}}
+% \end{align}
+
+\begin{figure}[!h] 
+%\onehalfspacing
+ \includegraphics[width=0.9\textwidth]{fig_all_extract.pdf} 
+ \caption{Demonstration of data extraction from different plot types}
+\label{fig:all_extract}
+\end{figure}
+
+\subsubsection{Mean and error plots} 
+metaDigitise prompts the user to enter groups names and allows the user to enter sample sizes, which are used in downstream processing.
+The user is then prompted to click on an error bar followed by the mean.
+Error bars above or below the mean can be clicked - sometimes one is clearer than the other.
+metaDigitise assumes that the error bars are symmetrical.
+Where the user has clicked the error is displayed in a different colour to the mean (Figure \ref{fig:all_extract}A).
+The user can subsequently add more groups or remove groups.
+
+\subsubsection{Box plots}
+ metaDigitise prompts the user to enter groups names and allows the user to enter sample sizes, which are used in downstream processing.
+The user is then prompted to click on the maximum, upper quartile, median, lower quartile and minimum.
+metaDigitise will check that the maximum is greater than the minimum, and return an warning if that is not the case.
+The user can subsequently add more groups or remove groups.
+
+\subsubsection{Scatter plots}
+ metaDigitise prompts the user to enter groups names and then to click on points.
+Points added by mistake can be deleted.
+The user can subsequently add groups, edit groups (add or remove points) or delete groups.
+Different groups are plotted in different colours and shapes, with a legend at the bottom of the figure (Figure \ref{fig:all_extract}C).
+The x and y coordinates are returned.
+
+\subsubsection{Histograms}
+metaDigitise prompts the user to enter groups names and then to click on the top corners of each bar.
+bars can subsequently be deleted.
+MetaDigitise then calculates the midpoint and the frequency for each bar.
+
+\subsection{Re-importing, Editing and Plotting Previously Digitised data}
+
+A particularly useful feature of `metaDigitise` is its ability to re-import, edit and re-plot previously digitised figures.
+Running metaDigitise brings up a series of prompts for the user using a main menu that provides access to a number of its features:
+
+\begin{verbatim}
+	Do you want to...
+1: Process new images
+2: Import existing data
+3: Edit existing data
+Selection: 
+\end{verbatim}
+
+The user simply enters in the numeric value that corresponds to what they would like to do.
+Choosing "Process new images" will permit users to digitise new figures, integrating previously completed digitisations along with newly digitised data together at the end of the session, or when the user decides to quit.
+If the user chooses "Import existing data", they have the option of either 1) importing all data or 2) importing a particular figure that has been digitised.
+If 2, then a list of files are provided to the user that they can select.
+
+Editing existing data allows users to easily re-plot or edit information or digitisations that have previously be done for any plot.
+This is accomplished by guiding the user through a new set of options:
+
+\begin{verbatim}
+Choose how you want to edit files:
+1: Cycle through images
+2: Choose specific file to edit
+3: Enter previously omitted sample sizes
+Selection: 
+\end{verbatim}
+
+If the user is unsure about the name of the specific figure they need to edit or simply want to just check the digitisations of figures they can choose "Cycle through images", which will bring up each figure, one by one, overlaying the calibrations, group names (if they exist), sample sizes (if they were entered) and the selected points.
+Alternatively, choosing option 2, will bring up a list of the completed files in the folder and the specific file can be chosen, at which point it will be replotted.
+Either of these options will cycle through a number of questions asking the user what they would like to edit:
+
+\begin{verbatim}
+Edit rotation? If yes, then the whole extraction will be redone (y/n) n
+Change plot type? If yes, then the whole extraction will be redone (y/n) n
+Variable entered as: Sepal length
+Rename Variables (y/n) n
+Edit calibration? (y/n) n
+Re-extract data (y/n) y
+Change group identifier? (y/n) n
+Add group, Delete group or Finish plot? a/d/f d
+1: setosa
+2: versicolor
+3: viriginica
+Selection: 2
+Add group, Delete group or Finish plot? a/d/f a
+\end{verbatim}
+
+A whole host of information can be edited from the rotation, plot type, the variable name that was provided, the calibration and even the digitisation of groups.
+When editing the metaDigitise object is re-written to the caldat folder and the edits are immediately integrated into the existing object once complete.
+
+\section{Additional Features}
+
+\subsection{Figure Rotation and Adjustment}
+Figure may have been extracted from old publications, for example from scanned images, and so are not perfectly orientated on the image.
+This will make the calibration of the points in the figure from the image problematic.
+metaDigitise allows users to rotate the image.
+By clicking two points on the x-axis, metaDigitse calculates the angle needed to rotate the image so the x-axis is horizontal, and rotates it.
+(Figure \ref{fig:rotate}A,B)
+
+\begin{figure}[!h] 
+%\onehalfspacing
+ \includegraphics[width=0.9\textwidth]{fig_rotate.pdf} 
+ \caption{Figure rotation. A) and B) show how non-aligned imaged can be realigned through user defined rotation. C) and D) show show figures can be re-orientated so as to aid data input.}
+\label{fig:rotate}
+\end{figure}
+
+Furthermore, in some figures mean and error, boxplots or histograms may be presented with horizontal bars.
+metaDigitise assumes that the bars are vertical, but allows the user to flip the image so that the bars are vertical (Figure \ref{fig:rotate}C,D).
+
+\subsection{Obtaining Raw Data}
+
+While metaDigitise provides users with the summary statistics by default, for all plot types, in many cases the user may actually be interested in obtaining the raw digitised data from scatter plots.
+This is very easy to do my changing the default `summary` argument from TRUE to FALSE in the metaDigitise function.
+Instead of providing the user with summary statistics it will return a list containing four slots for each of the figure types (mean error, box plot, histogram and scatter plots).
+An example of a data object returned from digitising figures is as follows:
+
+\begin{verbatim}
+str(data)
+List of 3
+ $ mean_error :List of 1
+  ..$ 002_Doe_2013_Fig1.png:'data.frame': 3 obs. of  5 variables:
+  .. ..$ id      : Factor w/ 3 levels "setosa","versicolor",..: 1 2 3
+  .. ..$ mean    : num [1:3] 5 5.93 6.59
+  .. ..$ error   : num [1:3] 0.111 0.148 0.178
+  .. ..$ n       : num [1:3] 50 50 50
+  .. ..$ variable: chr [1:3] "Sepal length" "Sepal length" "Sepal length"
+ $ hist       :List of 1
+  ..$ 003_Doe_2013_Fig3.png:'data.frame': 8 obs. of  3 variables:
+  .. ..$ midpoints: num [1:8] 4.3 4.5 4.7 4.9 5.1 ...
+  .. ..$ frequency: num [1:8] 4 5 7 12 11 6 2 3
+  .. ..$ variable : chr [1:8] "Sepal length" "Sepal length" "Sepal length" "Sepal length" ...
+ $ scatterplot:List of 1
+  ..$ 001_Anderson_1935_Fig1.png:'data.frame':  83 obs. of  8 variables:
+  .. ..$ id        : Factor w/ 2 levels "setosa","versicolor": 1 1 1 1 1 1 1 1 1 1 ...
+  .. ..$ x         : num [1:83] 2.3 2.9 3 3 3 ...
+  .. ..$ y         : num [1:83] 4.5 4.4 4.41 4.3 4.8 ...
+  .. ..$ group     : num [1:83] 1 1 1 1 1 1 1 1 1 1 ...
+  .. ..$ col       : Factor w/ 2 levels "red","green": 1 1 1 1 1 1 1 1 1 1 ...
+  .. ..$ pch       : num [1:83] 19 19 19 19 19 19 19 19 19 19 ...
+  .. ..$ y_variable: chr [1:83] "Sepal length (mm)" "Sepal length (mm)" "Sepal length (mm)" "Sepal length (mm)" ...
+  .. ..$ x_variable: chr [1:83] "Sepal width (mm)" "Sepal width (mm)" "Sepal width (mm)" "Sepal width (mm)" ...
+\end{verbatim}
+
+Here, the user can easily access the list of raw scatter plot data by simply extracting the scatter plot slot:
+
+\begin{verbatim}
+scatterplot <- data$scatterplot
+\end{verbatim}
 
 %check extraction
-
-\subsection{Editing data extraction}
-
 %Including addition of N later
@@
 %% scatterplot from airquality data
+We'll now turn to an example of how metaDigitise works using figures generated using the well known iris data set \ref{fig:all_extract}. Users can install metaDigitise from GitHub as follows:
+
+\begin{verbatim}
+	install.packages("devtools")
+	devtools::install_github("daniel1noble/metaDigitise")
+	library(metaDigitise)
+\end{verbatim}
+
+Assume that we have three figures, from two hypothetical studies, that we would like to extract summary statistics from.
+Each hypothetical study quantifies sepal length and width on a number of iris species.
+
+It then prompts the user to select whether they want to extract from new images, import previously extracted data or edit previously extracted data.
+Extracting from new images takes the user through the steps outlined below, figure by figure, giving the user the option to quit the process after each figure.
+If this process has already been started when the user starts metaDigitise, then it starts where the user left off.
+Similarly if the user adds new images to the parent directory after all other images are extracted (or part way through), the new images will be integrated into the extraction procedure.
+After extraction (or upon quitting the function), the extracted data will be returned.
+%% explain summary etc
 

### 2017-11-07 R/edit_metaDigitise.R
hopefully solved directory issue

@@ -19,6 +19,7 @@ bulk_edit <- function(dir, summary=TRUE){
 	if(Q==1){
 		for(i in filepaths){
 			object <- readRDS(i)
+			object$image_file <- paste0(dir,filename(object$image_file))
 			plot(object)
 			editQ <- user_options("\nEdit file? y/n ", c("y","n"))
 			if(editQ == "y") {
@@ -37,6 +38,7 @@ bulk_edit <- function(dir, summary=TRUE){
 			cat_matrix(files, 3)
 			edit_file <- user_options("\nSelect number of file to edit ", 1:length(files))
 			object <- readRDS(filepaths[as.numeric(edit_file)])
+			object$image_file <- paste0(dir,filename(object$image_file))
 			object <- edit_metaDigitise(object)
 			saveRDS(object, file=filepaths[as.numeric(edit_file)])
 			editQ <- readline("\nEdit more files? y/n ")
@@ -52,6 +54,7 @@ bulk_edit <- function(dir, summary=TRUE){
 			N_files <- filepaths[needed]
 			for(i in N_files){
 				object <- readRDS(i)
+				object$image_file <- paste0(dir,filename(object$image_file))
 				plot(object)
 				object$raw_data <- enter_N(object$raw_data)
 				object$entered_N <- TRUE

### 2017-11-07 R/metaDigitise_functions.R

@@ -11,7 +11,8 @@
 internal_digitise <- function(image_file, plot_type=NULL, cex=1){
 		
 	output <- list()
-	output$image_file <- paste0("./",filename(image_file))
+	output$image_name <- filename(image_file)
+	output$image_file <-image_file
 
 	output$cex <- cex
 

### 2017-11-07 R/edit_metaDigitise.R
fix to enter N

@@ -58,6 +58,7 @@ bulk_edit <- function(dir, summary=TRUE){
 				plot(object)
 				object$raw_data <- enter_N(object$raw_data)
 				object$entered_N <- TRUE
+				object$processed_data <- process_data(object)
 				saveRDS(object, file=i)
 			}
 		}

### 2017-11-07 tests/testthat/test-metaDigitise.R
Adding tests for various metaDigitise functions

@@ -1,0 +1,56 @@
+context("Check various metaDigitise.R functions...")
+
+
+
+test_that("Checking setup_calibration_dir works as expected..", {
+	dir <- paste0(tempdir(), "/")
+	setup_calibration_dir(dir)
+	expect_equal(list.files(dir), "caldat", info = "caldat setup correctly")
+})
+
+test_that("Checking dir_details works as expected..", {
+	dir <- paste0(tempdir(), "/")
+	setup_calibration_dir(dir)
+	list <- dir_details(dir)
+
+	expect_equal(length(list$images), 0, info = "No images in directory as expected..")
+	expect_equal(list$paths, dir, info = "Directory in paths matches..")
+	expect_equal(list$cal_dir, paste0(dir, "caldat/"), info = "Directory in paths to cal_dat folder matches as expected..")
+
+})
+
+test_that("Checking specify_type works as expected..", {
+	with_mock(
+		readline = function(...) "m",
+		expect_equal(specify_type(), "mean_error", info = "specify_type not working correctly does not match mean_error")
+	)
+
+	with_mock(
+		readline = function(...) "s",
+		expect_equal(specify_type(), "scatterplot", info = "specify_type not working correctly does not match scatterplot")
+	)
+
+	with_mock(
+		readline = function(...) "b",
+		expect_equal(specify_type(), "boxplot", info = "specify_type not working correctly does not match boxplot")
+	)
+	
+})
+
+
+mock_object <- list(
+	image_file = "./image.png",
+	flip=FALSE,
+	rotate=0,
+	plot_type="mean_error",
+	variable="y",
+	calpoints = data.frame(x=c(0,0),y=c(0,100)), 
+	point_vals = c(1,2), 
+	entered_N=TRUE,
+	raw_data = data.frame(id=rep("control",2), x=c(60,60), y=c(75,50), n=rep(20,2)),
+	knownN = NULL,
+	error_type="sd",
+	processed_data=data.frame(id=as.factor("control"),mean=1.5, error=0.25, n=20, variable="y", stringsAsFactors = FALSE)	
+	)
+
+class(mock_object) <- 'metaDigitise'

### 2017-11-07 tests/testthat/test-metaDigitise.R
Some test fixes

@@ -32,7 +32,7 @@ test_that("Checking specify_type works as expected..", {
 
 	with_mock(
 		readline = function(...) "b",
-		expect_equal(specify_type(), "boxplot", info = "specify_type not working correctly does not match boxplot")
+		expect_equal(metaDigitise::specify_type(), "boxplot", info = "specify_type not working correctly does not match boxplot")
 	)
 	
 })

### 2017-11-07 tests/testthat/test-metaDigitise.R
More tests for extract digitised

@@ -54,3 +54,18 @@ mock_object <- list(
 	)
 
 class(mock_object) <- 'metaDigitise'
+
+
+test_that("Checking extract_digitised works as expected..", {
+	
+	ob_sum <- extract_digitised(list(mock_object), summary = TRUE)
+	ob_raw <- extract_digitised(list(mock_object), summary = FALSE)
+
+	expect_equal(class(ob_sum), "data.frame", info = "Summary object is indeed a data frame..")
+	expect_equal(ob_sum$mean, 1.5, info = "Summary object mean matches..")
+	expect_equal(class(ob_raw), "list", info = "Raw object is indeed a list..")
+	expect_equal(names(ob_raw), "image.png", info = "Raw object name is correct..")
+	expect_equal(names(ob_raw), "image.png", info = "Raw object name is correct..")
+	expect_equal(ob_raw[[1]]$mean, 1.5, info = "List object mean matches..")
+
+})

### 2017-11-07 tests/testthat/test-import_metaDigitise.R
Added more tests for import functions

@@ -1,0 +1,72 @@
+context("Check import_digitise.R functions...")
+
+mock_object1 <- list(
+	image_file = "./image1.png",
+	flip=FALSE,
+	rotate=0,
+	plot_type="mean_error",
+	variable="y",
+	calpoints = data.frame(x=c(0,0),y=c(0,100)), 
+	point_vals = c(1,2), 
+	entered_N=TRUE,
+	raw_data = data.frame(id=rep("control",2), x=c(60,60), y=c(75,50), n=rep(20,2)),
+	knownN = NULL,
+	error_type="sd",
+	processed_data=data.frame(id=as.factor("control"),mean=1.5, error=0.25, n=20, variable="y", stringsAsFactors = FALSE)	
+	)
+
+class(mock_object1) <- 'metaDigitise'
+
+mock_object2 <- list(
+	image_file = "./image2.png",
+	flip=FALSE,
+	rotate=0,
+	plot_type="mean_error",
+	variable="y",
+	calpoints = data.frame(x=c(0,0),y=c(0,100)), 
+	point_vals = c(1,2), 
+	entered_N=TRUE,
+	raw_data = data.frame(id=rep("control",2), x=c(60,60), y=c(75,50), n=rep(20,2)),
+	knownN = NULL,
+	error_type="sd",
+	processed_data=data.frame(id=as.factor("control"),mean=4, error=2, n=20, variable="y", stringsAsFactors = FALSE)	
+	)
+
+class(mock_object2) <- 'metaDigitise'
+
+list <- list(mock_object1=mock_object1, mock_object2=mock_object2)
+tmp_dir <- paste0(tempdir(), "/")
+dir.create(paste0(tmp_dir, "/caldat/"))
+
+list.files(paste0(tmp_dir, "/caldat/"))
+list.files(tmp_dir)
+
+saveRDS(mock_object1, file = paste0(tmp_dir, "/caldat/", "mock_object1"))
+saveRDS(mock_object2, file = paste0(tmp_dir, "/caldat/", "mock_object2"))
+
+doneCalFiles <- paste0(tmp_dir, "/caldat/", list.files(paste0(tmp_dir, "/caldat/")))
+names <- list("mock_object1", "mock_object2")
+	
+
+test_that("Checking load_metaDigitise works as expected..", {
+	loaded <- load_metaDigitise(doneCalFiles, names)
+	summaries <- lapply(loaded, function(x) summary(x))
[…]

### 2017-11-07 tests/testthat/test-import_metaDigitise.R
Testing input_menu

@@ -70,3 +70,19 @@ test_that("Checking ordered_lists works as expected..", {
 
 	expect_equal(names(list), "mean_error", info = "Checked that plot type names are set up correct in ordered lists")
 })
+
+user_options_tester_func <- function(...) {
+	with_mock(
+		readline = function(question) "a",
+		user_options (...)
+		)
+}
+
+test_that("Checking import_menu works as expected..", {
+	with_mock(
+			menu = function(...) 1,
+			tmp <- import_menu(tmp_dir, summary = TRUE), 
+			expect_equal(tmp$mean[1], 1.5, info="Object 1 loaded and matches mean after import_menu..."),
+			expect_equal(tmp$mean[2], 4, info="Object 2 loaded and matches mean after import_menu...")
+	)
+})

### 2017-11-07 docs/manuscript.tex

@@
 Meta-analysis relies foremost on data extracted from primary literature – more specifically statistics that have been reported in the text or tables of research papers.
-However, necessary summary statistics are often presented in figures, and so, these need to be manually extracted using digitsiing programs to be used in the final analysis.
+However, necessary summary statistics are often presented in figures, and so, these need to be manually extracted using digitising programs to be used in the final analysis.
 Although there are several existing tools to perform tasks like this, these tools are not made for this specific purpose (i.e.
@@
 Specifically, they do not differentiate between common plot types that are use to present data, meaning that they require a large amount of downstream data manipulation.
-For example, extracting often only provides the raw data or values from plots.
+For example, many stand alone programs only provide the raw data or values from plots.
 This means that these data need to be imported into other programs to calculate correlations (if scatter plots) or errors on means (from mean error plots) need to be back calculated by hand to convert them to a common type (e.g., standard errors to standard deviations).
-Digitising programs often do not allow for different grouping of points to be made, and frequently do not integrate metadata (such as variable names, group names and sample sizes) at the time of data extraction.
+Digitising programs also do not easily allow for different grouping of points to be made, and frequently do not integrate metadata (such as variable names, group names and sample sizes) at the time of data extraction.
 Additionally, existing stand-alone software does not allow easy import into commonly used statistical software (such as R), or try to optimize the research pipelines to facilitate editing and reproducibility.
@@
 This facilitates reproducibility should colleagues want to see the digitisations or integrate previously digitised figures with newly digitised figures.
-All users need to provide them with is the project or image folder.
+All users need to achieve this is to provide others with is the project or image folder.
 
@@
 The data from completed figures will automatically be written to a special `caldat/` folder that is created when first executing the metaDigitise function.
-Digitised data along with metadata for figures are stored in caldat/ for later use and editing (see below)
+Digitised data along with metadata for figures are stored in caldat/ for later use and editing (see below).
 
@@
 %% scatterplot from airquality data
-We'll now turn to an example of how metaDigitise works using figures generated using the well known iris data set \ref{fig:all_extract}. Users can install metaDigitise from GitHub as follows:
+We'll now turn to an example of how metaDigitise works using figures generated using the well known iris data set. Users can install metaDigitise from GitHub as follows:
 
 \begin{verbatim}
-	install.packages("devtools")
-	devtools::install_github("daniel1noble/metaDigitise")
-	library(metaDigitise)
+	> install.packages("devtools")
+	> devtools::install_github("daniel1noble/metaDigitise")
+	> library(metaDigitise)
 \end{verbatim}
 
-Assume that we have three figures, from two hypothetical studies, that we would like to extract summary statistics from.
-Each hypothetical study quantifies sepal length and width on a number of iris species.
+Assume that we would like to extract summary statistics from studies measuring sepal length or width in iris species for a fictitious project.
+There are a few studies that only present these data in figures.
+As we read papers found from a systematic search we can add in figures with relevant data to a "figs\_to\_extract" folder as follows
 
-It then prompts the user to select whether they want to extract from new images, import previously extracted data or edit previously extracted data.
-Extracting from new images takes the user through the steps outlined below, figure by figure, giving the user the option to quit the process after each figure.
-If this process has already been started when the user starts metaDigitise, then it starts where the user left off.
-Similarly if the user adds new images to the parent directory after all other images are extracted (or part way through), the new images will be integrated into the extraction procedure.
-After extraction (or upon quitting the function), the extracted data will be returned.
-%% explain summary etc
+\begin{lstlisting}
+	*figs_to_extract
+      + 001_Anderson_1935_Fig1.png
+\end{lstlisting}
+
+\begin{figure}[!h] 
+%\onehalfspacing
+ \includegraphics[width=0.9\textwidth]{001_Anderson_1935_Fig1.png} 
+ \caption{Example scatterplot (001\_Anderson\_1935\_Fig1.png) of sepal length and width for two species of iris (setosa and versicolor)}
+\label{fig:rotate}
+\end{figure}
+
+Here, the naming of the files placed in the folder will contain the paper number, first author and the figure number to keep data uniquely associated with figures.
+We can set our working directory to "figs\_to\_extract" and use metaDigitise to begin processing the image from the first paper
+
+\begin{verbatim}
+	> setwd("~/figs_to_extract")
+	> digitised_data <- metaDigitise(".", summary = TRUE)
+\end{verbatim}
+
+As soon metaDigitise is executed, it will bring the user to the main menu (see "Re-importing, Editing and Plotting Previously Digitised data" above).
+When the user selects option one ("Process new images"), we are asked whether we have different types of plot(s) in the folder.
+This question is most relevant when there are lots of different figures in the folder because it will then ask us for the type of figure as they are cycled through.
+
+\begin{lstlisting}
+	Are all plot types the same? (diff/same)
+\end{lstlisting}
+
+metaDigitise then asks the user whether the figure needs to be rotated or flipped.
+This can be needed when box plots and mean and error plots are not orientated correctly.
+In some cases, older papers can give slightly off angled images which can be corrected by rotating.
+So, in this prompt the user has three options: `f` for "Flip", `r` for "rotate" or `c` for "continue.
+
+\begin{lstlisting}
+	mean_error and boxplots should be vertically orientated
+	       _ 
+	       |  
+	  I.E. o    NOT  |-o-|
+	       |
+	       _
+
+	If they are not then chose flip to correct this.
+
+	If figures are wonky, chose rotate.
+
+	Otherwise chose continue
+
+	Flip, rotate or continue f/r/c c
+'\end{lstlisting}
+
+After this, `metaDigistise` will ask the user to specify the plot type.
+Depending on the figure, the user can specify that it is a figure containing the mean and error (`m`), a box plot (`b`), a scatter plot (`s`) or a histogram (`h`).
+If the user has specified `diff` instead of `same` in response to the question about whether the plot types are the same or different, this question will pop up for each plot, but will only be asked once if plots are all the same.
+
+\begin{lstlisting}
+	Please specify the plot_type as either: mean and error, box plot, scatter plot or histogram m/b/s/h:
+\end{lstlisting}
+
+After selecting the figure type a new set of prompts will come up that will ask the user first what the y and x-axis variables are.
+This is useful as users can keep track of the different variables across figures and papers.
+Here, the user can just add this information in to the R console.
+Once complete, details on how to calibrate the x and y-axis appear, so that the relevant statistics / data can be correctly calculated.
+When working with a plot of mean and standard errors, the x-axis is rather useless in terms of calibration so `metaDigitise` just asks the user to calibrate the y-axis.
+
+\begin{lstlisting}
+What is the y variable? Sepal Length (mm)
+
+What is the x variable? Sepal Width (mm)
+
+On the Figure, click IN ORDER: 
+      y1, y2 , x1, x2
+
+    Step 1 ----> Click on y1
+  |
+  |
+  |
+  |
+  y1
+  |_________________________
+  ....
+
+    Step 3 ----> Click on x1
+  |
+  |
+  |
+  |
+  |
+  |_____x1__________________
+
+  ....
+ \end{lstlisting}
+
+The user can just follow the instructions on screen step-by-step (instructions above have been truncated by `...` to simplify), and in the order specified.
+Before moving on, the user is forced to check whether or not the calibration has been set up correctly.
+If `n` is chosen because something needs to be fixed then the user can re-calibrate.
+
+\begin{lstlisting}
+What is the value of y1 ?
+4.5
+What is the value of y2 ?
+7
+What is the value of x1 ?
+2
+What is the value of x2 ?
+4
+
+Re-calibrate? (y/n) 
+\end{lstlisting}
+
+Often, plots might contain multiple groups that the meta-analyst wants to extract from.
+`metaDigitise` handles this nicely by prompting the user to enter the group first, followed by the digitisation of this groups data.
+After digitising the first group, and having exited (i.e., hit `esc` from plot window), `mteaDigitise` will ask the user whether they would like to add another group.
+Users can continually add groups (`a`), or simply continue to the next plot (`c` - if another plot exists).
+The number of groups are not really limited and users can just keep adding in groups to accommodate the different numbers that may be presented across figures (although it can get complicated with too many).
+
+\begin{lstlisting}
+	Follow instructions below, to exit point adding or removing:
+
+	 - Windows: right click on the plot area and choose 'Stop'!
+
+	 - X11: hit any mouse button other than the left one.
+
+	 - quartz/OS X: hit ESC
+
+	Group identifier: setosa
+
+	Click on points you want to add.
+	If you want to remove a point, or are finished with a
+	group, exit (see above), then follow prompts
+
+	Add points, delete points or continue? a/d/c c
+
+\end{lstlisting}
+
+Once we are done digitising all the groups our plot will look something like this:
+
+\begin{figure}[!h] 
+%\onehalfspacing
+ \includegraphics[width=0.9\textwidth]{001_Anderson_1935_Fig1_digitised.png} 
+ \caption{Digitisation of sepal length and width for two species of iris (setosa and versicolor).
+Names of the variables and calibration (in blue) are plotted alongside the digitised points (green = versicolor; red = setosa).
+The sample sizes for each group are provided on the lower part of the plot.
+All figures are clearly labeled at the top to remind users of the filename and plot type.
+This reduces errors throughout the digitisation process.}
+\label{fig:rotate}
+\end{figure}
+
+When completed metaDigitise will write the digitised data to the caldat directory, such that our new directory structure is as follows
+
+\begin{lstlisting}
+	*figs_to_extract
+	  + caldat
+	  	+ 001_Anderson_1935_Fig1
+      + 001_Anderson_1935_Fig1.png
+\end{lstlisting}
+
+Users can access the metaDigitise object created (001\_Anderson\_1935\_Fig1) at any time using the metaDigitse function.
+In the R console, the summarised data for the digitised figure can be printed on screen or even written to a .csv file:
+
+\begin{lstlisting}
+data
+                   filename   group_id         variable       mean      sd  n         r  plot_type
+1 iris_scatter_multigrp.png     setosa  Sepal Width (mm) 3.420332 0.4034259 39 0.7524909 scatterplot
+2 iris_scatter_multigrp.png     setosa Sepal Length (mm) 4.998267 0.3831294 39 0.7524909 scatterplot
+3 iris_scatter_multigrp.png versicolor  Sepal Width (mm) 2.766857 0.3234723 44 0.5157938 scatterplot
+4 iris_scatter_multigrp.png versicolor Sepal Length (mm) 5.936769 0.5292535 44 0.5157938 scatterplot
+\end{lstlisting}
+
+The mean for each of the two variables, along with the two species, are provided.
+Since this is a scatterplot, the user also gets the Person's correlation coefficient between sepal length and width for each species.
+These match reasonably well with the actual means of sepal length and width for each of the species in the full `iris` dataset:
+
+\begin{lstlisting}
+     Species meanSL meanSW
+1     setosa  5.006  3.428
+2 versicolor  5.936  2.770
+\end{lstlisting}
+
+One thing anyone with a familiarity with the iris dataset will notice is that the sample sizes for each of these species (which are n = 50 each) are quite a bit lower.
+This is an example of some of the challenges when extracting data from scatter plots, often data points will overlap with each other making it impossible (without having the real data) to know whether this is a problem.
+However, a meta-analyst will probably realise that the sample sizes here conflict with what is reported in the paper.
+Hence, metaDigitise also provides the user with  options to input the sample sizes directly (see Editing section above), even for scatter plots and histograms where, strictly speaking, this should not be necessary.
+Nonetheless, it is important to recognise the impact that overlapping points can have on summary statistics, particularly its effects on standard deviation (SD) and standard error (SE).
+Here, the mean point estimates are nearly bang on, but the SD's are slightly over-estimated:
+
+\begin{lstlisting}
+     Species    meanSL    meanSW
+1     setosa 0.3524897 0.3790644
+2 versicolor 0.5161711 0.3137983
+\end{lstlisting}
+
+\subsection{Adding new figures}
+
+Users can add additional figures as new papers with relevant information are found.
+Each figure should be in its own file with unique naming, even if a single paper has multiple figures for extraction.
+For example, another paper on different populations (and one new species) of iris contained two additional figures where important data could be extracted.
+These figures can simply be named accordingly and added directly to the same extraction folder:
+
+\begin{lstlisting}
+	*figs_to_extract
+	  + caldat
+	  	+ 001_Anderson_1935_Fig1
+      + 001_Anderson_1935_Fig1.png
+      + 002_Doe_2013_Fig1.png
+      + 002_Doe_2013_Fig3.png
+\end{lstlisting}
+
+We have already processed one figure (001\_Anderson\_1935\_Fig1.png) and we can tell because it has digitised data (caldat/001\_Anderson\_1935\_Fig1), but now we have our two new figures that have not yet been digitised.
+This example will nicely demonstrate how users can easily pick up from where they left off and how all previous data gets re-integrated.
+It will also demonstrate how different plot types are handled.
+All we have to do to begin, is again, provide the directory where all the figures are located:
+
+\begin{verbatim}
+	> setwd("~/figs_to_extract")
+	> digitised_data <- metaDigitise(".", summary = TRUE)
+\end{verbatim}
+
+The user gets the same set of prompts and simply chooses option one to digitise the new figures.
+This time, 001\_Anderson\_1935\_Fig1.png, is ignored and the new plots cycle on screen.
+First for 002\_Doe\_2013\_Fig1.png and then 002\_Doe\_2013\_Fig3.png.
+Since there are a few new figures, we answer the first question in the R console as "diff":
+
+\begin{lstlisting}
+	Are all plot types the same? (diff/same) diff
+
+**** NEW PLOT ****
+
+mean_error and boxplots should be vertically orientated
+       _ 
+       |  
+  I.E. o    NOT  |-o-|
+       |
+       _
+
+If they are not then chose flip to correct this.
+
+If figures are wonky, chose rotate.
+
+Otherwise chose continue
+
+Flip, rotate or continue f/r/c c
+Please specify the plot_type as either: mean and error, box plot, scatter plot or histogram m/b/s/h: m
+\end{lstlisting}
+
+Here, we specify the new plot type as `m` for 002\_Doe\_2013\_Fig1.png because we have a plot of the mean and error of sepal length for each of the three species.
+We're then prompted a bit differently from our scatter plot as the x-axis is not needed for calibration:
+
+\begin{lstlisting}
+What is the variable? Sepal length
+
+On the Figure, click IN ORDER: 
+      y1, y2
+
+    Step 1 ----> Click on y1
+  |
+  |
+  |
+  |
+  y1
+  |_________________________
+
+    Step 2 ----> Click on y2
+  |
+  y2
+  |
+  |
+  |
+  |_________________________
+
+What is the value of y1 ?
+5
+What is the value of y2 ?
+6.5
+
+Re-calibrate? (y/n) n
+
+Enter sample sizes? y/n y
+
+Group identifier: setosa
+
+Group sample size: 50
+
+Click on Error Bar, followed by the Mean
+
+Add group, Delete group or Finish plot? a/d/f a
+\end{lstlisting}
+Again, metaDigitsie will simply guide the user through digitising each of these figures describing to them exactly what needs to be done.
+At any point if mistakes are made the user can choose relevant options to edit or correct things.
+This can continues for each plot so long as the user would like to continue after completing a single plot:
+
+\begin{lstlisting}
+Do you want continue: 1 plots out of 2 plots remaining (y/n) y
+\end{lstlisting}
+
+This continues until users have completed all non-digitised figures in the folder, at which point metaDigitise concatenates the new data with previously digitised data in the object:
+
+\begin{lstlisting}
+data
+               filename        group_id         variable     mean        sd   n         r    plot_type
+1 001_Anderson_1935_Fig1.png     setosa  Sepal Width (mm) 3.420332 0.4034259 39 0.7524909 scatterplot
+2 001_Anderson_1935_Fig1.png     setosa Sepal Length (mm) 4.998267 0.3831294 39 0.7524909 scatterplot
+3 001_Anderson_1935_Fig1.png versicolor  Sepal Width (mm) 2.766857 0.3234723 44 0.5157938 scatterplot
+4 001_Anderson_1935_Fig1.png versicolor Sepal Length (mm) 5.936769 0.5292535 44 0.5157938 scatterplot
+5      002_Doe_2013_Fig1.png     setosa      Sepal length 5.000336 0.7828656 50        NA  mean_error
+6      002_Doe_2013_Fig1.png versicolor      Sepal length 5.931340 1.0466177 50        NA  mean_error
+7      002_Doe_2013_Fig1.png viriginica      Sepal length 6.588705 1.2608173 50        NA  mean_error
+8      003_Doe_2013_Fig3.png       <NA>      Sepal length 4.948472 0.3624212 50        NA  histogram
+\end{lstlisting}
 
@@
 
+\section{Conclusions and Future Enhancements}
+
+mmetaDigitise is available through GitHub.
+WHile it is incredibly flexible already, our future plans include building in abilities to deal with log-transformed axes, including arguments for calculating standard deviations from 95\% confidence intervals and standard errors using the t-distribution to correct the t-value for small sample sizes (currently assumed t-values are 1.96 as is normal), dealing with asymmetric error bars and the possibility of zooming in plots such that greater accuracy can be achieved when digitising.
+We hope to also provide options for assessing inter-observer reliability in the future.
 \end{document}

### 2017-11-19 docs/manuscript.tex
working on manuscript

@@
-\documentclass{article}
-\usepackage[utf8]{inputenc}
-\usepackage{geometry} 
-\geometry{a4paper}  
-%\usepackage[parfill]{parskip} 
-\usepackage{graphicx}	
-\usepackage{amsmath}
-\usepackage{fullpage}
-\usepackage{setspace} 
-\usepackage{lineno}
-\usepackage[none]{hyphenat}
+\documentclass[article]{jss}
+
+%% -- LaTeX packages and custom commands ---------------------------------------
+
+%% recommended packages
+\usepackage{thumbpdf,lmodern}
+
+% \usepackage[utf8]{inputenc}
+% \usepackage{geometry} 
+% \geometry{a4paper}  
+% %\usepackage[parfill]{parskip} 
+% \usepackage{graphicx}	
+% \usepackage{amsmath}
+% \usepackage{fullpage}
+% \usepackage{setspace} 
+% \usepackage{lineno}
+% \usepackage[none]{hyphenat}
  \usepackage{listings}
+% \usepackage[round]{natbib}
 
-\usepackage[round]{natbib}
+%% new custom commands
+\newcommand{\class}[1]{`\code{#1}'}
+\newcommand{\fct}[1]{\code{#1()}}
 
-\title{Reproducible, flexible and high throughput data extraction from primary literature: The metaDigitise R package}
-\author{Joel L. Pick, Shinichi Nakagawa \& Daniel W.A. Noble}
-% \date{}
+%% -- Article metainformation (author, title, ...) -----------------------------
+
+%% - \author{} with primary affiliation
+%% - \Plainauthor{} without affiliations
+%% - Separate authors by \And or \AND (in \author) or by comma (in \Plainauthor).
+%% - \AND starts a new line, \And does not.
+\author{Joel L. Pick\\University of New South Wales
+   \And Shinichi Nakagawa\\University of New South Wales
+   \AND Daniel W.A. Noble\\University of New South Wales
+   }
+\Plainauthor{Joel L. Pick, Shinichi Nakagawa, Daniel W.A. Noble}
+
+%% - \title{} in title case
+%% - \Plaintitle{} without LaTeX markup (if any)
+%% - \Shorttitle{} with LaTeX markup (if any), used as running title
+\title{Reproducible, flexible and high throughput data extraction from primary literature: The \pkg{metaDigitise} \proglang{R} package}
+\Plaintitle{Reproducible, flexible and high throughput data extraction from primary literature: The metaDigitise R package}
+\Shorttitle{Data extraction with \pkg{metaDigitise}}
+
+%% - \Abstract{} almost as usual
+\Abstract{
+
+}
+
+%% - \Keywords{} with LaTeX markup, at least one required
+%% - \Plainkeywords{} without LaTeX markup (if necessary)
+%% - Should be comma-separated and in sentence case.
+\Keywords{meta-analysis, comparative analysis, data extraction, \proglang{R}, reproducibility, figures, images, summary statistics}
+\Plainkeywords{meta-analysis, comparative analysis, data extraction, R, reproducibility, figures, images, summary statistics}
+
+%% - \Address{} of at least one author
+%% - May contain multiple affiliations for each author
+%%   (in extra lines, separated by \emph{and}\\).
+%% - May contain multiple authors for the same affiliation
+%%   (in the same first line, separated by comma).
+\Address{
+  Joel L. Pick, Shinichi Nakagawa, Daniel W.A. Noble\\
+  Ecology and Evolution Research Center\\
+  \emph{and}\\
+  School of Biological, Earth and Environmental Sciences\\
+  University of New South Wales\\
+  b\\
+  b\\
+  E-mail: \email{joel.l.pick@gmail.com}\\
+  URL: \url{https://eeecon.uibk.ac.at/~zeileis/}
+}
 
 \begin{document}
-\doublespacing
-%\raggedright
 
-\maketitle
-Prepared for: Journal of Statistical Software
-
-\section*{Abstract}
-
-Keywords: meta-analysis, comparative analysis, data extraction, R, reproducibility, figures, images, summary statistics
-
-\clearpage
-\section*{Introduction}
-
-Meta-analysis is becoming increasingly common in many research fields.
-They make use of various summary statistics from studies to generate effect size estimates and their sampling error variance to understand whether particular experimental manipulations, or relationships between variables, show strong overall effects and attempt to explain variation among these effects (Nakagawa et al.
+\section{Introduction}
+In many different contexts, researchers need to make use of data presented in primary literature.
+Most notably, this includes meta-analysis, which is becoming increasingly common in many research fields.
+Meta-analysis uses effect size estimates and their sampling variance, taken from many studies, to understand whether particular effects are common across studies and to explain variation among these effects (Nakagawa et al.
 2017).
-Meta-analysis relies foremost on data extracted from primary literature – more specifically statistics that have been reported in the text or tables of research papers.
-However, necessary summary statistics are often presented in figures, and so, these need to be manually extracted using digitising programs to be used in the final analysis.
-Although there are several existing tools to perform tasks like this, these tools are not made for this specific purpose (i.e.
+Meta-analysis therefore relies foremost on data extracted from primary literature and more specifically statistics that have been reported in the text or tables of research papers.
+These summary statistics are, however, frequently presented in figures and so need to be manually extracted using digitising programs.
+Although there are several existing tools to perform tasks like this (EXAMPLES), these tools are not designed this specific purpose (i.e.
 a meta-analysis).
@@
 Digitising programs also do not easily allow for different grouping of points to be made, and frequently do not integrate metadata (such as variable names, group names and sample sizes) at the time of data extraction.
-Additionally, existing stand-alone software does not allow easy import into commonly used statistical software (such as R), or try to optimize the research pipelines to facilitate editing and reproducibility.
+Additionally, existing stand-alone software does not allow easy import into commonly used statistical software (such as \proglang{R}), or try to optimize the research pipelines to facilitate editing and reproducibility.
 These are major issues because extracting from figures can be an incredibly time-consuming process.
-Furthermore, meta-analyses can have a major impact on research fields, and should be expanded upon as more research comes to light.
-Having a tool that facilitates reproducibility in meta-analyses will increase transparency and go a long way to resolving the reproducibility crises we are seeing in many fields.
+Furthermore, although meta-analysis is an important tool in consolidating the data from many studies, relying on transparency of those studies, many of the processes involved in data extraction are opaque and difficult to reproduce, making extending studies problematic.
+Having a tool that facilitates reproducibility in meta-analyses will increase transparency and go a long way to resolving the reproducibility crises we are seeing in many fields (REF).
 
-Here, we present an interactive R package, `metaDigitise`, which is designed for large scale data extraction from figures, specifically catering to the the needs of meta-analysts.
+Here, we present an interactive \proglang{R} package, \pkg{metaDigitise}, which is designed for large scale data extraction from figures, specifically catering to the the needs of meta-analysts.
 To this end, we provide tools specific to data extraction from common plot types (mean and error plots, box plots, scatter plots and histograms, see Figure \ref{fig:plot_type}).
-`metaDigitise` operates within the R environment making data extraction, analysis and export more streamlined.
+\pkg{metaDigitise} operates within the \proglang{R} environment making data extraction, analysis and export more streamlined.
 It also provides users with options to conduct the necessary calculations on raw data immediately after extraction so that comparable summary statistics can be obtained quickly.
-Summaries will condense multiple figures into data frames or lists (depending on the type of figures) and these objects can easily be exported from R, or if using the raw data, analysed in any way the user desires.
-Conveniently, when needing to process many figures at different times `metaDigitise` will only import figures not already completed within a directory.
-This makes it easy to add new figures at anytime.
-`metaDigitise` has also been built for reproducibility in mind.
-It has functions that allow users to redraw their digitisations on figures, correct anything and access the raw calibration data which is written automatically for each figure that is digitised into a special folder within the directory.
+Summaries will condense multiple figures into data frames or lists (depending on the type of figures) and these objects can easily be exported from \proglang{R}, or if using the raw data, analysed in any way the user desires.
+Conveniently, when needing to process many figures at different times \pkg{metaDigitise} will only import figures not already completed within a directory.
+This makes it easy to add new figures at any time.
+\pkg{metaDigitise} has also been built for reproducibility in mind.
+It has functions that allow users to redraw their digitisations on figures, make corrections and access the raw calibration data which is written automatically for each figure that is digitised into a special folder within the directory.
 This makes sharing figure digitisation and reproducing the work of others simple and easy, and allows meta-analysts to update meta-analyses more easily.
 
-\begin{figure}[ht] 
+\begin{figure}[b!] 
 %\onehalfspacing
  \includegraphics[width=0.9\textwidth]{fig_plot_type.pdf} 
- \caption{Four plot types that metaDigitise is designed to extract data from: A) mean and error plot, B) box plot, C) scatter plot and D) histogram.
+ \caption{Four plot types that \pkg{metaDigitise} is designed to extract data from: A) mean and error plot, B) box plot, C) scatter plot and D) histogram.
 Data is taken from the iris dataset in R.
@@
 
-\section*{Directory Structure and Reproducibility}
+\section{Directory Structure, Image Processing and Reproducibility}
 
-The `metaDigitise` package is quite flexible.
-There is one main function in the package, metaDigitise, which interactively takes the user through the process of extracting figures from graphs.
-`metaDigitise` can work on a directory with figures (currently .png, .jpg, .tiff, .pdf images can be used) from many different papers and that are of different types (see below).
-However, users can get creative in how they set up the directories of figures to facilitate extraction.
-For example, one might have 3–4 figures from a single paper that need extracting and the user may want to focus on a single paper at a time while the information about a paper is on hand.
-This could be done by simply setting up a file structure as follows and then using `metaDigitise` with path names (i.e., directory) for each papers folder:
-
-\begin{lstlisting}
-	* Main project directory
-		+ Paper1_P1
-			+ Figure1.png
-			+ Figure2.png
-			+ Figure3.png
-		+ Paper2_P2
-			+ Figure1.png
-			+ Figure2.png
-			+ Figure3.png
-\end{lstlisting}
-
-An alternative directory structure (and probably the most flexible) would be to simply have a set of different figures with an informative and relevant naming scheme to make it easy to identify the paper and figure the data come from.
+The \pkg{metaDigitise} package is designed to be flexible, yet simple to use.
+There is one main function in the package, \fct{metaDigitise}, which interactively takes the user through the process of extracting data from figures.
+\fct{metaDigitise} was created with the idea that the user would likely have multiple images to extract from.
+It therefore operates in the same way whether the user has one or multiple images.
+\fct{metaDigitise} is designed to work on a directory containing images of figures copied from primary literature, in .png, .jpg, .tiff, .pdf format.
+This directory is sepcified to \fct{metaDigitise} through the \code{dir} argument.
+The user is free to set their own broad directory structure (e.g.
+one directory for all images or one directory for each paper extracted from).
+We would recommend having all files for one project in a single directory with an informative and unambiguous naming scheme for images to make it easy to identify the paper and figure the data come from.
 This cuts out the need to change directories constantly.
@@
 
-\begin{lstlisting}
+\begin{CodeChunk}
+\begin{CodeOutput}
 * Main project directory
     + FiguresToExtract
-			+ P1_Figure1_trait1.png
-			+ P1_Figure2_trait2.png
-			+ P1_Figure3_trait3.png
-			+ P2_Figure1_trait1.png
-			+ P2_Figure2_trait2.png
-			+ P2_Figure3_trait3.png
-\end{lstlisting}
+        + Paper1_Figure1_trait1.png
+        + Paper1_Figure2_trait2.png
+        + Paper1_Figure3_trait3.png
+        + Paper2_Figure1_trait1.png
+        + Paper2_Figure2_trait2.png
+        + Paper2_Figure3_trait3.png
+\end{CodeOutput}
+\end{CodeChunk}
 
-The above directory structure is probably the easiest in combination with a clear and unambiguous naming scheme for each figure.
-Even if only figures from a single paper are digitised, one paper at a time, an overall figure directory will work perfectly because `metaDigitise` will only cycle through incomplete figures, so figures can be added at anytime.
+It is important to for the user to think about their directory structure early on in this process (also more generally in the context of their entire project), especially if they plan to share the extractions with collaborators or when publishing the project.
 
-Nonetheless, how users set up their directory is really up to them.
-However, it is important for users to think carefully about reproducibility at this stage.
-Would they like to share the entire project folder with colleagues?
-or would they prefer to simply share the image folder?
-The answers to these questions are important because relative path names are stored in `metaDigitise` objects, meaning that the directory structure (currently at least) needs to be the same for colleagues to re-load previously digitised objects.
-Therefore, the working directory needs to be set to either the "main project directory" or the "FiguresToExtract" directory and relative path names (e.g., users should specify dir = "./FiguresToExtract/" if setting the working directory to "Main project directory") used as opposed to absolute paths.
-This facilitates reproducibility should colleagues want to see the digitisations or integrate previously digitised figures with newly digitised figures.
-All users need to achieve this is to provide others with is the project or image folder.
+When \fct{metaDigitise} is run, it recognizes all the images in a directory and automatically imports them one by one, allowing the user to click and enter relevant information about a figure as they go.
+This expedites digitising figures by preventing users from having to constantly change directories and / or open new images.
+The data from a completed image is automatically saved in an .RDS file to a \code{caldat} directory, that is created within the parent directory when first executing the \fct{metaDigitise} function.
+These files enable re-plotting and editing of images at a later point (see below), but do not need to be directly accessed by the user.
 
-\section*{High-throughput Digitisation of Figure(s) within a Directory}
-metaDigitise was created with the idea that the user would likely have multiple images to extract from.
-It therefore operates in the same way whether the user has one or multiple images.
-It assumes that the user has put all the figures needed for extraction in one parent directory (see above).
-Users can extract single figures (if this is all they have) using the `metaDigitise` function with a path name to the directory with the file.
-However, often many figures need extracting from a single paper or set of papers.
-`metaDigitise` will also handle these situations seamlessly by simply cycling through all figures within a directory.
-This is useful because it expedites digitising figures by preventing users from having to constantly change directories and / or open new images.
-It operates by recognizing all the figures in a folder and bringing up each figure automatically allowing the user to click and enter relevant information about a figure as they go.
-All this information is then stored in a data frame or list at the end of the process, saving quite a bit of time.
-Users can stop mid-way through a folder by simply exiting after the last plot they have digitised.
-The data from completed figures will automatically be written to a special `caldat/` folder that is created when first executing the metaDigitise function.
-Digitised data along with metadata for figures are stored in caldat/ for later use and editing (see below).
+A particularly powerful and flexible aspect of \fct{metaDigitise} is its ability to identify images that have been previously digitised and only import images that have not been digitised in subsequent calls of the function.
+This means both that all figures do not need to be extracted at one time and that new figures can be added as the project develops.
+After each image is extracted, the user is asked whether they wish to continue or quit the extraction process.
+Upon rerunning \fct{metaDigitise}, previously digitised figures are simply ignored during processing, but their data re-integrated within the final output after new files are completed.
 
-In order to correctly extract data from figures metaDigitise always requires the user to calibrate the axes in the figure.
-To do this, the user is required to click on two known points on the axis in question, and then enter the value of those points in the figure (Figure \ref{fig:calibrate}).
-Using this information, metaDigitise then calculates the value of any clicked points in terms of the figure axes.
+After completing all images, or upon quitting, the processed data (in a form specified by the user) is then returned. 
+%% how about asking whether summary or not at end? instead of having it in the arguements?
+From all plot types, \fct{metaDigitise} summarises the data from a figure as a mean, standard deviation and sample size, for each identified group within the plot (should multiple groups exist).
+These are the summary statistics needed to create many of the relevant effect sizes and sampling error for a meta-analysis.
+%% are there any hwere additional info is needed??
+In the case of scatter plots, \fct{metaDigitise} also returns the correlation coefficient between the points within each identified group.
+
+\section{Diverse Plot Types}
+\pkg{metaDigitise} recognises four main types of plot; Mean and error plots, box plots, scatter plots and histograms, shown in Figure \ref{fig:plot_type}.
+Each of these can be processed together and integrated into a single output.
+Alternatively, users can keep like figures together and process them separately.
+
+%% where to put calibration???
+
+In order to correctly extract data from figures \fct{metaDigitise} always requires the user to calibrate the axes in the figure.
+To do this, the user is required to click on two known points on the axis in question, and then enter the value of those points in the figure
+%(Figure \ref{fig:calibrate})
+.
+Using this information, \fct{metaDigitise} then calculates the value of any clicked points in terms of the figure axes.
 In the case of mean and error plots and box plots, it calibrates only the y axis (assuming the x axis is redundant).
@@
 
-\begin{figure}[!h] 
-%\onehalfspacing
- \includegraphics[width=0.9\textwidth]{fig_calibrate.pdf} 
- \caption{Axis calibration. The user defines two points on each axis and labels them according to the values shown in the figure.}
-\label{fig:calibrate}
-\end{figure}
-
-A particularly powerful and flexible aspect of metaDigitise is its ability to recognize figures that are not already completed and only bring up those figures that have not been digitised in subsequent calls of the function.
-This means that new figures can be added at any time as the project develops and previously digitised figures are simply ignore during processing, but their data re-integrated within the final output after new files are completed.
-
-From all plot types, metaDigitise summarises the data from a figure or set of figures as a mean, standard deviation and sample size, for each identified group within the plot (should multiple groups exist).
-These are the summary statistics needed to create many of the relevant effect sizes and sampling error for a meta-analysis.
-In the case of scatter plots, metaDigitise also returns the correlation coefficient between the points within each identified group.
-
-\subsection{Diverse Plot Types}
-metaDigitise recognises four main types of plot; Mean and error plots, box plots, scatter plots and histograms, shown in Figure \ref{fig:plot_type}.
-Each of these can be processed together and integrated into a single output.
-Alternatively, users can keep like figures together and process them separately.
-
-%%% Equations for mean and sd
-%% pooled stats: 
-% \begin{align}
-% \mu _{X}&={\frac {\sum _{i}{N_{X_{i}}\mu _{X_{i}}}}{\sum _{i}{N_{X_{i}}}}}
-% \\
-% \sigma _{X}&={\sqrt {{\frac {1}{\sum _{i}{N_{X_{i}}-1}}}\left(\sum _{i}{\left[(N_{X_{i}}-1)\sigma _{X_{i}}^{2}+N_{X_{i}}\mu _{X_{i}}^{2}\right]}-\left[\sum _{i}{N_{X_{i}}}\right]\mu _{X}^{2}\right)}}
-% \end{align}
+% \begin{figure}[!h] 
+%  \includegraphics[width=0.9\textwidth]{fig_calibrate.pdf} 
+%  \caption{Axis calibration. The user defines two points on each axis and labels them according to the values shown in the figure.}
+% \label{fig:calibrate}
+% \end{figure}
 
 \begin{figure}[!h] 
-%\onehalfspacing
  \includegraphics[width=0.9\textwidth]{fig_all_extract.pdf} 
@@
 
-\subsubsection{Mean and error plots} 
-metaDigitise prompts the user to enter groups names and allows the user to enter sample sizes, which are used in downstream processing.
+\subsection{Mean and error plots} 
+\fct{metaDigitise} prompts the user to enter groups names and allows the user to enter sample sizes ($n$), which are used in downstream processing.
 The user is then prompted to click on an error bar followed by the mean.
 Error bars above or below the mean can be clicked - sometimes one is clearer than the other.
-metaDigitise assumes that the error bars are symmetrical.
+\fct{metaDigitise} assumes that the error bars are symmetrical.
 Where the user has clicked the error is displayed in a different colour to the mean (Figure \ref{fig:all_extract}A).
-The user can subsequently add more groups or remove groups.
+The user can subsequently add more groups, edit groups or remove groups.
+Finally the user is asked what type of error was used in the figure: standard deviation (SD, $\sigma$), standard error (SE) or 95\% confidence intervals (CI95).
+Standard deviation is calculated from standard error as
+\begin{equation}
+\sigma = SE * \sqrt{n}
+\end{equation}
+and from 95\% confidence intervals as
+\begin{equation}
+\sigma = \frac{CI}{1.96} * \sqrt{n}
+\end{equation}
+If the user does not enter a sample size at the time of data extraction (if for example the information is not readily available) the SD is not calculated.
+This can be entered at a later time, however (see below).
 
-\subsubsection{Box plots}
- metaDigitise prompts the user to enter groups names and allows the user to enter sample sizes, which are used in downstream processing.
-The user is then prompted to click on the maximum, upper quartile, median, lower quartile and minimum.
-metaDigitise will check that the maximum is greater than the minimum, and return an warning if that is not the case.
-The user can subsequently add more groups or remove groups.
+\subsection{Box plots}
+As with mean and error plots, \fct{metaDigitise} prompts the user to enter groups names and allows the user to enter sample sizes ($n$), which are used in downstream processing.
+The user is then prompted to click on the maximum ($b$), upper quartile ($q_3$), median ($m$), lower quartile ($q_1$) and minimum ($a$).
+\fct{metaDigitise} will check that the maximum is greater than the minimum, and return an warning if that is not the case.
+The user can subsequently add, edit or remove groups.
+From the extracted data, the mean ($\mu$) and SD are calculated as
 
-\subsubsection{Scatter plots}
- metaDigitise prompts the user to enter groups names and then to click on points.
+\begin{equation}
+\mu = \frac{a + 2q_1 + 2m + 2q_3 +b}{8}
+\end{equation}
+
+\begin{equation}
+\sigma = \frac{b-a}{4\Phi^{-1}(\frac{n-0.375}{n+0.25})} + \frac{q_3-q_1}{4\Phi^{-1}(\frac{0.75n-0.125}{n+0.25})}
+\end{equation}
+
+, where $\Phi^{-1}(z)$ is the upper zth percentile of the standard normal distribution, following \citet{Wan2014}.
+As with mean and error plots, if the user does not enter a sample size at the time of data extraction the SD is not calculated.
+
+\subsection{Scatter plots}
+ \fct{metaDigitise} prompts the user to enter groups names and then to click on points.
 Points added by mistake can be deleted.
@@
 Different groups are plotted in different colours and shapes, with a legend at the bottom of the figure (Figure \ref{fig:all_extract}C).
-The x and y coordinates are returned.
+Mean, SD and n are calculated from the clicked points, for each group.
+In the case that the sample size from the clicked points does not match a known sample size (e.g.
+if there are overlaid points), the user can enter a alternate sample size.
 
 \subsubsection{Histograms}
-metaDigitise prompts the user to enter groups names and then to click on the top corners of each bar.
+\pkg{metaDigitise} prompts the user to enter groups names and then to click on the top corners of each bar.
 bars can subsequently be deleted.
-MetaDigitise then calculates the midpoint and the frequency for each bar.
+\pkg{MetaDigitise} then calculates the midpoint and the frequency for each bar.
 
-\subsection{Re-importing, Editing and Plotting Previously Digitised data}
+\section{Extracting Data From Plots}
+%% scatterplot from airquality data
 
-A particularly useful feature of `metaDigitise` is its ability to re-import, edit and re-plot previously digitised figures.
-Running metaDigitise brings up a series of prompts for the user using a main menu that provides access to a number of its features:
+We'll now demonstrate how \fct{metaDigitise} works using figures generated from the well known iris data set. Users can install the \pkg{metaDigitise} package from GitHub as follows:
 
-\begin{verbatim}
+\begin{CodeChunk}
+\begin{CodeInput}
+R> install.packages("devtools")
+R> devtools::install_github("daniel1noble/metaDigitise")
+R> library(metaDigitise)
+\end{CodeInput}
+\end{CodeChunk}
+
+Assume that we would like to extract summary statistics from studies measuring sepal length or width in iris species for a fictitious project.
+There are a few studies that only present these data in figures.
+As we read papers found from a systematic search we can add in figures with relevant data to a "FiguresToExtract" folder as follows
+
+\begin{CodeChunk}
+\begin{CodeOutput}
+*FiguresToExtract
+    + 001_Anderson_1935_Fig1.png
+\end{CodeOutput}
+\end{CodeChunk}
+
+\begin{figure}[!h] 
+ \includegraphics[width=0.9\textwidth]{001_Anderson_1935_Fig1.png} 
+ \caption{Example scatterplot (001\_Anderson\_1935\_Fig1.png) of sepal length and width for two species of iris (setosa and versicolor)}
+\label{fig:rotate}
+\end{figure}
+
+Here, the naming of the files placed in the folder will contain the paper number, first author and the figure number to keep data uniquely associated with figures.
+We can use \fct{metaDigitise} to begin processing the image from the first paper.
+Running \fct{metaDigitise} brings up a series of prompts for the user using a main menu that provides access to a number of its features:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> digitised_data <- metaDigitise("~/FiguresToExtract", summary = TRUE)
+\end{CodeInput}
+\begin{CodeOutput}
 	Do you want to...
@@
 Selection: 
-\end{verbatim}
+\end{CodeOutput}
+\end{CodeChunk}
 
-The user simply enters in the numeric value that corresponds to what they would like to do.
-Choosing "Process new images" will permit users to digitise new figures, integrating previously completed digitisations along with newly digitised data together at the end of the session, or when the user decides to quit.
-If the user chooses "Import existing data", they have the option of either 1) importing all data or 2) importing a particular figure that has been digitised.
+The user simply enters in the numeric value that corresponds to what they would like to do. In this case we want to "Process new images".
+
+As soon as \fct{metaDigitise} is executed, it will bring the user to the main menu (see "Re-importing, Editing and Plotting Previously Digitised data" above).
+When the user selects option one ("Process new images"), we are asked whether we have different types of plot(s) in the folder.
+This question is most relevant when there are lots of different figures in the folder because it will then ask us for the type of figure as they are cycled through.
+
+\begin{CodeChunk}
+\begin{CodeOutput}
+Are all plot types the same? (diff/same)
+\end{CodeOutput}
+\end{CodeChunk}
+
+\fct{metaDigitise} then asks the user whether the figure needs to be rotated or flipped.
+This can be needed when box plots and mean and error plots are not orientated correctly.
+In some cases, older papers can give slightly off angled images which can be corrected by rotating.
+So, in this prompt the user has three options: \code{f} for "Flip", \code{r} for "rotate" or \code{c} for "continue.
+
+\begin{CodeChunk}
+\begin{CodeOutput}
+mean_error and boxplots should be vertically orientated
+       _ 
+       |  
+  I.E. o    NOT  |-o-|
+       |
+       _
+
+If they are not then chose flip to correct this.
+
+If figures are wonky, chose rotate.
+
+Otherwise chose continue
+
+Flip, rotate or continue f/r/c 
+\end{CodeOutput}
+\begin{CodeInput}
+c
+\end{CodeInput}
+\end{CodeChunk}
+
+After this, \fct{metaDigitise} will ask the user to specify the plot type. Depending on the figure, the user can specify that it is a figure containing the mean and error (\code{m}), a box plot (\code
+{b}), a scatter plot (\code{s}) or a histogram (\code{h}).
+If the user has specified \code{diff} instead of \code{same} in response to the question about whether the plot types are the same or different, this question will pop up for each plot, but will only be asked once if plots are all the same.
+
+\begin{CodeChunk}
+\begin{CodeOutput}
+Please specify the plot_type as either: mean and error, box plot,
+ scatter plot or histogram m/b/s/h:
+\end{CodeOutput}
+\end{CodeChunk}
+
+After selecting the figure type a new set of prompts will come up that will ask the user first what the y and x-axis variables are.
+This is useful as users can keep track of the different variables across figures and papers.
+Here, the user can just add this information in to the \proglang{R} console.
+Once complete, details on how to calibrate the x and y-axis appear, so that the relevant statistics / data can be correctly calculated.
+When working with a plot of mean and standard errors, the x-axis is rather useless in terms of calibration so \pkg{metaDigitise} just asks the user to calibrate the y-axis.
+
+\begin{CodeChunk}
+\begin{CodeOutput}
+What is the y variable? 
+\end{CodeOutput}
+\begin{CodeInput}
+R> Sepal Length (mm)
+\end{CodeInput}
+
+\begin{CodeOutput}
+What is the x variable? 
+\end{CodeOutput}
+\begin{CodeInput}
+R> Sepal Width (mm)
+\end{CodeInput}
+
+\begin{CodeOutput}
+On the Figure, click IN ORDER: 
+      y1, y2 , x1, x2
+
+    Step 1 ----> Click on y1
+  |
+  |
+  |
+  |
+  y1
+  |_________________________
+  ....
+
+    Step 3 ----> Click on x1
+  |
+  |
+  |
+  |
+  |
+  |_____x1__________________
+
+  ....
+\end{CodeOutput}
+\end{CodeChunk}
+
+The user can just follow the instructions on screen step-by-step (instructions above have been truncated by `...` to simplify), and in the order specified.
+Before moving on, the user is forced to check whether or not the calibration has been set up correctly.
+If `n` is chosen because something needs to be fixed then the user can re-calibrate.
+
+\begin{CodeChunk}
+\begin{CodeOutput}
+What is the value of y1 ?
+\end{CodeOutput}
+\begin{CodeInput}
+R> 4.5
+\end{CodeInput}
+\begin{CodeOutput}
+What is the value of y2 ?
+\end{CodeOutput}
+\begin{CodeInput}
+R> 7
+\end{CodeInput}
+\begin{CodeOutput}
+What is the value of x1 ?
+\end{CodeOutput}
+\begin{CodeInput}
+R> 2
+\end{CodeInput}
+\begin{CodeOutput}
+What is the value of x2 ?
+\end{CodeOutput}
+\begin{CodeInput}
+R> 4
+\end{CodeInput}
+\begin{CodeOutput}
+Re-calibrate? (y/n) 
+\end{CodeOutput}
+\end{CodeChunk}
+
+Often, plots might contain multiple groups that the meta-analyst wants to extract from.
+\fct{metaDigitise} handles this nicely by prompting the user to enter the group first, followed by the digitisation of this groups data.
+After digitising the first group, and having exited (i.e., hit `esc` from plot window), \fct{metaDigitise} will ask the user whether they would like to add another group.
+Users can continually add groups (\code{a}), or simply continue to the next plot (\code{c} - if another plot exists).
+%%% This needs checking - think it is now different
+The number of groups are not really limited and users can just keep adding in groups to accommodate the different numbers that may be presented across figures (although it can get complicated with too many).
+
+\begin{CodeChunk}
+\begin{CodeOutput}
+Follow instructions below, to exit point adding or removing:
+
+ - Windows: right click on the plot area and choose 'Stop'!
+
+ - X11: hit any mouse button other than the left one.
+
+ - quartz/OS X: hit ESC
+
+Group identifier:
+\end{CodeOutput}
+\begin{CodeInput}
+R> setosa
+\end{CodeInput}
+\begin{CodeOutput}
+Click on points you want to add.
+If you want to remove a point, or are finished with a
+group, exit (see above), then follow prompts
+
+Add points, delete points or continue? a/d/c 
+\end{CodeOutput}
+\begin{CodeInput}
+c
+\end{CodeInput}
+\end{CodeChunk}
+
+Once we are done digitising all the groups our plot will look something like Figure \ref{fig:scatter_extract}
+
+\begin{figure}[!h] 
+ \includegraphics[width=0.9\textwidth]{001_Anderson_1935_Fig1_digitised.png} 
+ \caption{Digitisation of sepal length and width for two species of iris (setosa and versicolor).
+Names of the variables and calibration (in blue) are plotted alongside the digitised points (green = versicolor; red = setosa).
+The sample sizes for each group are provided on the lower part of the plot.
+All figures are clearly labeled at the top to remind users of the filename and plot type.
+This reduces errors throughout the digitisation process.}
+\label{fig:scatter_extract}
+\end{figure}
+
+When completed \pkg{metaDigitise} will write the digitised data to the caldat directory, such that our new directory structure is as follows
+
+\begin{CodeChunk}
+\begin{CodeOutput}
+*figs_to_extract
+    + caldat
+        + 001_Anderson_1935_Fig1
+    + 001_Anderson_1935_Fig1.png
+\end{CodeOutput}
+\end{CodeChunk}
+
+Users can access the \pkg{metaDigitise} object created (001\_Anderson\_1935\_Fig1) at any time using the metaDigitse function.
+In the \proglang{R} console, the summarised data for the digitised figure can be printed on screen or even written to a .csv file:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> digitised_data
+\end{CodeInput}
+\begin{CodeOutput}
+ filename   group_id         variable       mean      sd  n         r  plot_type
+1 001_Anderson_1935_Fig1.png     setosa  Sepal Width (mm) 3.420332 0.4034259 39 0.7524909 scatterplot
+2 001_Anderson_1935_Fig1.png     setosa Sepal Length (mm) 4.998267 0.3831294 39 0.7524909 scatterplot
+3 001_Anderson_1935_Fig1.png versicolor  Sepal Width (mm) 2.766857 0.3234723 44 0.5157938 scatterplot
+4 001_Anderson_1935_Fig1.png versicolor Sepal Length (mm) 5.936769 0.5292535 44 0.5157938 scatterplot
+\end{CodeOutput}
+\end{CodeChunk}
+
+The mean for each of the two variables, along with the two species, are provided.
+Since this is a scatterplot, the user also gets the Person's correlation coefficient between sepal length and width for each species.
+These match reasonably well with the actual means of sepal length and width for each of the species in the full `iris` dataset:
+
+\begin{CodeChunk}
+\begin{CodeOutput}
+     Species meanSL meanSW
+1     setosa  5.006  3.428
+2 versicolor  5.936  2.770
+\end{CodeOutput}
+\end{CodeChunk}
+
+One thing anyone with a familiarity with the iris dataset will notice is that the sample sizes for each of these species (which are n = 50 each) are quite a bit lower.
+This is an example of some of the challenges when extracting data from scatter plots, often data points will overlap with each other making it impossible (without having the real data) to know whether this is a problem.
+However, a meta-analyst will probably realise that the sample sizes here conflict with what is reported in the paper.
+Hence, \pkg{metaDigitise} also provides the user with  options to input the sample sizes directly (see Editing section above), even for scatter plots and histograms where, strictly speaking, this should not be necessary.
+Nonetheless, it is important to recognise the impact that overlapping points can have on summary statistics, particularly its effects on standard deviation (SD) and standard error (SE).
+Here, the mean point estimates are nearly bang on, but the SD's are slightly over-estimated:
+
+\begin{CodeChunk}
+\begin{CodeOutput}
+     Species    meanSL    meanSW
+1     setosa 0.3524897 0.3790644
+2 versicolor 0.5161711 0.3137983
+\end{CodeOutput}
+\end{CodeChunk}
+
+\subsection{Adding new figures}
+
+Users can add additional figures as new papers with relevant information are found.
+Each figure should be in its own file with unique naming, even if a single paper has multiple figures for extraction.
+For example, another paper on different populations (and one new species) of iris contained two additional figures where important data could be extracted.
+These figures can simply be named accordingly and added directly to the same extraction folder:
+
+\begin{CodeChunk}
+\begin{CodeOutput}
+*figs_to_extract
+	+ caldat
+	  	+ 001_Anderson_1935_Fig1
+    + 001_Anderson_1935_Fig1.png
+    + 002_Doe_2013_Fig1.png
+    + 002_Doe_2013_Fig3.png
+\end{CodeOutput}
+\end{CodeChunk}
+
+We have already processed one figure (001\_Anderson\_1935\_Fig1.png) and we can tell because it has digitised data (caldat/001\_Anderson\_1935\_Fig1), but now we have our two new figures that have not yet been digitised.
+This example will nicely demonstrate how users can easily pick up from where they left off and how all previous data gets re-integrated.
+It will also demonstrate how different plot types are handled.
+All we have to do to begin, is again, provide the directory where all the figures are located:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> setwd("~/figs_to_extract")
+R> digitised_data <- metaDigitise(".", summary = TRUE)
+\end{CodeInput}
+\end{CodeChunk}
+
+The user gets the same set of prompts and simply chooses option one to digitise the new figures.
+This will permit users to digitise new figures, and will integrate previously completed digitisations along with newly digitised data together at the end of the session, or when the user decides to quit.
+This time, 001\_Anderson\_1935\_Fig1.png, is ignored and the new plots cycle on screen.
+First for 002\_Doe\_2013\_Fig1.png and then 002\_Doe\_2013\_Fig3.png.
+Since there are a few new figures, we answer the first question in the \proglang{R} console as "diff":
+
+\begin{CodeChunk}
+\begin{CodeOutput}
+Are all plot types the same? (diff/same) 
+\end{CodeOutput}
+\begin{CodeInput}
+R> diff
+\end{CodeInput}
+\begin{CodeOutput}
+**** NEW PLOT ****
+
+mean_error and boxplots should be vertically orientated
+       _ 
+       |  
+  I.E. o    NOT  |-o-|
+       |
+       _
+
+If they are not then chose flip to correct this.
+
+If figures are wonky, chose rotate.
+
+Otherwise chose continue
+
+Flip, rotate or continue f/r/c 
+\end{CodeOutput}
+\begin{CodeInput}
+R> c
+\end{CodeInput}
+\begin{CodeOutput}
+Please specify the plot_type as either: mean and error, box plot,
+scatter plot or histogram m/b/s/h: 
+\end{CodeOutput}
+\begin{CodeInput}
+R> m
+\end{CodeInput}
+\end{CodeChunk}
+
+Here, we specify the new plot type as \code{m} for 002\_Doe\_2013\_Fig1.png because we have a plot of the mean and error of sepal length for each of the three species.
+We're then prompted a bit differently from our scatter plot as the x-axis is not needed for calibration:
+
+\begin{CodeChunk}
+\begin{CodeOutput}
+What is the variable? 
+\end{CodeOutput}
+\begin{CodeInput}
+R> Sepal length
+\end{CodeInput}
+\begin{CodeOutput}
+On the Figure, click IN ORDER: 
+      y1, y2
+
+    Step 1 ----> Click on y1
+  |
+  |
+  |
+  |
+  y1
+  |_________________________
+
+    Step 2 ----> Click on y2
+  |
+  y2
+  |
+  |
+  |
+  |_________________________
+
+What is the value of y1 ?
+\end{CodeOutput}
+\begin{CodeInput}
+R> 5
+\end{CodeInput}
+\begin{CodeOutput}
+What is the value of y2 ?
+\end{CodeOutput}
+\begin{CodeInput}
+R> 6.5
+\end{CodeInput}
+\begin{CodeOutput}
+Re-calibrate? (y/n) 
+\end{CodeOutput}
+\begin{CodeInput}
+R> n
+\end{CodeInput}
+\begin{CodeOutput}
+Enter sample sizes? y/n 
+\end{CodeOutput}
+\begin{CodeInput}
+R> y
+\end{CodeInput}
+\begin{CodeOutput}
+Group identifier: 
+\end{CodeOutput}
+\begin{CodeInput}
+R> setosa
+\end{CodeInput}
+\begin{CodeOutput}
+Group sample size: 
+\end{CodeOutput}
+\begin{CodeInput}
+R> 50
+\end{CodeInput}
+\begin{CodeOutput}
+Click on Error Bar, followed by the Mean
+
+Add group, Delete group or Finish plot? a/d/f 
+\end{CodeOutput}
+\begin{CodeInput}
+R> a
+\end{CodeInput}
+\end{CodeChunk}
+
+Again, \fct{metaDigitise} will simply guide the user through digitising each of these figures describing to them exactly what needs to be done.
+At any point if mistakes are made the user can choose relevant options to edit or correct things.
+This can continues for each plot so long as the user would like to continue after completing a single plot:
+
+\begin{CodeChunk}
+\begin{CodeOutput}
+Do you want continue: 1 plots out of 2 plots remaining (y/n)
+\end{CodeOutput}
+\begin{CodeInput}
+R> y
+\end{CodeInput}
+\end{CodeChunk}
+
+This continues until users have completed all non-digitised figures in the folder, at which point \fct{metaDigitise} concatenates the new data with previously digitised data in the object:
+
+\begin{CodeChunk}
+\begin{CodeOutput}
+data
+               filename        group_id         variable     mean        sd   n         r    plot_type
+1 001_Anderson_1935_Fig1.png     setosa  Sepal Width (mm) 3.420332 0.4034259 39 0.7524909 scatterplot
+2 001_Anderson_1935_Fig1.png     setosa Sepal Length (mm) 4.998267 0.3831294 39 0.7524909 scatterplot
+3 001_Anderson_1935_Fig1.png versicolor  Sepal Width (mm) 2.766857 0.3234723 44 0.5157938 scatterplot
+4 001_Anderson_1935_Fig1.png versicolor Sepal Length (mm) 5.936769 0.5292535 44 0.5157938 scatterplot
+5      002_Doe_2013_Fig1.png     setosa      Sepal length 5.000336 0.7828656 50        NA  mean_error
+6      002_Doe_2013_Fig1.png versicolor      Sepal length 5.931340 1.0466177 50        NA  mean_error
+7      002_Doe_2013_Fig1.png viriginica      Sepal length 6.588705 1.2608173 50        NA  mean_error
+8      003_Doe_2013_Fig3.png       <NA>      Sepal length 4.948472 0.3624212 50        NA  histogram
+\end{CodeOutput}
+\end{CodeChunk}
+
+\section{Re-importing, Editing and Plotting Previously Digitised data}
+
+A particularly useful feature of \pkg{metaDigitise} is its ability to re-import, edit and re-plot previously digitised figures. We can do this from the initial options from \fct{metaDigitise}
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> metaDigitise("./FiguresToExtract")
+\end{CodeInput}
+\begin{CodeOutput}
+    Do you want to...
+1: Process new images
+2: Import existing data
+3: Edit existing data
+Selection: 
+\end{CodeOutput}
+\end{CodeChunk}
+
+If the user chooses "Import existing data", they have the option of either 1) importing data from all digitised images or 2) importing data from particular image that has been digitised.
 If 2, then a list of files are provided to the user that they can select.
-
 Editing existing data allows users to easily re-plot or edit information or digitisations that have previously be done for any plot.
@@
 
-\begin{verbatim}
+\begin{CodeChunk}
+\begin{CodeOutput}
 Choose how you want to edit files:
@@
 Selection: 
-\end{verbatim}
+\end{CodeOutput}
+\end{CodeChunk}
 
 If the user is unsure about the name of the specific figure they need to edit or simply want to just check the digitisations of figures they can choose "Cycle through images", which will bring up each figure, one by one, overlaying the calibrations, group names (if they exist), sample sizes (if they were entered) and the selected points.
+The user will then be given the choice to edit individual images.
 Alternatively, choosing option 2, will bring up a list of the completed files in the folder and the specific file can be chosen, at which point it will be replotted.
@@
 
-\begin{verbatim}
-Edit rotation? If yes, then the whole extraction will be redone (y/n) n
-Change plot type? If yes, then the whole extraction will be redone (y/n) n
-Variable entered as: Sepal length
-Rename Variables (y/n) n
-Edit calibration? (y/n) n
-Re-extract data (y/n) y
-Change group identifier? (y/n) n
-Add group, Delete group or Finish plot? a/d/f d
+\begin{CodeChunk}
+
+\begin{CodeOutput}
+Edit rotation? If yes, then the whole extraction will be redone (y/n) 
+\end{CodeOutput}
+\begin{CodeInput}
+R> n
+\end{CodeInput}
+\begin{CodeOutput}
+Change plot type? If yes, then the whole extraction will be redone (y/n) 
+\end{CodeOutput}
+\begin{CodeInput}
+R> n
+\end{CodeInput}
+\begin{CodeOutput}
+Variable entered as: 
+\end{CodeOutput}
+\begin{CodeInput}
+R> Sepal length
+\end{CodeInput}
+\begin{CodeOutput}
+Rename Variables (y/n) 
+\end{CodeOutput}
+\begin{CodeInput}
+R> n
+\end{CodeInput}
+\begin{CodeOutput}
+Edit calibration? (y/n) 
+\end{CodeOutput}
+\begin{CodeInput}
+R> n
+\end{CodeInput}
+\begin{CodeOutput}
+Re-extract data (y/n) 
+\end{CodeOutput}
+\begin{CodeInput}
+R> y
+\end{CodeInput}
+\begin{CodeOutput}
+Change group identifier? (y/n) 
+\end{CodeOutput}
+\begin{CodeInput}
+R> n
+\end{CodeInput}
+\begin{CodeOutput}
+Add group, Delete group or Finish plot? a/d/f 
+\end{CodeOutput}
+\begin{CodeInput}
+R> d
+\end{CodeInput}
+\begin{CodeOutput}
 1: setosa
@@
 3: viriginica
-Selection: 2
-Add group, Delete group or Finish plot? a/d/f a
-\end{verbatim}
+Selection: 
+\end{CodeOutput}
+\begin{CodeInput}
+R> 2
+\end{CodeInput}
+\begin{CodeOutput}
+Add group, Delete group or Finish plot? a/d/f 
+\end{CodeOutput}
+\begin{CodeInput}
+R> a
+\end{CodeInput}
+
+\end{CodeChunk}
 
 A whole host of information can be edited from the rotation, plot type, the variable name that was provided, the calibration and even the digitisation of groups.
-When editing the metaDigitise object is re-written to the caldat folder and the edits are immediately integrated into the existing object once complete.
+When editing the \code{metaDigitise} object is re-written to the caldat folder and the edits are immediately integrated into the existing object once complete.
 
@@
 This will make the calibration of the points in the figure from the image problematic.
-metaDigitise allows users to rotate the image.
+\fct{metaDigitise} allows users to rotate the image.
 By clicking two points on the x-axis, metaDigitse calculates the angle needed to rotate the image so the x-axis is horizontal, and rotates it.
@@
 
-\begin{figure}[!h] 
-%\onehalfspacing
+\begin{figure}[!b] 
  \includegraphics[width=0.9\textwidth]{fig_rotate.pdf} 
@@
 Furthermore, in some figures mean and error, boxplots or histograms may be presented with horizontal bars.
-metaDigitise assumes that the bars are vertical, but allows the user to flip the image so that the bars are vertical (Figure \ref{fig:rotate}C,D).
+\fct{metaDigitise} assumes that the bars are vertical, but allows the user to flip the image so that the bars are vertical (Figure \ref{fig:rotate}C,D).
 
@@
 
-While metaDigitise provides users with the summary statistics by default, for all plot types, in many cases the user may actually be interested in obtaining the raw digitised data from scatter plots.
-This is very easy to do my changing the default `summary` argument from TRUE to FALSE in the metaDigitise function.
+While \fct{metaDigitise} provides users with the summary statistics by default, for all plot types, in many cases the user may actually be interested in obtaining the raw digitised data from scatter plots.
+This is very easy to do my changing the default \code{summary} argument from TRUE to FALSE in \fct{metaDigitise}.
 Instead of providing the user with summary statistics it will return a list containing four slots for each of the figure types (mean error, box plot, histogram and scatter plots).
@@
 
-\begin{verbatim}
-str(data)
+\begin{CodeChunk}
+
+\begin{CodeInput}
+>R str(data)
+\end{CodeInput}
+
+\begin{CodeOutput}
 List of 3
@@
   .. ..$ x_variable: chr [1:83] "Sepal width (mm)" "Sepal width (mm)" "Sepal width (mm)" "Sepal width (mm)" ...
-\end{verbatim}
+\end{CodeOutput}
+\end{CodeChunk}
 
@@
 
-\begin{verbatim}
-scatterplot <- data$scatterplot
-\end{verbatim}
+\begin{CodeChunk}
+\begin{CodeInput}
+>R scatterplot <- data$scatterplot
+\end{CodeInput}
+\end{CodeChunk}
 
@@
 %Including addition of N later
-
-\section{Example}
-%% scatterplot from airquality data
-We'll now turn to an example of how metaDigitise works using figures generated using the well known iris data set. Users can install metaDigitise from GitHub as follows:
-
-\begin{verbatim}
-	> install.packages("devtools")
-	> devtools::install_github("daniel1noble/metaDigitise")
-	> library(metaDigitise)
-\end{verbatim}
-
-Assume that we would like to extract summary statistics from studies measuring sepal length or width in iris species for a fictitious project.
-There are a few studies that only present these data in figures.
-As we read papers found from a systematic search we can add in figures with relevant data to a "figs\_to\_extract" folder as follows
-
-\begin{lstlisting}
-	*figs_to_extract
-      + 001_Anderson_1935_Fig1.png
-\end{lstlisting}
-
-\begin{figure}[!h] 
-%\onehalfspacing
- \includegraphics[width=0.9\textwidth]{001_Anderson_1935_Fig1.png} 
- \caption{Example scatterplot (001\_Anderson\_1935\_Fig1.png) of sepal length and width for two species of iris (setosa and versicolor)}
-\label{fig:rotate}
-\end{figure}
-
-Here, the naming of the files placed in the folder will contain the paper number, first author and the figure number to keep data uniquely associated with figures.
-We can set our working directory to "figs\_to\_extract" and use metaDigitise to begin processing the image from the first paper
-
-\begin{verbatim}
-	> setwd("~/figs_to_extract")
-	> digitised_data <- metaDigitise(".", summary = TRUE)
-\end{verbatim}
-
-As soon metaDigitise is executed, it will bring the user to the main menu (see "Re-importing, Editing and Plotting Previously Digitised data" above).
-When the user selects option one ("Process new images"), we are asked whether we have different types of plot(s) in the folder.
-This question is most relevant when there are lots of different figures in the folder because it will then ask us for the type of figure as they are cycled through.
-
-\begin{lstlisting}
-	Are all plot types the same? (diff/same)
-\end{lstlisting}
-
-metaDigitise then asks the user whether the figure needs to be rotated or flipped.
-This can be needed when box plots and mean and error plots are not orientated correctly.
-In some cases, older papers can give slightly off angled images which can be corrected by rotating.
-So, in this prompt the user has three options: `f` for "Flip", `r` for "rotate" or `c` for "continue.
-
-\begin{lstlisting}
-	mean_error and boxplots should be vertically orientated
-	       _ 
-	       |  
-	  I.E. o    NOT  |-o-|
-	       |
-	       _
-
-	If they are not then chose flip to correct this.
-
-	If figures are wonky, chose rotate.
-
-	Otherwise chose continue
-
-	Flip, rotate or continue f/r/c c
-'\end{lstlisting}
-
-After this, `metaDigistise` will ask the user to specify the plot type.
-Depending on the figure, the user can specify that it is a figure containing the mean and error (`m`), a box plot (`b`), a scatter plot (`s`) or a histogram (`h`).
-If the user has specified `diff` instead of `same` in response to the question about whether the plot types are the same or different, this question will pop up for each plot, but will only be asked once if plots are all the same.
-
-\begin{lstlisting}
-	Please specify the plot_type as either: mean and error, box plot, scatter plot or histogram m/b/s/h:
-\end{lstlisting}
-
-After selecting the figure type a new set of prompts will come up that will ask the user first what the y and x-axis variables are.
-This is useful as users can keep track of the different variables across figures and papers.
-Here, the user can just add this information in to the R console.
-Once complete, details on how to calibrate the x and y-axis appear, so that the relevant statistics / data can be correctly calculated.
-When working with a plot of mean and standard errors, the x-axis is rather useless in terms of calibration so `metaDigitise` just asks the user to calibrate the y-axis.
-
-\begin{lstlisting}
-What is the y variable? Sepal Length (mm)
-
-What is the x variable? Sepal Width (mm)
-
-On the Figure, click IN ORDER: 
-      y1, y2 , x1, x2
-
-    Step 1 ----> Click on y1
-  |
-  |
-  |
-  |
-  y1
-  |_________________________
-  ....
-
-    Step 3 ----> Click on x1
-  |
-  |
-  |
-  |
-  |
-  |_____x1__________________
-
-  ....
- \end{lstlisting}
-
-The user can just follow the instructions on screen step-by-step (instructions above have been truncated by `...` to simplify), and in the order specified.
-Before moving on, the user is forced to check whether or not the calibration has been set up correctly.
-If `n` is chosen because something needs to be fixed then the user can re-calibrate.
-
-\begin{lstlisting}
-What is the value of y1 ?
-4.5
-What is the value of y2 ?
-7
-What is the value of x1 ?
-2
-What is the value of x2 ?
-4
-
-Re-calibrate? (y/n) 
-\end{lstlisting}
-
-Often, plots might contain multiple groups that the meta-analyst wants to extract from.
-`metaDigitise` handles this nicely by prompting the user to enter the group first, followed by the digitisation of this groups data.
-After digitising the first group, and having exited (i.e., hit `esc` from plot window), `mteaDigitise` will ask the user whether they would like to add another group.
-Users can continually add groups (`a`), or simply continue to the next plot (`c` - if another plot exists).
-The number of groups are not really limited and users can just keep adding in groups to accommodate the different numbers that may be presented across figures (although it can get complicated with too many).
-
-\begin{lstlisting}
-	Follow instructions below, to exit point adding or removing:
-
-	 - Windows: right click on the plot area and choose 'Stop'!
-
-	 - X11: hit any mouse button other than the left one.
-
-	 - quartz/OS X: hit ESC
-
-	Group identifier: setosa
-
-	Click on points you want to add.
-	If you want to remove a point, or are finished with a
-	group, exit (see above), then follow prompts
-
-	Add points, delete points or continue? a/d/c c
-
-\end{lstlisting}
-
-Once we are done digitising all the groups our plot will look something like this:
-
-\begin{figure}[!h] 
-%\onehalfspacing
- \includegraphics[width=0.9\textwidth]{001_Anderson_1935_Fig1_digitised.png} 
- \caption{Digitisation of sepal length and width for two species of iris (setosa and versicolor).
-Names of the variables and calibration (in blue) are plotted alongside the digitised points (green = versicolor; red = setosa).
-The sample sizes for each group are provided on the lower part of the plot.
-All figures are clearly labeled at the top to remind users of the filename and plot type.
-This reduces errors throughout the digitisation process.}
-\label{fig:rotate}
-\end{figure}
-
-When completed metaDigitise will write the digitised data to the caldat directory, such that our new directory structure is as follows
-
-\begin{lstlisting}
-	*figs_to_extract
-	  + caldat
-	  	+ 001_Anderson_1935_Fig1
-      + 001_Anderson_1935_Fig1.png
-\end{lstlisting}
-
-Users can access the metaDigitise object created (001\_Anderson\_1935\_Fig1) at any time using the metaDigitse function.
-In the R console, the summarised data for the digitised figure can be printed on screen or even written to a .csv file:
-
-\begin{lstlisting}
-data
-                   filename   group_id         variable       mean      sd  n         r  plot_type
-1 iris_scatter_multigrp.png     setosa  Sepal Width (mm) 3.420332 0.4034259 39 0.7524909 scatterplot
-2 iris_scatter_multigrp.png     setosa Sepal Length (mm) 4.998267 0.3831294 39 0.7524909 scatterplot
-3 iris_scatter_multigrp.png versicolor  Sepal Width (mm) 2.766857 0.3234723 44 0.5157938 scatterplot
-4 iris_scatter_multigrp.png versicolor Sepal Length (mm) 5.936769 0.5292535 44 0.5157938 scatterplot
-\end{lstlisting}
-
-The mean for each of the two variables, along with the two species, are provided.
-Since this is a scatterplot, the user also gets the Person's correlation coefficient between sepal length and width for each species.
-These match reasonably well with the actual means of sepal length and width for each of the species in the full `iris` dataset:
-
-\begin{lstlisting}
-     Species meanSL meanSW
-1     setosa  5.006  3.428
-2 versicolor  5.936  2.770
-\end{lstlisting}
-
-One thing anyone with a familiarity with the iris dataset will notice is that the sample sizes for each of these species (which are n = 50 each) are quite a bit lower.
-This is an example of some of the challenges when extracting data from scatter plots, often data points will overlap with each other making it impossible (without having the real data) to know whether this is a problem.
-However, a meta-analyst will probably realise that the sample sizes here conflict with what is reported in the paper.
-Hence, metaDigitise also provides the user with  options to input the sample sizes directly (see Editing section above), even for scatter plots and histograms where, strictly speaking, this should not be necessary.
-Nonetheless, it is important to recognise the impact that overlapping points can have on summary statistics, particularly its effects on standard deviation (SD) and standard error (SE).
-Here, the mean point estimates are nearly bang on, but the SD's are slightly over-estimated:
-
-\begin{lstlisting}
-     Species    meanSL    meanSW
-1     setosa 0.3524897 0.3790644
-2 versicolor 0.5161711 0.3137983
-\end{lstlisting}
-
-\subsection{Adding new figures}
-
-Users can add additional figures as new papers with relevant information are found.
-Each figure should be in its own file with unique naming, even if a single paper has multiple figures for extraction.
-For example, another paper on different populations (and one new species) of iris contained two additional figures where important data could be extracted.
-These figures can simply be named accordingly and added directly to the same extraction folder:
-
-\begin{lstlisting}
-	*figs_to_extract
-	  + caldat
-	  	+ 001_Anderson_1935_Fig1
-      + 001_Anderson_1935_Fig1.png
-      + 002_Doe_2013_Fig1.png
-      + 002_Doe_2013_Fig3.png
-\end{lstlisting}
-
-We have already processed one figure (001\_Anderson\_1935\_Fig1.png) and we can tell because it has digitised data (caldat/001\_Anderson\_1935\_Fig1), but now we have our two new figures that have not yet been digitised.
-This example will nicely demonstrate how users can easily pick up from where they left off and how all previous data gets re-integrated.
-It will also demonstrate how different plot types are handled.
-All we have to do to begin, is again, provide the directory where all the figures are located:
-
-\begin{verbatim}
-	> setwd("~/figs_to_extract")
-	> digitised_data <- metaDigitise(".", summary = TRUE)
-\end{verbatim}
-
-The user gets the same set of prompts and simply chooses option one to digitise the new figures.
-This time, 001\_Anderson\_1935\_Fig1.png, is ignored and the new plots cycle on screen.
-First for 002\_Doe\_2013\_Fig1.png and then 002\_Doe\_2013\_Fig3.png.
-Since there are a few new figures, we answer the first question in the R console as "diff":
-
-\begin{lstlisting}
-	Are all plot types the same? (diff/same) diff
-
-**** NEW PLOT ****
-
-mean_error and boxplots should be vertically orientated
-       _ 
-       |  
-  I.E. o    NOT  |-o-|
-       |
-       _
-
-If they are not then chose flip to correct this.
-
-If figures are wonky, chose rotate.
-
-Otherwise chose continue
-
-Flip, rotate or continue f/r/c c
-Please specify the plot_type as either: mean and error, box plot, scatter plot or histogram m/b/s/h: m
-\end{lstlisting}
-
-Here, we specify the new plot type as `m` for 002\_Doe\_2013\_Fig1.png because we have a plot of the mean and error of sepal length for each of the three species.
-We're then prompted a bit differently from our scatter plot as the x-axis is not needed for calibration:
-
-\begin{lstlisting}
-What is the variable? Sepal length
-
-On the Figure, click IN ORDER: 
-      y1, y2
-
-    Step 1 ----> Click on y1
-  |
-  |
-  |
-  |
-  y1
-  |_________________________
-
-    Step 2 ----> Click on y2
-  |
-  y2
-  |
-  |
-  |
-  |_________________________
-
-What is the value of y1 ?
-5
-What is the value of y2 ?
-6.5
-
-Re-calibrate? (y/n) n
-
-Enter sample sizes? y/n y
-
-Group identifier: setosa
-
-Group sample size: 50
-
-Click on Error Bar, followed by the Mean
-
-Add group, Delete group or Finish plot? a/d/f a
-\end{lstlisting}
-Again, metaDigitsie will simply guide the user through digitising each of these figures describing to them exactly what needs to be done.
-At any point if mistakes are made the user can choose relevant options to edit or correct things.
-This can continues for each plot so long as the user would like to continue after completing a single plot:
-
-\begin{lstlisting}
-Do you want continue: 1 plots out of 2 plots remaining (y/n) y
-\end{lstlisting}
-
-This continues until users have completed all non-digitised figures in the folder, at which point metaDigitise concatenates the new data with previously digitised data in the object:
-
-\begin{lstlisting}
-data
-               filename        group_id         variable     mean        sd   n         r    plot_type
-1 001_Anderson_1935_Fig1.png     setosa  Sepal Width (mm) 3.420332 0.4034259 39 0.7524909 scatterplot
-2 001_Anderson_1935_Fig1.png     setosa Sepal Length (mm) 4.998267 0.3831294 39 0.7524909 scatterplot
-3 001_Anderson_1935_Fig1.png versicolor  Sepal Width (mm) 2.766857 0.3234723 44 0.5157938 scatterplot
-4 001_Anderson_1935_Fig1.png versicolor Sepal Length (mm) 5.936769 0.5292535 44 0.5157938 scatterplot
-5      002_Doe_2013_Fig1.png     setosa      Sepal length 5.000336 0.7828656 50        NA  mean_error
-6      002_Doe_2013_Fig1.png versicolor      Sepal length 5.931340 1.0466177 50        NA  mean_error
-7      002_Doe_2013_Fig1.png viriginica      Sepal length 6.588705 1.2608173 50        NA  mean_error
-8      003_Doe_2013_Fig3.png       <NA>      Sepal length 4.948472 0.3624212 50        NA  histogram
-\end{lstlisting}
 
@@
 
-mmetaDigitise is available through GitHub.
+\pkg{metaDigitise} is available through GitHub.
 WHile it is incredibly flexible already, our future plans include building in abilities to deal with log-transformed axes, including arguments for calculating standard deviations from 95\% confidence intervals and standard errors using the t-distribution to correct the t-value for small sample sizes (currently assumed t-values are 1.96 as is normal), dealing with asymmetric error bars and the possibility of zooming in plots such that greater accuracy can be achieved when digitising.
 We hope to also provide options for assessing inter-observer reliability in the future.
+
+\section*{Acknowledgments}
+We thanks the i-deel group at UNSW for useful feedback, and ...
+
+\bibliography{metaDigitise}
+
 \end{document}

### 2017-11-20 docs/manuscript.tex
manuscript changes

@@
 %% recommended packages
-\usepackage{thumbpdf,lmodern}
+\usepackage{thumbpdf,lmodern,amssymb}
 
@@
 
+%%%---------------------------------
+%%%---------------------------------
+%%%---------------------------------
+
 \section{Introduction}
@@
 Meta-analysis therefore relies foremost on data extracted from primary literature and more specifically statistics that have been reported in the text or tables of research papers.
-These summary statistics are, however, frequently presented in figures and so need to be manually extracted using digitising programs.
-Although there are several existing tools to perform tasks like this (EXAMPLES), these tools are not designed this specific purpose (i.e.
-a meta-analysis).
-Specifically, they do not differentiate between common plot types that are use to present data, meaning that they require a large amount of downstream data manipulation.
-For example, many stand alone programs only provide the raw data or values from plots.
-This means that these data need to be imported into other programs to calculate correlations (if scatter plots) or errors on means (from mean error plots) need to be back calculated by hand to convert them to a common type (e.g., standard errors to standard deviations).
-Digitising programs also do not easily allow for different grouping of points to be made, and frequently do not integrate metadata (such as variable names, group names and sample sizes) at the time of data extraction.
+These summary statistics are also, however, frequently presented in figures and so need to be manually extracted using digitising programs.
+Although there are several existing tools to perform tasks like this (e.g.
+DataThief (REF), GraphClick (REF), WebPlotDigitizer (REF))
+, these tools are not designed this specific purpose (i.e. a meta-analysis). 
+%% need to spell out clear points
+%% 1 - dont manipulate data/allow different plot types
+Firstly, they typically only provide the user with calibrated x,y coordinates from imported figures, and do not differentiate between common plot types that are use to present data.
+This means that a large amount of downstream data manipulation is subsequently required.
+For example, data are frequently presented in mean and error plots (Figure \ref{fig:plot_type}A), for which the user wants a mean and error estimate for each group presented in the figure.
+With existing programs, x,y coordinates of means and errors are returned, to which the user must manually discern between mean and error coordinates and assign points to groups.
+The error then needs to be calculated as the deviation from the mean, and then transformed to a standard deviation, depending on the type of error presented.
+%% 2- adding metadata
+Secondly, digitising programs do not easily allow the integration of metadata at the time of data extraction, such as  grouping points or adding variable names, group names and sample sizes.
+%% 3 - not in statistical software
 Additionally, existing stand-alone software does not allow easy import into commonly used statistical software (such as \proglang{R}), or try to optimize the research pipelines to facilitate editing and reproducibility.
@@
 \pkg{metaDigitise} operates within the \proglang{R} environment making data extraction, analysis and export more streamlined.
+%% need to be careful abiou th use of the term 'raw data' as internally this means uncalibrated data
 It also provides users with options to conduct the necessary calculations on raw data immediately after extraction so that comparable summary statistics can be obtained quickly.
@@
 
+\begin{table}[t!]
+\centering
+\begin{tabular}{lcccc}
+\hline
+Function                  & metaDigitise & GraphClick & DataThief  & WebPlotDigitizer\\
+\hline
+Groups                    & \checkmark   & \checkmark & $\times$   & \checkmark  \\
+Scatterplots              & \checkmark   & \checkmark & \checkmark & \checkmark  \\
+Mean and error plots      & \checkmark   &            & \checkmark & \\
+Boxplots                  & \checkmark   &            & $\times$    & \\
+Histograms                & \checkmark   && $\times$    &\\
+Graph rotation            & \checkmark   &&&\\
+Asymmetric error bars     & $\times$     &&&\\
+Log axis                  & $\times$     &&&\\
+Dates                     & $\times$     && \checkmark &\\
+Entry of metadata         & \checkmark   &&&\\
+Calculation of mean/sd/n  & \checkmark   & $\times$    &            & \\
+Automated point detection & $\times$     &            &            & \\
+\hline
+\end{tabular}
+
+\caption{\label{tab:comparison} Comparison of functionality between different digitisation softwares.}
+
+\end{table}
+
+%%%---------------------------------
+%%%---------------------------------
+%%%---------------------------------
+
 \section{Directory Structure, Image Processing and Reproducibility}
@@
 \fct{metaDigitise} is designed to work on a directory containing images of figures copied from primary literature, in .png, .jpg, .tiff, .pdf format.
-This directory is sepcified to \fct{metaDigitise} through the \code{dir} argument.
+This directory is specified to \fct{metaDigitise} through the \code{dir} argument.
 The user is free to set their own broad directory structure (e.g.
@@
 In the case of scatter plots, \fct{metaDigitise} also returns the correlation coefficient between the points within each identified group.
+
+%%%---------------------------------
+%%%---------------------------------
+%%%---------------------------------
 
@@
 \end{equation}
-If the user does not enter a sample size at the time of data extraction (if for example the information is not readily available) the SD is not calculated.
+If the user does not enter a sample size at the time of data extraction (if, for example, the information is not readily available) the SD is not calculated.
 This can be entered at a later time, however (see below).
@@
 Different groups are plotted in different colours and shapes, with a legend at the bottom of the figure (Figure \ref{fig:all_extract}C).
-Mean, SD and n are calculated from the clicked points, for each group.
+Mean, SD and sample size (n) are calculated from the clicked points, for each group.
 In the case that the sample size from the clicked points does not match a known sample size (e.g.
@@
 \subsubsection{Histograms}
-\pkg{metaDigitise} prompts the user to enter groups names and then to click on the top corners of each bar.
+\fct{metaDigitise} prompts the user to enter groups names and then to click on the top corners of each bar.
 bars can subsequently be deleted.
-\pkg{MetaDigitise} then calculates the midpoint and the frequency for each bar.
+\fct{metaDigitise} then calculates the midpoint and the frequency for each bar.
+
+%%%---------------------------------
+%%%---------------------------------
+%%%---------------------------------
 
@@
 
+% maybe something about automated point detection - from our experience, as calibration needs to be done for each plot, and often points are not detected, it is much more reliable, and sometimes even quicker to manuslly select the points
+
 \section*{Acknowledgments}

### 2017-11-20 docs/manuscript.tex
Small changes

--- 2017-11-19 docs/manuscript.tex
@@
   Joel L. Pick, Shinichi Nakagawa, Daniel W.A. Noble\\
-  Ecology and Evolution Research Center\\
+  Ecology and Evolution Research Centre\\
   \emph{and}\\
@@
   University of New South Wales\\
-  b\\
-  b\\
+  Kensington, NSW 2052\\
+  Sydney, AUSTRALIA\\
   E-mail: \email{joel.l.pick@gmail.com}\\
@@
 \section{Introduction}
-In many different contexts, researchers need to make use of data presented in primary literature.
+In many different contexts, researchers need to make use of data presented from primary literature.
 Most notably, this includes meta-analysis, which is becoming increasingly common in many research fields.
-Meta-analysis uses effect size estimates and their sampling variance, taken from many studies, to understand whether particular effects are common across studies and to explain variation among these effects (Nakagawa et al.
-2017).
-Meta-analysis therefore relies foremost on data extracted from primary literature and more specifically statistics that have been reported in the text or tables of research papers.
-These summary statistics are, however, frequently presented in figures and so need to be manually extracted using digitising programs.
-Although there are several existing tools to perform tasks like this (EXAMPLES), these tools are not designed this specific purpose (i.e.
-a meta-analysis).
-Specifically, they do not differentiate between common plot types that are use to present data, meaning that they require a large amount of downstream data manipulation.
-For example, many stand alone programs only provide the raw data or values from plots.
-This means that these data need to be imported into other programs to calculate correlations (if scatter plots) or errors on means (from mean error plots) need to be back calculated by hand to convert them to a common type (e.g., standard errors to standard deviations).
-Digitising programs also do not easily allow for different grouping of points to be made, and frequently do not integrate metadata (such as variable names, group names and sample sizes) at the time of data extraction.
-Additionally, existing stand-alone software does not allow easy import into commonly used statistical software (such as \proglang{R}), or try to optimize the research pipelines to facilitate editing and reproducibility.
+Meta-analysis uses effect size estimates and their sampling variance, taken from many studies, to understand whether observed effects are common across studies and to explain variation among these effects \citep{Naka2017, Glass1976, Koricheva2013, Borenstein2009}.
+Meta-analysis therefore relies on statistics that have been reported, for example, in text or tables of research papers.
+However, summary statistics are very frequently presented in figures, and so, meta-analysts need to manually extract statistical information using digitising programs.
+Although there are several existing tools to perform these tasks (e.g., graphClick, dataThief), they are not designed for the unique requirements of meta-analysis.
+Specifically, they do not differentiate between common plot types that can be used to present data, meaning that they require a large amount of downstream data manipulation after digitising, data from figures.
+For example, many stand alone programs only provide the raw data or specific values from digitised plots.
+This means that these data need to be imported into other programs to calculate correlations (if scatter plots) or errors on means (from mean error plots).
+Additionally, errors often need to be converted to usable types (e.g., standard deviations from standard errors) manually.
+Digitising programs also do not easily distinguish different groups of points within a plot, and frequently do not integrate metadata (such as variable names, group names and sample sizes) during data extraction.
+Existing stand-alone software also do not allow easy import into commonly used statistical software (such as \proglang{R}), or try to optimize the research pipelines to facilitate editing and reproducibility.
 These are major issues because extracting from figures can be an incredibly time-consuming process.
-Furthermore, although meta-analysis is an important tool in consolidating the data from many studies, relying on transparency of those studies, many of the processes involved in data extraction are opaque and difficult to reproduce, making extending studies problematic.
-Having a tool that facilitates reproducibility in meta-analyses will increase transparency and go a long way to resolving the reproducibility crises we are seeing in many fields (REF).
+Furthermore, many of the processes involved in data extraction during a meta-analysis can be opaque and difficult to reproduce, making updating of meta-analyses problematic.
+Having a tool that facilitates reproducibility in meta-analyses will increase transparency and go a long way to resolving the reproducibility crises we are seeing in many fields \citep{peng_reproducible_2006, peng_reproducible_2011, sandve_ten_2013}.
 
@@
 Conveniently, when needing to process many figures at different times \pkg{metaDigitise} will only import figures not already completed within a directory.
-This makes it easy to add new figures at any time.
+This makes it easy to add new figures at any point during the projects lifetime.
 \pkg{metaDigitise} has also been built for reproducibility in mind.

### 2017-11-20 docs/manuscript.tex

--- 2017-11-20 docs/manuscript.tex
@@
 \section{Introduction}
+
 In many different contexts, researchers need to make use of data presented in primary literature.
 Most notably, this includes meta-analysis, which is becoming increasingly common in many research fields.
-Meta-analysis uses effect size estimates and their sampling variance, taken from many studies, to understand whether particular effects are common across studies and to explain variation among these effects (Nakagawa et al.
-2017).
+Meta-analysis uses effect size estimates and their sampling variance, taken from many studies, to understand whether particular effects are common across studies and to explain variation among these effects \citep{Naka2017, Glass1976, Koricheva2013, Borenstein2009}.
 Meta-analysis therefore relies foremost on data extracted from primary literature and more specifically statistics that have been reported in the text or tables of research papers.
@@
 Furthermore, although meta-analysis is an important tool in consolidating the data from many studies, relying on transparency of those studies, many of the processes involved in data extraction are opaque and difficult to reproduce, making extending studies problematic.
-Having a tool that facilitates reproducibility in meta-analyses will increase transparency and go a long way to resolving the reproducibility crises we are seeing in many fields (REF).
+Having a tool that facilitates reproducibility in meta-analyses will increase transparency and go a long way to resolving the reproducibility crises we are seeing in many fields \citep{peng_reproducible_2006, peng_reproducible_2011, sandve_ten_2013}.
 

### 2017-11-20 docs/manuscript.tex
manuscript updates

--- 2017-11-20 docs/manuscript.tex
@@
 
-\subsubsection{Histograms}
-\fct{metaDigitise} prompts the user to enter groups names and then to click on the top corners of each bar.
-bars can subsequently be deleted.
-\fct{metaDigitise} then calculates the midpoint and the frequency for each bar.
+\subsection{Histograms}
+\fct{metaDigitise} prompts the user to click on the top corners of each bar. Bars can subsequently be deleted. 
+For each bar a midpoint (mean x coordinates) and a frequency (mean y coordinates, rounded to the nearest integer) is calculated.
+The sample size is calculated as the sum of the frequencies., the mean as
+\begin{equation}
+\sigma = SE * \sqrt{n}
+\end{equation}
+and from 95\% confidence intervals as
+\begin{equation}
+\sigma = \frac{CI}{1.96} * \sqrt{n}
+\end{equation}
+
+Using the x coordinates, \fct{metaDigitise} then calculates the midpoint and the frequency for each bar.
 

### 2017-11-20 docs/manuscript.tex
manuscript editing (added table)

@@
 %% recommended packages
-\usepackage{thumbpdf,lmodern,amssymb}
+\usepackage{thumbpdf,lmodern,amssymb,rotating}
 
-% \usepackage[utf8]{inputenc}
+% \usepackage[utf8]{inputenc},
 % \usepackage{geometry} 
@@
 % \usepackage[none]{hyphenat}
- \usepackage{listings}
+% \usepackage{listings}
 % \usepackage[round]{natbib}
@@
 Although there are several existing tools to perform tasks like this (e.g.
-DataThief (REF), GraphClick (REF), WebPlotDigitizer (REF))
-, these tools are not designed this specific purpose (i.e. a meta-analysis). 
+DataThief \citep{DataThief}, GraphClick \citep{GraphClick}, WebPlotDigitizer \citep{WebPlotDigitizer}), these tools are not designed this specific purpose (i.e.
+a meta-analysis).
 %% need to spell out clear points
@@
 \end{figure}
-
-\begin{table}[t!]
-\centering
-\begin{tabular}{lcccc}
-\hline
-Function                  & metaDigitise & GraphClick & DataThief  & WebPlotDigitizer\\
-\hline
-Groups                    & \checkmark   & \checkmark & $\times$   & \checkmark  \\
-Scatterplots              & \checkmark   & \checkmark & \checkmark & \checkmark  \\
-Mean and error plots      & \checkmark   &            & \checkmark & \\
-Boxplots                  & \checkmark   &            & $\times$    & \\
-Histograms                & \checkmark   && $\times$    &\\
-Graph rotation            & \checkmark   &&&\\
-Asymmetric error bars     & $\times$     &&&\\
-Log axis                  & $\times$     &&&\\
-Dates                     & $\times$     && \checkmark &\\
-Entry of metadata         & \checkmark   &&&\\
-Calculation of mean/sd/n  & \checkmark   & $\times$    &            & \\
-Automated point detection & $\times$     &            &            & \\
-\hline
-\end{tabular}
-
-\caption{\label{tab:comparison} Comparison of functionality between different digitisation softwares.}
-
-\end{table}
 
@@
 
-\begin{figure}[!h] 
- \includegraphics[width=0.9\textwidth]{fig_all_extract.pdf} 
- \caption{Demonstration of data extraction from different plot types}
-\label{fig:all_extract}
-\end{figure}
-
 \subsection{Mean and error plots} 
@@
 In the case that the sample size from the clicked points does not match a known sample size (e.g.
-if there are overlaid points), the user can enter a alternate sample size.
+if there are overlaid points), the user can enter an alternate sample size.
 
@@
 \fct{metaDigitise} prompts the user to click on the top corners of each bar. Bars can subsequently be deleted. 
-For each bar a midpoint (mean x coordinates) and a frequency (mean y coordinates, rounded to the nearest integer) is calculated.
-The sample size is calculated as the sum of the frequencies., the mean as
+For each bar a midpoint (m; mean x coordinates) and a frequency (f; mean y coordinates, rounded to the nearest integer) is calculated. The sample size, mean and SD were calculated as:
 \begin{equation}
-\sigma = SE * \sqrt{n}
-\end{equation}
-and from 95\% confidence intervals as
-\begin{equation}
-\sigma = \frac{CI}{1.96} * \sqrt{n}
+n = \sum_{i=1}^n{f_i}
 \end{equation}
 
-Using the x coordinates, \fct{metaDigitise} then calculates the midpoint and the frequency for each bar.
+\begin{equation}
+\mu = \frac{\sum_{i=1}^n{m_i f_i}}{n}
+\end{equation}
+
+\begin{equation}
+\sigma = \sqrt{\frac{\sum_{i=1}^n{(m_i f_i - \mu f_i)^2}}{n-1}}
+\end{equation}
+
+As with the scatterplots, If the sample size from the extracted data does not match a known sample size, the user can enter an alternate sample size.
+
+\begin{figure}[!h] 
+ \includegraphics[width=0.9\textwidth]{fig_all_extract.pdf} 
+ \caption{Demonstration of data extraction from different plot types}
+\label{fig:all_extract}
+\end{figure}
 
@@
 
-We'll now demonstrate how \fct{metaDigitise} works using figures generated from the well known iris data set. Users can install the \pkg{metaDigitise} package from GitHub as follows:
+We will now demonstrate how \fct{metaDigitise} works using figures generated from the well known iris data set. Users can install the \pkg{metaDigitise} package from GitHub as follows:
 
@@
 
-\section{Conclusions and Future Enhancements}
+\section{Discussion and Conclusions}
 
-\pkg{metaDigitise} is available through GitHub.
+%% discuss specifically everything it can't do, and jsutify it slightly
+
+\begin{sidewaystable}[t!]
+\centering
+\begin{tabular}{lccccccc}
+\hline
+Function                  & metaDigitise & GraphClick$^1$ & DataThief$^2$  & DigitizeIt$^3$ & WebPlotDigitizer$^4$ & metagear$^5$ & digitize$^6$\\
+\hline
+Scatterplots              & \checkmark   & \checkmark & \checkmark & \checkmark & \checkmark     & \checkmark$^7$ & \checkmark \\
+Mean and error plots      & \checkmark   & \checkmark & \checkmark & $\times$   & $\times$       & \checkmark$^7$ & $\times$    \\
+Boxplots                  & \checkmark   & $\times$   & $\times$   & $\times$   & $\times$       & $\times$       & $\times$    \\
+Histograms                & \checkmark   & $\times$   & $\times$   & $\times$   & \checkmark$^7$ & $\times$       & $\times$    \\
+Graph rotation$^8$        & \checkmark   & \checkmark & \checkmark & \checkmark & \checkmark     & $\times$       & $\times$    \\
+Groups                    & \checkmark   & \checkmark & $\times$   & \checkmark & \checkmark     & $\times$       & $\times$    \\
+Entry of metadata         & \checkmark   & $\times$   & $\times$   & $\times$   & $\times$       & $\times$       & $\times$    \\
+Summarising data          & \checkmark   & $\times$   & $\times$   & $\times$   & $\times$       & $\times$       & $\times$    \\
+Multiple image processing & \checkmark   & $\times$   & $\times$   & $\times$   & $\times$       & $\times$       & $\times$    \\
+Editing digitisation$^9$  & \checkmark   & \checkmark & \checkmark & $\times$   & \checkmark     & $\times$       & $\times$    \\
+Automated point detection & $\times$     & \checkmark &     ?      & \checkmark & \checkmark     & \checkmark     & $\times$    \\
+Line extraction           & $\times$     & \checkmark & \checkmark & \checkmark & \checkmark     & $\times$       & $\times$    \\
+Zoom                      & $\times$     & \checkmark & \checkmark & \checkmark &       ?        & $\times$       & $\times$    \\
+Log axis                  & $\times$ *   & \checkmark & \checkmark & \checkmark & \checkmark     & $\times$       & $\times$    \\
+Dates                     & $\times$ *   & $\times$   & \checkmark & $\times$   & \checkmark     & $\times$       & $\times$    \\
+Asymmetric error bars     & $\times$     & $\times$   & \checkmark & $\times$   & $\times$       & $\times$       & $\times$    \\
+Freeware                  & \checkmark$^{10}$ & \checkmark$^{11}$ & \checkmark$^{11}$ & $\times$$^{11}$ & \checkmark$^{11}$ & \checkmark$^{10}$ & \checkmark$^{10}$\\
+\hline
+
+%http://datathief.org/DatathiefManual.pdf
+%http://www.digitizeit.de/#Features
+%https://automeris.io/WebPlotDigitizer/userManual.pdf
+
+\end{tabular}
+
+$^1$ \citet{GraphClick}
+$^2$ \citet{DataThief}
+$^3$ \citet{DigitizeIt}
+$^4$ \citet{WebPlotDigitizer}
+$^5$ \citet{Lajeunesse2016}
+$^6$ \citet{Poisot2011}
+$^7$ Only automated, no manual extraction.
+$^8$ Or handles rotated graphs. 
+$^9$ Allows saving, re-plotting and editing of data extraction.
+$^{10}$ \proglang{R} package.
+$^{11}$ Standalone software.
+
+\caption{\label{tab:comparison} Comparison of functionality between different digitisation softwares.}
+
+\end{sidewaystable}
+
+%\pkg{metaDigitise} is available through GitHub.
 WHile it is incredibly flexible already, our future plans include building in abilities to deal with log-transformed axes, including arguments for calculating standard deviations from 95\% confidence intervals and standard errors using the t-distribution to correct the t-value for small sample sizes (currently assumed t-values are 1.96 as is normal), dealing with asymmetric error bars and the possibility of zooming in plots such that greater accuracy can be achieved when digitising.
@@
 
-% maybe something about automated point detection - from our experience, as calibration needs to be done for each plot, and often points are not detected, it is much more reliable, and sometimes even quicker to manuslly select the points
+Although \pkg{metaDigitise} is already very flexible, and provides functionality not seen in any other package, from comparison with other packages (Table \ref{tab:comparison}) it is clear that there are some functions that it does not perform.
+A notable feature that \pkg{metaDigitise} lacks is automated point detection.
+This is available in several packages (Table \ref{tab:comparison}).
+However, from our experience of using these functions, as calibration (for point detection) needs to be done for each plot individually, and often some points are not detected, it is much more reliable, and sometimes even quicker to manually select the points.
+From the testing of our package (see above) figures can be extracted in around 1-2 minutes, including entry of metadata.
+We do not think that current automation methods give substantial benefits to this.
+
+Another feature that \pkg{metaDigitise} (currently) lacks, if a zoom function.
+This may enable users to gain greater accuracy when clicking on points.
+However, from our own experience, if you are using a reasonably sized screen then the accuracy is already high from these programs (see results above for example), and not much gain is to be had from zooming in on points.
+
+\pkg{metaDigitise} also does not extract lines from figures, as many other packages do.
+In our own experience, this is not particularly useful for meta-analysis, although we recognise that it may be useful in other fields.
+In this case we would recommend extracting data as a scatter plot, and clicking along the line in question.
+A model can then be fitted to these points to estimate the parameters needed.
+
+Finally, \pkg{metaDigitise} does not allow for asymmetric error bars.
+At present this is a deliberate omission, as it is not clear how best to derive SD from such data, given also that such asymmetric error bars may represent different things in different figures.
 

### 2017-11-20 docs/manuscript.tex
update manuscript

@@
 Firstly, they typically only provide the user with calibrated x,y coordinates from imported figures, and do not differentiate between common plot types that are use to present data.
-This means that a large amount of downstream data manipulation is subsequently required.
+This means that a large amount of downstream data manipulation is subsequently required, that differs for different plots types.
 For example, data are frequently presented in mean and error plots (Figure \ref{fig:plot_type}A), for which the user wants a mean and error estimate for each group presented in the figure.
@@
 Secondly, digitising programs do not easily allow the integration of metadata at the time of data extraction, such as grouping points or adding variable names, group names and sample sizes.
-%% 3 - not in statistical software
-Additionally, existing stand-alone software does not allow easy import into commonly used statistical software (such as \proglang{R}), or try to optimize the research pipelines to facilitate editing and reproducibility.
+This makes the downstream calculations more laborious, as the information has to be added later, in most cases using different software.
+%% 3 - in same statistical software and high throughput
+Finally, existing programs do not import a set of images and allow the user to systematically work through them.
+Instead they require the user to manually import images one by one, and export data into individual files, that need to be imported and edited using different software.
+In essence, existing software does not provide an optimized research pipeline to facilitate data extraction, editing and reproducibility.
 These are major issues because extracting from figures can be an incredibly time-consuming process.
@@
 Having a tool that facilitates reproducibility in meta-analyses will increase transparency and go a long way to resolving the reproducibility crises we are seeing in many fields \citep{peng_reproducible_2006, peng_reproducible_2011, sandve_ten_2013}.
+
+\begin{figure}[bt!] 
+\centering
+ \includegraphics[width=0.75\textwidth]{fig_plot_type.pdf} 
+ \caption{Four plot types that \pkg{metaDigitise} is designed to extract data from: A) mean and error plot, B) box plot, C) scatter plot and D) histogram.
+Data is taken from the iris dataset in R.
+A and B are plotted with the whole dataset, C and D are just the data for the species setosa.}
+\label{fig:plot_type}
+\end{figure}
 
@@
 %% need to be careful abiou th use of the term 'raw data' as internally this means uncalibrated data
-It also provides users with options to conduct the necessary calculations on raw data immediately after extraction so that comparable summary statistics can be obtained quickly.
-Summaries will condense multiple figures into data frames or lists (depending on the type of figures) and these objects can easily be exported from \proglang{R}, or if using the raw data, analysed in any way the user desires.
+It also provides users with options to conduct the necessary calculations on processed data immediately after extraction so that comparable summary statistics can be obtained quickly.
+\pkg{metaDigitise} condenses data extracted from multiple figures into a single data frame which can be can easily be used in downstream analysis within \proglang{R} or exported from \proglang{R}.
+Processed data can also be easily extracted and analysed in any way the user desires.
 Conveniently, when needing to process many figures at different times \pkg{metaDigitise} will only import figures not already completed within a directory.
@@
 This makes sharing figure digitisation and reproducing the work of others simple and easy, and allows meta-analysts to update meta-analyses more easily.
-
-\begin{figure}[b!] 
-%\onehalfspacing
- \includegraphics[width=0.9\textwidth]{fig_plot_type.pdf} 
- \caption{Four plot types that \pkg{metaDigitise} is designed to extract data from: A) mean and error plot, B) box plot, C) scatter plot and D) histogram.
-Data is taken from the iris dataset in R.
-A and B are plotted with the whole dataset, C and D are just the data for the species setosa.}
-\label{fig:plot_type}
-\end{figure}
 
@@
 In order to correctly extract data from figures \fct{metaDigitise} always requires the user to calibrate the axes in the figure.
-To do this, the user is required to click on two known points on the axis in question, and then enter the value of those points in the figure
-%(Figure \ref{fig:calibrate})
-.
+To do this, the user is required to click on two known points on the axis in question, and then enter the value of those points in the figure.
 Using this information, \fct{metaDigitise} then calculates the value of any clicked points in terms of the figure axes.
@@
 For scatter plots and histograms both axes are calibrated.
-
-% \begin{figure}[!h] 
-%  \includegraphics[width=0.9\textwidth]{fig_calibrate.pdf} 
-%  \caption{Axis calibration. The user defines two points on each axis and labels them according to the values shown in the figure.}
-% \label{fig:calibrate}
-% \end{figure}
 
@@
 \begin{figure}[!h]
- \includegraphics[width=0.9\textwidth]{fig_all_extract.pdf} 
+\centering 
+ \includegraphics[width=0.75\textwidth]{fig_all_extract.pdf} 
  \caption{Demonstration of data extraction from different plot types}
@@
 
-Assume that we would like to extract summary statistics from studies measuring sepal length or width in iris species for a fictitious project.
+Assume that the user would like to extract summary statistics from studies measuring sepal length or width in iris species for a fictitious project.
 There are a few studies that only present these data in figures.
-As we read papers found from a systematic search we can add in figures with relevant data to a "FiguresToExtract" folder as follows
+As the user readds papers found from a systematic search, they add figures with relevant data to a "FiguresToExtract" folder as follows
 
@@
 \begin{figure}[!h] 
- \includegraphics[width=0.9\textwidth]{001_Anderson_1935_Fig1.png} 
+ \includegraphics[width=0.75\textwidth]{001_Anderson_1935_Fig1.png} 
  \caption{Example scatterplot (001\_Anderson\_1935\_Fig1.png) of sepal length and width for two species of iris (setosa and versicolor)}
-\label{fig:rotate}
+\label{fig:Anderson}
 \end{figure}
@@
 Here, the naming of the files placed in the folder will contain the paper number, first author and the figure number to keep data uniquely associated with figures.
-We can use \fct{metaDigitise} to begin processing the image from the first paper.
+At first there is one figure in the folder, shown in Figure \ref{fig:Anderson}.
 Running \fct{metaDigitise} brings up a series of prompts for the user using a main menu that provides access to a number of its features:
@@
 
-The user simply enters in the numeric value that corresponds to what they would like to do. In this case we want to "Process new images".
-
-As soon as \fct{metaDigitise} is executed, it will bring the user to the main menu (see "Re-importing, Editing and Plotting Previously Digitised data" above).
-When the user selects option one ("Process new images"), we are asked whether we have different types of plot(s) in the folder.
-This question is most relevant when there are lots of different figures in the folder because it will then ask us for the type of figure as they are cycled through.
+The user simply enters in the numeric value that corresponds to what they would like to do.
+In this case they want to "Process new images".
+The user is then asked whether there is different types of plot(s) in the folder.
+This question is most relevant when there are lots of different figures in the folder because it will then ask the user for the type of figure as they are cycled through.
 
@@
 
-The user can just follow the instructions on screen step-by-step (instructions above have been truncated by `...` to simplify), and in the order specified.
+The user can just follow the instructions on screen step-by-step (instructions above have been truncated by `...' to simplify), and in the order specified.
 Before moving on, the user is forced to check whether or not the calibration has been set up correctly.
-If `n` is chosen because something needs to be fixed then the user can re-calibrate.
+If \code{n} is chosen because something needs to be fixed then the user can re-calibrate.
 
@@
 \fct{metaDigitise} handles this nicely by prompting the user to enter the group first, followed by the digitisation of this groups data.
-After digitising the first group, and having exited (i.e., hit `esc` from plot window), \fct{metaDigitise} will ask the user whether they would like to add another group.
-Users can continually add groups (\code{a}), or simply continue to the next plot (\code{c} - if another plot exists).
+After digitising the first group, and having exited (e.g., hit `esc' on mac from plot window), \fct{metaDigitise} will ask the user whether they would like to add another group.
+Users can continually add groups (\code{a}), delete groups (\code{d}), edit group (\code{e}) or finish plot and continue to the next one (\code{f} - if another plot exists).
 %%% This needs checking - think it is now different
@@
 \begin{CodeOutput}
-*figs_to_extract
+*FiguresToExtract
     + caldat
@@
 \begin{CodeOutput}
-*figs_to_extract
+*FiguresToExtract
 	+ caldat
@@
 
-We have already processed one figure (001\_Anderson\_1935\_Fig1.png) and we can tell because it has digitised data (caldat/001\_Anderson\_1935\_Fig1), but now we have our two new figures that have not yet been digitised.
+The user has already processed one figure (001\_Anderson\_1935\_Fig1.png).
+We can tell this because the caldat folder has digitised data in it (caldat/001\_Anderson\_1935\_Fig1).
+Now the user has two new figures that have not yet been digitised.
 This example will nicely demonstrate how users can easily pick up from where they left off and how all previous data gets re-integrated.
@@
 \begin{CodeInput}
-R> setwd("~/figs_to_extract")
-R> digitised_data <- metaDigitise(".", summary = TRUE)
+R> digitised_data <- metaDigitise("./FiguresToExtract", summary = TRUE)
 \end{CodeInput}
@@
 First for 002\_Doe\_2013\_Fig1.png and then 002\_Doe\_2013\_Fig3.png.
-Since there are a few new figures, we answer the first question in the \proglang{R} console as "diff":
+Since there are a few new figures, the user answers the first question in the \proglang{R} console as "diff":
 
@@
 
-Here, we specify the new plot type as \code{m} for 002\_Doe\_2013\_Fig1.png because we have a plot of the mean and error of sepal length for each of the three species.
-We're then prompted a bit differently from our scatter plot as the x-axis is not needed for calibration:
+Here, the user specifies the new plot type as \code{m} for 002\_Doe\_2013\_Fig1.png because the user has a plot of the mean and error of sepal length for each of the three species.
+The user is then prompted a bit differently from our scatter plot as the x-axis is not needed for calibration:
 
@@
 
-\subsection{Obtaining Raw Data}
+\subsection{Obtaining Processed Data}
 
-While \fct{metaDigitise} provides users with the summary statistics by default, for all plot types, in many cases the user may actually be interested in obtaining the raw digitised data from scatter plots.
+While \fct{metaDigitise} provides users with the summary statistics by default, for all plot types, in many cases the user may actually be interested in obtaining the processed digitised data from scatter plots (i.e.
+calibrated points).
 This is very easy to do my changing the default \code{summary} argument from TRUE to FALSE in \fct{metaDigitise}.
@@
 
-Here, the user can easily access the list of raw scatter plot data by simply extracting the scatter plot slot:
+Here, the user can easily access the list of processed scatter plot data by simply extracting the scatter plot slot:
 
@@
 
-%check extraction
+\subsection{Checking Extraction}
+
+\subsection{Post-hoc adding smaple sizes}
 %Including addition of N later
@@
 \begin{sidewaystable}[t!]
-\centering
+{\centering
 \begin{tabular}{lccccccc}
@@
 Multiple image processing & \checkmark   & $\times$   & $\times$   & $\times$   & $\times$       & $\times$       & $\times$    \\
-Editing digitisation$^9$  & \checkmark   & \checkmark & \checkmark & $\times$   & \checkmark     & $\times$       & $\times$    \\
+Reproducable$^9$  & \checkmark   & \checkmark & \checkmark & $\times$   & \checkmark     & $\times$       & $\times$    \\
 Automated point detection & $\times$     & \checkmark &     ?      & \checkmark & \checkmark     & \checkmark     & $\times$    \\
@@
 \end{tabular}
-
+}
+{\footnotesize
+\\
 $^1$ \citet{GraphClick}
@@
 $^6$ \citet{Poisot2011}
-$^7$ Only automated, no manual extraction.
-$^8$ Or handles rotated graphs. 
-$^9$ Allows saving, re-plotting and editing of data extraction.
-$^{10}$ \proglang{R} package.
-$^{11}$ Standalone software.
+\\$^7$ Only automated, no manual extraction.
+\\$^8$ Or handles rotated graphs. 
+\\$^9$ Allows saving, re-plotting and editing of data extraction.
+\\$^{10}$ \proglang{R} package.
+\\$^{11}$ Standalone software.
+}
 

### 2017-11-21 docs/manuscript.tex
Edits of MS and added in some more paragraphs here and there

@@
   Ecology and Evolution Research Centre\\
-  \emph{and}\\
   School of Biological, Earth and Environmental Sciences\\
@@
 Although there are several existing tools to perform tasks like this (e.g.
-DataThief \citep{DataThief}, GraphClick \citep{GraphClick}, WebPlotDigitizer \citep{WebPlotDigitizer}), these tools are not designed this specific purpose (i.e.
-a meta-analysis).
+DataThief \citep{DataThief}, GraphClick \citep{GraphClick}, WebPlotDigitizer \citep{WebPlotDigitizer}), these tools are not designed specifically for meta-analysis.
 %% need to spell out clear points
@@
 %% 2- adding metadata
-Secondly, digitising programs do not easily allow the integration of metadata at the time of data extraction, such as grouping points or adding variable names, group names and sample sizes.
+Secondly, digitising programs do not easily allow the integration of metadata at the time of data extraction, such as experimental group or variable names, and sample sizes.
 This makes the downstream calculations more laborious, as the information has to be added later, in most cases using different software.
@@
 These are major issues because extracting from figures can be an incredibly time-consuming process.
-Furthermore, although meta-analysis is an important tool in consolidating the data from many studies, relying on transparency of those studies, many of the processes involved in data extraction are opaque and difficult to reproduce, making extending studies problematic.
-Having a tool that facilitates reproducibility in meta-analyses will increase transparency and go a long way to resolving the reproducibility crises we are seeing in many fields \citep{peng_reproducible_2006, peng_reproducible_2011, sandve_ten_2013}.
+Furthermore, although meta-analysis is an important tool in consolidating the data from many studies many of the processes involved in data extraction are opaque and difficult to reproduce, making extending studies problematic.
+Having a tool that facilitates reproducibility in meta-analyses will increase transparency and go a long way to resolving the reproducibility crises we are seeing in many fields \citep{peng_reproducible_2006, peng_reproducible_2011, sandve_ten_2013, Parker2016}.
 
@@
 It also provides users with options to conduct the necessary calculations on processed data immediately after extraction so that comparable summary statistics can be obtained quickly.
-\pkg{metaDigitise} condenses data extracted from multiple figures into a single data frame which can be can easily be used in downstream analysis within \proglang{R} or exported from \proglang{R}.
-Processed data can also be easily extracted and analysed in any way the user desires.
+\pkg{metaDigitise} condenses summary data extracted from multiple figures into a single data frame which can be can easily be exported.
+Processed data (i.e., raw data) can also be easily extracted and analysed in any way the user desires in downstream analysis within \proglang{R}.
 Conveniently, when needing to process many figures at different times \pkg{metaDigitise} will only import figures not already completed within a directory.
@@
 
-It is important to for the user to think about their directory structure early on in this process (also more generally in the context of their entire project), especially if they plan to share the extractions with collaborators or when publishing the project.
+It is important for the user to think about their directory structure early on in this process (also more generally in the context of their entire project), especially if they plan to share the extractions with collaborators or when publishing the project.
 
@@
 The data from a completed image is automatically saved in an .RDS file to a \code{caldat} directory, that is created within the parent directory when first executing the \fct{metaDigitise} function.
-These files enable re-plotting and editing of images at a later point (see below), but do not need to be directly accessed by the user.
+These files enable re-plotting and editing of images at a later point (see below).
 
 A particularly powerful and flexible aspect of \fct{metaDigitise} is its ability to identify images that have been previously digitised and only import images that have not been digitised in subsequent calls of the function.
-This means both that all figures do not need to be extracted at one time and that new figures can be added as the project develops.
+This means that all figures do not need to be extracted at one time and that new figures can be added as the project develops.
 After each image is extracted, the user is asked whether they wish to continue or quit the extraction process.
-Upon rerunning \fct{metaDigitise}, previously digitised figures are simply ignored during processing, but their data re-integrated within the final output after new files are completed.
+Upon rerunning \fct{metaDigitise}, previously digitised figures are simply ignored during processing, but their data is re-integrated within the final output after new files are completed automatically.
 
@@
 \subsection{Mean and error plots} 
-\fct{metaDigitise} prompts the user to enter groups names and allows the user to enter sample sizes ($n$), which are used in downstream processing.
+\fct{metaDigitise} prompts the user to enter group names and allows the user to enter sample sizes ($n$), which are used in downstream processing.
 The user is then prompted to click on an error bar followed by the mean.
@@
 \subsection{Box plots}
-As with mean and error plots, \fct{metaDigitise} prompts the user to enter groups names and allows the user to enter sample sizes ($n$), which are used in downstream processing.
+As with mean and error plots, \fct{metaDigitise} prompts the user to enter group names and allows the user to enter sample sizes ($n$), which are used in downstream processing.
 The user is then prompted to click on the maximum ($b$), upper quartile ($q_3$), median ($m$), lower quartile ($q_1$) and minimum ($a$).
-\fct{metaDigitise} will check that the maximum is greater than the minimum, and return an warning if that is not the case.
+\fct{metaDigitise} will check that the maximum is greater than the minimum, and return a warning if that is not the case.
 The user can subsequently add, edit or remove groups.
@@
 Mean, SD and sample size (n) are calculated from the clicked points, for each group.
-In the case that the sample size from the clicked points does not match a known sample size (e.g.
+Where the sample size from the clicked points does not match a known sample size (e.g.
 if there are overlaid points), the user can enter an alternate sample size.
@@
 \fct{metaDigitise} prompts the user to click on the top corners of each bar. Bars can subsequently be deleted. 
-For each bar a midpoint (m; mean x coordinates) and a frequency (f; mean y coordinates, rounded to the nearest integer) is calculated. The sample size, mean and SD were calculated as:
+For each bar a midpoint (m; mean x coordinates) and a frequency (f; mean y coordinates, rounded to the nearest integer) is calculated. The sample size, mean and SD are calculated as:
 \begin{equation}
@@
 There are a few studies that only present these data in figures.
-As the user readds papers found from a systematic search, they add figures with relevant data to a "FiguresToExtract" folder as follows
+As the user reads papers found from a systematic search, they add figures with relevant data to a "FiguresToExtract" folder as follows
 
@@
 \begin{CodeInput}
-R> digitised_data <- metaDigitise("~/FiguresToExtract", summary = TRUE)
+R> setwd("~/FiguresToExtract")
+R> digitised_data <- metaDigitise(".", summary = TRUE)
 \end{CodeInput}
@@
 \begin{CodeInput}
-c
+R> c
 \end{CodeInput}
@@
  scatter plot or histogram m/b/s/h:
+ R> s
 \end{CodeOutput}
@@
 Re-calibrate? (y/n) 
+R> n
 \end{CodeOutput}
@@
 Often, plots might contain multiple groups that the meta-analyst wants to extract from.
-\fct{metaDigitise} handles this nicely by prompting the user to enter the group first, followed by the digitisation of this groups data.
+\fct{metaDigitise} handles this nicely by prompting the user to enter the group first, followed by digitisation of this groups data.
 After digitising the first group, and having exited (e.g., hit `esc' on mac from plot window), \fct{metaDigitise} will ask the user whether they would like to add another group.
-Users can continually add groups (\code{a}), delete groups (\code{d}), edit group (\code{e}) or finish plot and continue to the next one (\code{f} - if another plot exists).
+Users can continually add groups (\code{a}), delete groups (\code{d}), edit groups (\code{e}) or finish a plot and continue to the next one (\code{f} - if another plot exists).
 %%% This needs checking - think it is now different
@@
 One thing anyone with a familiarity with the iris dataset will notice is that the sample sizes for each of these species (which are n = 50 each) are quite a bit lower.
-This is an example of some of the challenges when extracting data from scatter plots, often data points will overlap with each other making it impossible (without having the real data) to know whether this is a problem.
+This is an example of some of the challenges when extracting data from scatter plots.
+Often data points will overlap with each other making it impossible (without having the real data) to know whether this is a problem.
 However, a meta-analyst will probably realise that the sample sizes here conflict with what is reported in the paper.
-Hence, \pkg{metaDigitise} also provides the user with  options to input the sample sizes directly (see Editing section above), even for scatter plots and histograms where, strictly speaking, this should not be necessary.
+Hence, \pkg{metaDigitise} also provides the user with  options to input the sample sizes directly (see Editing section below), even for scatter plots and histograms where, strictly speaking, this should not be necessary.
 Nonetheless, it is important to recognise the impact that overlapping points can have on summary statistics, particularly its effects on standard deviation (SD) and standard error (SE).
@@
 \begin{CodeInput}
-R> digitised_data <- metaDigitise("./FiguresToExtract", summary = TRUE)
+R> setwd("~/FiguresToExtract")
+R> digitised_data <- metaDigitise(".", summary = TRUE)
 \end{CodeInput}
@@
 
-The user gets the same set of prompts and simply chooses option one to digitise the new figures.
+The user gets the same set of prompts and simply chooses option one.
 This will permit users to digitise new figures, and will integrate previously completed digitisations along with newly digitised data together at the end of the session, or when the user decides to quit.
@@
 First for 002\_Doe\_2013\_Fig1.png and then 002\_Doe\_2013\_Fig3.png.
-Since there are a few new figures, the user answers the first question in the \proglang{R} console as "diff":
+Since there are a few different figure types, the user answers the first question in the \proglang{R} console as "diff":
 
@@
 Again, \fct{metaDigitise} will simply guide the user through digitising each of these figures describing to them exactly what needs to be done.
-At any point if mistakes are made the user can choose relevant options to edit or correct things.
-This can continues for each plot so long as the user would like to continue after completing a single plot:
+At any point if mistakes are made the user can choose relevant options to edit or correct things before ending the figure.
+This process continues for each plot so long as the user would like to continue and after completing a single plot the user is always prompted as follows:
 
@@
 \begin{CodeInput}
-R> metaDigitise("./FiguresToExtract")
+R> setwd("~/FiguresToExtract")
+R> digitised_data <- metaDigitise(".")
 \end{CodeInput}
@@
 
-A whole host of information can be edited from the rotation, plot type, the variable name that was provided, the calibration and even the digitisation of groups.
+A whole host of information can be edited including the rotation, plot type, the variable name(s) that were provided, the calibration and even the digitisation of groups.
 When editing the \code{metaDigitise} object is re-written to the caldat folder and the edits are immediately integrated into the existing object once complete.
@@
 \subsection{Figure Rotation and Adjustment}
-Figure may have been extracted from old publications, for example from scanned images, and so are not perfectly orientated on the image.
+Figures may have been extracted from old publications, for example from scanned images, and so are not perfectly orientated on the image.
 This will make the calibration of the points in the figure from the image problematic.
@@
  \includegraphics[width=0.9\textwidth]{fig_rotate.pdf} 
- \caption{Figure rotation. A) and B) show how non-aligned imaged can be realigned through user defined rotation. C) and D) show show figures can be re-orientated so as to aid data input.}
+ \caption{Figure rotation. A) and B) show how non-aligned images can be realigned through user defined rotation. C) and D) show how figures can be re-orientated so as to aid data input.}
 \label{fig:rotate}
@@
 
-Furthermore, in some figures mean and error, boxplots or histograms may be presented with horizontal bars.
-\fct{metaDigitise} assumes that the bars are vertical, but allows the user to flip the image so that the bars are vertical (Figure \ref{fig:rotate}C,D).
+Furthermore, some figures, including mean and error, boxplots or histograms, may be presented with horizontal bars.
+\fct{metaDigitise} assumes that the bars are vertical, but allows the user to flip the image so that the bars are vertical if provided horizontally (Figure \ref{fig:rotate}C,D).
 
@@
 
-\subsection{Checking Extraction}
-
-\subsection{Post-hoc adding smaple sizes}
+\subsection{Adding sample sizes to previous Digitisations}
 %Including addition of N later
@@
 %% x people doing y plots
+
+We evaluated the consistency of digitisation using the \pkg{metaDigitise} by simulating a series of data and using these data in each of the four different plot types (scatterplot, mean and error, histogram and boxplots).
+Each variable was plotted twice and figures were modified slightly to give users a sense that they were digitising new data.
+We randomised the order of these same 14 figures (using the same data), and provdied the entire figure directory to a total of 15 indpendent digitisers.
+Digitisers variaed in their level of experience from those having done meta-analyses or comparative work to those not having any experience doing so.
+We got users to digitise all 14 figures and then send the data on these to evaluate interobserver variablity for each figure type and to assess the accuracy and differences in digitising data from the same data across different figure types.
 
@@
 
-Although \pkg{metaDigitise} is already very flexible, and provides functionality not seen in any other package, from comparison with other packages (Table \ref{tab:comparison}) it is clear that there are some functions that it does not perform.
+Although \pkg{metaDigitise} is already very flexible, and provides functionality not seen in any other packages (Table \ref{tab:comparison}) it is clear that there are some functions that it does not perform.
 A notable feature that \pkg{metaDigitise} lacks is automated point detection.
 This is available in several packages (Table \ref{tab:comparison}).
-However, from our experience of using these functions, as calibration (for point detection) needs to be done for each plot individually, and often some points are not detected, it is much more reliable, and sometimes even quicker to manually select the points.
-From the testing of our package (see above) figures can be extracted in around 1-2 minutes, including entry of metadata.
-We do not think that current automation methods give substantial benefits to this.
+However, from our experience of using these functions, manual digitising is more reliable and often equally as fast.
+Particularly given that calibration (for point detection) needs to be done for each plot individually in any case.
+Additionally, auto-detection often misses many points which then subsequently need to be manually added.
+Bssed on tests of \pkg{metaDigitise} (see above) figures can be extracted in around 1-2 minutes, including the entry of metadata.
+As a result, we do not belive that current automated point detection provides substantial time saving benefits.
 
-Another feature that \pkg{metaDigitise} (currently) lacks, if a zoom function.
+Another feature that \pkg{metaDigitise} (currently) lacks, is an ability to zoom in plots.
 This may enable users to gain greater accuracy when clicking on points.
-However, from our own experience, if you are using a reasonably sized screen then the accuracy is already high from these programs (see results above for example), and not much gain is to be had from zooming in on points.
+However, from our own experience (and indeed the results from our inter-observer reliability), if you are using a reasonably sized screen then the accuracy is already high from these programs (see results above for example), and there is not much gain is to be had from zooming in on points in many circumstances.
 
-\pkg{metaDigitise} also does not extract lines from figures, as many other packages do.
+In contrast to some other packages, \pkg{metaDigitise} currently also does not extract lines from figures.
 In our own experience, this is not particularly useful for meta-analysis, although we recognise that it may be useful in other fields.
-In this case we would recommend extracting data as a scatter plot, and clicking along the line in question.
+In this case, we would recommend extracting data as a scatter plot, and clicking along the line in question.
 A model can then be fitted to these points to estimate the parameters needed.
 
-Finally, \pkg{metaDigitise} does not allow for asymmetric error bars.
+Finally, \pkg{metaDigitise} currently also does not allow for asymmetric error bars.
 At present this is a deliberate omission, as it is not clear how best to derive SD from such data, given also that such asymmetric error bars may represent different things in different figures.
 
+Increasing the reproducibility of figure extraction for meta-analysis and making the laborious process more streamlined, flexible and integrated with existing statistical software will go a long way in facilitating the production of high quality meta-analytic studies that can be updated in the future.
+We belive that \pkg{metaDigitise} will improve this research synthesis pipeline, and will hopefully become an integral package that can be added to the meta-analysts toolkit.
+
 \section*{Acknowledgments}
-We thanks the i-deel group at UNSW for useful feedback, and ...
+We thanks the I-DEEL group at UNSW for increadibly useful feedback, and a host of colleagues for testing, providing feedback and digitising including: Rose O'Dea, Fonti Kar, Malgorzata Lagisz, Julia Riley, Diego Barneche, Erin Macartney, Gihan Samarasinghe.
+DWAN was supported by an ARC Discovery Early Career Research Award (DE150101774) and UNSW Vice Chancellors Fellowship and SN an Australian Research Council Future Fellowship (FT130100268).
 

### 2017-11-21 docs/manuscript.tex
Added awknoledgements

@@
 \section*{Acknowledgments}
-We thanks the I-DEEL group at UNSW for increadibly useful feedback, and a host of colleagues for testing, providing feedback and digitising including: Rose O'Dea, Fonti Kar, Malgorzata Lagisz, Julia Riley, Diego Barneche, Erin Macartney, Gihan Samarasinghe.
+We thanks the I-DEEL group at UNSW for increadibly useful feedback, and a host of colleagues for testing, providing feedback and digitising including: Rose O'Dea, Fonti Kar, Malgorzata Lagisz, Julia Riley, Diego Barneche, Erin Macartney, Ivan Beltran, Gihan Samarasinghe.
 DWAN was supported by an ARC Discovery Early Career Research Award (DE150101774) and UNSW Vice Chancellors Fellowship and SN an Australian Research Council Future Fellowship (FT130100268).

### 2017-11-21 docs/analysis.R
Adding some script for analysing between observer digitising

@@ -1,0 +1,13 @@
+
+#Grab all the files that are relevant
+	files <- list.files("./docs/sim_figs/", recursive = TRUE, pattern = ".csv")
+	paths <- paste0("./docs/sim_figs/", files)
+	names <- gsub(".+/", "", x = files)
+
+# Import all the data
+	datalist <- lapply(paths, function(x) read.csv(x))
+	names(datalist) <- names
+
+# Change list to data and get the unique plot names for analysis 
+	data <- ldply(datalist)
+	data$plot <- gsub("[0-9]+_", "", data$filename)

### 2017-11-21 docs/manuscript.tex
Added some small MS changes thanking Dax

--- 2017-11-21 docs/manuscript.tex
@@
 \section*{Acknowledgments}
-We thanks the I-DEEL group at UNSW for increadibly useful feedback, and a host of colleagues for testing, providing feedback and digitising including: Rose O'Dea, Fonti Kar, Malgorzata Lagisz, Julia Riley, Diego Barneche, Erin Macartney, Gihan Samarasinghe.
-DWAN was supported by an ARC Discovery Early Career Research Award (DE150101774) and UNSW Vice Chancellors Fellowship and SN an Australian Research Council Future Fellowship (FT130100268).
+We thanks the I-DEEL group at UNSW for increadibly useful feedback, and a host of colleagues for testing, providing feedback and digitising including: Rose O'Dea, Fonti Kar, Malgorzata Lagisz, Julia Riley, Diego Barneche, Erin Macartney, Ivan Beltran, Gihan Samarasinghe, Dax Kellie.
+DWAN was supported by an Australian Research Council Discovery Early Career Research Award (DE150101774) and UNSW Vice Chancellors Fellowship and SN an Australian Research Council Future Fellowship (FT130100268).
 

### 2017-11-21 docs/analysis.R

@@ -1,4 +1,5 @@
 
+rm(list=ls())
 #Grab all the files that are relevant
 	files <- list.files("./docs/sim_figs/", recursive = TRUE, pattern = ".csv")
 	paths <- paste0("./docs/sim_figs/", files)
@@ -11,3 +12,27 @@
 # Change list to data and get the unique plot names for analysis 
 	data <- ldply(datalist)
 	data$plot <- gsub("[0-9]+_", "", data$filename)
+	data$dat <- gsub("[0-9][0-9][0-9]", "", data$plot)
+
+# Model
+  # mean
+	modX <- lmer(mean ~ dat + (1|.id), REML = FALSE, data = data)
+	summary(modX)
+
+	modX2 <- lm(mean ~ dat, data = data)
+
+	anova(modX, modX2)
+
+	# SD
+	modSD <- lmer(sd ~ dat + (1|.id), REML = FALSE, data = data)
+	summary(modSD)
+
+	modSD2 <- lm(sd ~ dat, data = data)
+	anova(modSD, modSD2)
+
+	# r
+	modr <- lmer(r ~ 1 + (1|.id), REML = FALSE, data = data)
+	summary(modr)
+
+	modr2 <- lm(r ~ 1, data = data)
+	anova(modr, modr2)

### 2017-11-21 docs/analysis.R
Updated MS

@@ -36,3 +36,7 @@ rm(list=ls())
 
 	modr2 <- lm(r ~ 1, data = data)
 	anova(modr, modr2)
+
+	# Check out BLUPS
+	coef(modr)
+	ranef(modr)

### 2017-11-22 R/H_extract.R
histogram editing bug fix

@@ -11,7 +11,7 @@ histogram_extract <- function(edit=FALSE, raw_data = data.frame(), ...){
 		if(idQ=="y"){
 			group_id <- readline("\nGroup identifier: ")
 			raw_data$id <- group_id
-		}
+		}else{group_id<-raw_data$id}
 	}else{
 		group_id <- readline("\nGroup identifier: ")
 	}

### 2017-11-22 docs/manuscript.tex

@@
 \section*{Acknowledgments}
-We thanks the I-DEEL group at UNSW for increadibly useful feedback, and a host of colleagues for testing, providing feedback and digitising including: Rose O'Dea, Fonti Kar, Malgorzata Lagisz, Julia Riley, Diego Barneche, Erin Macartney, Ivan Beltran, Gihan Samarasinghe, Dax Kellie.
+We thanks the I-DEEL group at UNSW for increadibly useful feedback, and a host of colleagues for testing, providing feedback and digitising including: Rose O'Dea, Fonti Kar, Malgorzata Lagisz, Julia Riley, Diego Barneche, Erin Macartney, Ivan Beltran, Gihan Samarasinghe, Dax Kellie, Jonathan Noble.
 DWAN was supported by an Australian Research Council Discovery Early Career Research Award (DE150101774) and UNSW Vice Chancellors Fellowship and SN an Australian Research Council Future Fellowship (FT130100268).

### 2017-11-22 docs/sim_figs/analysis (Autosaved).R
Updates to MS

@@ -1,0 +1,84 @@
+
+rm(list=ls())
+#Grab all the files that are relevant
+	
+	setwd("~/Dropbox/sim_figs")
+
+	files <- list.files(".", recursive = TRUE, pattern = ".csv")
+	paths <- paste0("./", files)
+	names <- gsub(".+/", "", x = files)
+
+# Import all the data
+
+	datalist <- lapply(paths, function(x) cbind(read.csv(x, stringsAsFactors=FALSE), id=gsub(".+/", "", x = x)))
+#	names(datalist) <- names
+
+# Change list to data and get the unique plot names for analysis 
+#ldply
+	data <- do.call(rbind,datalist)
+	rownames(data) <- 1:nrow(data)
+	data$axis <- ifelse(data$axis=="",substr(data$filename,nchar(data$filename)-4,nchar(data$filename)-4),data$axis)
+	data$group <- ifelse(is.na(data$group), "all",data$group)
+
+head(data)
+
+#	data$plot <- gsub("[0-9]+_", "", data$filename)
+#	data$dat <- gsub("[0-9][0-9][0-9]", "", data$plot)
+
+
+##generated data
+	set.seed(86)
+	x <- rnorm(20, 3, 1)
+	grp <- rep(c(1,0), each = 10)
+	y <- rnorm(20, 4 + 2*grp + 1.2*x + -0.5*(x*grp), 1)
+
+	gen_dat <- data.frame(dat=c(x,y), axis=rep(c("x","y"),each=20), group=rep(grp,2))
+
+	gen_means <- rbind(aggregate(dat~group+axis,gen_dat,mean), cbind(group="all", aggregate(dat~axis,gen_dat,mean)))
+	colnames(gen_means)[3]<-"gen_mean"
+	gen_sd <- rbind(aggregate(dat~group+axis,gen_dat,sd), cbind(group="all", aggregate(dat~axis,gen_dat,sd)))
+	colnames(gen_sd)[3]<-"gen_sd"
+
+#gen_r <- 
+
+	merge_dat <- merge(merge(data,gen_means), gen_sd)
+
+
+par(mfcol=c(2,4))
+for(i in unique(merge_dat$plot_type)) {
+	within(subset(merge_dat, plot_type==i),{ 
+		hist((mean-gen_mean)/mean, col="grey", main=paste("mean -",i), breaks=50)
[…]

### 2017-11-22 docs/sim_figs/analysis (Autosaved).R
Fixing some weird file structure changes

@@ -1,84 +1,0 @@
-
-rm(list=ls())
-#Grab all the files that are relevant
-	
-	setwd("~/Dropbox/sim_figs")
-
-	files <- list.files(".", recursive = TRUE, pattern = ".csv")
-	paths <- paste0("./", files)
-	names <- gsub(".+/", "", x = files)
-
-# Import all the data
-
-	datalist <- lapply(paths, function(x) cbind(read.csv(x, stringsAsFactors=FALSE), id=gsub(".+/", "", x = x)))
-#	names(datalist) <- names
-
-# Change list to data and get the unique plot names for analysis 
-#ldply
-	data <- do.call(rbind,datalist)
-	rownames(data) <- 1:nrow(data)
-	data$axis <- ifelse(data$axis=="",substr(data$filename,nchar(data$filename)-4,nchar(data$filename)-4),data$axis)
-	data$group <- ifelse(is.na(data$group), "all",data$group)
-
-head(data)
-
-#	data$plot <- gsub("[0-9]+_", "", data$filename)
-#	data$dat <- gsub("[0-9][0-9][0-9]", "", data$plot)
-
-
-##generated data
-	set.seed(86)
-	x <- rnorm(20, 3, 1)
-	grp <- rep(c(1,0), each = 10)
-	y <- rnorm(20, 4 + 2*grp + 1.2*x + -0.5*(x*grp), 1)
-
-	gen_dat <- data.frame(dat=c(x,y), axis=rep(c("x","y"),each=20), group=rep(grp,2))
-
-	gen_means <- rbind(aggregate(dat~group+axis,gen_dat,mean), cbind(group="all", aggregate(dat~axis,gen_dat,mean)))
-	colnames(gen_means)[3]<-"gen_mean"
-	gen_sd <- rbind(aggregate(dat~group+axis,gen_dat,sd), cbind(group="all", aggregate(dat~axis,gen_dat,sd)))
-	colnames(gen_sd)[3]<-"gen_sd"
-
-#gen_r <- 
-
-	merge_dat <- merge(merge(data,gen_means), gen_sd)
-
-
-par(mfcol=c(2,4))
-for(i in unique(merge_dat$plot_type)) {
-	within(subset(merge_dat, plot_type==i),{ 
-		hist((mean-gen_mean)/mean, col="grey", main=paste("mean -",i), breaks=50)
[…]

### 2017-11-22 R/S_extract.R
small bug fix

@@ -7,6 +7,7 @@ delete_group <- function(raw_data){
 	ids <- unique(raw_data$id)
 	remove <- menu(ids)
 	raw_data <- subset(raw_data, id != ids[remove])
+	raw_data$id <- droplevels(raw_data$id)
 	return(raw_data)
 }
 

### 2017-11-22 docs/manuscript.tex

--- 2017-11-21 docs/manuscript.tex
@@
 * Main project directory
-    + FiguresToExtract
+    + FiguresToExtract/
         + Paper1_Figure1_trait1.png
@@
 \begin{CodeOutput}
-*FiguresToExtract
+*FiguresToExtract/
     + 001_Anderson_1935_Fig1.png
@@
 \begin{CodeOutput}
-*FiguresToExtract
-    + caldat
+*FiguresToExtract/
+    + caldat/
         + 001_Anderson_1935_Fig1
@@
                   filename   group_id         variable  mean  sd  n    r  plot_type
-1 001_Anderson_1935_Fig1.png     setosa  Sepal Width (mm) 3.420332 0.4034259 39 0.7524909 scatterplot
-2 001_Anderson_1935_Fig1.png     setosa Sepal Length (mm) 4.998267 0.3831294 39 0.7524909 scatterplot
-3 001_Anderson_1935_Fig1.png versicolor  Sepal Width (mm) 2.766857 0.3234723 44 0.5157938 scatterplot
-4 001_Anderson_1935_Fig1.png versicolor Sepal Length (mm) 5.936769 0.5292535 44 0.5157938 scatterplot
+001_Anderson_1935_Fig1.png     setosa  Sepal Width (mm) 3.42 0.40 39 0.75 scatterplot
+001_Anderson_1935_Fig1.png     setosa Sepal Length (mm) 5.00 0.38 39 0.75 scatterplot
+001_Anderson_1935_Fig1.png versicolor  Sepal Width (mm) 2.77 0.32 44 0.52 scatterplot
+001_Anderson_1935_Fig1.png versicolor Sepal Length (mm) 5.94 0.53 44 0.52 scatterplot
 \end{CodeOutput}
@@
 \begin{CodeOutput}
-*FiguresToExtract
-	+ caldat
+*FiguresToExtract/
+	+ caldat/
 	  	+ 001_Anderson_1935_Fig1
@@
                filename      group_id         variable  mean  sd   n    r   plot_type
-1 001_Anderson_1935_Fig1.png     setosa  Sepal Width (mm) 3.420332 0.4034259 39 0.7524909 scatterplot
-2 001_Anderson_1935_Fig1.png     setosa Sepal Length (mm) 4.998267 0.3831294 39 0.7524909 scatterplot
-3 001_Anderson_1935_Fig1.png versicolor  Sepal Width (mm) 2.766857 0.3234723 44 0.5157938 scatterplot
-4 001_Anderson_1935_Fig1.png versicolor Sepal Length (mm) 5.936769 0.5292535 44 0.5157938 scatterplot
-5      002_Doe_2013_Fig1.png     setosa      Sepal length 5.000336 0.7828656 50        NA  mean_error
-6      002_Doe_2013_Fig1.png versicolor      Sepal length 5.931340 1.0466177 50        NA  mean_error
-7      002_Doe_2013_Fig1.png viriginica      Sepal length 6.588705 1.2608173 50        NA  mean_error
-8      003_Doe_2013_Fig3.png       <NA>      Sepal length 4.948472 0.3624212 50        NA  histogram
+001_Anderson_1935_Fig1.png     setosa  Sepal Width (mm) 3.42 0.40 39 0.75 scatterplot
+001_Anderson_1935_Fig1.png     setosa Sepal Length (mm) 5.00 0.38 39 0.75 scatterplot
+001_Anderson_1935_Fig1.png versicolor  Sepal Width (mm) 2.77 0.32 44 0.52 scatterplot
+001_Anderson_1935_Fig1.png versicolor Sepal Length (mm) 5.94 0.53 44 0.52 scatterplot
+     002_Doe_2013_Fig1.png     setosa      Sepal length 5.00 0.78 50   NA  mean_error
+     002_Doe_2013_Fig1.png versicolor      Sepal length 5.93 1.05 50   NA  mean_error
+     002_Doe_2013_Fig1.png viriginica      Sepal length 6.59 1.26 50   NA  mean_error
+     003_Doe_2013_Fig3.png       <NA>      Sepal length 4.95 0.36 50   NA  histogram
 \end{CodeOutput}
@@
   .. ..$ frequency: num [1:8] 4 5 7 12 11 6 2 3
-  .. ..$ variable : chr [1:8] "Sepal length" "Sepal length" "Sepal length" "Sepal length" ...
+  .. ..$ variable : chr [1:8] "Sepal length" "Sepal length" ...
  $ scatterplot:List of 1
   ..$ 001_Anderson_1935_Fig1.png:'data.frame':  83 obs. of  8 variables:
-  .. ..$ id        : Factor w/ 2 levels "setosa","versicolor": 1 1 1 1 1 1 1 1 1 1 ...
+  .. ..$ id        : Factor w/ 2 levels "setosa","versicolor": 1 1 1 1 1 ...
   .. ..$ x         : num [1:83] 2.3 2.9 3 3 3 ...
@@
   .. ..$ group     : num [1:83] 1 1 1 1 1 1 1 1 1 1 ...
-  .. ..$ col       : Factor w/ 2 levels "red","green": 1 1 1 1 1 1 1 1 1 1 ...
+  .. ..$ col       : Factor w/ 2 levels "red","green": 1 1 1 1 1 1 1 1 1 ...
   .. ..$ pch       : num [1:83] 19 19 19 19 19 19 19 19 19 19 ...
-  .. ..$ y_variable: chr [1:83] "Sepal length (mm)" "Sepal length (mm)" "Sepal length (mm)" "Sepal length (mm)" ...
-  .. ..$ x_variable: chr [1:83] "Sepal width (mm)" "Sepal width (mm)" "Sepal width (mm)" "Sepal width (mm)" ...
+  .. ..$ y_variable: chr [1:83] "Sepal length (mm)" "Sepal length (mm)"  ...
+  .. ..$ x_variable: chr [1:83] "Sepal width (mm)" "Sepal width (mm)"    ...
 \end{CodeOutput}
@@
 
-\section{Testing}
+\section{Interobserver Variability and Validation}
 %% simulated data
 %% x people doing y plots
-
-We evaluated the consistency of digitisation using the \pkg{metaDigitise} by simulating a series of data and using these data in each of the four different plot types (scatterplot, mean and error, histogram and boxplots).
-Each variable was plotted twice and figures were modified slightly to give users a sense that they were digitising new data.
-We randomised the order of these same 14 figures (using the same data), and provdied the entire figure directory to a total of 15 indpendent digitisers.
-Digitisers variaed in their level of experience from those having done meta-analyses or comparative work to those not having any experience doing so.
-We got users to digitise all 14 figures and then send the data on these to evaluate interobserver variablity for each figure type and to assess the accuracy and differences in digitising data from the same data across different figure types.
+\subsection{Methods}
+We evaluated the consistency of digitisation using \pkg{metaDigitise} and tested how well digitisations estimate the true underlying data.
+This was accomplished by simulating a dataset with two different groups and two traits, and these data were then used to construct four different plot types (scatterplot, mean and error, histogram and boxplots).
+Each variable was plotted twice for a given plot type (figures were modified slightly to give users a sense that they were digitising new data) generating a total of 14 figures.
+We randomised the order of these figures (using the same data) and provdied the entire figure directory to a total of 15 indpendent digitisers.
+The same data was presented twice for each figure type to estimate between- and within- user variability (i.e., the intra-class correlation coefficient) in the data.
+Digitisers varied in their level of experience from those having done meta-analyses or comparative work to those not having any experience doing so.
+We got users to digitise all 14 figures and then send the data on these to evaluate interobserver variablity.
+
+We analysed these data using linear mixed effect models with user identify as a random effect.
+The overall model also included plot type to account for the different means across the variables and plots.
+We used the mean, standard deviation and correlation coeficients as response variables in three seperate models.
+If digitisations were consistent across all users then we would predict no significant between user variability in the data.
+We also qualitatively report how close the average digitised mean, SD and correlation coefficient were to the actual mean values of traits presented in the figures.
+
+\subsection{Results}
 
@@
 
-%\pkg{metaDigitise} is available through GitHub.
-WHile it is incredibly flexible already, our future plans include building in abilities to deal with log-transformed axes, including arguments for calculating standard deviations from 95\% confidence intervals and standard errors using the t-distribution to correct the t-value for small sample sizes (currently assumed t-values are 1.96 as is normal), dealing with asymmetric error bars and the possibility of zooming in plots such that greater accuracy can be achieved when digitising.
-We hope to also provide options for assessing inter-observer reliability in the future.
-
 Although \pkg{metaDigitise} is already very flexible, and provides functionality not seen in any other packages (Table \ref{tab:comparison}) it is clear that there are some functions that it does not perform.
@@
 \section*{Acknowledgments}
-We thanks the I-DEEL group at UNSW for increadibly useful feedback, and a host of colleagues for testing, providing feedback and digitising including: Rose O'Dea, Fonti Kar, Malgorzata Lagisz, Julia Riley, Diego Barneche, Erin Macartney, Ivan Beltran, Gihan Samarasinghe, Dax Kellie.
+We thank the I-DEEL group at UNSW for increadibly useful feedback, and a host of colleagues for testing, providing feedback and digitising including: Rose O'Dea, Fonti Kar, Malgorzata Lagisz, Julia Riley, Diego Barneche, Erin Macartney, Ivan Beltran, Gihan Samarasinghe, Dax Kellie, Jonathan Noble and Yian Noble.
 DWAN was supported by an Australian Research Council Discovery Early Career Research Award (DE150101774) and UNSW Vice Chancellors Fellowship and SN an Australian Research Council Future Fellowship (FT130100268).

### 2017-11-23 R/H_extract.R
Small asthetic changes

@@ -9,11 +9,11 @@ histogram_extract <- function(edit=FALSE, raw_data = data.frame(), ...){
 	if(edit){ 
 		idQ <- user_options("Change group identifier? (y/n) ",c("y","n"))
 		if(idQ=="y"){
-			group_id <- readline("\nGroup identifier: ")
+			group_id <- readline("\nGroup identifier: \n")
 			raw_data$id <- group_id
 		}else{group_id<-raw_data$id}
 	}else{
-		group_id <- readline("\nGroup identifier: ")
+		group_id <- readline("\nGroup identifier: \n")
 	}
 
 
@@ -24,7 +24,7 @@ histogram_extract <- function(edit=FALSE, raw_data = data.frame(), ...){
 		
 		if(histQ=="a"){
 			i=i+1
-			cat("Click on left then right upper corners of bar\n")
+			cat("Click on the left followed by the right upper corners of a single bar")
 			bar_points <- locator(2, type="o", col="red", lwd=2, pch=19)
 			raw_data <- rbind(raw_data, data.frame(id=group_id, x=bar_points$x,y=bar_points$y, bar=i))
 		}
@@ -35,7 +35,7 @@ histogram_extract <- function(edit=FALSE, raw_data = data.frame(), ...){
 		}
 
 		internal_redraw(raw_data=raw_data,...)
-		histQ <- readline("Add bar, Delete bar or Finish plot? a/d/f ")
+		histQ <- readline("Add bar, Delete bar or Finish plot? (a/d/f) \n")
 
 	}
 	return(raw_data)

### 2017-11-23 R/ME_Box_extraction.R

@@ -11,7 +11,7 @@ single_MB_extract <- function(plot_type){
 	if(plot_type=="boxplot"){
 		cat("\nClick on Max, Upper Q, Median, Lower Q, and Minimum\nIn that order\n\n")
 		group_points <- locator(5, type="o", col="red", lwd=2, pch=19)
-		if(group_points$y[1]<group_points$y[5]) warning("max is smaller than min - consider deleting and re-adding group", call. = FALSE, immediate. = TRUE)
+		if(group_points$y[1]<group_points$y[5]) warning("max is smaller than min - consider deleting and re-adding group\n", call. = FALSE, immediate. = TRUE)
 	}
 
 	return(group_points)
@@ -29,6 +29,7 @@ single_MB_extract <- function(plot_type){
 MB_extract <- function(edit=FALSE, plot_type, entered_N, raw_data = data.frame(), ...){
 
 	add_removeQ <- if(edit){ "b" }else{ "a" }
+	if(!edit) cat("\nIf there are multiple groups, enter unique group identifiers (otherwise press enter)")
 
 	while(add_removeQ != "f"){
 		
@@ -69,7 +70,7 @@ MB_extract <- function(edit=FALSE, plot_type, entered_N, raw_data = data.frame()
 
 
 		internal_redraw(plot_type=plot_type, raw_data=raw_data,...)
-		add_removeQ <- readline("Add group, Edit Group, Delete group or Finish plot? a/e/d/f ")
+		add_removeQ <- readline("Add group, Edit Group, Delete group or Finish plot? (a/e/d/f) \n")
 
 	}
 	return(raw_data)

### 2017-11-23 R/S_extract.R

@@ -57,7 +57,7 @@ edit_group <- function(raw_data, group_id,...){
 		}
 
 		internal_redraw(...,raw_data=rbind(raw_data, group_data), calibration=TRUE, points=TRUE)
-		add_removeQ <- readline("\nAdd points, delete points or continue? a/d/c ")
+		add_removeQ <- readline("\nAdd points, delete points or continue? a/d/c \n")
 	}
 
 	raw_data <- rbind(raw_data, group_data)		
@@ -101,7 +101,7 @@ group_scatter_extract <- function(edit=FALSE, raw_data = data.frame(), ...){
 		if(editQ == "d") raw_data <- delete_group(raw_data)
 	
 		internal_redraw(...,raw_data=raw_data, calibration=TRUE, points=TRUE)
-		editQ <- readline("\nAdd group, Edit group, Delete group, or Finish plot? a/e/d/f ")
+		editQ <- readline("\nAdd group, Edit group, Delete group, or Finish plot? a/e/d/f \n")
 	}
 	return(raw_data)
 }

### 2017-11-23 R/calibrate_coords.R

@@ -7,7 +7,7 @@ print_cal_instructions <- function(plot_type,...){
   cat( "\nOn the Figure, click IN ORDER: 
       y1, y2",if(!plot_type %in% c("mean_error","boxplot")) ", x1, x2", " \n
 
-    Step 1 ----> Click on y1
+    Step 1 ----> Click on known value on y axis - y1
   |
   |
   |
@@ -15,7 +15,7 @@ print_cal_instructions <- function(plot_type,...){
   y1
   |_________________________
 \n
-    Step 2 ----> Click on y2
+    Step 2 ----> Click on another known value on y axis - y2
   |
   y2
   |
@@ -24,7 +24,7 @@ print_cal_instructions <- function(plot_type,...){
   |_________________________
   \n",
 if(!plot_type %in% c("mean_error","boxplot")) "
-    Step 3 ----> Click on x1
+    Step 3 ----> Click on known value on x axis - x1
   |
   |
   |
@@ -32,7 +32,7 @@ if(!plot_type %in% c("mean_error","boxplot")) "
   |
   |_____x1__________________
 \n
-    Step 4 ----> Click on x2
+    Step 4 ----> Click on another known value on x axis - x2
   |
   |
   |
@@ -76,7 +76,7 @@ cal_coords <- function(plot_type,...) {
 getVals <- function(calpoints,...) {
   names <- if(nrow(calpoints)==2){c("y1","y2")}else{c("y1","y2","x1","x2")}
   vals <- NULL
-  for (i in names) vals[i] <- user_numeric(paste("What is the value of", i, "?\n"))
+  for (i in names) vals[i] <- user_numeric(paste("\nWhat is the value of", i, "?"))
   return(vals)
 }
 
@@ -96,7 +96,7 @@ user_calibrate <- function(object){
       object$point_vals <- do.call(getVals, object)
       do.call(internal_redraw, c(object, calibration=TRUE, points=FALSE))
     }
-    cal_Q <- readline("\nRe-calibrate? (y/n) ")
+    cal_Q <- readline("\nRe-calibrate? (y/n) \n")
   }
   return(object[c("calpoints","point_vals")])
   ##output object?

### 2017-11-23 R/graph_rotation.R

@@ -41,7 +41,7 @@ Otherwise chose continue\n
 			rotate <- rotate + f
 		}
 		internal_redraw(image_file, flip = flip, rotate = rotate)	
-		rotateQ <- readline("Flip, rotate or continue f/r/c ")
+		rotateQ <- readline("Flip, rotate or continue (f/r/c) \n")
 	}
 
 	out <- list(flip=flip, rotate=rotate)

### 2017-11-23 R/metaDigitise.R

@@ -44,7 +44,7 @@ process_new_files <- function(dir, summary = TRUE) {
 			       setup_calibration_dir(dir)
 			      done_details <- dir_details(dir)
 	     details <- get_notDone_file_details(dir)
-		   type <- user_options("Are all plot types the same? (diff/same)\n" , c("diff", "same"))
+		   type <- user_options("Are all plot types Different or the Same? (d/s)" , c("d", "s"))
 	
 	if(length(done_details$calibrations) >= 1){	
 		done_objects <- load_metaDigitise(done_details$doneCalFiles, done_details$names)
@@ -57,7 +57,7 @@ process_new_files <- function(dir, summary = TRUE) {
 		names = NULL
 	}
 
-	 plot_types <-  if (type == "diff") {NULL} else { specify_type() }
+	 plot_types <-  if (type == "d") {NULL} else { specify_type() }
 		 
 		 data_list <- list()
 
@@ -70,7 +70,7 @@ process_new_files <- function(dir, summary = TRUE) {
 				breakQ <-  user_options(paste("\n\nDo you want continue:", length(details$paths)- i, "plots out of", length(details$paths), "plots remaining (y/n) "), c("y","n"))
 				if(breakQ=="n") break
 			}else{
-				cat("Congratulations! Looks like you have finished digitising all figures in this directory. If you haven't please delete files from the caldat/ folder and try again!")
+				cat("Congratulations! Looks like you have finished digitising all figures in this directory.\n")
 			}
 		 }
 	
@@ -116,7 +116,7 @@ specify_type <- function(){
 		#user enters numeric value to specify the plot BEFORE moving on
 	 	pl_type <- NA
 	 	#while keeps asking the user the question until the input is one of the options
-		while(!pl_type %in% c("m","b","s","h")) pl_type <- readline("Please specify the plot_type as either: mean and error, box plot, scatter plot or histogram m/b/s/h: ")
+		while(!pl_type %in% c("m","b","s","h")) pl_type <- readline("Please specify the plot_type as either:\n\n m: Mean and error\n b: Box plot\n s: Scatter plot \n h: Histogram\n\n ")
 	
 	 	plot_type <- ifelse(pl_type == "m", "mean_error", ifelse(pl_type == "b", "boxplot",ifelse(pl_type == "s", "scatterplot","histogram")))
 	
@@ -173,7 +173,7 @@ get_notDone_file_details <- function(dir){
 	# Check whether there are new files still needing to be done.
 
 	if(length(details$calibrations) == length(details$name)) {
[…]

### 2017-11-23 R/metaDigitise_functions.R

@@ -38,7 +38,7 @@ internal_digitise <- function(image_file, plot_type=NULL, cex=1){
 
 	### N entered?
 	if(plot_type %in% c("mean_error","boxplot")) {
-		askN <- user_options("\nEnter sample sizes? y/n ",c("y","n"))
+		askN <- user_options("\nDo you know sample sizes? (y/n) ",c("y","n"))
 		output$entered_N <- ifelse(askN =="y", TRUE, FALSE)
 	}else{
 		output$entered_N <- TRUE

### 2017-11-23 R/misc_func.R

@@ -30,7 +30,7 @@ filename <- function(x) {
 user_options <- function(question, allowed_answers) {
 	input_bad <- TRUE
 	while(input_bad){
-		input <- readline(question)
+		input <- readline(paste(question,"\n"))
 		input_bad <- !input %in% allowed_answers
 		if(input_bad) cat("\n**** Invalid response ****\n")
 	}
@@ -45,7 +45,7 @@ user_options <- function(question, allowed_answers) {
 user_unique <- function(question, previous_answers) {
 	input_bad <- TRUE
 	while(input_bad){
-		input <- readline(question)
+		input <- readline(paste(question,"\n"))
 		input_bad <- input %in% previous_answers
 		if(input_bad) cat("\n**** Response must be unique ****\n")
 	}
@@ -59,7 +59,7 @@ user_unique <- function(question, previous_answers) {
 user_numeric <- function(question) {
 	input_bad <- TRUE
 	while(input_bad){
-    	input <- suppressWarnings(as.numeric(readline(question)))
+    	input <- suppressWarnings(as.numeric(readline(paste(question,"\n"))))
 		input_bad <- is.na(input)
 		if(input_bad) cat("\n**** Input must be numeric ****\n")
 	}
@@ -74,7 +74,7 @@ user_numeric <- function(question) {
 user_count <- function(question) {
 	input_bad <- TRUE
 	while(input_bad){
-		input <- suppressWarnings(as.numeric(readline(question)))
+		input <- suppressWarnings(as.numeric(readline(paste(question,"\n"))))
 		input_bad <- is.na(input)| input<1 | !is.wholenumber(input)
 		if(input_bad) cat("\n**** Input must be an integer above 0 ****\n")
 	}
@@ -89,11 +89,13 @@ user_count <- function(question) {
 
 ask_variable <- function(plot_type){
 	if(plot_type == "scatterplot"){
-		y_variable <- readline("\nWhat is the y variable? ")
-		x_variable <- readline("\nWhat is the x variable? ")
+		y_variable <- readline("\nWhat is the y variable? \n")
+		x_variable <- readline("\nWhat is the x variable? \n")
 		variable <- c(y=y_variable,x=x_variable)
+	}else if(plot_type == "histogram"){
+		variable <- c(readline("\nWhat is the x variable? \n"))
 	}else{
-		variable <- c(readline("\nWhat is the variable? "))
+		variable <- c(readline("\nWhat is the y variable? \n"))
 	}
 	return(variable)
 }

### 2017-11-23 R/transformStats.R

@@ -40,17 +40,6 @@ CI95_to_sd <- function(CI,n) {
 	CI/1.96 * sqrt(n)
 }
 
-# #' @title range_to_sd
-# #' @description Transforms range to standard deviation
-# #' @param CI Interval difference from the mean
-# #' @param n Sample Size
-# #' @return Returns vector of standard deviations
-# #' @author Joel Pick
-# #' @export
-
-# range_to_sd <- function(min,max,n) {
-# 	CI/1.96 * sqrt(n)
-# }
 
 
 #' @title rqm_to_mean

### 2017-11-23 R/H_extract.R
tidying prompts, and some bug fixes

@@ -5,7 +5,6 @@
 #' @description Extraction of data from histograms
 histogram_extract <- function(edit=FALSE, raw_data = data.frame(), ...){
 
-
 	if(edit){ 
 		idQ <- user_options("Change group identifier? (y/n) ",c("y","n"))
 		if(idQ=="y"){
@@ -13,6 +12,7 @@ histogram_extract <- function(edit=FALSE, raw_data = data.frame(), ...){
 			raw_data$id <- group_id
 		}else{group_id<-raw_data$id}
 	}else{
+		cat("\nIf there are multiple groups, enter unique group identifiers (otherwise press enter)")
 		group_id <- readline("\nGroup identifier: \n")
 	}
 

### 2017-11-23 R/S_extract.R

@@ -47,7 +47,9 @@ edit_group <- function(raw_data, group_id,...){
 		if(add_removeQ=="a"){		
 			cat("\nClick on points you want to add.\nIf you want to remove a point, or are finished with a\ngroup, exit (see above), then follow prompts\n")
 			select_points <- locator(type="p", lwd=2, col=cols[i], pch=pchs[i])
-			group_data <- rbind(group_data, data.frame(id=group_id, x=select_points$x, y=select_points$y, group=i, col=cols[i], pch=pchs[i]) )		
+			if(!is.null(select_points)){
+				group_data <- rbind(group_data, data.frame(id=group_id, x=select_points$x, y=select_points$y, group=i, col=cols[i], pch=pchs[i]) )
+			}		
 		}
 
 		if(add_removeQ=="d"){
@@ -58,7 +60,7 @@ edit_group <- function(raw_data, group_id,...){
 		}
 
 		internal_redraw(...,raw_data=rbind(raw_data, group_data), calibration=TRUE, points=TRUE)
-		add_removeQ <- readline("\nAdd points, delete points or continue? a/d/c \n")
+		add_removeQ <- readline("\nAdd or Delete points to this group, or Continue? (a/d/c) \n")
 	}
 
 	raw_data <- rbind(raw_data, group_data)		
@@ -83,6 +85,7 @@ group_scatter_extract <- function(edit=FALSE, raw_data = data.frame(), ...){
     sep = "\n\n")
 
 	editQ <- if(edit){ "b" }else{ "a" }
+	if(!edit) cat("\nIf there are multiple groups, enter unique group identifiers (otherwise press enter)")
 
 	while(editQ != "f"){
 	
@@ -102,7 +105,7 @@ group_scatter_extract <- function(edit=FALSE, raw_data = data.frame(), ...){
 		if(editQ == "d") raw_data <- delete_group(raw_data)
 	
 		internal_redraw(...,raw_data=raw_data, calibration=TRUE, points=TRUE)
-		editQ <- readline("\nAdd group, Edit group, Delete group, or Finish plot? a/e/d/f \n")
+		editQ <- readline("\nAdd group, Edit group, Delete group, or Finish plot? (a/e/d/f) \n")
 	}
 	return(raw_data)
 }

### 2017-11-23 R/edit_metaDigitise.R

@@ -12,7 +12,7 @@ bulk_edit <- function(dir, summary=TRUE){
 
 	if(length(filepaths)==0) stop("No files to edit", call.=FALSE)
 
-	cat("Choose how you want to edit files:\n")
+	cat("\nChoose how you want to edit files:\n")
 	Q <- menu(c("Cycle through images","Choose specific file to edit","Enter previously omitted sample sizes"))
 
 # cycle 
@@ -86,7 +86,7 @@ N_needed <- function(filepaths){
 enter_N <- function(raw_data,...){
 	ids <- 	unique(raw_data$id)
 	for (i in ids){
-		raw_data[raw_data$id==i,"n"] <- user_count(paste("Group \"", i,"\": Enter sample size "))
+		raw_data[raw_data$id==i,"n"] <- user_count(paste("\nGroup \"", i,"\": Enter sample size "))
 	}
 	return(raw_data)
 }
@@ -108,12 +108,12 @@ edit_metaDigitise <- function(object){
 
 	## ROTATION
 	rotQ <- user_options("\nEdit rotation? If yes, then the whole extraction will be redone (y/n) ", c("y","n"))
-	if(rotQ=="y") output <- metaDigitise(object$image_file)
+	if(rotQ=="y") output <- internal_digitise(object$image_file)
 
 
 	### plot type
 	ptQ <- user_options("\nChange plot type? If yes, then the whole extraction will be redone (y/n) ", c("y","n"))
-	if(ptQ=="y") output <- metaDigitise(object$image_file)
+	if(ptQ=="y") output <- internal_digitise(object$image_file)
 	
 
 	### variables

### 2017-11-23 R/import_metaDigitise.R

@@ -11,7 +11,7 @@ import_menu<-function(dir, summary){
 
 	if(length(filepaths) == 0) stop("No digitised files to import!", call. = FALSE)
 
-	cat("Import all extracted data or from one image:\n")
+	cat("\nImport all extracted data or from one image:\n")
 	Q <- menu(c("All","One"))
 
 # all 

### 2017-11-23 R/metaDigitise.R

@@ -23,7 +23,7 @@ metaDigitise<-function(dir, summary = TRUE){
 		dir <- paste0(dir, "/")
 	}
 
-	cat("Do you want to...\n")
+	cat("\nDo you want to...\n")
 	
 	Q <- menu(c("Process new images", "Import existing data", "Edit existing data"))
 	
@@ -44,7 +44,7 @@ process_new_files <- function(dir, summary = TRUE) {
 			       setup_calibration_dir(dir)
 			      done_details <- dir_details(dir)
 	     details <- get_notDone_file_details(dir)
-		   type <- user_options("Are all plot types Different or the Same? (d/s)" , c("d", "s"))
+		   type <- user_options("\nAre all plot types Different or the Same? (d/s)" , c("d", "s"))
 	
 	if(length(done_details$calibrations) >= 1){	
 		done_objects <- load_metaDigitise(done_details$doneCalFiles, done_details$names)
@@ -70,7 +70,7 @@ process_new_files <- function(dir, summary = TRUE) {
 				breakQ <-  user_options(paste("\n\nDo you want continue:", length(details$paths)- i, "plots out of", length(details$paths), "plots remaining (y/n) "), c("y","n"))
 				if(breakQ=="n") break
 			}else{
-				cat("Congratulations! Looks like you have finished digitising all figures in this directory.\n")
+				cat("\nCongratulations! Looks like you have finished digitising all figures in this directory.\n")
 			}
 		 }
 	
@@ -116,7 +116,7 @@ specify_type <- function(){
 		#user enters numeric value to specify the plot BEFORE moving on
 	 	pl_type <- NA
 	 	#while keeps asking the user the question until the input is one of the options
-		while(!pl_type %in% c("m","b","s","h")) pl_type <- readline("Please specify the plot_type as either:\n\n m: Mean and error\n b: Box plot\n s: Scatter plot \n h: Histogram\n\n ")
+		while(!pl_type %in% c("m","b","s","h")) pl_type <- readline("\nPlease specify the plot_type as either:\n\n m: Mean and error\n b: Box plot\n s: Scatter plot \n h: Histogram\n\n ")
 	
 	 	plot_type <- ifelse(pl_type == "m", "mean_error", ifelse(pl_type == "b", "boxplot",ifelse(pl_type == "s", "scatterplot","histogram")))
 	
@@ -173,7 +173,7 @@ get_notDone_file_details <- function(dir){
 	# Check whether there are new files still needing to be done.
 
[…]

### 2017-11-23 R/metaDigitise_functions.R

@@ -104,7 +104,8 @@ summary.metaDigitise<-function(object, ...){
 			group_id=pd$id,
 			variable=pd$variable,
 			mean=pd$mean,
-			sd=error_to_sd(error=pd$error, n=pd$n, error_type=object$error_type),
+			error=pd$error, 
+			error_type=object$error_type,
 			n=pd$n,
 			r=NA
 		)
@@ -116,7 +117,8 @@ summary.metaDigitise<-function(object, ...){
 			group_id=pd$id,
 			variable=pd$variable,
 			mean=rqm_to_mean(min=pd$min,LQ=pd$q1,median=pd$med,UQ=pd$q3,max=pd$max),
-			sd=rqm_to_sd(min=pd$min,LQ=pd$q1,UQ=pd$q3,max=pd$max,n=pd$n),
+			error = rqm_to_sd(min=pd$min,LQ=pd$q1,UQ=pd$q3,max=pd$max,n=pd$n),
+			error_type = "sd",
 			n=pd$n,
 			r=NA
 		)
@@ -129,7 +131,8 @@ summary.metaDigitise<-function(object, ...){
 						group_id=z$id[1],
 						variable=c(z$x_variable[1],z$y_variable[1]),
 						mean=apply(z[,c("x","y")],2,mean),
-						sd=apply(z[,c("x","y")],2,sd),
+						error = apply(z[,c("x","y")],2,sd),
+						error_type = "sd",
 					 	n=ifelse(is.null(object$knownN), nrow(z), object$knownN[z$id[1]]),
 					 	r=cor(z$x,z$y)
 					)
@@ -143,11 +146,13 @@ summary.metaDigitise<-function(object, ...){
 			group_id=pd$id[1],
 			variable=pd$variable[1],
 			mean=mean(hist_data),
-			sd=sd(hist_data),
+			error = sd(hist_data),
+			error_type = "sd",
 			n=ifelse(is.null(object$knownN), length(hist_data), object$knownN),
 			r=NA
 		)
 	}
+	out$sd <- with(out, error_to_sd(error=error,n=n,error_type=error_type))
 	out$plot_type <- object$plot_type
 	return(out)
 }

### 2017-11-23 R/transformStats.R

@@ -7,10 +7,12 @@
 #' @author Joel Pick
 #' @export
 
-error_to_sd <- function(error, n, error_type=c("se","CI95","sd")){
-	if(error_type=="se") sd <- se_to_sd(error, n)
-	if(error_type=="CI95") sd <- CI95_to_sd(error, n)
-	if(error_type=="sd") sd <- error
+error_to_sd <- function(error, n, error_type=c("se","CI95","sd",NA)){
+	 sd <- ifelse(error_type=="se", se_to_sd(error, n),
+			ifelse(error_type=="CI95", CI95_to_sd(error, n),
+			ifelse(error_type=="sd", error,
+			ifelse(is.na(error_type), NA
+			))))
 	return(sd)
 }
 

### 2017-11-23 tests/testthat/test-metaDigitise_functions.R
fixed tests

@@ -23,6 +23,6 @@ test_that("Checking summary.metaDigitise..", {
 	expect_equal(
 		summary.metaDigitise(mock_object),
 		data.frame(
-			filename=as.factor("image.png"), group_id=as.factor("control"), variable=as.factor("y"), mean=1.5, sd=0.25, n=20, r=NA, plot_type="mean_error", stringsAsFactors = FALSE)
+			filename=as.factor("image.png"), group_id=as.factor("control"), variable=as.factor("y"), mean=1.5, error=0.25, error_type=as.factor("sd"), n=20, r=NA, sd=0.25,plot_type="mean_error", stringsAsFactors = FALSE)
 	)
 })

### 2017-11-23 docs/manuscript.tex
Updated the documentation to make it jive with new prompts and data output

@@
 In this case they want to "Process new images".
-The user is then asked whether there is different types of plot(s) in the folder.
+The user is then asked whether there are different types of plot(s) in the folder.
 This question is most relevant when there are lots of different figures in the folder because it will then ask the user for the type of figure as they are cycled through.
@@
 \begin{CodeOutput}
-Are all plot types the same? (diff/same)
+Are all plot types Different or the Same? (d/s)
 \end{CodeOutput}
@@
 
-Flip, rotate or continue f/r/c 
+Flip, rotate or continue (f/r/c) 
 \end{CodeOutput}
@@
 \begin{CodeOutput}
-Please specify the plot_type as either: mean and error, box plot,
- scatter plot or histogram m/b/s/h:
+Please specify the plot_type as either:
+
+ m: Mean and error
+ b: Box plot
+ s: Scatter plot 
+ h: Histogram
+
  R> s
@@
 
-    Step 1 ----> Click on y1
+    Step 1 ----> Click on known value on y axis - y1
   |
@@
 
-    Step 3 ----> Click on x1
+    Step 3 ----> Click on known value on x axis - x1
   |
@@
 
-Add points, delete points or continue? a/d/c 
+Add points, delete points or continue? (a/d/c) 
 \end{CodeOutput}
@@
 \begin{CodeOutput}
-                  filename   group_id         variable  mean  sd  n    r  plot_type
-001_Anderson_1935_Fig1.png     setosa  Sepal Width (mm) 3.42 0.40 39 0.75 scatterplot
-001_Anderson_1935_Fig1.png     setosa Sepal Length (mm) 5.00 0.38 39 0.75 scatterplot
-001_Anderson_1935_Fig1.png versicolor  Sepal Width (mm) 2.77 0.32 44 0.52 scatterplot
-001_Anderson_1935_Fig1.png versicolor Sepal Length (mm) 5.94 0.53 44 0.52 scatterplot
+                  filename    group_id         variable  mean  error error_type  n     r    sd  plot_type
+001_Anderson_1935_Fig1.png      setosa  Sepal width (mm)  3.42  0.40  sd        39  0.75  0.40  scatterplot
+001_Anderson_1935_Fig1.png      setosa  Sepal length (mm) 5.00  0.38  sd        39  0.75  0.38  scatterplot
+001_Anderson_1935_Fig1.png  versicolor  Sepal width (mm)  2.77  0.32  sd        44  0.52  0.32  scatterplot
+001_Anderson_1935_Fig1.png  versicolor  Sepal length (mm) 5.95  0.53  sd        44  0.52  0.53  scatterplot
 \end{CodeOutput}
@@
 data
-               filename      group_id         variable  mean  sd   n    r   plot_type
-001_Anderson_1935_Fig1.png     setosa  Sepal Width (mm) 3.42 0.40 39 0.75 scatterplot
-001_Anderson_1935_Fig1.png     setosa Sepal Length (mm) 5.00 0.38 39 0.75 scatterplot
-001_Anderson_1935_Fig1.png versicolor  Sepal Width (mm) 2.77 0.32 44 0.52 scatterplot
-001_Anderson_1935_Fig1.png versicolor Sepal Length (mm) 5.94 0.53 44 0.52 scatterplot
-     002_Doe_2013_Fig1.png     setosa      Sepal length 5.00 0.78 50   NA  mean_error
-     002_Doe_2013_Fig1.png versicolor      Sepal length 5.93 1.05 50   NA  mean_error
-     002_Doe_2013_Fig1.png viriginica      Sepal length 6.59 1.26 50   NA  mean_error
-     003_Doe_2013_Fig3.png       <NA>      Sepal length 4.95 0.36 50   NA  histogram
+                  filename    group_id         variable   mean  error error_type n    r    sd   plot_type
+001_Anderson_1935_Fig1.png      setosa  Sepal width (mm)  3.42  0.40  sd        39  0.75  0.40  scatterplot
+001_Anderson_1935_Fig1.png      setosa  Sepal length (mm) 5.00  0.38  sd        39  0.75  0.38  scatterplot
+001_Anderson_1935_Fig1.png  versicolor  Sepal width (mm)  2.77  0.32  sd        44  0.52  0.32  scatterplot
+001_Anderson_1935_Fig1.png  versicolor  Sepal length (mm) 5.95  0.53  sd        44  0.52  0.53  scatterplot
+     002_Doe_2013_Fig1.png      setosa  Sepal length      5.00  0.11  se        50  NA    0.78  mean_error
+     002_Doe_2013_Fig1.png  viriginica  Sepal length      6.59  0.18  se        50  NA    1.26  mean_error
+     002_Doe_2013_Fig1.png  versicolor  Sepal length      5.94  0.14  se        50  NA    1.01  mean_error
+     003_Doe_2013_Fig3.png      catana  Sepal length      4.95  0.36  sd        50  NA    0.36  histogram
 \end{CodeOutput}
@@
 \begin{CodeOutput}
-Add group, Delete group or Finish plot? a/d/f 
+Add group, Delete group or Finish plot? (a/d/f) 
 \end{CodeOutput}

### 2017-11-24 docs/manuscript.tex
Updating the MS

@@
 %% x people doing y plots
-\subsection{Methods}
-We evaluated the consistency of digitisation using \pkg{metaDigitise} and tested how well digitisations estimate the true underlying data.
-This was accomplished by simulating a dataset with two different groups and two traits, and these data were then used to construct four different plot types (scatterplot, mean and error, histogram and boxplots).
+\subsection{Interobserver variability in digitisations}
+We evaluated the consistency of digitisation using \pkg{metaDigitise} by simulating a dataset with two different groups and two traits.
+These data were then used to construct four different plot types (scatterplot, mean and error, histogram and boxplots).
 Each variable was plotted twice for a given plot type (figures were modified slightly to give users a sense that they were digitising new data) generating a total of 14 figures.
-We randomised the order of these figures (using the same data) and provdied the entire figure directory to a total of 15 indpendent digitisers.
-The same data was presented twice for each figure type to estimate between- and within- user variability (i.e., the intra-class correlation coefficient) in the data.
+We randomised the order of these figures (using the same data) and provided the entire figure directory to a total of 15 indpendent digitisers.
+The same data was presented at least twice for each figure type to estimate between- and within- user variability (i.e., the intra-class correlation coefficient) in the data.
 Digitisers varied in their level of experience from those having done meta-analyses or comparative work to those not having any experience doing so.
-We got users to digitise all 14 figures and then send the data on these to evaluate interobserver variablity.
-
-We analysed these data using linear mixed effect models with user identify as a random effect.
-The overall model also included plot type to account for the different means across the variables and plots.
-We used the mean, standard deviation and correlation coeficients as response variables in three seperate models.
+We got users to digitise all 14 figures and then send the data on these to evaluate interobserver variablity across the plots.
+
+We analysed these data by first transformting each users mean, standard deviation and correlation coefficient generated from their digitisations to standardized mean differences (in otherwords, deviations were percentage differences from the true summary statistics).
+We then estimated intra-class correlation coefficients (ICC) (adjusted - accounting for covariates and unadjusted - not accounting for covariates) using linear mixed effect models with user identify as a random effect.
+We tested the significance of ICC estimates using liklihood ratio tests based on 1000 parametric bootstraps of the model.
+The overall model also included plot type, variable and group to account for possible biases across the variables and plots.
+We used standardised mean, standard deviation and correlation coeficients as response variables in seperate models using both \pkg{lme4} \citep{bates2015} and \pkg{MCMCglmm} \citep{hadfield2010}, estimating ICC using the \pkg{rptR} package \citep{stoffel2017} in R.
 If digitisations were consistent across all users then we would predict no significant between user variability in the data.
-We also qualitatively report how close the average digitised mean, SD and correlation coefficient were to the actual mean values of traits presented in the figures.
 
 \subsection{Results}
+
+  Across plot types we found no evidence for any inter-observer variability in digitisations for the mean (Unadjusted ICC = 0, 95 CI = 0 to 0.029, p = 1; Adjusted ICC = 0, 95 CI = 0 to 0.033, p = 1), standard deviation (Unadjusted ICC = ; Adjusted ICC = ) or correlation coefficient (Unadjusted ICC = ; Adjusted ICC = )
 

### 2017-11-24 docs/manuscript.tex
MS updated with some results

@@
 %% 1 - dont manipulate data/allow different plot types
+
 Firstly, they typically only provide the user with calibrated x,y coordinates from imported figures, and do not differentiate between common plot types that are use to present data.
@@
 %% 2- adding metadata
+
 Secondly, digitising programs do not easily allow the integration of metadata at the time of data extraction, such as experimental group or variable names, and sample sizes.
@@
 %% 3 - in same statistical software and high throughput
+
 Finally, existing programs do not import a set of images and allow the user to systematically work through them.
@@
 
+\subsection{Testing the accuracy of digitisations}
+To test how accurate \pkg{metaDigitise} is at matching points to their true values, we generated four random scatterplots and tested how well the digitised point in each scatterplot matches the actual, true point generated from simulations.
+We evaluated the accuracy using a Pearson's correlation coefficient between the digitised data from \pkg{metaDigitise} and the true known data value from the simulated data.
+
 \subsection{Results}
 
-  Across plot types we found no evidence for any inter-observer variability in digitisations for the mean (Unadjusted ICC = 0, 95 CI = 0 to 0.029, p = 1; Adjusted ICC = 0, 95 CI = 0 to 0.033, p = 1), standard deviation (Unadjusted ICC = ; Adjusted ICC = ) or correlation coefficient (Unadjusted ICC = ; Adjusted ICC = )
+Across plot types we found no evidence for any inter-observer variability in digitisations for the mean (Unadjusted ICC = 0, 95\% CI = 0 to 0.029, p = 1; Adjusted ICC = 0, 95\% CI = 0 to 0.033, p = 1), standard deviation (Unadjusted ICC = 0, 95\% CI = 0 to 0.033, p = 0.5; Adjusted ICC = 0.002, 95\% CI = 0 to 0.032, p = 1) or correlation coefficient (Unadjusted ICC = 0.053, 95\% CI = 0 to 0.296, p = 0.377; Adjusted ICC = 0.048, 95\% CI = 0 to 0.3, p = 0.374).
+Overall, there were very small differences between digitised values and the true values only deviating, on average 3.34\% (mean = 0.02\%, sd = 5\%, r = 5\%) for all three summary statistics.
+%% Note that unadjusted and adjusted values come from rptR and the deviation of percent are the intercept values (*100) for mean, sd and r. I've averaged these three for 3.34%
+
+Digitised data was essentially perfectly correlated with the true simulated data for both the x-variable (r = 0.9999915, t = 2137.4, df = 78, p-value < 2.2e-16) and y-variable (r = 0.9999892, t = 1897.8, df = 78, p-value < 2.2e-16) in scatterplots.
 

### 2017-11-24 docs/manuscript.tex
Finished MS

@@
 %Including addition of N later
+In many cases important information, such as sample sizes, may not be readily available or clear when digitising figures.
+In these circumstances users will have answered 'no' to the question about whether they have sample sizes or not while digitising.
+To expedite finding and adding in these sample sizes to do the necssary calculations (if for example a figure presented 95\% CI's or standard errors), \pkg{metaDigitise} has s specific edit option that allows users to enter in previously omitted sample sizes.
+It works by first identifying the missing sample sizes in the digitised output, re-plotting the relevant figure and then prompting the user to enter the sample sizes for the relevant groups in the figure, 1 by one.
+As an example, assume that we were missing sample sizes for two groups in 002\_Doe\_2013\_Fig1.png:
+
+\begin{CodeChunk}
+\begin{CodeOutput}
+                  filename    group_id         variable   mean  error error_type n    r    sd   plot_type
+     002_Doe_2013_Fig1.png      setosa  Sepal length      5.00  0.11  se        NA  NA    NA  mean_error
+     002_Doe_2013_Fig1.png  viriginica  Sepal length      6.59  0.18  se        NA  NA    NA  mean_error
+\end{CodeInput}
+\end{CodeChunk}
+
+Here, we can see that we are missing the sample sizes for setosa and viriginica, and as a result, sd is not calculated because \pkg{metaDigitise} needs this information to make the calculation.
+If the user found this information after contacting the authors for clarification then they can add these back in as follows:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+>R setwd("~/FiguresToExtract")
+>R digitised_data <- metaDigitise(".")
+\end{CodeInput}
+\end{CodeChunk}
+
+\begin{CodeChunk}
+\begin{CodeInput}
+Do you want to...
+
+1: Process new images
+2: Import existing data
+3: Edit existing data
+
+Selection: 3
+\end{CodeInput}
+\end{CodeChunk}
+
+\begin{CodeChunk}
+\begin{CodeInput}
+Choose how you want to edit files:
+
+1: Cycle through images
+2: Choose specific file to edit
+3: Enter previously omitted sample sizes
+
+Selection: 3
+\end{CodeInput}
+\end{CodeChunk}
+
+\pkg{metaDigitise} will replot the figure after this and list, only the groups missing data, for which the user can then update the data.
+This is then re-integrated back into the data automatically and the sd calculated.
+
+\begin{CodeChunk}
+\begin{CodeInput}
+Group " setosa ": Enter sample size  
+50
+
+Group " viriginica ": Enter sample size  
+50
+\end{CodeInput}
+\end{CodeChunk}
 

### 2017-11-24 R/metaDigitise.R
Added metaDigitise example that can be run

@@ -15,6 +15,32 @@
 #' # data <- metaDigitise(dir = "./example_figs/", summary = TRUE)
 #' # data
 #' @return A data frame or list containing the raw digitised data or the processed, summary statistics from the digitised data
+#' @examples
+#' \dontrun{# Generate a directory in current working directory and fill with figures
+#' dir.create("./prac_figs")
+#' set.seed(103)
+#' x <- rnorm(20,0,1)
+#' y <- rnorm(20,0,1)
+#' means <- c(mean(x),mean(y))
+#' ses <- c(sd(x)/sqrt(length(x))*1.96, sd(y)/sqrt(length(y))*1.96)
+#' png(filename = paste0("./prac_figs/mean_error.png"), width = 480, height = 480)
+#' plot(means, ylim = c(min(means-ses)-0.1,max(means+ses)+0.1), xlim=c(0.5,2.5), 
+#' xaxt="n", pch=19, cex=2, ylab="Variable +/- SE", xlab="Treatment", main="Mean Error")
+#' arrows(1:length(means),means+ses, 1:length(means), means-ses, code=3, angle=90, length=0.1)
+#' axis(1,1:length(means),names(means))
+#' dev.off()
+#' png(filename = paste0("boxplot.png"), width = 480, height = 480)
+#' boxplot(x,y, main="Boxplot")
+#' dev.off()
+#' png(filename = paste0("histogram.png"),width = 480, height = 480)
+#' hist(c(x,y), xlab= "variable", main="Histogram")
+#' dev.off()
+#' png(filename = paste0("scatterplot.png"), width = 480, height = 480)
+#' plot(x,y, main="Scatterplot")
+#' dev.off()
+#' 
+#' data <- metaDigitise("./prac_figs/")
+#' }
 #' @export
 
 metaDigitise<-function(dir, summary = TRUE){

### 2017-11-24 R/metaDigitise.R
Fixed example path names

@@ -29,13 +29,13 @@
 #' arrows(1:length(means),means+ses, 1:length(means), means-ses, code=3, angle=90, length=0.1)
 #' axis(1,1:length(means),names(means))
 #' dev.off()
-#' png(filename = paste0("boxplot.png"), width = 480, height = 480)
+#' png(filename = paste0("./prac_figs/boxplot.png"), width = 480, height = 480)
 #' boxplot(x,y, main="Boxplot")
 #' dev.off()
-#' png(filename = paste0("histogram.png"),width = 480, height = 480)
+#' png(filename = paste0("./prac_figs/histogram.png"),width = 480, height = 480)
 #' hist(c(x,y), xlab= "variable", main="Histogram")
 #' dev.off()
-#' png(filename = paste0("scatterplot.png"), width = 480, height = 480)
+#' png(filename = paste0("./prac_figs/scatterplot.png"), width = 480, height = 480)
 #' plot(x,y, main="Scatterplot")
 #' dev.off()
 #' 

### 2017-11-25 docs/manuscript.tex
manuscript editting

--- 2017-11-24 docs/manuscript.tex
@@
 \usepackage{thumbpdf,lmodern,amssymb,rotating}
-
-% \usepackage[utf8]{inputenc},
-% \usepackage{geometry} 
-% \geometry{a4paper}  
-% %\usepackage[parfill]{parskip} 
-% \usepackage{graphicx}	
-% \usepackage{amsmath}
-% \usepackage{fullpage}
-% \usepackage{setspace} 
-% \usepackage{lineno}
-% \usepackage[none]{hyphenat}
-% \usepackage{listings}
-% \usepackage[round]{natbib}
 
@@
   E-mail: \email{joel.l.pick@gmail.com}\\
-  URL: \url{https://eeecon.uibk.ac.at/~zeileis/}
+  URL: \url{https://github.com/daniel1noble/metaDigitise}
 }
@@
 Most notably, this includes meta-analysis, which is becoming increasingly common in many research fields.
-Meta-analysis uses effect size estimates and their sampling variance, taken from many studies, to understand whether particular effects are common across studies and to explain variation among these effects \citep{Naka2017, Glass1976, Koricheva2013, Borenstein2009}.
+Meta-analysis uses effect size estimates and their sampling variance, taken from many studies, to understand whether particular effects are common across studies and to explain variation among these effects \citep{Glass1976,Borenstein2009,Koricheva2013,Naka2017}.
 Meta-analysis therefore relies foremost on data extracted from primary literature and more specifically statistics that have been reported in the text or tables of research papers.
@@
 Although there are several existing tools to perform tasks like this (e.g.
-DataThief \citep{DataThief}, GraphClick \citep{GraphClick}, WebPlotDigitizer \citep{WebPlotDigitizer}), these tools are not designed specifically for meta-analysis.
+\proglang{DataThief} \citep{DataThief}, \proglang{GraphClick} \citep{GraphClick}, \proglang{WebPlotDigitizer} \citep{WebPlotDigitizer}), these tools are not designed specifically for meta-analysis.
 %% need to spell out clear points
 %% 1 - dont manipulate data/allow different plot types
+
 Firstly, they typically only provide the user with calibrated x,y coordinates from imported figures, and do not differentiate between common plot types that are use to present data.
@@
 Furthermore, although meta-analysis is an important tool in consolidating the data from many studies many of the processes involved in data extraction are opaque and difficult to reproduce, making extending studies problematic.
-Having a tool that facilitates reproducibility in meta-analyses will increase transparency and go a long way to resolving the reproducibility crises we are seeing in many fields \citep{peng_reproducible_2006, peng_reproducible_2011, sandve_ten_2013, Parker2016}.
+Having a tool that facilitates reproducibility in meta-analyses will increase transparency and go a long way to resolving the reproducibility crises we are seeing in many fields \citep{peng_reproducible_2006, peng_reproducible_2011, sandve_ten_2013, Parker2016,Ihle2017}.
 
@@
 This expedites digitising figures by preventing users from having to constantly change directories and / or open new images.
-The data from a completed image is automatically saved in an .RDS file to a \code{caldat} directory, that is created within the parent directory when first executing the \fct{metaDigitise} function.
+The data from a completed image is automatically saved as a \code{metaDigitise} object in an .RDS file to a \code{caldat} directory, that is created within the parent directory when first executing the \fct{metaDigitise} function.
 These files enable re-plotting and editing of images at a later point (see below).
@@
 Alternatively, users can keep like figures together and process them separately.
-
-%% where to put calibration???
 
@@
 At first there is one figure in the folder, shown in Figure \ref{fig:Anderson}.
-Running \fct{metaDigitise} brings up a series of prompts for the user using a main menu that provides access to a number of its features:
-
-\begin{CodeChunk}
-\begin{CodeInput}
-R> setwd("~/FiguresToExtract")
-R> digitised_data <- metaDigitise(".", summary = TRUE)
+Running \fct{metaDigitise} brings up a series of prompts for the user using a main menu that provides access to a number of its features (...
+here represents the user's project directory):
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> digitised_data <- metaDigitise(".../FiguresToExtract", summary = TRUE)
 \end{CodeInput}
@@
 In some cases, older papers can give slightly off angled images which can be corrected by rotating.
-So, in this prompt the user has three options: \code{f} for "Flip", \code{r} for "rotate" or \code{c} for "continue.
+So, in this prompt the user has three options: \code{f} for ``Flip'', \code{r} for ``rotate'' or \code{c} for ``continue''.
 
@@
 {b}), a scatter plot (\code{s}) or a histogram (\code{h}).
-If the user has specified \code{diff} instead of \code{same} in response to the question about whether the plot types are the same or different, this question will pop up for each plot, but will only be asked once if plots are all the same.
+If the user has specified \code{d} instead of \code{s} in response to the question about whether the plot types are the same or different, this question will pop up for each plot, but will only be asked once if plots are all the same.
 
@@
  h: Histogram
-
+\end{CodeOutput}
+\begin{CodeInput}
 R> s
-\end{CodeOutput}
+\end{CodeInput}
 \end{CodeChunk}
@@
 Once complete, details on how to calibrate the x and y-axis appear, so that the relevant statistics / data can be correctly calculated.
-When working with a plot of mean and standard errors, the x-axis is rather useless in terms of calibration so \pkg{metaDigitise} just asks the user to calibrate the y-axis.
+When working with a plot of mean and standard errors, the x-axis is rather useless in terms of calibration so \fct{metaDigitise} just asks the user to calibrate the y-axis.
 
@@
 Re-calibrate? (y/n) 
+\end{CodeOutput}
+\begin{CodeInput}
 R> n
-\end{CodeOutput}
+\end{CodeInput}
 \end{CodeChunk}
@@
 group, exit (see above), then follow prompts
-
-Add points, delete points or continue? (a/d/c) 
+\end{CodeOutput}
+\end{CodeChunk}
+
+To finish selecting points, the user can exit following the instructions (e.g. by pressing ESC on a mac). The user is then asked if they want to add or delete points from that group.
+
+\begin{CodeChunk}
+\begin{CodeOutput}
+Add or Delete points to this group, or Continue? (a/d/c) 
 \end{CodeOutput}
@@
 
-Once we are done digitising all the groups our plot will look something like Figure \ref{fig:scatter_extract}
+Once we are done digitising all the groups our plot will look something like Figure \ref{fig:scatter_extract}.
 
@@
 
-When completed \pkg{metaDigitise} will write the digitised data to the caldat directory, such that our new directory structure is as follows
+When completed \fct{metaDigitise} will write the digitised data as a \code{metaDigitise} object to a RDS file in the caldat directory, such that our new directory structure is as follows
 
@@
 
-Users can access the \pkg{metaDigitise} object created (001\_Anderson\_1935\_Fig1) at any time using the metaDigitse function.
+Users can access the \code{metaDigitise} object created (001\_Anderson\_1935\_Fig1) at any time using the metaDigitse function.
 In the \proglang{R} console, the summarised data for the digitised figure can be printed on screen or even written to a .csv file:
@@
 \end{CodeInput}
+{\scriptsize
 \begin{CodeOutput}
@@
 \end{CodeOutput}
+}
 \end{CodeChunk}
@@
 \begin{CodeInput}
-R> setwd("~/FiguresToExtract")
-R> digitised_data <- metaDigitise(".", summary = TRUE)
+R> digitised_data <- metaDigitise(".../FiguresToExtract", summary = TRUE)
 \end{CodeInput}
@@
 \begin{CodeOutput}
-Are all plot types the same? (diff/same) 
-\end{CodeOutput}
-\begin{CodeInput}
-R> diff
+Are all plot types Different or the Same? (d/s)
+\end{CodeOutput}
+\begin{CodeInput}
+R> d
 \end{CodeInput}
@@
 \begin{CodeOutput}
-Please specify the plot_type as either: mean and error, box plot,
-scatter plot or histogram m/b/s/h: 
+Please specify the plot_type as either:
+
+ m: Mean and error
+ b: Box plot
+ s: Scatter plot 
+ h: Histogram
 \end{CodeOutput}
@@
 \begin{CodeOutput}
-What is the variable? 
+What is the y variable? 
 \end{CodeOutput}
@@
 \begin{CodeOutput}
-Enter sample sizes? y/n 
+Do you know sample sizes? (y/n)  
 \end{CodeOutput}
@@
 \begin{CodeOutput}
+If there are multiple groups, enter unique group identifiers (otherwise press enter)
 Group identifier: 
@@
 
-Add group, Delete group or Finish plot? a/d/f 
+Add group, Edit Group, Delete group or Finish plot? (a/e/d/f) 
 \end{CodeOutput}
@@
 \begin{CodeChunk}
+{\scriptsize
 \begin{CodeOutput}
@@
 \begin{CodeInput}
-R> setwd("~/FiguresToExtract")
-R> digitised_data <- metaDigitise(".")
+R> digitised_data <- metaDigitise(".../FiguresToExtract")
 \end{CodeInput}
@@
 %Including addition of N later
-
-\section{Interobserver Variability and Validation}
-%% simulated data
-%% x people doing y plots
+In many cases important information, such as sample sizes, may not be readily available or clear when digitising figures.
+In these circumstances users will have answered `no' to the question about whether they have sample sizes or not while digitising.
+To expedite finding and adding in these sample sizes to do the necessary calculations (if for example a figure presented 95\% CI's or standard errors), \fct{metaDigitise} has s specific edit option that allows users to enter in previously omitted sample sizes.
+It works by first identifying the missing sample sizes in the digitised output, re-plotting the relevant figure and then prompting the user to enter the sample sizes for the relevant groups in the figure, one by one.
+As an example, assume that we were missing sample sizes for two groups in 002\_Doe\_2013\_Fig1.png:
+
+\begin{CodeChunk}
+{\scriptsize
+\begin{CodeOutput}
+            filename    group_id         variable   mean  error error_type n    r    sd   plot_type
+002_Doe_2013_Fig1.png      setosa  Sepal length      5.00  0.11  se        NA  NA    NA  mean_error
+002_Doe_2013_Fig1.png  viriginica  Sepal length      6.59  0.18  se        NA  NA    NA  mean_error
+\end{CodeOutput}
+}
+\end{CodeChunk}
+
+Here, we can see that we are missing the sample sizes for setosa and viriginica, and as a result, sd is not calculated because \fct{metaDigitise} needs this information to make the calculation.
+If the user found this information after contacting the authors for clarification then they can add these back in as follows:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> digitised_data <- metaDigitise(".../FiguresToExtract")
+\end{CodeInput}
+\end{CodeChunk}
+
+\begin{CodeChunk}
+\begin{CodeOutput}
+Do you want to...
+
+1: Process new images
+2: Import existing data
+3: Edit existing data
+
+Selection:
+\end{CodeOutput}
+\begin{CodeInput}
+R> 3
+\end{CodeInput}
+\end{CodeChunk}
+
+\begin{CodeChunk}
+\begin{CodeOutput}
+Choose how you want to edit files:
+
+1: Cycle through images
+2: Choose specific file to edit
+3: Enter previously omitted sample sizes
+
+Selection:
+\end{CodeOutput}
+\begin{CodeInput}
+>R 3
+\end{CodeInput}
+\end{CodeChunk}
+
+\fct{metaDigitise} will replot the figure after this and list, only the groups missing data, for which the user can then update the data.
+This is then re-integrated back into the data automatically and the sd calculated.
+
+\begin{CodeChunk}
+\begin{CodeOutput}
+Group " setosa ": Enter sample size  
+\end{CodeOutput}
+\begin{CodeInput}
+R> 50
+\end{CodeInput}
+\begin{CodeOutput}
+Group " viriginica ": Enter sample size
+\end{CodeOutput}
+\begin{CodeInput}
+R> 50
+\end{CodeInput}
+\end{CodeChunk}
+
+\section{Inter-observer Variability and Validation}
+
 \subsection{Interobserver variability in digitisations}
-We evaluated the consistency of digitisation using \pkg{metaDigitise} by simulating a dataset with two different groups and two traits.
-These data were then used to construct four different plot types (scatterplot, mean and error, histogram and boxplots).
-Each variable was plotted twice for a given plot type (figures were modified slightly to give users a sense that they were digitising new data) generating a total of 14 figures.
-We randomised the order of these figures (using the same data) and provided the entire figure directory to a total of 15 indpendent digitisers.
-The same data was presented at least twice for each figure type to estimate between- and within- user variability (i.e., the intra-class correlation coefficient) in the data.
-Digitisers varied in their level of experience from those having done meta-analyses or comparative work to those not having any experience doing so.
-We got users to digitise all 14 figures and then send the data on these to evaluate interobserver variablity across the plots.
-
-We analysed these data by first transformting each users mean, standard deviation and correlation coefficient generated from their digitisations to standardized mean differences (in otherwords, deviations were percentage differences from the true summary statistics).
-We then estimated intra-class correlation coefficients (ICC) (adjusted - accounting for covariates and unadjusted - not accounting for covariates) using linear mixed effect models with user identify as a random effect.
-We tested the significance of ICC estimates using liklihood ratio tests based on 1000 parametric bootstraps of the model.
-The overall model also included plot type, variable and group to account for possible biases across the variables and plots.
-We used standardised mean, standard deviation and correlation coeficients as response variables in seperate models using both \pkg{lme4} \citep{bates2015} and \pkg{MCMCglmm} \citep{hadfield2010}, estimating ICC using the \pkg{rptR} package \citep{stoffel2017} in R.
-If digitisations were consistent across all users then we would predict no significant between user variability in the data.
-
-\subsection{Results}
-
-  Across plot types we found no evidence for any inter-observer variability in digitisations for the mean (Unadjusted ICC = 0, 95 CI = 0 to 0.029, p = 1; Adjusted ICC = 0, 95 CI = 0 to 0.033, p = 1), standard deviation (Unadjusted ICC = ; Adjusted ICC = ) or correlation coefficient (Unadjusted ICC = ; Adjusted ICC = )
+In order to evaluate the consistency of digitisation using \pkg{metaDigitise} between users, we simulated a dataset of two traits with two different groups.
+These data were then used to construct plots of the four different types (scatterplot, mean and error, histogram and boxplots).
+Each variable was plotted twice for each given plot type (figures were modified slightly to give users a sense that they were digitising new data) generating a total of 14 figures.
+15 independent digitisers were provided with a directory with all 14 figures in a randomised order.
+Digitisers ran \pkg{metaDigistise} on their own computers, across different operating systems (including Mac, Windows and Linux).
+Digitisers varied in their level of experience, from people with experience of meta-analyses or comparative work to those without any science background.
+We asked users to digitise all 14 figures and collected the mean, standard deviation and correlation coefficient (for scatterplots) generated by \fct{metaDigitise} for every plot digitised.
+We transformed these data to standardized differences as
+
+\begin{equation}
+\frac{\theta - \hat{\theta}}{\hat{\theta}}
+\end{equation}
+
+where $\theta$ is the estimate value and $\hat{\theta}$ is the true value, meaning that deviations were percentage differences from the true summary statistics.
+The correlation coefficient deviation was not divided by the true value, as it is already on a standardised scale.
+This deviation can be seen as a measure of bias.
+The resulting data was used to assess between- and within- user variability (i.e., the intra-class correlation coefficient) in the data.
+This was done using linear mixed effect models with user identify as a random effect.
+Standardised mean, standard deviation and correlation coeficients were used as response variables in seperate models.
+Sampling variance for ICC estiamtes was generated based on 1000 parametric bootstraps of the model and the significance was tested using liklihood ratio tests.
+These models were run using the \pkg{lme4} \citep{bates2015} and \pkg{rptR} \citep{stoffel2017} packages in \proglang{R}.
+
+If digitisations were consistent across all users then we should find no significant between user variability in the data.
+Indeed, across plot types we found no evidence for any inter-observer variability in digitisations for the mean (ICC = 0, 95\% CI = 0 to 0.029, p = 1), standard deviation (ICC = 0, 95\% CI = 0 to 0.033, p = 0.5) or correlation coefficient (ICC = 0.053, 95\% CI = 0 to 0.296, p = 0.377).
+There were was little bias in between digitised and true values, on average 1.63\% (mean = 0.02\%, sd = 4.9\%, r = -0.03\%) and overall there were only small absolute differences between digitised and true values, deviating, on average 2.18\% (mean = 0.40\%, sd = 5.81\%, r = 0.33\%) for all three summary statistics.
+
+SD estimates from digitisations are clearly more prone to error than means or correaltion cefficients.
+This is due to the fact that SD estimation from the summary statistics derived from plots is more error prone, especially at small sample sizes \citep{Wan2014}.
+If the mean absoluate difference is calculated for each plot type, we can see that this effect is driven mainly by extraction from boxplots and histgrams (\% difference):
+\begin{CodeOutput}
+    boxplot   histogram  mean_error scatterplot 
+     15.805       5.210       1.500       0.433 
+\end{CodeOutput}
+
+\subsection{Testing the accuracy of digitisations}
+To test how accurate \pkg{metaDigitise} is at matching points to their true values, we generated four random scatterplots, each with 20 data points, and digitised these with \fct{metaDigitise}.
+This was done by one digitiser, as there is no detectable between user variation.
+Data digitised using \pkg{metaDigitise} was essentially perfectly correlated with the true simulated data for both the x-variable (Pearson's correlation; r = 0.9999915, t = 2137.4, df = 78, p-value < 0.001) and y-variable (r = 0.9999892, t = 1897.8, df = 78, p-value < 0.001).
 
@@
 
-%% discuss specifically everything it can't do, and jsutify it slightly
-
-\begin{sidewaystable}[t!]
+\begin{sidewaystable}[ph!]
 {\centering
@@
 Additionally, auto-detection often misses many points which then subsequently need to be manually added.
-Bssed on tests of \pkg{metaDigitise} (see above) figures can be extracted in around 1-2 minutes, including the entry of metadata.
-As a result, we do not belive that current automated point detection provides substantial time saving benefits.
-
-Another feature that \pkg{metaDigitise} (currently) lacks, is an ability to zoom in plots.
+Based on tests of \pkg{metaDigitise} (see above) figures can be extracted in around 1-2 minutes, including the entry of metadata.
+As a result, we do not belive that current automated point detection provides substantial benefits in terms of time or accuracy.
+
+Another feature that \pkg{metaDigitise} (currently) lacks, is an ability to zoom in on plots.
 This may enable users to gain greater accuracy when clicking on points.
@@
 
-Finally, \pkg{metaDigitise} currently also does not allow for asymmetric error bars.
+Finally, \pkg{metaDigitise} currently does not allow for asymmetric error bars.
 At present this is a deliberate omission, as it is not clear how best to derive SD from such data, given also that such asymmetric error bars may represent different things in different figures.
@@
 \section*{Acknowledgments}
-We thank the I-DEEL group at UNSW for increadibly useful feedback, and a host of colleagues for testing, providing feedback and digitising including: Rose O'Dea, Fonti Kar, Malgorzata Lagisz, Julia Riley, Diego Barneche, Erin Macartney, Ivan Beltran, Gihan Samarasinghe, Dax Kellie, Jonathan Noble and Yian Noble.
-DWAN was supported by an Australian Research Council Discovery Early Career Research Award (DE150101774) and UNSW Vice Chancellors Fellowship and SN an Australian Research Council Future Fellowship (FT130100268).
+We thank the I-DEEL group at UNSW for incredibly useful feedback, and a host of colleagues for testing, providing feedback and digitising including: Rose O'Dea, Fonti Kar, Malgorzata Lagisz, Julia Riley, Diego Barneche, Erin Macartney, Ivan Beltran, Gihan Samarasinghe, Dax Kellie, Jonathan Noble, Yian Noble and Alison Pick.
+JLP was supported by a Swiss National Science Foundation Early Mobility grant (P2ZHP3\_164962), DWAN was supported by an Australian Research Council Discovery Early Career Research Award (DE150101774) and UNSW Vice Chancellors Fellowship and SN an Australian Research Council Future Fellowship (FT130100268).
 

### 2017-11-29 R/H_extract.R
histogram improvements

@@ -19,14 +19,19 @@ histogram_extract <- function(edit=FALSE, raw_data = data.frame(), ...){
 
 	histQ <- if(edit){ "b" }else{ "a" }
 	i <- if(edit){ max(unique(raw_data$bar)) }else{ 0 }
-
+	bar_cols <- c("red","orange")
+	
 	while(histQ != "f"){
 		
 		if(histQ=="a"){
+			polygon(box_x,box_y, col="red", border=NA)
+			cat("\nClick on the left followed by the right upper corners of each bar\n Double click on red box in bottom left corner to exit extraction\n")
+		} 
+		while(histQ=="a"){
 			i=i+1
-			cat("Click on the left followed by the right upper corners of a single bar")
-			bar_points <- locator(2, type="o", col="red", lwd=2, pch=19)
-			raw_data <- rbind(raw_data, data.frame(id=group_id, x=bar_points$x,y=bar_points$y, bar=i))
+			bar_points <- locator(2, type="o", col=bar_cols[is.even(i)+1], lwd=2, pch=19)
+			if( mean(bar_points$x)<max(box_x) & mean(bar_points$y)<max(box_y) & mean(bar_points$x)>min(box_x) & mean(bar_points$y)>min(box_y)) histQ <- "b"
+			if(histQ=="a") raw_data <- rbind(raw_data, data.frame(id=group_id, x=bar_points$x,y=bar_points$y, bar=i))
 		}
 
 		if(histQ=="d"){

### 2017-11-29 R/misc_func.R

@@ -13,6 +13,14 @@ is.wholenumber <- function(x, tol = .Machine$double.eps^0.5)  abs(x - round(x)) 
 isNumeric <- function(x) !suppressWarnings(is.na(as.numeric(x)))
 
 
+#' @title is.even
+#' @param x integer value
+#' @description Checks whether a integer is even
+
+is.even <- function(x) x %% 2 == 0
+
+
+
 #' @title filename
 #' @param x filepath
 #' @description extracts filename from filepath

### 2017-11-29 R/redraw.R

@@ -97,11 +97,13 @@ redraw_points <- function(plot_type, raw_data, image_details, scalar){
 	}
 
 	if(plot_type=="histogram"& nrow(raw_data)>0){
+		bar_cols <- c("red","orange")
 		for(i in unique(raw_data$bar)){
 			bar_data <- subset(raw_data, bar==i)
-			points(y~x,bar_data, pch=19, col=point_col, cex=point_cex)
-			lines(y~x, bar_data, lwd=line_width, col=point_col)
-			text(mean(bar_data$x),mean(bar_data$y)+legend_pos,bar_data$bar[1], col=point_col, cex=text_cex)
+			bar_col <- bar_cols[is.even(i)+1]
+			points(y~x,bar_data, pch=19, col=bar_col, cex=point_cex)
+			lines(y~x, bar_data, lwd=line_width, col=bar_col)
+			text(mean(bar_data$x),mean(bar_data$y)+legend_pos,bar_data$bar[1], col=bar_col, cex=text_cex)
 		}
 	}
 }

### 2017-11-29 tests/testthat/test-misc_func.R

@@ -15,6 +15,13 @@ test_that("Checking isNumeric..", {
 		info = "isNumeric failed")
 })
 
+test_that("Checking is.even..", {
+	expect_equal(
+		is.even(x=1), 
+		FALSE, 
+		info = "is.even failed")
+})
+
 test_that("Checking filename..", {
 	expect_equal(
 		filename(x="~/Dropbox/0_postdoc/10_metaDigitise/example_figs/5_fig2a.png"), 

### 2017-11-29 R/H_extract.R
fixed bug

@@ -1,9 +1,10 @@
 #' @title histogram_extract
 #' @param edit logical; whether in edit mode 
 #' @param raw_data raw data
+#' @param calpoints The calibration points
 #' @param ... arguments to pass to internal_redraw
 #' @description Extraction of data from histograms
-histogram_extract <- function(edit=FALSE, raw_data = data.frame(), ...){
+histogram_extract <- function(edit=FALSE, raw_data = data.frame(), calpoints, ...){
 
 	if(edit){ 
 		idQ <- user_options("Change group identifier? (y/n) ",c("y","n"))
@@ -20,7 +21,10 @@ histogram_extract <- function(edit=FALSE, raw_data = data.frame(), ...){
 	histQ <- if(edit){ "b" }else{ "a" }
 	i <- if(edit){ max(unique(raw_data$bar)) }else{ 0 }
 	bar_cols <- c("red","orange")
-	
+	box_y <- c(mean(calpoints$y[3:4]), mean(calpoints$y[3:4]),0,0,mean(calpoints$y[3:4]))/2
+	box_x <- c(0,mean(calpoints$x[1:2]), mean(calpoints$x[1:2]),0,0)/2
+
+
 	while(histQ != "f"){
 		
 		if(histQ=="a"){
@@ -30,7 +34,10 @@ histogram_extract <- function(edit=FALSE, raw_data = data.frame(), ...){
 		while(histQ=="a"){
 			i=i+1
 			bar_points <- locator(2, type="o", col=bar_cols[is.even(i)+1], lwd=2, pch=19)
-			if( mean(bar_points$x)<max(box_x) & mean(bar_points$y)<max(box_y) & mean(bar_points$x)>min(box_x) & mean(bar_points$y)>min(box_y)) histQ <- "b"
+			if( mean(bar_points$x)<max(box_x) & mean(bar_points$y)<max(box_y) & mean(bar_points$x)>min(box_x) & mean(bar_points$y)>min(box_y)){
+					histQ <- "b"
+					i=i-1
+				}
 			if(histQ=="a") raw_data <- rbind(raw_data, data.frame(id=group_id, x=bar_points$x,y=bar_points$y, bar=i))
 		}
 
@@ -39,7 +46,7 @@ histogram_extract <- function(edit=FALSE, raw_data = data.frame(), ...){
 			raw_data <- subset(raw_data, bar != delQ)
 		}
 
-		internal_redraw(raw_data=raw_data,...)
[…]

### 2017-11-29 docs/manuscript.tex
Corrections to MS based on Shin's comments, some really good feedback

@@
 \Abstract{
-
+Research synthesis, especially, in the form of meta-analysis requires data extraction from primary studies.
+Meta-analysis synthesizes effect sizes, often calculated from summary statistics of studies.
+However, exact values of such statistics are commonly hidden in figures.
+The R package metaDigitise extracts descriptive statistics such as means, standard deviations and, if applicable, correlations from the four types of plots: 1) mean and error plots (e.g.
+bar graphs with standard errors), 2) box plots, 3) scatter plots and 4) histograms.
+The package interactively guides the user through data extraction process.
+Notably, it enables a large-scale extraction using image files, letting the user stop processing, and edit and add to the resulting data fame at any point.
+Further, it facilitates reproducible data extractions from plots with little inter-observer bias, thus, allowing a group of people to participate the extraction of data collaboratively.
 }
@@
 Meta-analysis uses effect size estimates and their sampling variance, taken from many studies, to understand whether particular effects are common across studies and to explain variation among these effects \citep{Glass1976,Borenstein2009,Koricheva2013,Naka2017}.
-Meta-analysis therefore relies foremost on data extracted from primary literature and more specifically statistics that have been reported in the text or tables of research papers.
+Meta-analysis therefore relies foremost on data extracted from primary literature, and more specifically, statistics that have been reported in the text or tables of research papers (e.g., means, standard deviations, correlation coefficients).
 These summary statistics are also, however, frequently presented in figures and so need to be manually extracted using digitising programs.
+While inferential statistics (e.g., t- and F-statistics) are often presented along side summary statistics and used to derive effect sizes, raw descriptive statistics are much more appropriate to use because sources of non-independence in experimental designs can be dealt with more easily using summary statistics \citep{Noble2017}.
 Although there are several existing tools to perform tasks like this (e.g.
@@
 This expedites digitising figures by preventing users from having to constantly change directories and / or open new images.
-The data from a completed image is automatically saved as a \code{metaDigitise} object in an .RDS file to a \code{caldat} directory, that is created within the parent directory when first executing the \fct{metaDigitise} function.
+The data from a completed image is automatically saved as a \code{metaDigitise} object in an .RDS file to a \code{caldat} directory that is created within the parent directory when first executing the \fct{metaDigitise} function.
 These files enable re-plotting and editing of images at a later point (see below).
@@
 A notable feature that \pkg{metaDigitise} lacks is automated point detection.
-This is available in several packages (Table \ref{tab:comparison}).
+Point detection is available in several packages (Table \ref{tab:comparison}).
 However, from our experience of using these functions, manual digitising is more reliable and often equally as fast.

### 2017-11-29 R/S_extract.R
scatterplot update

@@ -15,13 +15,18 @@ delete_group <- function(raw_data){
 #' @title edit_group
 #' @param raw_data data
 #' @param group_id group_id
+#' @param calpoints The calibration points
 #' @param ... other functions to pass to internal_redraw
 #' @description Edit group points in scatterplots
  
-edit_group <- function(raw_data, group_id,...){
+edit_group <- function(raw_data, group_id, calpoints, ...){
 	
 	cols <- rep(c("red", "green", "purple"),length.out=90)
 	pchs <- rep(rep(c(19, 17, 15),each=3),length.out=90)
+
+	box_y <- c(mean(calpoints$y[3:4]), mean(calpoints$y[3:4]),0,0,mean(calpoints$y[3:4]))/2
+	box_x <- c(0,mean(calpoints$x[1:2]), mean(calpoints$x[1:2]),0,0)/2
+
 
 	if(!is.null(group_id)) {
 		group_data <- data.frame()
@@ -44,22 +49,30 @@ edit_group <- function(raw_data, group_id,...){
 
 	while(add_removeQ!="c"){
 
-		if(add_removeQ=="a"){		
-			cat("\nClick on points you want to add.\nIf you want to remove a point, or are finished with a\ngroup, exit (see above), then follow prompts\n")
-			select_points <- locator(type="p", lwd=2, col=cols[i], pch=pchs[i])
-			if(!is.null(select_points)){
+		if(add_removeQ=="a"){
+			polygon(box_x,box_y, col="red", border=NA)
+			cat("\nClick on points you want to add.\nIf you want to remove a point, or are finished with a group, \nexit by clicking on red box in bottom left corner, then follow prompts\n")
+		} 
+		while(add_removeQ=="a"){
+			select_points <- locator(1,type="p", lwd=2, col=cols[i], pch=pchs[i])			
+			if( select_points$x<max(box_x) & select_points$y<max(box_y) & select_points$x>min(box_x) & select_points$y>min(box_y)) {
+				add_removeQ <- "b"
+			}
+			else{ 
 				group_data <- rbind(group_data, data.frame(id=group_id, x=select_points$x, y=select_points$y, group=i, col=cols[i], pch=pchs[i]) )
-			}		
+			}
 		}
 
 		if(add_removeQ=="d"){
-			cat("\nClick on points you want to delete\nOnce you are finished removing points exit (see above)\n")
[…]

### 2017-12-13 docs/manuscript.tex
Fixed up Discussion, various edits and inrto based on Shin's comments

@@
 Meta-analysis uses effect size estimates and their sampling variance, taken from many studies, to understand whether particular effects are common across studies and to explain variation among these effects \citep{Glass1976,Borenstein2009,Koricheva2013,Naka2017}.
-Meta-analysis therefore relies foremost on data extracted from primary literature, and more specifically, statistics that have been reported in the text or tables of research papers (e.g., means, standard deviations, correlation coefficients).
-These summary statistics are also, however, frequently presented in figures and so need to be manually extracted using digitising programs.
-While inferential statistics (e.g., t- and F-statistics) are often presented along side summary statistics and used to derive effect sizes, raw descriptive statistics are much more appropriate to use because sources of non-independence in experimental designs can be dealt with more easily using summary statistics \citep{Noble2017}.
+Meta-analysis therefore relies foremost on data extracted from primary literature, and more specifically, descriptive statistics (e.g., means, standard deviations, correlation coefficients) that have been reported in the text or tables of research papers.
+Descriptive statistics are also, however, frequently presented in figures and so need to be manually extracted using digitising programs.
+While inferential statistics (e.g., t- and F-statistics) are often presented along side descriptive statistics and can be used to derive effect sizes, descriptive statistics are much more appropriate to use because sources of non-independence in experimental designs can be dealt with more easily \citep{Noble2017}.
 Although there are several existing tools to perform tasks like this (e.g.
@@
 
-Firstly, they typically only provide the user with calibrated x,y coordinates from imported figures, and do not differentiate between common plot types that are use to present data.
-This means that a large amount of downstream data manipulation is subsequently required, that differs for different plots types.
+Firstly, they typically only provide the user with calibrated x,y coordinates from imported figures, and do not differentiate between common plot types that are used to present data.
+This means that a large amount of downstream data manipulation is subsequently required, that is different across plots types.
 For example, data are frequently presented in mean and error plots (Figure \ref{fig:plot_type}A), for which the user wants a mean and error estimate for each group presented in the figure.
@@
 From all plot types, \fct{metaDigitise} summarises the data from a figure as a mean, standard deviation and sample size, for each identified group within the plot (should multiple groups exist).
-These are the summary statistics needed to create many of the relevant effect sizes and sampling error for a meta-analysis.
+These are the descriptive statistics needed to create many of the relevant effect sizes and sampling error for a meta-analysis.
 %% are there any hwere additional info is needed??
@@
 Using this information, \fct{metaDigitise} then calculates the value of any clicked points in terms of the figure axes.
-In the case of mean and error plots and box plots, it calibrates only the y axis (assuming the x axis is redundant).
+In the case of mean and error plots and box plots, it calibrates only the y-axis (assuming the x-axis is redundant).
 For scatter plots and histograms both axes are calibrated.
@@
 
-As with the scatterplots, If the sample size from the extracted data does not match a known sample size, the user can enter an alternate sample size.
+As with the scatterplots, if the sample size from the extracted data does not match a known sample size, the user can enter an alternate sample size.
 
@@
 
-Assume that the user would like to extract summary statistics from studies measuring sepal length or width in iris species for a fictitious project.
+Assume that the user would like to extract descriptive statistics from studies measuring sepal length or width in iris species for a fictitious project.
 There are a few studies that only present these data in figures.
@@
 At first there is one figure in the folder, shown in Figure \ref{fig:Anderson}.
-Running \fct{metaDigitise} brings up a series of prompts for the user using a main menu that provides access to a number of its features (...
-here represents the user's project directory):
+Running \fct{metaDigitise} brings up a series of prompts for the user using a main menu that provides access to a number of its features ("..." here represents the user's path to the project directory):
 
@@
 First for 002\_Doe\_2013\_Fig1.png and then 002\_Doe\_2013\_Fig3.png.
-Since there are a few different figure types, the user answers the first question in the \proglang{R} console as "diff":
+Since there are a few different figure types, the user answers the first question in the \proglang{R} console as "d":
 
@@
 
-If the user chooses "Import existing data", they have the option of either 1) importing data from all digitised images or 2) importing data from particular image that has been digitised.
+If the user chooses "Import existing data", they have the option of either 1) importing data from all digitised images or 2) importing data from a single image that has been digitised.
 If 2, then a list of files are provided to the user that they can select.
@@
 
-Although \pkg{metaDigitise} is already very flexible, and provides functionality not seen in any other packages (Table \ref{tab:comparison}) it is clear that there are some functions that it does not perform.
+Although \pkg{metaDigitise} is already very flexible, and provides functionality not seen in any other package (Table \ref{tab:comparison}) it is clear that there are some functions that it does not perform.
 A notable feature that \pkg{metaDigitise} lacks is automated point detection.
@@
 Another feature that \pkg{metaDigitise} (currently) lacks, is an ability to zoom in on plots.
-This may enable users to gain greater accuracy when clicking on points.
-However, from our own experience (and indeed the results from our inter-observer reliability), if you are using a reasonably sized screen then the accuracy is already high from these programs (see results above for example), and there is not much gain is to be had from zooming in on points in many circumstances.
+Zooming may enable users to gain greater accuracy when clicking on points.
+However, from our own experience (and indeed the results from our inter-observer reliability), if you are using a reasonably sized screen then the accuracy is already high from these programs (see results above for example), and there is not much gain to be had from zooming in on points in many circumstances.
 
 In contrast to some other packages, \pkg{metaDigitise} currently also does not extract lines from figures.
-In our own experience, this is not particularly useful for meta-analysis, although we recognise that it may be useful in other fields.
-In this case, we would recommend extracting data as a scatter plot, and clicking along the line in question.
-A model can then be fitted to these points to estimate the parameters needed.
+In our own experience, line extraction is not particularly useful for meta-analysis, although we recognise that it may be useful in other fields.
+Should a user like to extract lines with \pkg{metaDigitise}, we would recommend extracting data as a scatter plot, and clicking along the line in question.
+A model can then be fitted to these points (setting the argument "summary = FALSE" in \pkg{metaDigitise} - will provide access to the raw data) to estimate the parameters needed.
 
@@
 At present this is a deliberate omission, as it is not clear how best to derive SD from such data, given also that such asymmetric error bars may represent different things in different figures.
-
-Increasing the reproducibility of figure extraction for meta-analysis and making the laborious process more streamlined, flexible and integrated with existing statistical software will go a long way in facilitating the production of high quality meta-analytic studies that can be updated in the future.
+While \pkg{metaDigitise} extracts descriptve statistics from figures, given these are most often presented in figures and are usually the most robust sources of information for caluclating effect size statitsics \citep{Noble2017}, users may also like to compare estimates from inferential statistics from those derived with descriptive statistics from a paper.
+Comparing effects sizes derived from both sources can be useful in identifying uncertainties and problems within a paper.
+In the future, we hope to provide functions to easily convert inferential statistics to standardised effect size estimates, which can seamlessly be integrated with summary statistics from \pkg{metaDigitise}, to calculate equivalent standardised effect size estimates and their sampling variance.
+
+Increasing the reproducibility of figure extraction for meta-analysis and making this laborious process more streamlined, flexible and integrated with existing statistical software will go a long way in facilitating the production of high quality meta-analytic studies that can be updated in the future.
 We belive that \pkg{metaDigitise} will improve this research synthesis pipeline, and will hopefully become an integral package that can be added to the meta-analysts toolkit.

### 2018-01-07 docs/manuscript.tex
manuscript updates

@@
 The package interactively guides the user through data extraction process.
-Notably, it enables a large-scale extraction using image files, letting the user stop processing, and edit and add to the resulting data fame at any point.
-Further, it facilitates reproducible data extractions from plots with little inter-observer bias, thus, allowing a group of people to participate the extraction of data collaboratively.
+Notably, it enables a large-scale extraction using image files, letting the user stop processing, edit and add to the resulting data fame at any point.
+Further, it facilitates reproducible data extraction from plots with little inter-observer bias, thus, allowing a group of people to participate the extraction of data collaboratively.
 }
@@
 Although there are several existing tools to perform tasks like this (e.g.
-\proglang{DataThief} \citep{DataThief}, \proglang{GraphClick} \citep{GraphClick}, \proglang{WebPlotDigitizer} \citep{WebPlotDigitizer}), these tools are not designed specifically for meta-analysis.
+\proglang{DataThief} \citep{DataThief}, \proglang{GraphClick} \citep{GraphClick}, \proglang{WebPlotDigitizer} \citep{WebPlotDigitizer}), these tools are not designed specifically for meta-analysis for three main reasons.
 %% need to spell out clear points
@@
 In essence, existing software does not provide an optimized research pipeline to facilitate data extraction, editing and reproducibility.
+
 These are major issues because extracting from figures can be an incredibly time-consuming process.
-Furthermore, although meta-analysis is an important tool in consolidating the data from many studies many of the processes involved in data extraction are opaque and difficult to reproduce, making extending studies problematic.
+Furthermore, although meta-analysis is an important tool in consolidating the data from multiple studies, many of the processes involved in data extraction are opaque and difficult to reproduce, making extending studies problematic.
 Having a tool that facilitates reproducibility in meta-analyses will increase transparency and go a long way to resolving the reproducibility crises we are seeing in many fields \citep{peng_reproducible_2006, peng_reproducible_2011, sandve_ten_2013, Parker2016,Ihle2017}.
@@
 \pkg{metaDigitise} condenses summary data extracted from multiple figures into a single data frame which can be can easily be exported.
-Processed data (i.e., raw data) can also be easily extracted and analysed in any way the user desires in downstream analysis within \proglang{R}.
+Processed data can also be easily extracted and analysed in any way the user desires in downstream analysis within \proglang{R}.
 Conveniently, when needing to process many figures at different times \pkg{metaDigitise} will only import figures not already completed within a directory.
@@
 This can be entered at a later time, however (see below).
+A function, \fct{error\_to\_sd}, that converts the different error types to SD is also available in the package.
 
@@
 \begin{equation}
-\mu = \frac{a + 2q_1 + 2m + 2q_3 +b}{8}
+\mu = \frac{(n+3)(a+b) + 2(n-1)(q_1 + m + q_3)}{8n}
 \end{equation}
+following \citet{Bland2015} and
+
+% \begin{equation}
+% \mu = \frac{a + 2q_1 + 2m + 2q_3 +b}{8}
+% \end{equation}
 
@@
 As with mean and error plots, if the user does not enter a sample size at the time of data extraction the SD is not calculated.
+Two functions, \fct{rqm\_to\_mean} and \fct{rqm\_to\_sd}, that convert box plot data to mean and SD respectively are also available in the package.
 
@@
 \begin{figure}[!h]
+\centering
  \includegraphics[width=0.75\textwidth]{001_Anderson_1935_Fig1.png} 
@@
 \fct{metaDigitise} handles this nicely by prompting the user to enter the group first, followed by digitisation of this groups data.
-After digitising the first group, and having exited (e.g., hit `esc' on mac from plot window), \fct{metaDigitise} will ask the user whether they would like to add another group.
+After digitising the first group, and having exited, \fct{metaDigitise} will ask the user whether they would like to add another group.
 Users can continually add groups (\code{a}), delete groups (\code{d}), edit groups (\code{e}) or finish a plot and continue to the next one (\code{f} - if another plot exists).
@@
 \begin{CodeOutput}
-Follow instructions below, to exit point adding or removing:
-
- - Windows: right click on the plot area and choose 'Stop'!
-
- - X11: hit any mouse button other than the left one.
-
- - quartz/OS X: hit ESC
-
+If there are multiple groups, enter unique group identifiers (otherwise press enter)
 Group identifier: 
@@
 Click on points you want to add.
-If you want to remove a point, or are finished with a
-group, exit (see above), then follow prompts
-\end{CodeOutput}
-\end{CodeChunk}
-
-To finish selecting points, the user can exit following the instructions (e.g. by pressing ESC on a mac). The user is then asked if they want to add or delete points from that group.
+If you want to remove a point, or are finished with a group, 
+exit by clicking on red box in bottom left corner, then follow prompts
+\end{CodeOutput}
+\end{CodeChunk}
+
+To finish selecting points, the user can exit by clicking on the red button that appears when extracting points. The user is then asked if they want to add or delete points from that group.
 
@@
 \begin{figure}[!h] 
- \includegraphics[width=0.9\textwidth]{001_Anderson_1935_Fig1_digitised.png} 
+\centering
+ \includegraphics[width=0.75\textwidth]{001_Anderson_1935_Fig1_digitised.png} 
  \caption{Digitisation of sepal length and width for two species of iris (setosa and versicolor).
@@
 The sample sizes for each group are provided on the lower part of the plot.
-All figures are clearly labeled at the top to remind users of the filename and plot type.
+All figures are clearly labelled at the top to remind users of the filename and plot type.
 This reduces errors throughout the digitisation process.}
@@
 
-Users can access the \code{metaDigitise} object created (001\_Anderson\_1935\_Fig1) at any time using the metaDigitse function.
+Users can access the \code{metaDigitise} object created (001\_Anderson\_1935\_Fig1) at any time using the \fct{metaDigitse} function.
 In the \proglang{R} console, the summarised data for the digitised figure can be printed on screen or even written to a .csv file:
@@
 Nonetheless, it is important to recognise the impact that overlapping points can have on summary statistics, particularly its effects on standard deviation (SD) and standard error (SE).
-Here, the mean point estimates are nearly bang on, but the SD's are slightly over-estimated:
+Here, the mean point estimates are nearly exactly the same as the true values, but the SD's are slightly over-estimated:
 
@@
 This will permit users to digitise new figures, and will integrate previously completed digitisations along with newly digitised data together at the end of the session, or when the user decides to quit.
-This time, 001\_Anderson\_1935\_Fig1.png, is ignored and the new plots cycle on screen.
-First for 002\_Doe\_2013\_Fig1.png and then 002\_Doe\_2013\_Fig3.png.
+This time, 001\_Anderson\_1935\_Fig1.png is ignored and the new plots cycle on screen; first 002\_Doe\_2013\_Fig1.png and then 002\_Doe\_2013\_Fig3.png.
 Since there are a few different figure types, the user answers the first question in the \proglang{R} console as "d":
@@
 \begin{figure}[!b] 
- \includegraphics[width=0.9\textwidth]{fig_rotate.pdf} 
+\centering
+ \includegraphics[width=0.75\textwidth]{fig_rotate.pdf} 
  \caption{Figure rotation. A) and B) show how non-aligned images can be realigned through user defined rotation. C) and D) show how figures can be re-orientated so as to aid data input.}
@@
 
-\subsection{Interobserver variability in digitisations}
+\subsection{Inter-observer variability in digitisations}
 In order to evaluate the consistency of digitisation using \pkg{metaDigitise} between users, we simulated a dataset of two traits with two different groups.
@@
 Indeed, across plot types we found no evidence for any inter-observer variability in digitisations for the mean (ICC = 0, 95\% CI = 0 to 0.029, p = 1), standard deviation (ICC = 0, 95\% CI = 0 to 0.033, p = 0.5) or correlation coefficient (ICC = 0.053, 95\% CI = 0 to 0.296, p = 0.377).
-There were was little bias in between digitised and true values, on average 1.63\% (mean = 0.02\%, sd = 4.9\%, r = -0.03\%) and overall there were only small absolute differences between digitised and true values, deviating, on average 2.18\% (mean = 0.40\%, sd = 5.81\%, r = 0.33\%) for all three summary statistics.
-
-SD estimates from digitisations are clearly more prone to error than means or correaltion cefficients.
-This is due to the fact that SD estimation from the summary statistics derived from plots is more error prone, especially at small sample sizes \citep{Wan2014}.
+There were was little bias between digitised and true values, on average 1.63\% (mean = 0.02\%, sd = 4.9\%, r = -0.03\%) and overall there were only small absolute differences between digitised and true values, deviating, on average 2.18\% (mean = 0.40\%, sd = 5.81\%, r = 0.33\%) across all three summary statistics.
+
+SD estimates from digitisations are clearly more prone to error than means or correlation coefficients.
 If the mean absoluate difference is calculated for each plot type, we can see that this effect is driven mainly by extraction from boxplots and histgrams (\% difference):
@@
 \end{CodeOutput}
+ This is because SD estimation from the summary statistics extracted from boxplots is more error prone, especially at small sample sizes \citep{Wan2014}.
 
@@
 Additionally, auto-detection often misses many points which then subsequently need to be manually added.
-Based on tests of \pkg{metaDigitise} (see above) figures can be extracted in around 1-2 minutes, including the entry of metadata.
+Based on tests of \pkg{metaDigitise} (see above), figures can be extracted in around 1-2 minutes, including the entry of metadata.
 As a result, we do not belive that current automated point detection provides substantial benefits in terms of time or accuracy.
@@
 Zooming may enable users to gain greater accuracy when clicking on points.
-However, from our own experience (and indeed the results from our inter-observer reliability), if you are using a reasonably sized screen then the accuracy is already high from these programs (see results above for example), and there is not much gain to be had from zooming in on points in many circumstances.
+However, from our own experience (and indeed from the results above), if you are using a reasonably sized screen then the accuracy is already high from these programs, and there is not much gain to be had from zooming in on points in many circumstances.
 
@@
 Should a user like to extract lines with \pkg{metaDigitise}, we would recommend extracting data as a scatter plot, and clicking along the line in question.
-A model can then be fitted to these points (setting the argument "summary = FALSE" in \pkg{metaDigitise} - will provide access to the raw data) to estimate the parameters needed.
+A model can then be fitted to these points (setting the argument "summary = FALSE" in \pkg{metaDigitise} - will provide access to the processed data) to estimate the parameters needed.
 
@@
 At present this is a deliberate omission, as it is not clear how best to derive SD from such data, given also that such asymmetric error bars may represent different things in different figures.
-While \pkg{metaDigitise} extracts descriptve statistics from figures, given these are most often presented in figures and are usually the most robust sources of information for caluclating effect size statitsics \citep{Noble2017}, users may also like to compare estimates from inferential statistics from those derived with descriptive statistics from a paper.
+
+Descriptve statistics are usually the most robust sources of information for calculating effect size statistics \citep{Noble2017}.
+These are most often presented in figures.
+Users may therefore also want to compare effect size estimates from inferential statistics with those derived from descriptive statistics (obtained for example using \pkg{metaDigitise}) from a paper.
+Comparing these different effects sizes can be useful in identifying uncertainties and problems within a paper.
+In the future, we hope to provide functions to easily convert inferential statistics to standardised effect size estimates, which can seamlessly be integrated with summary statistics from \pkg{metaDigitise}, to calculate equivalent standardised effect size estimates and their sampling variance.
+
+%While \pkg{metaDigitise} extracts descriptve statistics from figures, given these are most often presented in figures and are usually the most robust sources of information for calculating effect size statistics \citep{Noble2017}, users may also like to compare estimates from inferential statistics from those derived with descriptive statistics from a paper.
 Comparing effects sizes derived from both sources can be useful in identifying uncertainties and problems within a paper.

### 2018-01-07 R/transformStats.R

@@ -31,7 +31,7 @@ se_to_sd <- function(se, n) {
 
 
 #' @title CI95_to_sd
-#' @description Transforms standard error to standard deviation
+#' @description Transforms symmetrical confidence interval to standard deviation
 #' @param CI Interval difference from the mean
 #' @param n Sample Size
 #' @return Returns vector of standard deviations
@@ -51,10 +51,11 @@ CI95_to_sd <- function(CI,n) {
 #' @param median Median
 #' @param UQ Upper 75th quartile
 #' @param max Maximum value
+#' @param n Sample size
 #' @return Returns vector of mean
 #' @author Joel Pick
 #' @export
-rqm_to_mean <- function(min,LQ,median,UQ,max){
+rqm_to_mean <- function(min,LQ,median,UQ,max,n){
 	b <- max
 	a <- min
 	q3 <- UQ
@@ -63,6 +64,8 @@ rqm_to_mean <- function(min,LQ,median,UQ,max){
 	
 	##from Wan et al. 2014; equation 10
 	Xbar <- (a + 2*q1 + 2*m + 2*q3 + b)/8
+	## from Bland 2014;
+	#Xbar <- ( (n+3)*(a+b) + 2*(n-1)*(q_1 + m + q_3) )/8n
 	return(Xbar)
 }
 

### 2018-01-07 docs/manuscript.tex
Updates to MS. Still need to sort out some test failure with a dependency package

@@
 \Abstract{
-Research synthesis, especially, in the form of meta-analysis requires data extraction from primary studies.
+Research synthesis, especially in the form of meta-analysis, requires data extraction from primary studies.
 Meta-analysis synthesizes effect sizes, often calculated from summary statistics of studies.
 However, exact values of such statistics are commonly hidden in figures.
-The R package metaDigitise extracts descriptive statistics such as means, standard deviations and, if applicable, correlations from the four types of plots: 1) mean and error plots (e.g.
+The R package \pkg{metaDigitise} extracts descriptive statistics such as means, standard deviations and, if applicable, correlations from the four types of plots: 1) mean and error plots (e.g.
 bar graphs with standard errors), 2) box plots, 3) scatter plots and 4) histograms.
@@
 
-Firstly, they typically only provide the user with calibrated x,y coordinates from imported figures, and do not differentiate between common plot types that are used to present data.
+First, they typically only provide the user with calibrated x,y coordinates from imported figures, and do not differentiate between common plot types that are used to present data.
 This means that a large amount of downstream data manipulation is subsequently required, that is different across plots types.
@@
 
-Secondly, digitising programs do not easily allow the integration of metadata at the time of data extraction, such as experimental group or variable names, and sample sizes.
+Second, digitising programs do not easily allow the integration of metadata at the time of data extraction, such as experimental group or variable names, and sample sizes.
 This makes the downstream calculations more laborious, as the information has to be added later, in most cases using different software.
@@
 It also provides users with options to conduct the necessary calculations on processed data immediately after extraction so that comparable summary statistics can be obtained quickly.
-\pkg{metaDigitise} condenses summary data extracted from multiple figures into a single data frame which can be can easily be exported.
+\pkg{metaDigitise} condenses summary data extracted from multiple figures into a single data frame which can be can easily exported.
 Processed data can also be easily extracted and analysed in any way the user desires in downstream analysis within \proglang{R}.

### 2018-01-07 tests/testthat/test-metaDigitise.R
small fix just in case there are issues with test...sometimes weird files get written to temp folder, so extract the caldat from this list

@@ -1,11 +1,9 @@
 context("Check various metaDigitise.R functions...")
 
-
-
 test_that("Checking setup_calibration_dir works as expected..", {
 	dir <- paste0(tempdir(), "/")
 	setup_calibration_dir(dir)
-	expect_equal(list.files(dir), "caldat", info = "caldat setup correctly")
+	expect_equal(list.files(dir)[grep("caldat",list.files(dir))], "caldat", info = "caldat setup correctly")
 })
 
 test_that("Checking dir_details works as expected..", {

### 2018-01-08 docs/manuscript.tex
manuscript updates

@@
 Descriptive statistics are also, however, frequently presented in figures and so need to be manually extracted using digitising programs.
-While inferential statistics (e.g., t- and F-statistics) are often presented along side descriptive statistics and can be used to derive effect sizes, descriptive statistics are much more appropriate to use because sources of non-independence in experimental designs can be dealt with more easily \citep{Noble2017}.
+While inferential statistics (e.g., \textit{t}- and \textit{F}-statistics) are often presented along side descriptive statistics and can be used to derive effect sizes, descriptive statistics are much more appropriate to use because sources of non-independence in experimental designs can be dealt with more easily \citep{Noble2017}.
 Although there are several existing tools to perform tasks like this (e.g.
@@
 
-First, they typically only provide the user with calibrated x,y coordinates from imported figures, and do not differentiate between common plot types that are used to present data.
+First, they typically only provide the user with calibrated \textit{x,y} coordinates from imported figures, and do not differentiate between common plot types that are used to present data.
 This means that a large amount of downstream data manipulation is subsequently required, that is different across plots types.
 For example, data are frequently presented in mean and error plots (Figure \ref{fig:plot_type}A), for which the user wants a mean and error estimate for each group presented in the figure.
-With existing programs, x,y coordinates of means and errors are returned, to which the user must manually discern between mean and error coordinates and assign points to groups.
+With existing programs, \textit{x,y} coordinates of means and errors are returned, to which the user must manually discern between mean and error coordinates and assign points to groups.
 The error then needs to be calculated as the deviation from the mean, and then transformed to a standard deviation, depending on the type of error presented.
@@
 \begin{equation}
-\sigma = SE * \sqrt{n}
+\sigma = SE \sqrt{n}
 \end{equation}
@@
 \begin{equation}
-\sigma = \frac{CI}{1.96} * \sqrt{n}
+\sigma = \frac{CI}{1.96} \sqrt{n}
 \end{equation}
@@
 
-, where $\Phi^{-1}(z)$ is the upper zth percentile of the standard normal distribution, following \citet{Wan2014}.
+where $\Phi^{-1}(z)$ is the upper zth percentile of the standard normal distribution, following \citet{Wan2014}.
 As with mean and error plots, if the user does not enter a sample size at the time of data extraction the SD is not calculated.
@@
 Different groups are plotted in different colours and shapes, with a legend at the bottom of the figure (Figure \ref{fig:all_extract}C).
-Mean, SD and sample size (n) are calculated from the clicked points, for each group.
+Mean, SD and sample size are calculated from the clicked points, for each group.
 Where the sample size from the clicked points does not match a known sample size (e.g.
@@
 If digitisations were consistent across all users then we should find no significant between user variability in the data.
-Indeed, across plot types we found no evidence for any inter-observer variability in digitisations for the mean (ICC = 0, 95\% CI = 0 to 0.029, p = 1), standard deviation (ICC = 0, 95\% CI = 0 to 0.033, p = 0.5) or correlation coefficient (ICC = 0.053, 95\% CI = 0 to 0.296, p = 0.377).
-There were was little bias between digitised and true values, on average 1.63\% (mean = 0.02\%, sd = 4.9\%, r = -0.03\%) and overall there were only small absolute differences between digitised and true values, deviating, on average 2.18\% (mean = 0.40\%, sd = 5.81\%, r = 0.33\%) across all three summary statistics.
+Indeed, across plot types we found no evidence for any inter-observer variability in digitisations for the mean (ICC = 0, 95\% CI = 0 to 0.029, \textit{p} = 1), standard deviation (ICC = 0, 95\% CI = 0 to 0.033, \textit{p} = 0.5) or correlation coefficient (ICC = 0.053, 95\% CI = 0 to 0.296, \textit{p} = 0.377).
+There were was little bias between digitised and true values, on average 1.63\% (mean = 0.02\%, SD = 4.9\%, \textit{r} = -0.03\%) and overall there were only small absolute differences between digitised and true values, deviating, on average 2.18\% (mean = 0.40\%, SD = 5.81\%, \textit{r} = 0.33\%) across all three summary statistics.
 
@@
 This was done by one digitiser, as there is no detectable between user variation.
-Data digitised using \pkg{metaDigitise} was essentially perfectly correlated with the true simulated data for both the x-variable (Pearson's correlation; r = 0.9999915, t = 2137.4, df = 78, p-value < 0.001) and y-variable (r = 0.9999892, t = 1897.8, df = 78, p-value < 0.001).
+Data digitised using \pkg{metaDigitise} was essentially perfectly correlated with the true simulated data for both the \textit{x}-variable (Pearson's correlation; \textit{r} = 0.9999915, \textit{t} = 2137.4, \textit{df} = 78, \textit{p} < 0.001) and \textit{y}-variable (\textit{r} = 0.9999892, \textit{t} = 1897.8, \textit{df} = 78, \textit{p} < 0.001).
 
@@
 Reproducable$^9$  & \checkmark   & \checkmark & \checkmark & $\times$   & \checkmark     & $\times$       & $\times$    \\
-Automated point detection & $\times$     & \checkmark &     ?      & \checkmark & \checkmark     & \checkmark     & $\times$    \\
+Automated point detection & $\times$     & \checkmark &     $\times$      & \checkmark & \checkmark     & \checkmark     & $\times$    \\
 Line extraction           & $\times$     & \checkmark & \checkmark & \checkmark & \checkmark     & $\times$       & $\times$    \\
-Zoom                      & $\times$     & \checkmark & \checkmark & \checkmark &       ?        & $\times$       & $\times$    \\
-Log axis                  & $\times$ *   & \checkmark & \checkmark & \checkmark & \checkmark     & $\times$       & $\times$    \\
-Dates                     & $\times$ *   & $\times$   & \checkmark & $\times$   & \checkmark     & $\times$       & $\times$    \\
+Zoom                      & $\times$     & \checkmark & \checkmark & \checkmark & \checkmark     & $\times$       & $\times$    \\
+Log axis                  & $\times$   & \checkmark & \checkmark & \checkmark & \checkmark     & $\times$       & $\times$    \\
+Dates                     & $\times$   & $\times$   & \checkmark & $\times$   & \checkmark     & $\times$       & $\times$    \\
 Asymmetric error bars     & $\times$     & $\times$   & \checkmark & $\times$   & $\times$       & $\times$       & $\times$    \\

### 2018-01-14 docs/manuscript_plain.tex
preparing manuscript for submission

--- 2018-01-08 docs/manuscript.tex
@@
-\documentclass[article]{jss}
+\documentclass[12pt]{article}
 
@@
 %% recommended packages
-\usepackage{thumbpdf,lmodern,amssymb,rotating}
+\usepackage{amssymb,rotating,natbib,graphicx,fancyvrb,fullpage} 
+\usepackage[parfill]{parskip} 
+\usepackage{geometry}            % See geometry.pdf to learn the layout options. There are lots.
+\geometry{a4paper} 
+\usepackage[utf8]{inputenc}
+\usepackage{setspace}
+\usepackage{lineno}                                                                            
+\doublespacing
+\linenumbers
+\renewcommand\linenumberfont{\normalfont\small}
 
 %% new custom commands
+\newcommand{\code}[1]{\texttt{#1}}
 \newcommand{\class}[1]{`\code{#1}'}
-\newcommand{\fct}[1]{\code{#1()}}
-
-%% -- Article metainformation (author, title, ...) -----------------------------
-
-%% - \author{} with primary affiliation
-%% - \Plainauthor{} without affiliations
-%% - Separate authors by \And or \AND (in \author) or by comma (in \Plainauthor).
-%% - \AND starts a new line, \And does not.
-\author{Joel L. Pick\\University of New South Wales
-   \And Shinichi Nakagawa\\University of New South Wales
-   \AND Daniel W.A. Noble\\University of New South Wales
-   }
-\Plainauthor{Joel L. Pick, Shinichi Nakagawa, Daniel W.A. Noble}
-
-%% - \title{} in title case
-%% - \Plaintitle{} without LaTeX markup (if any)
-%% - \Shorttitle{} with LaTeX markup (if any), used as running title
-\title{Reproducible, flexible and high throughput data extraction from primary literature: The \pkg{metaDigitise} \proglang{R} package}
-\Plaintitle{Reproducible, flexible and high throughput data extraction from primary literature: The metaDigitise R package}
-\Shorttitle{Data extraction with \pkg{metaDigitise}}
+%\newcommand{\fct}[1]{#1}
+\newcommand{\fct}[1]{\texttt{#1()}}
+\newcommand{\pkg}[1]{{\fontseries{b}\selectfont #1}}
+\let\proglang=\textsf
+
+\begin{document}
+
+\DefineVerbatimEnvironment{Code}{Verbatim}{}
+\DefineVerbatimEnvironment{CodeInput}{Verbatim}{fontshape=sl}
+\DefineVerbatimEnvironment{CodeOutput}{Verbatim}{}
+\newenvironment{CodeChunk}{}{}
+
+\raggedright
+
+\textbf{Reproducible, flexible and high throughput data extraction from primary literature: The \pkg{metaDigitise} \proglang{R} package}
+
+Joel L. Pick$^{1,*}$, Shinichi Nakagawa$^1$, Daniel W.A. Noble$^1$
+
+$^1$
+  Ecology and Evolution Research Centre, School of Biological, Earth and Environmental Sciences,  University of New South Wales, Kensington, NSW 2052, Sydney, AUSTRALIA
+
+ $^*$Corresponding Author: joel.l.pick@gmail.com\\
 
 %% - \Abstract{} almost as usual
-\Abstract{
+\section*{Abstract}
 Research synthesis, especially in the form of meta-analysis, requires data extraction from primary studies.
@@
 Further, it facilitates reproducible data extraction from plots with little inter-observer bias, thus, allowing a group of people to participate the extraction of data collaboratively.
-}
-
-%% - \Keywords{} with LaTeX markup, at least one required
-%% - \Plainkeywords{} without LaTeX markup (if necessary)
-%% - Should be comma-separated and in sentence case.
-\Keywords{meta-analysis, comparative analysis, data extraction, \proglang{R}, reproducibility, figures, images, summary statistics}
-\Plainkeywords{meta-analysis, comparative analysis, data extraction, R, reproducibility, figures, images, summary statistics}
-
-%% - \Address{} of at least one author
-%% - May contain multiple affiliations for each author
-%%   (in extra lines, separated by \emph{and}\\).
-%% - May contain multiple authors for the same affiliation
-%%   (in the same first line, separated by comma).
-\Address{
-  Joel L. Pick, Shinichi Nakagawa, Daniel W.A. Noble\\
-  Ecology and Evolution Research Centre\\
-  School of Biological, Earth and Environmental Sciences\\
-  University of New South Wales\\
-  Kensington, NSW 2052\\
-  Sydney, AUSTRALIA\\
-  E-mail: \email{joel.l.pick@gmail.com}\\
-  URL: \url{https://github.com/daniel1noble/metaDigitise}
-}
-
-\begin{document}
+
+Keywords: meta-analysis, comparative analysis, data extraction, \proglang{R}, reproducibility, figures, images, summary statistics
 
@@
 
-\section{Introduction}
+\section*{Introduction}
 
@@
 
-\section{Directory Structure, Image Processing and Reproducibility}
+\section*{Directory Structure, Image Processing and Reproducibility}
 
@@
 
-\section{Diverse Plot Types}
+\section*{Diverse Plot Types}
 \pkg{metaDigitise} recognises four main types of plot; Mean and error plots, box plots, scatter plots and histograms, shown in Figure \ref{fig:plot_type}.
@@
 
-\subsection{Mean and error plots} 
+\subsection*{Mean and error plots} 
 \fct{metaDigitise} prompts the user to enter group names and allows the user to enter sample sizes ($n$), which are used in downstream processing.
@@
 
-\subsection{Box plots}
+\subsection*{Box plots}
 As with mean and error plots, \fct{metaDigitise} prompts the user to enter group names and allows the user to enter sample sizes ($n$), which are used in downstream processing.
@@
 
-\subsection{Scatter plots}
+\subsection*{Scatter plots}
 \fct{metaDigitise} prompts the user to enter groups names and then to click on points.
@@
 
-\subsection{Histograms}
+\subsection*{Histograms}
 \fct{metaDigitise} prompts the user to click on the top corners of each bar. Bars can subsequently be deleted. 
@@
 
-\section{Extracting Data From Plots}
+\section*{Extracting Data From Plots}
 %% scatterplot from airquality data
@@
 
-\subsection{Adding new figures}
+\subsection*{Adding new figures}
 
@@
 
-\section{Re-importing, Editing and Plotting Previously Digitised data}
+\section*{Re-importing, Editing and Plotting Previously Digitised data}
 
@@
 
-\section{Additional Features}
-
-\subsection{Figure Rotation and Adjustment}
+\section*{Additional Features}
+
+\subsection*{Figure Rotation and Adjustment}
 Figures may have been extracted from old publications, for example from scanned images, and so are not perfectly orientated on the image.
@@
 
-\subsection{Obtaining Processed Data}
+\subsection*{Obtaining Processed Data}
 
@@
 
-\subsection{Adding sample sizes to previous Digitisations}
+\subsection*{Adding sample sizes to previous Digitisations}
 %Including addition of N later
@@
 
-\section{Inter-observer Variability and Validation}
-
-\subsection{Inter-observer variability in digitisations}
+\section*{Inter-observer Variability and Validation}
+
+\subsection*{Inter-observer variability in digitisations}
 In order to evaluate the consistency of digitisation using \pkg{metaDigitise} between users, we simulated a dataset of two traits with two different groups.
@@
 Each variable was plotted twice for each given plot type (figures were modified slightly to give users a sense that they were digitising new data) generating a total of 14 figures.
-15 independent digitisers were provided with a directory with all 14 figures in a randomised order.
+14 independent digitisers were provided with a directory with all 14 figures in a randomised order.
 Digitisers ran \pkg{metaDigistise} on their own computers, across different operating systems (including Mac, Windows and Linux).
@@
 
-\subsection{Testing the accuracy of digitisations}
+\subsection*{Testing the accuracy of digitisations}
 To test how accurate \pkg{metaDigitise} is at matching points to their true values, we generated four random scatterplots, each with 20 data points, and digitised these with \fct{metaDigitise}.
@@
 
-\section{Discussion and Conclusions}
+\section*{Discussion and Conclusions}
 
@@
 
-%While \pkg{metaDigitise} extracts descriptve statistics from figures, given these are most often presented in figures and are usually the most robust sources of information for calculating effect size statistics \citep{Noble2017}, users may also like to compare estimates from inferential statistics from those derived with descriptive statistics from a paper.
-Comparing effects sizes derived from both sources can be useful in identifying uncertainties and problems within a paper.
-In the future, we hope to provide functions to easily convert inferential statistics to standardised effect size estimates, which can seamlessly be integrated with summary statistics from \pkg{metaDigitise}, to calculate equivalent standardised effect size estimates and their sampling variance.
-
 Increasing the reproducibility of figure extraction for meta-analysis and making this laborious process more streamlined, flexible and integrated with existing statistical software will go a long way in facilitating the production of high quality meta-analytic studies that can be updated in the future.
@@
 
+\bibliographystyle{FuncEcol.bst}
 \bibliography{metaDigitise}

### 2018-02-27 docs/JSS_submission/jss-article-tex/article.tex
Updating repo with revised MS from Joel

\documentclass[article]{jss}

%% -- LaTeX packages and custom commands ---------------------------------------

%% recommended packages
\usepackage{thumbpdf,lmodern}

%% another package (only for this demo article)
\usepackage{framed}

%% new custom commands
\newcommand{\class}[1]{`\code{#1}'}
\newcommand{\fct}[1]{\code{#1()}}

%% -- Article metainformation (author, title, ...) -----------------------------

%% - \author{} with primary affiliation
%% - \Plainauthor{} without affiliations
%% - Separate authors by \And or \AND (in \author) or by comma (in \Plainauthor).
%% - \AND starts a new line, \And does not.
\author{Achim Zeileis\\Universit\"at Innsbruck
   \And Second Author\\Plus Affiliation}
\Plainauthor{Achim Zeileis, Second Author}

%% - \title{} in title case
%% - \Plaintitle{} without LaTeX markup (if any)
%% - \Shorttitle{} with LaTeX markup (if any), used as running title
\title{A Short Demo Article: Regression Models for Count Data in \proglang{R}}
\Plaintitle{A Short Demo Article: Regression Models for Count Data in R}
\Shorttitle{A Short Demo Article in \proglang{R}}

%% - \Abstract{} almost as usual
\Abstract{
  This short article illustrates how to write a manuscript for the
  \emph{Journal of Statistical Software} (JSS) using its {\LaTeX} style files.
  Generally, we ask to follow JSS's style guide and FAQs precisely. Also,
  it is recommended to keep the {\LaTeX} code as simple as possible,
  i.e., avoid inclusion of packages/commands that are not necessary.
  For outlining the typical structure of a JSS article some brief text snippets
  are employed that have been inspired by \cite{Zeileis+Kleiber+Jackman:2008},
  discussing count data regression in \proglang{R}. Editorial comments and
  instructions are marked by vertical bars.
}

%% - \Keywords{} with LaTeX markup, at least one required
%% - \Plainkeywords{} without LaTeX markup (if necessary)
%% - Should be comma-separated and in sentence case.
\Keywords{JSS, style guide, comma-separated, not capitalized, \proglang{R}}
\Plainkeywords{JSS, style guide, comma-separated, not capitalized, R}

%% - \Address{} of at least one author
%% - May contain multiple affiliations for each author
%%   (in extra lines, separated by \emph{and}\\).
%% - May contain multiple authors for the same affiliation
%%   (in the same first line, separated by comma).
\Address{
  Achim Zeileis\\
  Journal of Statistical Software\\
  \emph{and}\\
  Department of Statistics\\
  Faculty of Economics and Statistics\\
  Universit\"at Innsbruck\\
  Universit\"atsstr.~15\\
  6020 Innsbruck, Austria\\
  E-mail: \email{Achim.Zeileis@R-project.org}\\
  URL: \url{https://eeecon.uibk.ac.at/~zeileis/}
}

\begin{document}

%% -- Introduction -------------------------------------------------------------

%% - In principle "as usual".
%% - But should typically have some discussion of both _software_ and _methods_.
%% - Use \proglang{}, \pkg{}, and \code{} markup throughout the manuscript.
%% - If such markup is in (sub)section titles, a plain text version has to be
%%   added as well.
%% - All software mentioned should be properly \cite-d.
%% - All abbreviations should be introduced.
%% - Unless the expansions of abbreviations are proper names (like "Journal
%%   of Statistical Software" above) they should be in sentence case (like
%%   "generalized linear models" below).

\section[Introduction: Count data regression in R]{Introduction: Count data regression in \proglang{R}} \label{sec:intro}

\begin{leftbar}
The introduction is in principle ``as usual''. However, it should usually embed
both the implemented \emph{methods} and the \emph{software} into the respective
relevant literature. For the latter both competing and complementary software
should be discussed (within the same software environment and beyond), bringing
out relative (dis)advantages. All software mentioned should be properly
\verb|\cite{}|d. (See also Appendix~\ref{app:bibtex} for more details on
\textsc{Bib}{\TeX}.)

For writing about software JSS requires authors to use the markup
\verb|\proglang{}| (programming languages and large programmable systems),
\verb|\pkg{}| (software packages), \verb|\code{}| (functions, commands,
arguments, etc.). If there is such markup in (sub)section titles (as above), a
plain text version has to be provided in the {\LaTeX} command as well. Below we
also illustrate how abbrevations should be introduced and citation commands can
be employed. See the {\LaTeX} code for more details.
\end{leftbar}

Modeling count variables is a common task in economics and the social sciences.
The classical Poisson regression model for count data is often of limited use in
these disciplines because empirical count data sets typically exhibit
overdispersion and/or an excess number of zeros. The former issue can be
addressed by extending  the plain Poisson regression model in various
directions: e.g., using sandwich covariances or estimating an additional
dispersion parameter (in a so-called quasi-Poisson model). Another more formal
way is to use a negative binomial (NB) regression. All of these models belong to
the family of generalized linear models (GLMs). However, although these models
typically can capture overdispersion rather well, they are in many applications
not sufficient for  modeling excess zeros. Since \cite{Mullahy:1986} there is
increased interest in zero-augmented models that address this issue by a second
model component capturing zero counts. An overview of count data models in
econometrics, including  hurdle and zero-inflated models, is provided in
\cite{Cameron+Trivedi:2013}.

In \proglang{R} \citep{R}, GLMs are provided by the model fitting functions
\fct{glm} in the \pkg{stats} package and \fct{glm.nb} in the \pkg{MASS} package
\citep[][Chapter~7.4]{Venables+Ripley:2002} along with associated methods for
diagnostics and inference. The manuscript that this document is based on
\citep{Zeileis+Kleiber+Jackman:2008} then introduced hurdle and zero-inflated
count models in the functions \fct{hurdle} and \fct{zeroinfl} in the \pkg{pscl}
package \citep{Jackman:2015}. Of course, much more software could be discussed
here, including (but not limited to) generalized additive models for count data
as available in the \proglang{R} packages \pkg{mgcv} \cite{Wood:2006}, 
\pkg{gamlss} \citep{Stasinopoulos+Rigby:2007}, or \pkg{VGAM} \citep{Yee:2009}.

%% -- Manuscript ---------------------------------------------------------------

%% - In principle "as usual" again.
%% - When using equations (e.g., {equation}, {eqnarray}, {align}, etc.
%%   avoid empty lines before and after the equation (which would signal a new
%%   paragraph.
%% - When describing longer chunks of code that are _not_ meant for execution
%%   (e.g., a function synopsis or list of arguments), the environment {Code}
%%   is recommended. Alternatively, a plain {verbatim} can also be used.
%%   (For executed code see the next section.)

\section{Models and software} \label{sec:models}

The basic Poisson regression model for count data is a special case of the GLM
framework \cite{McCullagh+Nelder:1989}. It describes the dependence of a count
response variable $y_i$ ($i = 1, \dots, n$) by assuming a Poisson distribution
$y_i \sim \mathrm{Pois}(\mu_i)$. The dependence of the conditional mean
$\E[y_i \, | \, x_i] = \mu_i$ on the regressors $x_i$ is then specified via a
log link and a linear predictor
%
\begin{equation} \label{eq:mean}
\log(\mu_i) \quad = \quad x_i^\top \beta,
\end{equation}
%
where the regression coefficients $\beta$ are estimated by maximum likelihood
(ML) using the iterative weighted least squares (IWLS) algorithm.

\begin{leftbar}
Note that around the \verb|{equation}| above there should be no spaces (avoided
in the {\LaTeX} code by \verb|%| lines) so that ``normal'' spacing is used and
not a new paragraph started.
\end{leftbar}

\proglang{R} provides a very flexible implementation of the general GLM
framework in the function \fct{glm} \citep{Chambers+Hastie:1992} in the
\pkg{stats} package. Its most important arguments are
\begin{Code}
glm(formula, data, subset, na.action, weights, offset,
  family = gaussian, start = NULL, control = glm.control(...),
  model = TRUE, y = TRUE, x = FALSE, ...)
\end{Code}
where \code{formula} plus \code{data} is the now standard way of specifying
regression relationships in \proglang{R}/\proglang{S} introduced in
\cite{Chambers+Hastie:1992}. The remaining arguments in the first line
(\code{subset}, \code{na.action}, \code{weights}, and \code{offset}) are also
standard  for setting up formula-based regression models in
\proglang{R}/\proglang{S}. The arguments in the second line control aspects
specific to GLMs while the arguments in the last line specify which components
are returned in the fitted model object (of class \class{glm} which inherits
from \class{lm}). For further arguments to \fct{glm} (including alternative
specifications of starting values) see \code{?glm}. For estimating a Poisson
model \code{family = poisson} has to be specified.

\begin{leftbar}
As the synopsis above is a code listing that is not meant to be executed,
one can use either the dedicated \verb|{Code}| environment or a simple
\verb|{verbatim}| environment for this. Again, spaces before and after should be
avoided.

Finally, there might be a reference to a \verb|{table}| such as
Table~\ref{tab:overview}. Usually, these are placed at the top of the page
(\verb|[t!]|), centered (\verb|\centering|), with a caption below the table,
column headers and captions in sentence style, and if possible avoiding vertical
lines.
\end{leftbar}

\begin{table}[t!]
\centering
\begin{tabular}{lllp{7.4cm}}
\hline
Type           & Distribution & Method   & Description \\ \hline
GLM            & Poisson      & ML       & Poisson regression: classical GLM,
                                           estimated by maximum likelihood (ML) \\
               &              & Quasi    & ``Quasi-Poisson regression'':
                                           same mean function, estimated by
                                           quasi-ML (QML) or equivalently
                                           generalized estimating equations (GEE),
                                           inference adjustment via estimated
                                           dispersion parameter \\
               &              & Adjusted & ``Adjusted Poisson regression'':
                                           same mean function, estimated by
                                           QML/GEE, inference adjustment via
                                           sandwich covariances\\
               & NB           & ML       & NB regression: extended GLM,
                                           estimated by ML including additional
                                           shape parameter \\ \hline
Zero-augmented & Poisson      & ML       & Zero-inflated Poisson (ZIP),
                                           hurdle Poisson \\
               & NB           & ML       & Zero-inflated NB (ZINB),
                                           hurdle NB \\ \hline
\end{tabular}
\caption{\label{tab:overview} Overview of various count regression models. The
table is usually placed at the top of the page (\texttt{[t!]}), centered
(\texttt{centering}), has a caption below the table, column headers and captions
are in sentence style, and if possible vertical lines should be avoided.}
\end{table}

%% -- Illustrations ------------------------------------------------------------

%% - Virtually all JSS manuscripts list source code along with the generated
%%   output. The style files provide dedicated environments for this.
%% - In R, the environments {Sinput} and {Soutput} - as produced by Sweave() or
%%   or knitr using the render_sweave() hook - are used (without the need to
%%   load Sweave.sty).
%% - Equivalently, {CodeInput} and {CodeOutput} can be used.
%% - The code input should use "the usual" command prompt in the respective
%%   software system.
%% - For R code, the prompt "R> " should be used with "+  " as the
%%   continuation prompt.
%% - Comments within the code chunks should be avoided - these should be made
%%   within the regular LaTeX text.

\section{Illustrations} \label{sec:illustrations}

For a simple illustration of basic Poisson and NB count regression the
\code{quine} data from the \pkg{MASS} package is used. This provides the number
of \code{Days} that children were absent from school in Australia in a
particular year, along with several covariates that can be employed as regressors.
The data can be loaded by
%
\begin{CodeChunk}
\begin{CodeInput}
R> data("quine", package = "MASS")
\end{CodeInput}
\end{CodeChunk}
%
and a basic frequency distribution of the response variable is displayed in
Figure~\ref{fig:quine}.

\begin{leftbar}
For code input and output, the style files provide dedicated environments.
Either the ``agnostic'' \verb|{CodeInput}| and \verb|{CodeOutput}| can be used
or, equivalently, the environments \verb|{Sinput}| and \verb|{Soutput}| as
produced by \fct{Sweave} or \pkg{knitr} when using the \code{render_sweave()}
hook. Please make sure that all code is properly spaced, e.g., using
\code{y = a + b * x} and \emph{not} \code{y=a+b*x}. Moreover, code input should
use ``the usual'' command prompt in the respective software system. For
\proglang{R} code, the prompt \code{"R> "} should be used with \code{"+  "} as
the continuation prompt. Generally, comments within the code chunks should be
avoided -- and made in the regular {\LaTeX} text instead. Finally, empty lines
before and after code input/output should be avoided (see above).
\end{leftbar}

\begin{figure}[t!]
\centering
\includegraphics{article-visualization}
\caption{\label{fig:quine} Frequency distribution for number of days absent
from school.}
\end{figure}

As a first model for the \code{quine} data, we fit the basic Poisson regression
model. (Note that JSS prefers when the second line of code is indented by two
spaces.)
%
\begin{CodeChunk}
\begin{CodeInput}
R> m_pois <- glm(Days ~ (Eth + Sex + Age + Lrn)^2, data = quine,
+    family = poisson)
\end{CodeInput}
\end{CodeChunk}
%
To account for potential overdispersion we also consider a negative binomial
GLM.
%
\begin{CodeChunk}
\begin{CodeInput}
R> library("MASS")
R> m_nbin <- glm.nb(Days ~ (Eth + Sex + Age + Lrn)^2, data = quine)
\end{CodeInput}
\end{CodeChunk}
%
In a comparison with the BIC the latter model is clearly preferred.
%
\begin{CodeChunk}
\begin{CodeInput}
R> BIC(m_pois, m_nbin)
\end{CodeInput}
\begin{CodeOutput}
       df      BIC
m_pois 18 2046.851
m_nbin 19 1157.235
\end{CodeOutput}
\end{CodeChunk}
%
Hence, the full summary of that model is shown below.
%
\begin{CodeChunk}
\begin{CodeInput}
R> summary(m_nbin)
\end{CodeInput}
\begin{CodeOutput}
Call:
glm.nb(formula = Days ~ (Eth + Sex + Age + Lrn)^2, data = quine, 
    init.theta = 1.60364105, link = log)

Deviance Residuals: 
    Min       1Q   Median       3Q      Max  
-3.0857  -0.8306  -0.2620   0.4282   2.0898

Coefficients: (1 not defined because of singularities)
            Estimate Std. Error z value Pr(>|z|)    
(Intercept)  3.00155    0.33709   8.904  < 2e-16 ***
EthN        -0.24591    0.39135  -0.628  0.52977    
SexM        -0.77181    0.38021  -2.030  0.04236 *  
AgeF1       -0.02546    0.41615  -0.061  0.95121    
AgeF2       -0.54884    0.54393  -1.009  0.31296    
AgeF3       -0.25735    0.40558  -0.635  0.52574    
LrnSL        0.38919    0.48421   0.804  0.42153    
EthN:SexM    0.36240    0.29430   1.231  0.21818    
EthN:AgeF1  -0.70000    0.43646  -1.604  0.10876    
EthN:AgeF2  -1.23283    0.42962  -2.870  0.00411 ** 
EthN:AgeF3   0.04721    0.44883   0.105  0.91622    
EthN:LrnSL   0.06847    0.34040   0.201  0.84059    
SexM:AgeF1   0.02257    0.47360   0.048  0.96198    
SexM:AgeF2   1.55330    0.51325   3.026  0.00247 ** 
SexM:AgeF3   1.25227    0.45539   2.750  0.00596 ** 
SexM:LrnSL   0.07187    0.40805   0.176  0.86019    
AgeF1:LrnSL -0.43101    0.47948  -0.899  0.36870    
AgeF2:LrnSL  0.52074    0.48567   1.072  0.28363    
AgeF3:LrnSL       NA         NA      NA       NA    
---
Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1

(Dispersion parameter for Negative Binomial(1.6036) family taken to be 1)

    Null deviance: 235.23  on 145  degrees of freedom
Residual deviance: 167.53  on 128  degrees of freedom
AIC: 1100.5

Number of Fisher Scoring iterations: 1

              Theta:  1.604 
          Std. Err.:  0.214

 2 x log-likelihood:  -1062.546 
\end{CodeOutput}
\end{CodeChunk}

%% -- Summary/conclusions/discussion -------------------------------------------

\section{Summary and discussion} \label{sec:summary}

\begin{leftbar}
As usual \dots
\end{leftbar}

%% -- Optional special unnumbered sections -------------------------------------

\section*{Computational details}

\begin{leftbar}
If necessary or useful, information about certain computational details
such as version numbers, operating systems, or compilers could be included
in an unnumbered section. Also, auxiliary packages (say, for visualizations,
maps, tables, \dots) that are not cited in the main text can be credited here.
\end{leftbar}

The results in this paper were obtained using
\proglang{R}~3.4.1 with the
\pkg{MASS}~7.3.47 package. \proglang{R} itself
and all packages used are available from the Comprehensive
\proglang{R} Archive Network (CRAN) at
\url{https://CRAN.R-project.org/}.

\section*{Acknowledgments}

\begin{leftbar}
All acknowledgments (note the AE spelling) should be collected in this
unnumbered section before the references. It may contain the usual information
about funding and feedback from colleagues/reviewers/etc. Furthermore,
information such as relative contributions of the authors may be added here
(if any).
\end{leftbar}

%% -- Bibliography -------------------------------------------------------------
%% - References need to be provided in a .bib BibTeX database.
%% - All references should be made with \cite, \citet, \citep, \citealp etc.
%%   (and never hard-coded). See the FAQ for details.
%% - JSS-specific markup (\proglang, \pkg, \code) should be used in the .bib.
%% - Titles in the .bib should be in title case.
%% - DOIs should be included where available.

\bibliography{refs}

%% -- Appendix (if any) --------------------------------------------------------
%% - After the bibliography with page break.
%% - With proper section titles and _not_ just "Appendix".

\newpage

\begin{appendix}

\section{More technical details} \label{app:technical}

\begin{leftbar}
Appendices can be included after the bibliography (with a page break). Each
section within the appendix should have a proper section title (rather than
just \emph{Appendix}).

For more technical style details, please check out JSS's style FAQ at
\url{https://www.jstatsoft.org/pages/view/style#frequently-asked-questions}
which includes the following topics:
\begin{itemize}
  \item Title vs.\ sentence case.
  \item Graphics formatting.
  \item Naming conventions.
  \item Turning JSS manuscripts into \proglang{R} package vignettes.
  \item Trouble shooting.
  \item Many other potentially helpful details\dots
\end{itemize}
\end{leftbar}

\section[Using BibTeX]{Using \textsc{Bib}{\TeX}} \label{app:bibtex}

\begin{leftbar}
References need to be provided in a \textsc{Bib}{\TeX} file (\code{.bib}). All
references should be made with \verb|\cite|, \verb|\citet|, \verb|\citep|,
\verb|\citealp| etc.\ (and never hard-coded). This commands yield different
formats of author-year citations and allow to include additional details (e.g.,
pages, chapters, \dots) in brackets. In case you are not familiar with these
commands see the JSS style FAQ for details.

Cleaning up \textsc{Bib}{\TeX} files is a somewhat tedious task -- especially
when acquiring the entries automatically from mixed online sources. However,
it is important that informations are complete and presented in a consistent
style to avoid confusions. JSS requires the following format.
\begin{itemize}
  \item JSS-specific markup (\verb|\proglang|, \verb|\pkg|, \verb|\code|) should
    be used in the references.
  \item Titles should be in title case.
  \item Journal titles should not be abbreviated and in title case.
  \item DOIs should be included where available.
  \item Software should be properly cited as well. For \proglang{R} packages
    \code{citation("pkgname")} typically provides a good starting point.
\end{itemize}
\end{leftbar}

\end{appendix}

%% -----------------------------------------------------------------------------

\end{document}

### 2018-02-27 docs/MEE_submission/manuscript_MEE.tex
Updating repo with revised MS from Joel

\documentclass[12pt]{article}

\usepackage{geometry}
\geometry{a4paper} 
\usepackage{fullpage}

\usepackage{amssymb,rotating,natbib,graphicx,fancyvrb}

\usepackage[parfill]{parskip} 
\usepackage[utf8]{inputenc}

%% double spacing
\usepackage{setspace}
 \doublespacing

% line numbers
\usepackage{lineno}
\linenumbers
\renewcommand\linenumberfont{\normalfont\small}

%% new custom commands for code formatting
\newcommand{\code}[1]{\texttt{#1}}
\newcommand{\class}[1]{`\code{#1}'}
%\newcommand{\fct}[1]{#1}
\newcommand{\fct}[1]{\texttt{#1()}}
\newcommand{\pkg}[1]{{\fontseries{b}\selectfont #1}}
\let\proglang=\textsf

% remove page numbers
%\pagenumbering{gobble}

\begin{document}

\DefineVerbatimEnvironment{Code}{Verbatim}{}
\DefineVerbatimEnvironment{CodeInput}{Verbatim}{fontshape=sl}
\DefineVerbatimEnvironment{CodeOutput}{Verbatim}{}
\newenvironment{CodeChunk}{}{}

\raggedright

\textbf{Reproducible, flexible and high throughput data extraction from primary literature: The \pkg{metaDigitise} \proglang{R} package}

Joel L. Pick$^{1,*}$, Shinichi Nakagawa$^1$, Daniel W.A. Noble$^1$

$^1$
  Ecology and Evolution Research Centre, School of Biological, Earth and Environmental Sciences,  University of New South Wales, Kensington, NSW 2052, Sydney, AUSTRALIA

 $^*$Corresponding Author: joel.l.pick@gmail.com\\

%% - \Abstract{} almost as usual
\section*{Abstract}
Research synthesis, especially in the form of meta-analysis, requires data extraction from primary studies. Meta-analysis synthesizes effect sizes, often calculated from summary statistics of studies. However, exact values of such statistics are commonly hidden in figures. The R package \pkg{metaDigitise} extracts descriptive statistics such as means, standard deviations and, if applicable, correlations from the four types of plots: 1) mean/error plots (e.g. bar graphs with standard errors), 2) box plots, 3) scatter plots and 4) histograms. The package interactively guides the user through data extraction process. Notably, it enables a large-scale extraction using image files, letting the user stop processing, edit and add to the resulting data fame at any point. Further, it facilitates reproducible data extraction from plots with little inter-observer bias, thus, allowing a group of people to participate the extraction of data collaboratively.

 Keywords: meta-analysis, comparative analysis, data extraction, \proglang{R}, reproducibility, figures, images, summary statistics

%%%---------------------------------
%%%---------------------------------
%%%---------------------------------

\section{Introduction}

In many different contexts, researchers need to make use of data presented in primary literature. Most notably, this includes meta-analysis, which is becoming increasingly common in many research fields. Meta-analysis uses effect size estimates and their sampling variance, taken from many studies, to understand whether particular effects are common across studies and to explain variation among these effects \citep{Glass1976,Borenstein2009,Koricheva2013,Naka2017}. Meta-analysis therefore relies foremost on data extracted from primary literature, and more specifically, descriptive statistics (e.g. means, standard deviations, sample sizes, correlation coefficients) that have been reported in the text or tables of research papers. Descriptive statistics are also, however, frequently presented in figures and so need to be manually extracted using digitising programs. While inferential statistics (e.g., \textit{t}- and \textit{F}-statistics) are often presented along side descriptive statistics and can be used to derive effect sizes, descriptive statistics are much more appropriate to use because sources of non-independence in experimental designs can be dealt with more easily \citep{Noble2017}. Although there are several existing tools to perform tasks like this (e.g. \proglang{DataThief} \citep{DataThief}, \proglang{GraphClick} \citep{GraphClick}, \proglang{WebPlotDigitizer} \citep{WebPlotDigitizer}), these tools are not appropriate for the needs of meta-analysis for three main reasons.

First, they typically only provide the user with calibrated \textit{x,y} coordinates from imported figures, and do not differentiate between common plot types that are used to present data. This means that a large amount of downstream data manipulation is required, that is different across plots types. For example, data are frequently presented in mean/error plots (Figure \ref{fig:all_extract}A), for which the user wants a mean and error estimate for each group presented in the figure. With existing programs, \textit{x,y} coordinates of means and errors are returned, to which the user must manually discern between mean and error coordinates and assign points to groups. The error then needs to be calculated as the deviation from the mean, and then transformed to a standard deviation, depending on the type of error presented. Second, digitising programs do not easily allow the integration of metadata at the time of data extraction, such as experimental group or variable names, and sample sizes. This makes the downstream calculations more laborious, as the information has to be added later, in most cases using different software. Finally, existing programs do not import a set of images and allow the user to systematically work through them. Instead they require the user to manually import images one by one, and export data into individual files, that need to be imported and edited using different software. In essence, existing software does not provide an optimized research pipeline to facilitate data extraction, editing and reproducibility.

These present major issues because extracting from figures can be an incredibly time-consuming process. Furthermore, although meta-analysis is an important tool in consolidating the data from multiple studies, many of the processes involved in data extraction are opaque and difficult to reproduce, making extending studies problematic. Having a tool that facilitates reproducibility in meta-analyses will increase transparency and aid in resolving the reproducibility crises seen in many fields \citep{peng_reproducible_2006, peng_reproducible_2011, sandve_ten_2013, Parker2016,Ihle2017}.

%%% This paragraph could be shorter as it is all said in the main text
Here, we present an interactive \proglang{R} package, \pkg{metaDigitise} (available at xxx), which is designed for large scale, reproducible data extraction from figures, specifically catering to the the needs of meta-analysts. To this end, we provide tools specific to data extraction from common plot types (mean/error plots, box plots, scatter plots and histograms, see Figure \ref{fig:all_extract}). \pkg{metaDigitise} operates within the \proglang{R} environment making data extraction, analysis and export more streamlined. The necessary calculations are carried out on processed data immediately after extraction so that comparable summary statistics can be obtained quickly. Summary data from multiple figures is returned into a single data frame which can be can easily exported or use in downstream analysis within \proglang{R}. Calibrated data is automatically saved for each digitised figure. Users can therefore redraw their digitisations on figures, make corrections and access calibration and proceeded data. This makes sharing figure digitisation and reproducing the work of others simple and easy, and allows meta-analysts to update meta-analyses more efficiently.

%% 724 words

%%%---------------------------------
%%%---------------------------------
%%%---------------------------------

\section{Directory Structure and Reproducibility}
The \pkg{metaDigitise} package was created with the idea that users would have multiple images to extract from and therefore operates in the same way whether the user has one or multiple images. There is one main function, \fct{metaDigitise}, which interactively takes the user through the process of extracting data from figures. \fct{metaDigitise} works on a directory containing images of figures copied from primary literature, in .png, .jpg, .tiff, .pdf format, specified to \fct{metaDigitise} through the \code{dir} argument. The user needs to think carefully about their directory structure early on in their project, especially if they plan to share the extractions with collaborators or publish the project. Although different directory structures may be used, we would recommend having all files for one project in a single directory with an informative and unambiguous naming scheme for images to help identify the paper and figure that data come from, for example:

\begin{CodeChunk}
\begin{CodeOutput}
* Main project directory
    + FiguresToExtract
        + Paper1_Figure1_trait1.png
        + Paper1_Figure2_trait2.png
        + Paper2_Figure1_trait3.png
\end{CodeOutput}
\end{CodeChunk}

When \fct{metaDigitise} is run, it recognizes all the images in a directory and automatically imports them one by one, allowing the user to extract the relevant information about a figure as they go. Having all figures in one directory therefore expedites digitising figures by preventing users from having to constantly change directories and / or open new images. The data from each completed image is automatically saved as a \code{metaDigitise} object in a separate .RDS file to a \code{caldat} directory that is created within the parent directory when first executing \fct{metaDigitise}. These files enable re-plotting and editing of images at a later point (see below).

\fct{metaDigitise} also identifies images within a directory that have been previously digitised and only import images that have not been digitised in previous calls of the function. All figures do not, therefore, need to be extracted at one time and new figures can be added to the directory as the project develops. After each image is extracted, the user is asked whether they wish to continue or quit the extraction process. Upon rerunning \fct{metaDigitise}, previously digitised figures are ignored during processing, but their data is automatically integrated into the final output.

This directory structure allows the complete digitisation process to be reproduced at a later stage, shared with collaborators and presented as supplementary materials for a publication. As long as all the images and the caldat directory are still in the directory, \fct{metaDigitise} will be able to reproduce all figure extractions, regardless of the computer it is run on. For an analysis to be updated, new figures can simply be added to the directory and \fct{metaDigitise} run to incorporate the new data.

%% 460 words
%%%---------------------------------
%%%---------------------------------
%%%---------------------------------

\section{Image Processing}
Running \fct{metaDigitise} presents the user with three options; `Process new images', `Import existing data' or `Edit existing data'. Selecting `Process New Images' starts the digitisation process on images within the directory that have not preciously been digitised; the other functions are discussed below.

For all plot types, \fct{metaDigitise} requires the user to calibrate the axes in the figure, by clicking on two known points on the axis in question, and entering the value of those points (Figure \ref{fig:all_extract}). \fct{metaDigitise} then calculates the value of any clicked points in terms of the figure axes. For mean/error and box plots, only the y-axis is calibrated (Figure \ref{fig:all_extract}A,B), assuming the x-axis is redundant. For scatter plots and histograms both axes are calibrated (Figure \ref{fig:all_extract}C,D).

As figures may have been copied from older, scanned publications, they may not be perfectly orientated. This makes calibration of the points in the figure problematic. \fct{metaDigitise} allows users to rotate the image (Figure \ref{fig:rotate}A,B). Furthermore, mean/error plots, box plots and histograms, may be presented with horizontal bars. \fct{metaDigitise} assumes that bars are vertical, but allows the user to flip the image to make the bars are vertical (Figure \ref{fig:rotate}C,D).

\pkg{metaDigitise} recognises four main types of plot; Mean/error plots, box plots, scatter plots and histograms, shown in Figure \ref{fig:all_extract}. All plot types can be extracted in a single call of \fct{metaDigitise} and integrated into one output. Alternatively, users can process different plot types separately, using separate directories. All four plot type are extracted slightly differently (outlined below). After completing all images, or upon quitting, either summarised or processed data is returned (specified by the user). From all plot types, summarised data consists of a mean, standard deviation and sample size, for each identified group within the plot (should multiple groups exist). In the case of scatter plots, the correlation coefficient between the points within each identified group is also returned. Alternatively choosing processed data will return a list with a slots for each of the four figure types, containing the calibrated points that the user has clicked. This may be particularly useful in the case of scatter plots.

\subsection{Mean/Error and Box Plots} 
\fct{metaDigitise} handles mean/error and box plots in a very similar way. For each mean/box, the user is enters group names and sample sizes. If the user does not enter a sample size at the time of data extraction (if, for example, the information is not readily available) a standard deviation (SD) is not calculated. This can, however, be entered at a later time (see below). For mean/error plots, the user clicks on an error bar and the mean. Error bars above or below the mean can be clicked, as sometimes one is clearer than the other. \fct{metaDigitise} assumes that the error bars are symmetrical. This is deliberate as it is not clear how best to derive SD from asymmetrical error bars, not least as they represent different things in different figures. Points are displayed where the user has clicked, with the error in a different colour to the mean (Figure \ref{fig:all_extract}A). The user also enters the type of error used in the figure: standard deviation (SD), standard error (SE) or 95\% confidence intervals (CI95). For box plots, the user clicks on the maximum, upper quartile, median, lower quartile and minimum. \fct{metaDigitise} will return a warning if the maximum is not greater than the minimum. For both plot types, the user can add, edit or remove groups. Three functions, \fct{error\_to\_sd}, \fct{rqm\_to\_mean} and \fct{rqm\_to\_sd}, that convert different error types to SD, box plot data to mean and box plot data SD, respectively, are also available in the package (see SM for further details of these conversions).

\subsection{Scatter plots}
Users can extract points from multiple groups from scatter plots. Points added by mistake can be deleted. The user can add more groups, edit groups (add or remove points) or delete groups. Different groups are plotted in different colours and shapes to enable them to be distinguished, with a legend at the bottom of the figure (Figure \ref{fig:all_extract}C). Mean, SD and sample size are calculated from the clicked points, for each group. Often data points will overlap with each other making it impossible to know whether points have been missed. However, a user may realise that the sample size from the digitisation conflicts with what is reported in the paper. For example, in Figure \ref{fig:all_extract}C only 49 points have been clicked when the sample size is known to be 50. Hence, \pkg{metaDigitise} also provides the user with the option to input a known sample sizes directly. Nonetheless, it is important to recognise the impact that overlapping points can have on summary statistics, and in particular on sampling variance.

\subsection{Histograms}
The user clicks on the top corners of each bar. A line is drawn, in alternate colours, at the top of these bars (Figure \ref{fig:all_extract}D). These are numbered to allow the the user to edit them. As with scatter plots, if the sample size from the extracted data does not match a known sample size, the user can enter an alternate sample size. The calculation of mean and SD from this data is shown in the SM.

%% 1000 words

%%%---------------------------------
%%%---------------------------------
%%%---------------------------------

\section{Importing and Editing Previously Digitised data}
\pkg{metaDigitise} is also able to re-import, edit and re-plot previously digitised figures. When running \fct{metaDigitise}, the user can choose to `Import existing data', which returns previously digitised data, for single or all images. Users can also choose to `Edit existing data' which allows them to re-plot or edit information or digitisations that have previously be done.

\subsection{Adding Sample Sizes to Previous Digitisations}
In many cases important information, such as sample sizes, may not be readily available when digitising figures. Such information does not, therefore, have to be added a the time of digitisation. To expedite finding and adding these sample sizes at a later point, \fct{metaDigitise} has a specific edit option that allows users to enter previously omitted sample sizes. This first identifies missing sample sizes in the digitised output, re-plots the relevant figures and prompts the user to enter the sample sizes for the relevant groups in the figure, one by one.

%% 150 words

%%%---------------------------------
%%%---------------------------------
%%%---------------------------------

\section{Software Validation}

\subsection{Inter-observer variability in digitisations}

In order to evaluate the consistency of digitisation using \pkg{metaDigitise} between users, we simulated a dataset of two variables with two groups ($n=10$ within groups). Each variable was plotted twice for each plot type (figures were modified slightly to give users a sense that they were digitising new data) generating a total of 14 figures. 14 independent digitisers (including the authors) were provided with a directory with all 14 figures in a randomised order. Digitisers ran \pkg{metaDigistise} on their own computers, across different operating systems (including Mac, Windows and Linux). Digitisers varied in their level of experience, from people with experience of meta-analyses or comparative work to those without any science background. We asked users to digitise all 14 figures and collected the mean, standard deviation and correlation coefficient ($r$, for scatter plots) generated by \fct{metaDigitise} for every plot digitised ($n=28$ per digitiser per metric, $n=4$ for $r$).

As a measure of bias, we calculated the percentage differences from the true summary statistics as

\begin{equation}
\frac{\theta - \hat{\theta}}{\hat{\theta}}
\end{equation}

where $\theta$ is the estimate and $\hat{\theta}$ is the true value. The deviation from the true value of $r$ was not further standardised, as it is already on a standardised scale. We also took the absolute values of these standardised differences as a measure of precision. The resulting data was used to assess between- and within- user variability (i.e., the intra-class correlation coefficient - ICC). This was done using linear mixed effect models with user identify as a random effect using \pkg{lme4} \citep{bates2015} in \proglang{R}. Standardised mean, standard deviation and correlation coefficients were used as response variables in separate models. Sampling variance for ICC estimates was generated based on 1000 parametric bootstraps of the model and the significance was tested using likelihood ratio tests, using \pkg{rptR} \citep{stoffel2017}.

If digitisations were consistent across users then we should find no significant between user variability in the data. Indeed, across plot types we found no evidence for any inter-observer variability in digitisations for the mean (ICC = 0, 95\% CI = 0 to 0.029, \textit{p} = 1), standard deviation (ICC = 0, 95\% CI = 0 to 0.033, \textit{p} = 0.5) or correlation coefficient (ICC = 0.053, 95\% CI = 0 to 0.296, \textit{p} = 0.377). There were was little bias between digitised and true values, on average 1.63\% (mean = 0.02\%, SD = 4.9\%, \textit{r} = -0.03\%) and overall there were only small absolute differences between digitised and true values, deviating, on average 2.18\% (mean = 0.40\%, SD = 5.81\%, \textit{r} = 0.33\%) across all three summary statistics. SD estimates from digitisations are clearly more prone to error than means or correlation coefficients. If the mean absolute difference is calculated for each plot type, we can see that this effect is driven mainly by extraction from box plots and histograms (\% difference; box plot: 15.805, histogram: 5.210, mean/error: 1.500, scatter plot: 0.433). SD estimation from box plot summary statistics is known to be more error prone, especially at small sample sizes \citep{Wan2014}.

\subsection{Testing the accuracy of digitisations}
To test how accurate \pkg{metaDigitise} is at matching clicked points to their true values, we generated four random scatterplots, each with 20 data points, and digitised these with \fct{metaDigitise}. This was done by one digitiser (J.L.P.), as there is no detectable between user variation. Data digitised using \pkg{metaDigitise} was essentially perfectly correlated with the true simulated data for both the \textit{x}-variable (Pearson's correlation; \textit{r} = 0.9999915, \textit{t} = 2137.4, \textit{df} = 78, $p < 0.001$) and \textit{y}-variable (\textit{r} = 0.9999892, \textit{t} = 1897.8, \textit{df} = 78, $p < 0.001$).

%% 552 words

%%%---------------------------------
%%%---------------------------------
%%%---------------------------------

\section{Limitations and Future Extensions}

Although \pkg{metaDigitise} is already very flexible, and provides functionality not seen in any other package (Table S1), it is clear that there are some functions that it does not perform. Notably \pkg{metaDigitise} lacks is automated point detection, available in several packages (Table S1). However, from our experience, manual digitising is more reliable and often equally as fast. Given the variation in image quality, calibration for automatic point detection needs to be done for each plot individually. Additionally, auto-detection often misses points which then need to be manually added. Based on tests of \pkg{metaDigitise} (see above), figures can be extracted in around 1-2 minutes, including the entry of metadata. As a result, we do not believe that current automated point detection techniques provides substantial benefits in terms of time or accuracy.

\pkg{metaDigitise} also (currently) lacks the ability to zoom in on plots. Zooming may enable users to gain greater accuracy when clicking on points. However, from our own experience (and indeed from the results above), if you are using a reasonably sized screen then the accuracy is already high, and so there is not much gain to be had from zooming in on points.

In contrast to some other packages, \pkg{metaDigitise} does not extract lines from figures. In our own experience, line extraction is not particularly useful for meta-analysis, although we recognise that it may be useful in other fields. Should a user like to extract lines with \pkg{metaDigitise}, we would recommend extracting data as a scatter plot, and clicking along the line in question. A model can then be fitted to these points (accessed by choosing to return processed rather than summary data) to estimate the parameters needed.

Descriptive statistics are usually the most robust sources of information for calculating effect size statistics \citep{Noble2017}. These are most often presented in figures. Users may therefore also want to compare effect size estimates from inferential statistics with those derived from descriptive statistics (obtained for example using \pkg{metaDigitise}) from a paper. Comparing these different effects sizes can be useful in identifying uncertainties and problems within a paper. In the future, we hope to provide functions to easily convert inferential statistics to standardised effect size estimates, which can seamlessly be integrated with summary statistics from \pkg{metaDigitise}, to calculate equivalent standardised effect size estimates and sampling variance.

%% 416 words

%%%---------------------------------
%%%---------------------------------
%%%---------------------------------

\section{Conclusions}
Increasing the reproducibility of figure extraction for meta-analysis and making this laborious process more streamlined, flexible and integrated with existing statistical software will go a long way in facilitating the production of high quality meta-analytic studies that can be updated in the future. We believe that \pkg{metaDigitise} will improve this research synthesis pipeline, and will hopefully become an integral package that can be added to the meta-analysts toolkit.

\section*{Acknowledgments}
We thank the I-DEEL group and colleagues at UNSW for for testing, providing feedback and digitising including: Rose O'Dea, Fonti Kar, Malgorzata Lagisz, Julia Riley, Diego Barneche, Erin Macartney, Ivan Beltran, Gihan Samarasinghe, Dax Kellie, Jonathan Noble, Yian Noble and Alison Pick. J.L.P. was supported by a Swiss National Science Foundation Early Mobility grant (P2ZHP3\_164962), D.W.A.N. was supported by an Australian Research Council Discovery Early Career Research Award (DE150101774) and UNSW Vice Chancellors Fellowship and S.N. an Australian Research Council Future Fellowship (FT130100268).

\section*{Author Contributions}
J.L.P. and D.W.A.N. conceived the study and J.L.P., S.N. and D.W.A.N. developed the idea. J.L.P. and D.W.A.N. developed the R-package. J.L.P. and D.W.A.N. wrote the first draft of the paper and J.L.P., S.N. and D.W.A.N. contributed substantially to subsequent revisions of the manuscript and gave final approval for publication.

\bibliographystyle{FuncEcol.bst}
\bibliography{metaDigitise}

\section*{Figures}

\begin{figure}[!h]
\centering 
 \includegraphics[width=0.75\textwidth]{fig_all_extract.pdf} 
 \caption{Four plot types that \pkg{metaDigitise} is designed to extract data from: A) mean/error plot, B) box plot, C) scatter plot and D) histogram. Data is taken from the iris dataset in R. A and B are plotted with the whole dataset, C and D are just the data for the species \textit{setosa}. Digitisation of the images is shown. All figures are clearly labelled at the top to remind users of the filename and plot type. This reduces errors throughout the digitisation process. Names of the variables and calibration (in blue) are plotted alongside the digitised points. In A) and B), user entered group names and sample sizes are displayed beside the relevant points. In C) the names and sample sizes for each group are shown below the figure.}
\label{fig:all_extract}
\end{figure}

\begin{figure}[!b] 
\centering
 \includegraphics[width=0.75\textwidth]{fig_rotate.pdf} 
 \caption{Figure rotation. A) and B) show how non-aligned images can be realigned through user defined rotation. C) and D) show how figures can be re-orientated so as to aid data input.}
\label{fig:rotate}
\end{figure}

\end{document}

### 2018-02-27 docs/MEE_submission/supplements.tex
Updating repo with revised MS from Joel

\documentclass[12pt]{article}

\usepackage{geometry}
\geometry{a4paper} 
\usepackage{fullpage}

\usepackage{amssymb,rotating,natbib,graphicx,fancyvrb}

\usepackage[parfill]{parskip} 
\usepackage[utf8]{inputenc}

%% double spacing
\usepackage{setspace}
 \onehalfspacing

% line numbers
% \usepackage{lineno}
% \linenumbers
% \renewcommand\linenumberfont{\normalfont\small}

%% new custom commands for code formatting
\newcommand{\code}[1]{\texttt{#1}}
\newcommand{\class}[1]{`\code{#1}'}
%\newcommand{\fct}[1]{#1}
\newcommand{\fct}[1]{\texttt{#1()}}
\newcommand{\pkg}[1]{{\fontseries{b}\selectfont #1}}
\let\proglang=\textsf

% remove page numbers
%\pagenumbering{gobble}

\title{Reproducible, flexible and high throughput data extraction from primary literature: The \pkg{metaDigitise} \proglang{R} package \vskip15pt Supplementary Materials}

\author{Joel L. Pick, Shinichi Nakagawa \& Daniel W.A. Noble}

\begin{document}

\DefineVerbatimEnvironment{Code}{Verbatim}{}
\DefineVerbatimEnvironment{CodeInput}{Verbatim}{fontshape=sl}
\DefineVerbatimEnvironment{CodeOutput}{Verbatim}{}
\newenvironment{CodeChunk}{}{}

\setcounter{equation}{0}
\setcounter{subsection}{0}

\renewcommand{\thetable}{S\arabic{table}}%
\renewcommand{\theequation}{S\arabic{equation}}%
\renewcommand{\thesection}{S\arabic{section}}%
\renewcommand{\thesubsection}{S\arabic{section}.\arabic{subsection}}%

\maketitle

\clearpage

\section{Comparison with existing software}
\begin{sidewaystable}[ph!]
{\centering
\begin{tabular}{lccccccc}
\hline
Function                  & metaDigitise & GraphClick$^1$ & DataThief$^2$  & DigitizeIt$^3$ & WebPlotDigitizer$^4$ & metagear$^5$ & digitize$^6$\\
\hline
Scatterplots              & \checkmark   & \checkmark & \checkmark & \checkmark & \checkmark     & \checkmark$^7$ & \checkmark \\
Mean/error plots      & \checkmark   & \checkmark & \checkmark & $\times$   & $\times$       & \checkmark$^7$ & $\times$    \\
Boxplots                  & \checkmark   & $\times$   & $\times$   & $\times$   & $\times$       & $\times$       & $\times$    \\
Histograms                & \checkmark   & $\times$   & $\times$   & $\times$   & \checkmark$^7$ & $\times$       & $\times$    \\
Graph rotation$^8$        & \checkmark   & \checkmark & \checkmark & \checkmark & \checkmark     & $\times$       & $\times$    \\
Groups                    & \checkmark   & \checkmark & $\times$   & \checkmark & \checkmark     & $\times$       & $\times$    \\
Entry of metadata         & \checkmark   & $\times$   & $\times$   & $\times$   & $\times$       & $\times$       & $\times$    \\
Summarising data          & \checkmark   & $\times$   & $\times$   & $\times$   & $\times$       & $\times$       & $\times$    \\
Multiple image processing & \checkmark   & $\times$   & $\times$   & $\times$   & $\times$       & $\times$       & $\times$    \\
Reproducable$^9$  & \checkmark   & \checkmark & \checkmark & $\times$   & \checkmark     & $\times$       & $\times$    \\
Automated point detection & $\times$     & \checkmark &     $\times$      & \checkmark & \checkmark     & \checkmark     & $\times$    \\
Line extraction           & $\times$     & \checkmark & \checkmark & \checkmark & \checkmark     & $\times$       & $\times$    \\
Zoom                      & $\times$     & \checkmark & \checkmark & \checkmark & \checkmark     & $\times$       & $\times$    \\
Log axis                  & $\times$   & \checkmark & \checkmark & \checkmark & \checkmark     & $\times$       & $\times$    \\
Dates                     & $\times$   & $\times$   & \checkmark & $\times$   & \checkmark     & $\times$       & $\times$    \\
Asymmetric error bars     & $\times$     & $\times$   & \checkmark & $\times$   & $\times$       & $\times$       & $\times$    \\
Freeware                  & \checkmark$^{10}$ & \checkmark$^{11}$ & \checkmark$^{11}$ & $\times$$^{11}$ & \checkmark$^{11}$ & \checkmark$^{10}$ & \checkmark$^{10}$\\
\hline

%http://datathief.org/DatathiefManual.pdf
%http://www.digitizeit.de/#Features
%https://automeris.io/WebPlotDigitizer/userManual.pdf

\end{tabular}
}
{\footnotesize
\\
$^1$ \citet{GraphClick}
$^2$ \citet{DataThief}
$^3$ \citet{DigitizeIt}
$^4$ \citet{WebPlotDigitizer}
$^5$ \citet{Lajeunesse2016}
$^6$ \citet{Poisot2011}
\\$^7$ Only automated, no manual extraction.
\\$^8$ Or handles rotated graphs. 
\\$^9$ Allows saving, re-plotting and editing of data extraction.
\\$^{10}$ \proglang{R} package.
\\$^{11}$ Standalone software.
}

\caption{\label{tab:comparison} Comparison of functionality between different digitisation softwares.}

\end{sidewaystable}

\clearpage

\section{Derivation of mean, standard deviation and sample size from different plot types}

\subsection{Mean/Error Plots}
The standard deviation is calculated depending on the type of error presented. The user can choose from standard deviation (SD, $\sigma$), standard error (SE) or 95\% confidence intervals (CI95). Standard deviation is calculated from standard error as
\begin{equation}
\sigma = SE \sqrt{n}
\end{equation}
and from 95\% confidence intervals as
\begin{equation}
\sigma = \frac{CI}{1.96} \sqrt{n}
\end{equation}

\subsection{Box Plots} 
The mean ($\mu$) and SD are calculated sing the maximum ($b$), upper quartile ($q_3$), median ($m$), lower quartile ($q_1$) and minimum ($a$) as 
\begin{equation}
\mu = \frac{(n+3)(a+b) + 2(n-1)(q_1 + m + q_3)}{8n}
\end{equation}
following \citet{Bland2015} and
\begin{equation}
\sigma = \frac{b-a}{4\Phi^{-1}(\frac{n-0.375}{n+0.25})} + \frac{q_3-q_1}{4\Phi^{-1}(\frac{0.75n-0.125}{n+0.25})}
\end{equation}
where $\Phi^{-1}(z)$ is the upper zth percentile of the standard normal distribution, following \citet{Wan2014}.

\subsection{Histograms} 
For each bar, the user click two point (the top of the bar). Using these points, a midpoint (m; mean x coordinates) and a frequency (f; mean y coordinates, rounded to the nearest integer) is calculated for each bar. The sample size, mean and SD are calculated as:
\begin{equation}
n = \sum_{i=1}^n{f_i}
\end{equation}

\begin{equation}
\mu = \frac{\sum_{i=1}^n{m_i f_i}}{n}
\end{equation}

\begin{equation}
\sigma = \sqrt{\frac{\sum_{i=1}^n{(m_i f_i - \mu f_i)^2}}{n-1}}
\end{equation}

\section{Tutorial}

\bibliographystyle{FuncEcol.bst}
\bibliography{metaDigitise}

\end{document}

### 2018-02-27 docs/MEE_submission/figures.R

@@ -1,0 +1,150 @@
+
+data(iris)
+head(iris)
+data(airquality)
+head(airquality)
+
+#plot(Temp~Wind,airquality, subset = Month %in% c(5,6), pch=c(19,1)[as.factor(airquality$Month)])
+
+cex.mtext = 2
+setosa <- subset(iris, Species == "setosa")
+
+setwd("/Users/joelpick/Dropbox/0_postdoc/10_metaDigitise")
+
+library(metaDigitise)
+
+## first plot
+
+setEPS()
+pdf("./docs/fig_plot_type.pdf", height=9, width=9)
+
+par(mfrow=c(2,2), mar=c(6,5,2,1), cex.lab=1.5)
+
+means <- with(iris,tapply(Sepal.Length, Species, mean))
+ses <- with(iris,tapply(Sepal.Length, Species, function(x) sd(x)/sqrt(length(x))))*1.96
+
+plot(means, ylim = c(min(means-ses)-0.1,max(means+ses)+0.1), xlim=c(0.5,3.5), xaxt="n", pch=19, cex=2, ylab="Sepal Length", xlab="Species")
+mtext("A", 3, adj=0, line=0, cex=cex.mtext)
+arrows(1:length(means),means+ses, 1:length(means), means-ses, code=3, angle=90, length=0.1)
+axis(1,1:length(means),names(means))
+
+boxplot(Sepal.Length ~ Species, iris, ylab="Sepal Length", xlab="Species", col="grey", range=0)
+mtext("B", 3, adj=0, line=0, cex=cex.mtext)
+
+
+set.seed(5)
+plot(Sepal.Length~jitter(Petal.Length),setosa, pch=19, ylab="Sepal Length", xlab="Petal Length")
+mtext("C", 3, adj=0, line=0, cex=cex.mtext)
+
+hist(setosa$Sepal.Length,, main="", col="grey", xlab="Sepal Length")
+mtext("D", 3, adj=0, line=0, cex=cex.mtext)
+
+dev.off()
+
+
+
+
+### make pngs
+png("./docs/images/iris_mean_error.png")
+par(mar=c(6,5,1,1), cex.lab=1.5)
+plot(means, ylim = c(min(means-ses)-0.1,max(means+ses)+0.1), xlim=c(0.5,3.5), xaxt="n", pch=19, cex=2, ylab="Sepal Length", xlab="Species")
[…]

### 2018-02-27 docs/MEE_submission/manuscript_MEE.tex
small MS change

@@
 The user needs to think carefully about their directory structure early on in their project, especially if they plan to share the extractions with collaborators or publish the project.
-Although different directory structures may be used, we would recommend having all files for one project in a single directory with an informative and unambiguous naming scheme for images to help identify the paper and figure that data come from, for example:
-
-\begin{CodeChunk}
-\begin{CodeOutput}
-* Main project directory
-    + FiguresToExtract
-        + Paper1_Figure1_trait1.png
-        + Paper1_Figure2_trait2.png
-        + Paper2_Figure1_trait3.png
-\end{CodeOutput}
-\end{CodeChunk}
+Although different directory structures may be used, we would recommend having all files for one project in a single directory with an informative and unambiguous naming scheme for images to help identify the paper and figure that data come from (e.g.
+paper_figure_trait.png).
 

### 2018-02-27 docs/MEE_submission/manuscript_MEE.tex
Streamlining MEE MS, down to 3200 words

@@
 \section*{Abstract}
-Research synthesis, especially in the form of meta-analysis, requires data extraction from primary studies.
-Meta-analysis synthesizes effect sizes, often calculated from summary statistics of studies.
+1.
+Research synthesis requires data extraction from primary studies with effect sizes for meta-analyses being calculated from summary statistics.
 However, exact values of such statistics are commonly hidden in figures.
-The R package \pkg{metaDigitise} extracts descriptive statistics such as means, standard deviations and, if applicable, correlations from the four types of plots: 1) mean/error plots (e.g.
+2.
+Extracting summary statistics from figures can be a slow process that is not easily reproducible.
+Additionally, current software lacks an ability to incorporate important meta-data (e.g., sample sizes, treatment names) about experiments.
+Furthermore, extraction software is often independent of analytical software preventing the automatic calculation of effect size statistics and making it more challenging to implement user defined analytical pipelines.
+3.
+Here we present the R package \pkg{metaDigitise} which extracts descriptive statistics such as means, standard deviations and correlations from the four plot types: 1) mean/error plots (e.g.
 bar graphs with standard errors), 2) box plots, 3) scatter plots and 4) histograms.
-The package interactively guides the user through data extraction process.
-Notably, it enables a large-scale extraction using image files, letting the user stop processing, edit and add to the resulting data fame at any point.
-Further, it facilitates reproducible data extraction from plots with little inter-observer bias, thus, allowing a group of people to participate the extraction of data collaboratively.
-
+\pkg{metaDigitise} is user-friendly and easy to learn as it interactively guides the user through the data extraction process.
+Notably, it enables large-scale extraction by automatically loading image files, letting the user stop processing, edit and add to the resulting data fame at any point.
+4.
+Digitised data can be easily re-plotted and checked, facilitating reproducible data extraction from plots with little inter-observer bias.
+Such approaches allow collaborative extraction of data.
+We hope that by making the process of figure extraction more flexible and easy to conduct it will improve transparency and quality of meta-analyses in the future.
+%% 230 words
  Keywords: meta-analysis, comparative analysis, data extraction, \proglang{R}, reproducibility, figures, images, summary statistics
@@
 Meta-analysis uses effect size estimates and their sampling variance, taken from many studies, to understand whether particular effects are common across studies and to explain variation among these effects \citep{Glass1976,Borenstein2009,Koricheva2013,Naka2017}.
-Meta-analysis therefore relies foremost on data extracted from primary literature, and more specifically, descriptive statistics (e.g.
-means, standard deviations, sample sizes, correlation coefficients) that have been reported in the text or tables of research papers.
+Meta-analysis relies on descriptive statistics (e.g.
+means, standard deviations, sample sizes, correlation coefficients) extracted from primary literature that have been reported in the text or tables of research papers.
 Descriptive statistics are also, however, frequently presented in figures and so need to be manually extracted using digitising programs.
-While inferential statistics (e.g., \textit{t}- and \textit{F}-statistics) are often presented along side descriptive statistics and can be used to derive effect sizes, descriptive statistics are much more appropriate to use because sources of non-independence in experimental designs can be dealt with more easily \citep{Noble2017}.
+While inferential statistics (e.g., \textit{t}- and \textit{F}-statistics) are often presented along side descriptive statistics, and can be used to derive effect sizes, descriptive statistics are much more appropriate to use because sources of non-independence in experimental designs can be dealt with more easily \citep{Noble2017}.
+
 Although there are several existing tools to perform tasks like this (e.g.
 \proglang{DataThief} \citep{DataThief}, \proglang{GraphClick} \citep{GraphClick}, \proglang{WebPlotDigitizer} \citep{WebPlotDigitizer}), these tools are not appropriate for the needs of meta-analysis for three main reasons.
-
 First, they typically only provide the user with calibrated \textit{x,y} coordinates from imported figures, and do not differentiate between common plot types that are used to present data.
@@
 For example, data are frequently presented in mean/error plots (Figure \ref{fig:all_extract}A), for which the user wants a mean and error estimate for each group presented in the figure.
-With existing programs, \textit{x,y} coordinates of means and errors are returned, to which the user must manually discern between mean and error coordinates and assign points to groups.
-The error then needs to be calculated as the deviation from the mean, and then transformed to a standard deviation, depending on the type of error presented.
+With existing programs, from \textit{x,y} coordinates users must manually discern between mean and error coordinates and assign points to groups.
+Error then needs to be calculated as the deviation from the mean, and then transformed to a standard deviation, depending on the type of error presented.
 Second, digitising programs do not easily allow the integration of metadata at the time of data extraction, such as experimental group or variable names, and sample sizes.
 This makes the downstream calculations more laborious, as the information has to be added later, in most cases using different software.
-Finally, existing programs do not import a set of images and allow the user to systematically work through them.
+Finally, existing programs do not automatically import new images upon completion to allow the user to systematically work through them.
 Instead they require the user to manually import images one by one, and export data into individual files, that need to be imported and edited using different software.
-In essence, existing software does not provide an optimized research pipeline to facilitate data extraction, editing and reproducibility.
 
-These present major issues because extracting from figures can be an incredibly time-consuming process.
+These present major issues because extracting from figures is an incredibly time-consuming process and existing software does not provide an optimized research pipeline to facilitate data extraction and editing.
 Furthermore, although meta-analysis is an important tool in consolidating the data from multiple studies, many of the processes involved in data extraction are opaque and difficult to reproduce, making extending studies problematic.
@@
 %%% This paragraph could be shorter as it is all said in the main text
-Here, we present an interactive \proglang{R} package, \pkg{metaDigitise} (available at xxx), which is designed for large scale, reproducible data extraction from figures, specifically catering to the the needs of meta-analysts.
+Here, we present an interactive \proglang{R} package, \pkg{metaDigitise} (available at https://github.com/daniel1noble/metaDigitise), which is designed for large scale, reproducible data extraction from figures, specifically catering to the the needs of meta-analysts.
 To this end, we provide tools specific to data extraction from common plot types (mean/error plots, box plots, scatter plots and histograms, see Figure \ref{fig:all_extract}).
@@
 Summary data from multiple figures is returned into a single data frame which can be can easily exported or use in downstream analysis within \proglang{R}.
-Calibrated data is automatically saved for each digitised figure.
+Completed digitisations are automatically saved for each figure.
 Users can therefore redraw their digitisations on figures, make corrections and access calibration and proceeded data.
-This makes sharing figure digitisation and reproducing the work of others simple and easy, and allows meta-analysts to update meta-analyses more efficiently.
+This makes sharing figure digitisation and reproducing the work of others simple and easy, and allows meta-analyses to be updated more efficiently.
 
-%% 724 words
+%% 647 words
 
@@
 \fct{metaDigitise} works on a directory containing images of figures copied from primary literature, in .png, .jpg, .tiff, .pdf format, specified to \fct{metaDigitise} through the \code{dir} argument.
-The user needs to think carefully about their directory structure early on in their project, especially if they plan to share the extractions with collaborators or publish the project.
+The user needs to think carefully about their directory structure early on in their project.
 Although different directory structures may be used, we would recommend having all files for one project in a single directory with an informative and unambiguous naming scheme for images to help identify the paper and figure that data come from (e.g.
@@
 
-\fct{metaDigitise} also identifies images within a directory that have been previously digitised and only import images that have not been digitised in previous calls of the function.
+\fct{metaDigitise} identifies images within a directory that have been previously digitised and only import images that have not been digitised in previous calls of the function.
 All figures do not, therefore, need to be extracted at one time and new figures can be added to the directory as the project develops.
@@
 
-%% 460 words
+%% 414 words
 %%%---------------------------------
@@
 Alternatively, users can process different plot types separately, using separate directories.
-All four plot type are extracted slightly differently (outlined below).
-After completing all images, or upon quitting, either summarised or processed data is returned (specified by the user).
-From all plot types, summarised data consists of a mean, standard deviation and sample size, for each identified group within the plot (should multiple groups exist).
+All four plot types are extracted slightly differently (outlined below).
+After completing images, or upon quitting, either summarised or processed data is returned (specified by the user).
+Summarised data consists of a mean, standard deviation and sample size, for each identified group within the plot (should multiple groups exist).
 In the case of scatter plots, the correlation coefficient between the points within each identified group is also returned.
-Alternatively choosing processed data will return a list with a slots for each of the four figure types, containing the calibrated points that the user has clicked.
+Alternatively choosing processed data will return a list with slots for each of the four figure types, containing the calibrated points that the user has clicked.
 This may be particularly useful in the case of scatter plots.
@@
 \fct{metaDigitise} assumes that the error bars are symmetrical.
-This is deliberate as it is not clear how best to derive SD from asymmetrical error bars, not least as they represent different things in different figures.
 Points are displayed where the user has clicked, with the error in a different colour to the mean (Figure \ref{fig:all_extract}A).
@@
 For box plots, the user clicks on the maximum, upper quartile, median, lower quartile and minimum.
-\fct{metaDigitise} will return a warning if the maximum is not greater than the minimum.
 For both plot types, the user can add, edit or remove groups.
@@
 Users can extract points from multiple groups from scatter plots.
-Points added by mistake can be deleted.
-The user can add more groups, edit groups (add or remove points) or delete groups.
 Different groups are plotted in different colours and shapes to enable them to be distinguished, with a legend at the bottom of the figure (Figure \ref{fig:all_extract}C).
 Mean, SD and sample size are calculated from the clicked points, for each group.
-Often data points will overlap with each other making it impossible to know whether points have been missed.
-However, a user may realise that the sample size from the digitisation conflicts with what is reported in the paper.
+Data points may overlap with each other making it impossible to know whether points have been missed.
+However, this will result in the sample size of digitised groups conflicting with what is reported in the paper.
 For example, in Figure \ref{fig:all_extract}C only 49 points have been clicked when the sample size is known to be 50.
-Hence, \pkg{metaDigitise} also provides the user with the option to input a known sample sizes directly.
+Hence, \pkg{metaDigitise} also provides the user with the option to input known sample sizes directly.
 Nonetheless, it is important to recognise the impact that overlapping points can have on summary statistics, and in particular on sampling variance.
@@
 \subsection{Histograms}
-The user clicks on the top corners of each bar.
-A line is drawn, in alternate colours, at the top of these bars (Figure \ref{fig:all_extract}D).
-These are numbered to allow the the user to edit them.
+The user clicks on the top corners of each bar and alternating colours are used across bars (Figure \ref{fig:all_extract}D).
+Bars are numbered to allow the the user to edit them.
 As with scatter plots, if the sample size from the extracted data does not match a known sample size, the user can enter an alternate sample size.
@@
 
-%% 1000 words
+%% 779 words
 
@@
 \pkg{metaDigitise} is also able to re-import, edit and re-plot previously digitised figures.
-When running \fct{metaDigitise}, the user can choose to `Import existing data', which returns previously digitised data, for single or all images.
+When running \fct{metaDigitise}, the user can choose to `Import existing data', which returns previously digitised data.
 Users can also choose to `Edit existing data' which allows them to re-plot or edit information or digitisations that have previously be done.
+Points added by mistake can be deleted.
+The user can add more groups, edit groups (add or remove points) or delete groups and this is automatically re-incorporated in the data.
 
@@
 In many cases important information, such as sample sizes, may not be readily available when digitising figures.
-Such information does not, therefore, have to be added a the time of digitisation.
+Such information does not need to be added a the time of digitisation.
 To expedite finding and adding these sample sizes at a later point, \fct{metaDigitise} has a specific edit option that allows users to enter previously omitted sample sizes.
@@
 
-%% 150 words
+%% 183 words
 
@@
 Each variable was plotted twice for each plot type (figures were modified slightly to give users a sense that they were digitising new data) generating a total of 14 figures.
-14 independent digitisers (including the authors) were provided with a directory with all 14 figures in a randomised order.
+Fourteen independent digitisers (including the authors) were provided with a directory with all 14 figures in a randomised order.
 Digitisers ran \pkg{metaDigistise} on their own computers, across different operating systems (including Mac, Windows and Linux).
-Digitisers varied in their level of experience, from people with experience of meta-analyses or comparative work to those without any science background.
+Digitisers varied in their level of experience, from people with experience of meta-analyses to those without any science background.
 We asked users to digitise all 14 figures and collected the mean, standard deviation and correlation coefficient ($r$, for scatter plots) generated by \fct{metaDigitise} for every plot digitised ($n=28$ per digitiser per metric, $n=4$ for $r$).
@@
 The resulting data was used to assess between- and within- user variability (i.e., the intra-class correlation coefficient - ICC).
-This was done using linear mixed effect models with user identify as a random effect using \pkg{lme4} \citep{bates2015} in \proglang{R}.
+This was done using linear mixed effect models with user identity as a random effect using \pkg{lme4} \citep{bates2015} in \proglang{R}.
 Standardised mean, standard deviation and correlation coefficients were used as response variables in separate models.
@@
 
-If digitisations were consistent across users then we should find no significant between user variability in the data.
-Indeed, across plot types we found no evidence for any inter-observer variability in digitisations for the mean (ICC = 0, 95\% CI = 0 to 0.029, \textit{p} = 1), standard deviation (ICC = 0, 95\% CI = 0 to 0.033, \textit{p} = 0.5) or correlation coefficient (ICC = 0.053, 95\% CI = 0 to 0.296, \textit{p} = 0.377).
+Across plot types we found no evidence for any inter-observer variability in digitisations for the mean (ICC = 0, 95\% CI = 0 to 0.029, \textit{p} = 1), standard deviation (ICC = 0, 95\% CI = 0 to 0.033, \textit{p} = 0.5) or correlation coefficient (ICC = 0.053, 95\% CI = 0 to 0.296, \textit{p} = 0.377).
 There were was little bias between digitised and true values, on average 1.63\% (mean = 0.02\%, SD = 4.9\%, \textit{r} = -0.03\%) and overall there were only small absolute differences between digitised and true values, deviating, on average 2.18\% (mean = 0.40\%, SD = 5.81\%, \textit{r} = 0.33\%) across all three summary statistics.
@@
 
-%% 552 words
+%% 577 words
 
@@
 
-Although \pkg{metaDigitise} is already very flexible, and provides functionality not seen in any other package (Table S1), it is clear that there are some functions that it does not perform.
-Notably \pkg{metaDigitise} lacks is automated point detection, available in several packages (Table S1).
+Although \pkg{metaDigitise} is already very flexible, and provides functionality not seen in any other package (see Table S1), it is clear that there are some functions that it does not perform.
+Notably \pkg{metaDigitise} lacks automated point detection.
 However, from our experience, manual digitising is more reliable and often equally as fast.
@@
 
-In contrast to some other packages, \pkg{metaDigitise} does not extract lines from figures.
-In our own experience, line extraction is not particularly useful for meta-analysis, although we recognise that it may be useful in other fields.
+In contrast to some other packages (Table S1), \pkg{metaDigitise} does not extract lines from figures.
+Line extraction may not particularly useful for most meta-analyses, although we recognise that it may be useful in other fields.
 Should a user like to extract lines with \pkg{metaDigitise}, we would recommend extracting data as a scatter plot, and clicking along the line in question.
@@
 
-Descriptive statistics are usually the most robust sources of information for calculating effect size statistics \citep{Noble2017}.
-These are most often presented in figures.
-Users may therefore also want to compare effect size estimates from inferential statistics with those derived from descriptive statistics (obtained for example using \pkg{metaDigitise}) from a paper.
-Comparing these different effects sizes can be useful in identifying uncertainties and problems within a paper.
-In the future, we hope to provide functions to easily convert inferential statistics to standardised effect size estimates, which can seamlessly be integrated with summary statistics from \pkg{metaDigitise}, to calculate equivalent standardised effect size estimates and sampling variance.
-
-%% 416 words
+%% 275 words
 
@@
 
+% 153
 \section*{Author Contributions}
@@
 
+% 51
 \bibliographystyle{FuncEcol.bst}

### 2018-02-28 docs/MEE_submission/manuscript_MEE.tex
Added small changes to MS, 3100 words

@@
 Extracting summary statistics from figures can be a slow process that is not easily reproducible.
-Additionally, current software lacks an ability to incorporate important meta-data (e.g., sample sizes, treatment names) about experiments.
-Furthermore, extraction software is often independent of analytical software preventing the automatic calculation of effect size statistics and making it more challenging to implement user defined analytical pipelines.
+Additionally, current software lacks an ability to incorporate important meta-data (e.g., sample sizes, treatment / variable names) about experiments and is not integrated with other software to streamline analysis pipelines.
 3.
@@
 Digitised data can be easily re-plotted and checked, facilitating reproducible data extraction from plots with little inter-observer bias.
-Such approaches allow collaborative extraction of data.
-We hope that by making the process of figure extraction more flexible and easy to conduct it will improve transparency and quality of meta-analyses in the future.
-%% 230 words
+We hope that by making the process of figure extraction more flexible and easy to conduct it will improve the transparency and quality of meta-analyses in the future.
+%% 209 words
  Keywords: meta-analysis, comparative analysis, data extraction, \proglang{R}, reproducibility, figures, images, summary statistics
@@
 
-%% 414 words
+%% 410 words
 %%%---------------------------------
@@
 
-%% 779 words
+%% 777 words
 
@@
 
-%% 183 words
+%% 171 words
 
@@
 
-%% 577 words
+%% 571 words
 
@@
 
-%% 275 words
+%% 271 words
 
@@
 
+% 68
 \section*{Acknowledgments}

### 2018-03-01 docs/MEE_submission/manuscript_MEE.tex
Updated supplement and MEE MS to try and cut down a bit

@@
 
-\subsection{Inter-observer variability in digitisations}
+In order to evaluate the consistency of digitisation using \pkg{metaDigitise} between users, we simulated a dataset of two variables with two groups.
+The same simulated datasets were given to 14 different digitisers to compare the inter-observer variability in digitisations.
+We also used simulated data to test the accuracy of digitisations with respect to known values (See SM for more details on how simulations were set up).
 
-In order to evaluate the consistency of digitisation using \pkg{metaDigitise} between users, we simulated a dataset of two variables with two groups ($n=10$ within groups).
-Each variable was plotted twice for each plot type (figures were modified slightly to give users a sense that they were digitising new data) generating a total of 14 figures.
-Fourteen independent digitisers (including the authors) were provided with a directory with all 14 figures in a randomised order.
-Digitisers ran \pkg{metaDigistise} on their own computers, across different operating systems (including Mac, Windows and Linux).
-Digitisers varied in their level of experience, from people with experience of meta-analyses to those without any science background.
-We asked users to digitise all 14 figures and collected the mean, standard deviation and correlation coefficient ($r$, for scatter plots) generated by \fct{metaDigitise} for every plot digitised ($n=28$ per digitiser per metric, $n=4$ for $r$).
-
-As a measure of bias, we calculated the percentage differences from the true summary statistics as
-
-\begin{equation}
-\frac{\theta - \hat{\theta}}{\hat{\theta}}
-\end{equation}
-
-where $\theta$ is the estimate and $\hat{\theta}$ is the true value.
-The deviation from the true value of $r$ was not further standardised, as it is already on a standardised scale.
-We also took the absolute values of these standardised differences as a measure of precision.
-The resulting data was used to assess between- and within- user variability (i.e., the intra-class correlation coefficient - ICC).
-This was done using linear mixed effect models with user identity as a random effect using \pkg{lme4} \citep{bates2015} in \proglang{R}.
-Standardised mean, standard deviation and correlation coefficients were used as response variables in separate models.
-Sampling variance for ICC estimates was generated based on 1000 parametric bootstraps of the model and the significance was tested using likelihood ratio tests, using \pkg{rptR} \citep{stoffel2017}.
-
-Across plot types we found no evidence for any inter-observer variability in digitisations for the mean (ICC = 0, 95\% CI = 0 to 0.029, \textit{p} = 1), standard deviation (ICC = 0, 95\% CI = 0 to 0.033, \textit{p} = 0.5) or correlation coefficient (ICC = 0.053, 95\% CI = 0 to 0.296, \textit{p} = 0.377).
-There were was little bias between digitised and true values, on average 1.63\% (mean = 0.02\%, SD = 4.9\%, \textit{r} = -0.03\%) and overall there were only small absolute differences between digitised and true values, deviating, on average 2.18\% (mean = 0.40\%, SD = 5.81\%, \textit{r} = 0.33\%) across all three summary statistics.
+Across the plot types we found no evidence for any inter-observer variability in digitisations for the mean (ICC = 0, 95\% CI = 0 to 0.029, \textit{p} = 1), standard deviation (ICC = 0, 95\% CI = 0 to 0.033, \textit{p} = 0.5) or correlation coefficient (ICC = 0.053, 95\% CI = 0 to 0.296, \textit{p} = 0.377).
+There were was also little bias between digitised and true values, on average 1.63\% (mean = 0.02\%, SD = 4.9\%, \textit{r} = -0.03\%) and overall there were only small absolute differences between digitised and true values, deviating, on average 2.18\% (mean = 0.40\%, SD = 5.81\%, \textit{r} = 0.33\%) across all three summary statistics.
 SD estimates from digitisations are clearly more prone to error than means or correlation coefficients.
@@
 
-\subsection{Testing the accuracy of digitisations}
-To test how accurate \pkg{metaDigitise} is at matching clicked points to their true values, we generated four random scatterplots, each with 20 data points, and digitised these with \fct{metaDigitise}.
-This was done by one digitiser (J.L.P.), as there is no detectable between user variation.
-Data digitised using \pkg{metaDigitise} was essentially perfectly correlated with the true simulated data for both the \textit{x}-variable (Pearson's correlation; \textit{r} = 0.9999915, \textit{t} = 2137.4, \textit{df} = 78, $p < 0.001$) and \textit{y}-variable (\textit{r} = 0.9999892, \textit{t} = 1897.8, \textit{df} = 78, $p < 0.001$).
+\pkg{metaDigitise} was extremely accurate at matching clicked points to their true values essentially being perfectly 
+correlated with the true simulated data for both the \textit{x}-variable (Pearson's correlation; \textit{r} = 0.9999915, \textit{t} = 2137.4, \textit{df} = 78, $p < 0.001$) and \textit{y}-variable (\textit{r} = 0.9999892, \textit{t} = 1897.8, \textit{df} = 78, $p < 0.001$) in scatterplots.
 

### 2018-03-01 docs/MEE_submission/supplements.tex
Updated supplement and MEE MS to try and cut down a bit

@@
 
+\section{Tutorial}
+\subsection{Extracting Data From Plots}
+%% scatterplot from airquality data
+
+We will now demonstrate how \fct{metaDigitise} works using figures generated from the well known iris data set. Users can install the \pkg{metaDigitise} package from GitHub as follows:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> install.packages("devtools")
+R> devtools::install_github("daniel1noble/metaDigitise")
+R> library(metaDigitise)
+\end{CodeInput}
+\end{CodeChunk}
+
+Assume that the user would like to extract descriptive statistics from studies measuring sepal length or width in iris species for a fictitious project.
+There are a few studies that only present these data in figures.
+As the user reads papers found from a systematic search, they add figures with relevant data to a "FiguresToExtract" folder as follows
+
+\begin{CodeChunk}
+\begin{CodeOutput}
+*FiguresToExtract/
+    + 001_Anderson_1935_Fig1.png
+\end{CodeOutput}
+\end{CodeChunk}
+
+\begin{figure}[!h]
+\centering
+ \includegraphics[width=0.75\textwidth]{001_Anderson_1935_Fig1.png} 
+ \caption{Example scatterplot (001\_Anderson\_1935\_Fig1.png) of sepal length and width for two species of iris (setosa and versicolor)}
+\label{fig:Anderson}
+\end{figure}
+
+Here, the naming of the files placed in the folder will contain the paper number, first author and the figure number to keep data uniquely associated with figures.
+At first there is one figure in the folder, shown in Figure \ref{fig:Anderson}.
+Running \fct{metaDigitise} brings up a series of prompts for the user using a main menu that provides access to a number of its features ("..." here represents the user's path to the project directory):
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> digitised_data <- metaDigitise(".../FiguresToExtract", summary = TRUE)
+\end{CodeInput}
+\begin{CodeOutput}
+	Do you want to...
+1: Process new images
+2: Import existing data
+3: Edit existing data
+Selection: 
+\end{CodeOutput}
+\end{CodeChunk}
+
+The user simply enters in the numeric value that corresponds to what they would like to do.
+In this case they want to "Process new images".
+The user is then asked whether there are different types of plot(s) in the folder.
+This question is most relevant when there are lots of different figures in the folder because it will then ask the user for the type of figure as they are cycled through.
+
+\begin{CodeChunk}
+\begin{CodeOutput}
+Are all plot types Different or the Same? (d/s)
+\end{CodeOutput}
+\end{CodeChunk}
+
+\fct{metaDigitise} then asks the user whether the figure needs to be rotated or flipped.
+This can be needed when box plots and mean and error plots are not orientated correctly.
+In some cases, older papers can give slightly off angled images which can be corrected by rotating.
+So, in this prompt the user has three options: \code{f} for ``Flip'', \code{r} for ``rotate'' or \code{c} for ``continue''.
+
+\begin{CodeChunk}
+\begin{CodeOutput}
+mean_error and boxplots should be vertically orientated
+       _ 
+       |  
+  I.E. o    NOT  |-o-|
+       |
+       _
+
+If they are not then chose flip to correct this.
+
+If figures are wonky, chose rotate.
+
+Otherwise chose continue
+
+Flip, rotate or continue (f/r/c) 
+\end{CodeOutput}
+\begin{CodeInput}
+R> c
+\end{CodeInput}
+\end{CodeChunk}
+
+After this, \fct{metaDigitise} will ask the user to specify the plot type. Depending on the figure, the user can specify that it is a figure containing the mean and error (\code{m}), a box plot (\code
+{b}), a scatter plot (\code{s}) or a histogram (\code{h}).
+If the user has specified \code{d} instead of \code{s} in response to the question about whether the plot types are the same or different, this question will pop up for each plot, but will only be asked once if plots are all the same.
+
+\begin{CodeChunk}
+\begin{CodeOutput}
+Please specify the plot_type as either:
+
+ m: Mean and error
+ b: Box plot
+ s: Scatter plot 
+ h: Histogram
+\end{CodeOutput}
+\begin{CodeInput}
+R> s
+\end{CodeInput}
+\end{CodeChunk}
+
+After selecting the figure type a new set of prompts will come up that will ask the user first what the y and x-axis variables are.
+This is useful as users can keep track of the different variables across figures and papers.
+Here, the user can just add this information in to the \proglang{R} console.
+Once complete, details on how to calibrate the x and y-axis appear, so that the relevant statistics / data can be correctly calculated.
+When working with a plot of mean and standard errors, the x-axis is rather useless in terms of calibration so \fct{metaDigitise} just asks the user to calibrate the y-axis.
+
+\begin{CodeChunk}
+\begin{CodeOutput}
+What is the y variable? 
+\end{CodeOutput}
+\begin{CodeInput}
+R> Sepal Length (mm)
+\end{CodeInput}
+
+\begin{CodeOutput}
+What is the x variable? 
+\end{CodeOutput}
+\begin{CodeInput}
+R> Sepal Width (mm)
+\end{CodeInput}
+
+\begin{CodeOutput}
+On the Figure, click IN ORDER: 
+      y1, y2 , x1, x2
+
+    Step 1 ----> Click on known value on y axis - y1
+  |
+  |
+  |
+  |
+  y1
+  |_________________________
+  ....
+
+    Step 3 ----> Click on known value on x axis - x1
+  |
+  |
+  |
+  |
+  |
+  |_____x1__________________
+
+  ....
+\end{CodeOutput}
+\end{CodeChunk}
+
+The user can just follow the instructions on screen step-by-step (instructions above have been truncated by `...' to simplify), and in the order specified.
+Before moving on, the user is forced to check whether or not the calibration has been set up correctly.
+If \code{n} is chosen because something needs to be fixed then the user can re-calibrate.
+
+\begin{CodeChunk}
+\begin{CodeOutput}
+What is the value of y1 ?
+\end{CodeOutput}
+\begin{CodeInput}
+R> 4.5
+\end{CodeInput}
+\begin{CodeOutput}
+What is the value of y2 ?
+\end{CodeOutput}
+\begin{CodeInput}
+R> 7
+\end{CodeInput}
+\begin{CodeOutput}
+What is the value of x1 ?
+\end{CodeOutput}
+\begin{CodeInput}
+R> 2
+\end{CodeInput}
+\begin{CodeOutput}
+What is the value of x2 ?
+\end{CodeOutput}
+\begin{CodeInput}
+R> 4
+\end{CodeInput}
+\begin{CodeOutput}
+Re-calibrate? (y/n) 
+\end{CodeOutput}
+\begin{CodeInput}
+R> n
+\end{CodeInput}
+\end{CodeChunk}
+
+Often, plots might contain multiple groups that the meta-analyst wants to extract from.
+\fct{metaDigitise} handles this nicely by prompting the user to enter the group first, followed by digitisation of this groups data.
+After digitising the first group, and having exited, \fct{metaDigitise} will ask the user whether they would like to add another group.
+Users can continually add groups (\code{a}), delete groups (\code{d}), edit groups (\code{e}) or finish a plot and continue to the next one (\code{f} - if another plot exists).
+%%% This needs checking - think it is now different
+The number of groups are not really limited and users can just keep adding in groups to accommodate the different numbers that may be presented across figures (although it can get complicated with too many).
+
+\begin{CodeChunk}
+\begin{CodeOutput}
+If there are multiple groups, enter unique group identifiers (otherwise press enter)
+Group identifier: 
+\end{CodeOutput}
+\begin{CodeInput}
+R> setosa
+\end{CodeInput}
+\begin{CodeOutput}
+Click on points you want to add.
+If you want to remove a point, or are finished with a group, 
+exit by clicking on red box in bottom left corner, then follow prompts
+\end{CodeOutput}
+\end{CodeChunk}
+
+To finish selecting points, the user can exit by clicking on the red button that appears when extracting points. The user is then asked if they want to add or delete points from that group.
+
+\begin{CodeChunk}
+\begin{CodeOutput}
+Add or Delete points to this group, or Continue? (a/d/c) 
+\end{CodeOutput}
+\begin{CodeInput}
+R> c
+\end{CodeInput}
+\end{CodeChunk}
+
+Once we are done digitising all the groups our plot will look something like Figure \ref{fig:scatter_extract}.
+
+\begin{figure}[!h] 
+\centering
+ \includegraphics[width=0.75\textwidth]{001_Anderson_1935_Fig1_digitised.png} 
+ \caption{Digitisation of sepal length and width for two species of iris (setosa and versicolor).
+Names of the variables and calibration (in blue) are plotted alongside the digitised points (green = versicolor; red = setosa).
+The sample sizes for each group are provided on the lower part of the plot.
+All figures are clearly labelled at the top to remind users of the filename and plot type.
+This reduces errors throughout the digitisation process.}
+\label{fig:scatter_extract}
+\end{figure}
+
+When completed \fct{metaDigitise} will write the digitised data as a \code{metaDigitise} object to a RDS file in the caldat directory, such that our new directory structure is as follows
+
+\begin{CodeChunk}
+\begin{CodeOutput}
+*FiguresToExtract/
+    + caldat/
+        + 001_Anderson_1935_Fig1
+    + 001_Anderson_1935_Fig1.png
+\end{CodeOutput}
+\end{CodeChunk}
+
+Users can access the \code{metaDigitise} object created (001\_Anderson\_1935\_Fig1) at any time using the \fct{metaDigitse} function.
+In the \proglang{R} console, the summarised data for the digitised figure can be printed on screen or even written to a .csv file:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> digitised_data
+\end{CodeInput}
+{\scriptsize
+\begin{CodeOutput}
+                  filename    group_id         variable  mean  error error_type  n     r    sd  plot_type
+001_Anderson_1935_Fig1.png      setosa  Sepal width (mm)  3.42  0.40  sd        39  0.75  0.40  scatterplot
+001_Anderson_1935_Fig1.png      setosa  Sepal length (mm) 5.00  0.38  sd        39  0.75  0.38  scatterplot
+001_Anderson_1935_Fig1.png  versicolor  Sepal width (mm)  2.77  0.32  sd        44  0.52  0.32  scatterplot
+001_Anderson_1935_Fig1.png  versicolor  Sepal length (mm) 5.95  0.53  sd        44  0.52  0.53  scatterplot
+\end{CodeOutput}
+}
+\end{CodeChunk}
+
+The mean for each of the two variables, along with the two species, are provided.
+Since this is a scatterplot, the user also gets the Person's correlation coefficient between sepal length and width for each species.
+These match reasonably well with the actual means of sepal length and width for each of the species in the full `iris` dataset:
+
+\begin{CodeChunk}
+\begin{CodeOutput}
+     Species meanSL meanSW
+1     setosa  5.006  3.428
+2 versicolor  5.936  2.770
+\end{CodeOutput}
+\end{CodeChunk}
+
+One thing anyone with a familiarity with the iris dataset will notice is that the sample sizes for each of these species (which are n = 50 each) are quite a bit lower.
+This is an example of some of the challenges when extracting data from scatter plots.
+Often data points will overlap with each other making it impossible (without having the real data) to know whether this is a problem.
+However, a meta-analyst will probably realise that the sample sizes here conflict with what is reported in the paper.
+Hence, \pkg{metaDigitise} also provides the user with  options to input the sample sizes directly (see Editing section below), even for scatter plots and histograms where, strictly speaking, this should not be necessary.
+Nonetheless, it is important to recognise the impact that overlapping points can have on summary statistics, particularly its effects on standard deviation (SD) and standard error (SE).
+Here, the mean point estimates are nearly exactly the same as the true values, but the SD's are slightly over-estimated:
+
+\begin{CodeChunk}
+\begin{CodeOutput}
+     Species    meanSL    meanSW
+1     setosa 0.3524897 0.3790644
+2 versicolor 0.5161711 0.3137983
+\end{CodeOutput}
+\end{CodeChunk}
+
+\subsection{Adding new figures}
+
+Users can add additional figures as new papers with relevant information are found.
+Each figure should be in its own file with unique naming, even if a single paper has multiple figures for extraction.
+For example, another paper on different populations (and one new species) of iris contained two additional figures where important data could be extracted.
+These figures can simply be named accordingly and added directly to the same extraction folder:
+
+\begin{CodeChunk}
+\begin{CodeOutput}
+*FiguresToExtract/
+    + caldat/
+        + 001_Anderson_1935_Fig1
+    + 001_Anderson_1935_Fig1.png
+    + 002_Doe_2013_Fig1.png
+    + 002_Doe_2013_Fig3.png
+\end{CodeOutput}
+\end{CodeChunk}
+
+The user has already processed one figure (001\_Anderson\_1935\_Fig1.png).
+We can tell this because the caldat folder has digitised data in it (caldat/001\_Anderson\_1935\_Fig1).
+Now the user has two new figures that have not yet been digitised.
+This example will nicely demonstrate how users can easily pick up from where they left off and how all previous data gets re-integrated.
+It will also demonstrate how different plot types are handled.
+All we have to do to begin, is again, provide the directory where all the figures are located:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> digitised_data <- metaDigitise("~/FiguresToExtract", summary = TRUE)
+\end{CodeInput}
+\end{CodeChunk}
+
+The user gets the same set of prompts and simply chooses option one.
+This will permit users to digitise new figures, and will integrate previously completed digitisations along with newly digitised data together at the end of the session, or when the user decides to quit.
+This time, 001\_Anderson\_1935\_Fig1.png is ignored and the new plots cycle on screen; first 002\_Doe\_2013\_Fig1.png and then 002\_Doe\_2013\_Fig3.png.
+Since there are a few different figure types, the user answers the first question in the \proglang{R} console as "d":
+
+\begin{CodeChunk}
+\begin{CodeOutput}
+Are all plot types Different or the Same? (d/s)
+\end{CodeOutput}
+\begin{CodeInput}
+R> d
+\end{CodeInput}
+\begin{CodeOutput}
+**** NEW PLOT ****
+
+mean_error and boxplots should be vertically orientated
+       _ 
+       |  
+  I.E. o    NOT  |-o-|
+       |
+       _
+
+If they are not then chose flip to correct this.
+
+If figures are wonky, chose rotate.
+
+Otherwise chose continue
+
+Flip, rotate or continue (f/r/c) 
+\end{CodeOutput}
+\begin{CodeInput}
+R> c
+\end{CodeInput}
+\begin{CodeOutput}
+Please specify the plot_type as either:
+
+ m: Mean and error
+ b: Box plot
+ s: Scatter plot 
+ h: Histogram
+\end{CodeOutput}
+\begin{CodeInput}
+R> m
+\end{CodeInput}
+\end{CodeChunk}
+
+Here, the user specifies the new plot type as \code{m} for 002\_Doe\_2013\_Fig1.png because the user has a plot of the mean and error of sepal length for each of the three species.
+The user is then prompted a bit differently from our scatter plot as the x-axis is not needed for calibration:
+
+\begin{CodeChunk}
+\begin{CodeOutput}
+What is the y variable? 
+\end{CodeOutput}
+\begin{CodeInput}
+R> Sepal length
+\end{CodeInput}
+\begin{CodeOutput}
+On the Figure, click IN ORDER: 
+      y1, y2
+
+    Step 1 ----> Click on y1
+  |
+  |
+  |
+  |
+  y1
+  |_________________________
+
+    Step 2 ----> Click on y2
+  |
+  y2
+  |
+  |
+  |
+  |_________________________
+
+What is the value of y1 ?
+\end{CodeOutput}
+\begin{CodeInput}
+R> 5
+\end{CodeInput}
+\begin{CodeOutput}
+What is the value of y2 ?
+\end{CodeOutput}
+\begin{CodeInput}
+R> 6.5
+\end{CodeInput}
+\begin{CodeOutput}
+Re-calibrate? (y/n) 
+\end{CodeOutput}
+\begin{CodeInput}
+R> n
+\end{CodeInput}
+\begin{CodeOutput}
+Do you know sample sizes? (y/n)  
+\end{CodeOutput}
+\begin{CodeInput}
+R> y
+\end{CodeInput}
+\begin{CodeOutput}
+If there are multiple groups, enter unique group identifiers (otherwise press enter)
+Group identifier: 
+\end{CodeOutput}
+\begin{CodeInput}
+R> setosa
+\end{CodeInput}
+\begin{CodeOutput}
+Group sample size: 
+\end{CodeOutput}
+\begin{CodeInput}
+R> 50
+\end{CodeInput}
+\begin{CodeOutput}
+Click on Error Bar, followed by the Mean
+
+Add group, Edit Group, Delete group or Finish plot? (a/e/d/f) 
+\end{CodeOutput}
+\begin{CodeInput}
+R> a
+\end{CodeInput}
+\end{CodeChunk}
+
+Again, \fct{metaDigitise} will simply guide the user through digitising each of these figures describing to them exactly what needs to be done.
+At any point if mistakes are made the user can choose relevant options to edit or correct things before ending the figure.
+This process continues for each plot so long as the user would like to continue and after completing a single plot the user is always prompted as follows:
+
+\begin{CodeChunk}
+\begin{CodeOutput}
+Do you want continue: 1 plots out of 2 plots remaining (y/n)
+\end{CodeOutput}
+\begin{CodeInput}
+R> y
+\end{CodeInput}
+\end{CodeChunk}
+
+This continues until users have completed all non-digitised figures in the folder, at which point \fct{metaDigitise} concatenates the new data with previously digitised data in the object:
+
+\begin{CodeChunk}
+{\scriptsize
+\begin{CodeOutput}
+data
+                  filename    group_id         variable   mean  error error_type n    r    sd   plot_type
+001_Anderson_1935_Fig1.png      setosa  Sepal width (mm)  3.42  0.40  sd        39  0.75  0.40  scatterplot
+001_Anderson_1935_Fig1.png      setosa  Sepal length (mm) 5.00  0.38  sd        39  0.75  0.38  scatterplot
+001_Anderson_1935_Fig1.png  versicolor  Sepal width (mm)  2.77  0.32  sd        44  0.52  0.32  scatterplot
+001_Anderson_1935_Fig1.png  versicolor  Sepal length (mm) 5.95  0.53  sd        44  0.52  0.53  scatterplot
+     002_Doe_2013_Fig1.png      setosa  Sepal length      5.00  0.11  se        50  NA    0.78  mean_error
+     002_Doe_2013_Fig1.png  viriginica  Sepal length      6.59  0.18  se        50  NA    1.26  mean_error
+     002_Doe_2013_Fig1.png  versicolor  Sepal length      5.94  0.14  se        50  NA    1.01  mean_error
+     003_Doe_2013_Fig3.png      catana  Sepal length      4.95  0.36  sd        50  NA    0.36  histogram
+\end{CodeOutput}
+}
+\end{CodeChunk}
+
+\subsection{Re-importing, Editing and Plotting Previously Digitised data}
+
+A particularly useful feature of \pkg{metaDigitise} is its ability to re-import, edit and re-plot previously digitised figures. We can do this from the initial options from \fct{metaDigitise}
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> digitised_data <- metaDigitise(".../FiguresToExtract")
+\end{CodeInput}
+\begin{CodeOutput}
+    Do you want to...
+1: Process new images
+2: Import existing data
+3: Edit existing data
+Selection: 
+\end{CodeOutput}
+\end{CodeChunk}
+
+If the user chooses "Import existing data", they have the option of either 1) importing data from all digitised images or 2) importing data from a single image that has been digitised.
+If 2, then a list of files are provided to the user that they can select.
+Editing existing data allows users to easily re-plot or edit information or digitisations that have previously be done for any plot.
+This is accomplished by guiding the user through a new set of options:
+
+\begin{CodeChunk}
+\begin{CodeOutput}
+Choose how you want to edit files:
+1: Cycle through images
+2: Choose specific file to edit
+3: Enter previously omitted sample sizes
+Selection: 
+\end{CodeOutput}
+\end{CodeChunk}
+
+If the user is unsure about the name of the specific figure they need to edit or simply want to just check the digitisations of figures they can choose "Cycle through images", which will bring up each figure, one by one, overlaying the calibrations, group names (if they exist), sample sizes (if they were entered) and the selected points.
+The user will then be given the choice to edit individual images.
+Alternatively, choosing option 2, will bring up a list of the completed files in the folder and the specific file can be chosen, at which point it will be replotted.
+Either of these options will cycle through a number of questions asking the user what they would like to edit:
+
+\begin{CodeChunk}
+
+\begin{CodeOutput}
+Edit rotation? If yes, then the whole extraction will be redone (y/n) 
+\end{CodeOutput}
+\begin{CodeInput}
+R> n
+\end{CodeInput}
+\begin{CodeOutput}
+Change plot type? If yes, then the whole extraction will be redone (y/n) 
+\end{CodeOutput}
+\begin{CodeInput}
+R> n
+\end{CodeInput}
+\begin{CodeOutput}
+Variable entered as: 
+\end{CodeOutput}
+\begin{CodeInput}
+R> Sepal length
+\end{CodeInput}
+\begin{CodeOutput}
+Rename Variables (y/n) 
+\end{CodeOutput}
+\begin{CodeInput}
+R> n
+\end{CodeInput}
+\begin{CodeOutput}
+Edit calibration? (y/n) 
+\end{CodeOutput}
+\begin{CodeInput}
+R> n
+\end{CodeInput}
+\begin{CodeOutput}
+Re-extract data (y/n) 
+\end{CodeOutput}
+\begin{CodeInput}
+R> y
+\end{CodeInput}
+\begin{CodeOutput}
+Change group identifier? (y/n) 
+\end{CodeOutput}
+\begin{CodeInput}
+R> n
+\end{CodeInput}
+\begin{CodeOutput}
+Add group, Delete group or Finish plot? (a/d/f) 
+\end{CodeOutput}
+\begin{CodeInput}
+R> d
+\end{CodeInput}
+\begin{CodeOutput}
+1: setosa
+2: versicolor
+3: viriginica
+Selection: 
+\end{CodeOutput}
+\begin{CodeInput}
+R> 2
+\end{CodeInput}
+\begin{CodeOutput}
+Add group, Delete group or Finish plot? (a/d/f) 
+\end{CodeOutput}
+\begin{CodeInput}
+R> a
+\end{CodeInput}
+
+\end{CodeChunk}
+
+A whole host of information can be edited including the rotation, plot type, the variable name(s) that were provided, the calibration and even the digitisation of groups.
+When editing the \code{metaDigitise} object is re-written to the caldat folder and the edits are immediately integrated into the existing object once complete.
+
+\subsection{Obtaining Processed Data}
+
+While \fct{metaDigitise} provides users with the summary statistics by default, for all plot types, in many cases the user may actually be interested in obtaining the processed digitised data from scatter plots (i.e.
+calibrated points).
+This is very easy to do my changing the default \code{summary} argument from TRUE to FALSE in \fct{metaDigitise}.
+Instead of providing the user with summary statistics it will return a list containing four slots for each of the figure types (mean error, box plot, histogram and scatter plots).
+An example of a data object returned from digitising figures is as follows:
+
+\begin{CodeChunk}
+
+\begin{CodeInput}
+>R str(data)
+\end{CodeInput}
+
+\begin{CodeOutput}
+List of 3
+ $ mean_error :List of 1
+  ..$ 002_Doe_2013_Fig1.png:'data.frame': 3 obs. of  5 variables:
+  .. ..$ id      : Factor w/ 3 levels "setosa","versicolor",..: 1 2 3
+  .. ..$ mean    : num [1:3] 5 5.93 6.59
+  .. ..$ error   : num [1:3] 0.111 0.148 0.178
+  .. ..$ n       : num [1:3] 50 50 50
+  .. ..$ variable: chr [1:3] "Sepal length" "Sepal length" "Sepal length"
+ $ hist       :List of 1
+  ..$ 003_Doe_2013_Fig3.png:'data.frame': 8 obs. of  3 variables:
+  .. ..$ midpoints: num [1:8] 4.3 4.5 4.7 4.9 5.1 ...
+  .. ..$ frequency: num [1:8] 4 5 7 12 11 6 2 3
+  .. ..$ variable : chr [1:8] "Sepal length" "Sepal length" ...
+ $ scatterplot:List of 1
+  ..$ 001_Anderson_1935_Fig1.png:'data.frame':  83 obs. of  8 variables:
+  .. ..$ id        : Factor w/ 2 levels "setosa","versicolor": 1 1 1 1 1 ...
+  .. ..$ x         : num [1:83] 2.3 2.9 3 3 3 ...
+  .. ..$ y         : num [1:83] 4.5 4.4 4.41 4.3 4.8 ...
+  .. ..$ group     : num [1:83] 1 1 1 1 1 1 1 1 1 1 ...
+  .. ..$ col       : Factor w/ 2 levels "red","green": 1 1 1 1 1 1 1 1 1 ...
+  .. ..$ pch       : num [1:83] 19 19 19 19 19 19 19 19 19 19 ...
+  .. ..$ y_variable: chr [1:83] "Sepal length (mm)" "Sepal length (mm)"  ...
+  .. ..$ x_variable: chr [1:83] "Sepal width (mm)" "Sepal width (mm)"    ...
+\end{CodeOutput}
+\end{CodeChunk}
+
+Here, the user can easily access the list of processed scatter plot data by simply extracting the scatter plot slot:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+>R scatterplot <- data$scatterplot
+\end{CodeInput}
+\end{CodeChunk}
+
+\subsection{Adding sample sizes to previous Digitisations}
+%Including addition of N later
+In many cases important information, such as sample sizes, may not be readily available or clear when digitising figures.
+In these circumstances users will have answered `no' to the question about whether they have sample sizes or not while digitising.
+To expedite finding and adding in these sample sizes to do the necessary calculations (if for example a figure presented 95\% CI's or standard errors), \fct{metaDigitise} has s specific edit option that allows users to enter in previously omitted sample sizes.
+It works by first identifying the missing sample sizes in the digitised output, re-plotting the relevant figure and then prompting the user to enter the sample sizes for the relevant groups in the figure, one by one.
+As an example, assume that we were missing sample sizes for two groups in 002\_Doe\_2013\_Fig1.png:
+
+\begin{CodeChunk}
+{\scriptsize
+\begin{CodeOutput}
+            filename    group_id         variable   mean  error error_type n    r    sd   plot_type
+002_Doe_2013_Fig1.png      setosa  Sepal length      5.00  0.11  se        NA  NA    NA  mean_error
+002_Doe_2013_Fig1.png  viriginica  Sepal length      6.59  0.18  se        NA  NA    NA  mean_error
+\end{CodeOutput}
+}
+\end{CodeChunk}
+
+Here, we can see that we are missing the sample sizes for setosa and viriginica, and as a result, sd is not calculated because \fct{metaDigitise} needs this information to make the calculation.
+If the user found this information after contacting the authors for clarification then they can add these back in as follows:
+
+\begin{CodeChunk}
+\begin{CodeInput}
+R> digitised_data <- metaDigitise(".../FiguresToExtract")
+\end{CodeInput}
+\end{CodeChunk}
+
+\begin{CodeChunk}
+\begin{CodeOutput}
+Do you want to...
+
+1: Process new images
+2: Import existing data
+3: Edit existing data
+
+Selection:
+\end{CodeOutput}
+\begin{CodeInput}
+R> 3
+\end{CodeInput}
+\end{CodeChunk}
+
+\begin{CodeChunk}
+\begin{CodeOutput}
+Choose how you want to edit files:
+
+1: Cycle through images
+2: Choose specific file to edit
+3: Enter previously omitted sample sizes
+
+Selection:
+\end{CodeOutput}
+\begin{CodeInput}
+>R 3
+\end{CodeInput}
+\end{CodeChunk}
+
+\fct{metaDigitise} will replot the figure after this and list, only the groups missing data, for which the user can then update the data.
+This is then re-integrated back into the data automatically and the sd calculated.
+
+\begin{CodeChunk}
+\begin{CodeOutput}
+Group " setosa ": Enter sample size  
+\end{CodeOutput}
+\begin{CodeInput}
+R> 50
+\end{CodeInput}
+\begin{CodeOutput}
+Group " viriginica ": Enter sample size
+\end{CodeOutput}
+\begin{CodeInput}
+R> 50
+\end{CodeInput}
+\end{CodeChunk}
+
+\subsection{Figure Rotation and Adjustment}
+Figures may have been extracted from old publications, for example from scanned images, and so are not perfectly orientated on the image.
+This will make the calibration of the points in the figure from the image problematic.
+\fct{metaDigitise} allows users to rotate the image.
+By clicking two points on the x-axis, metaDigitse calculates the angle needed to rotate the image so the x-axis is horizontal, and rotates it.
+(Figure \ref{fig:rotate}A,B)
+
+\begin{figure}[!b] 
+\centering
+ \includegraphics[width=0.75\textwidth]{fig_rotate.pdf} 
+ \caption{Figure rotation. A) and B) show how non-aligned images can be realigned through user defined rotation. C) and D) show how figures can be re-orientated so as to aid data input.}
+\label{fig:rotate}
+\end{figure}
+
+Furthermore, some figures, including mean and error, boxplots or histograms, may be presented with horizontal bars.
+\fct{metaDigitise} assumes that the bars are vertical, but allows the user to flip the image so that the bars are vertical if provided horizontally (Figure \ref{fig:rotate}C,D).
+
+\section{Derivation of mean, standard deviation and sample size from different plot types}
+
+\subsection{Mean/Error Plots}
+The standard deviation is calculated depending on the type of error presented.
+The user can choose from standard deviation (SD, $\sigma$), standard error (SE) or 95\% confidence intervals (CI95).
+Standard deviation is calculated from standard error as
+\begin{equation}
+\sigma = SE \sqrt{n}
+\end{equation}
+and from 95\% confidence intervals as
+\begin{equation}
+\sigma = \frac{CI}{1.96} \sqrt{n}
+\end{equation}
+
+\subsection{Box Plots} 
+The mean ($\mu$) and SD are calculated sing the maximum ($b$), upper quartile ($q_3$), median ($m$), lower quartile ($q_1$) and minimum ($a$) as 
+\begin{equation}
+\mu = \frac{(n+3)(a+b) + 2(n-1)(q_1 + m + q_3)}{8n}
+\end{equation}
+following \citet{Bland2015} and
+\begin{equation}
+\sigma = \frac{b-a}{4\Phi^{-1}(\frac{n-0.375}{n+0.25})} + \frac{q_3-q_1}{4\Phi^{-1}(\frac{0.75n-0.125}{n+0.25})}
+\end{equation}
+where $\Phi^{-1}(z)$ is the upper zth percentile of the standard normal distribution, following \citet{Wan2014}.
+
+\subsection{Histograms} 
+For each bar, the user click two point (the top of the bar).
+Using these points, a midpoint (m; mean x coordinates) and a frequency (f; mean y coordinates, rounded to the nearest integer) is calculated for each bar.
+The sample size, mean and SD are calculated as:
+\begin{equation}
+n = \sum_{i=1}^n{f_i}
+\end{equation}
+
+\begin{equation}
+\mu = \frac{\sum_{i=1}^n{m_i f_i}}{n}
+\end{equation}
+
+\begin{equation}
+\sigma = \sqrt{\frac{\sum_{i=1}^n{(m_i f_i - \mu f_i)^2}}{n-1}}
+\end{equation}
+
+\section{Software Validation: Details of Simulation}
+
+\subsection{Inter-observer variability in digitisations}
+In order to evaluate the consistency of digitisation using \pkg{metaDigitise} between users, we simulated a dataset of two variables with two groups($n=10$ within groups).
+Each variable was plotted twice for each plot type (figures were modified slightly to give users a sense that they were digitising new data) generating a total of 14 figures.
+Fourteen independent digitisers (including the authors) were provided with a directory with all 14 figures in a randomised order.
+Digitisers ran \pkg{metaDigistise} on their own computers, across different operating systems (including Mac, Windows and Linux).
+Digitisers varied in their level of experience, from people with experience of meta-analyses to those without any science background.
+We asked users to digitise all 14 figures and collected the mean, standard deviation and correlation coefficient ($r$, for scatter plots) generated by \fct{metaDigitise} for every plot digitised ($n=28$ per digitiser per metric, $n=4$ for $r$).
+
+As a measure of bias, we calculated the percentage differences from the true summary statistics as
+
+\begin{equation}
+\frac{\theta - \hat{\theta}}{\hat{\theta}}
+\end{equation}
+
+where $\theta$ is the estimate and $\hat{\theta}$ is the true value.
+The deviation from the true value of $r$ was not further standardised, as it is already on a standardised scale.
+We also took the absolute values of these standardised differences as a measure of precision.
+The resulting data was used to assess between- and within- user variability (i.e., the intra-class correlation coefficient - ICC).
+This was done using linear mixed effect models with user identity as a random effect using \pkg{lme4} \citep{bates2015} in \proglang{R}.
+Standardised mean, standard deviation and correlation coefficients were used as response variables in separate models.
+Sampling variance for ICC estimates was generated based on 1000 parametric bootstraps of the model and the significance was tested using likelihood ratio tests, using \pkg{rptR} \citep{stoffel2017}.
+
+Across plot types we found no evidence for any inter-observer variability in digitisations for the mean (ICC = 0, 95\% CI = 0 to 0.029, \textit{p} = 1), standard deviation (ICC = 0, 95\% CI = 0 to 0.033, \textit{p} = 0.5) or correlation coefficient (ICC = 0.053, 95\% CI = 0 to 0.296, \textit{p} = 0.377).
+There were was little bias between digitised and true values, on average 1.63\% (mean = 0.02\%, SD = 4.9\%, \textit{r} = -0.03\%) and overall there were only small absolute differences between digitised and true values, deviating, on average 2.18\% (mean = 0.40\%, SD = 5.81\%, \textit{r} = 0.33\%) across all three summary statistics.
+SD estimates from digitisations are clearly more prone to error than means or correlation coefficients.
+If the mean absolute difference is calculated for each plot type, we can see that this effect is driven mainly by extraction from box plots and histograms (\% difference; box plot: 15.805, histogram: 5.210, mean/error: 1.500, scatter plot: 0.433).
+SD estimation from box plot summary statistics is known to be more error prone, especially at small sample sizes \citep{Wan2014}.
+
+\subsection{Testing the accuracy of digitisations}
+To test how accurate \pkg{metaDigitise} is at matching clicked points to their true values, we generated four random scatterplots, each with 20 data points, and digitised these with \fct{metaDigitise}.
+This was done by one digitiser (J.L.P.), as there is no detectable between user variation.
+Data digitised using \pkg{metaDigitise} was essentially perfectly correlated with the true simulated data for both the \textit{x}-variable (Pearson's correlation; \textit{r} = 0.9999915, \textit{t} = 2137.4, \textit{df} = 78, $p < 0.001$) and \textit{y}-variable (\textit{r} = 0.9999892, \textit{t} = 1897.8, \textit{df} = 78, $p < 0.001$).
 \section{Comparison with existing software}
@@
 
-\section{Derivation of mean, standard deviation and sample size from different plot types}
-
-\subsection{Mean/Error Plots}
-The standard deviation is calculated depending on the type of error presented.
-The user can choose from standard deviation (SD, $\sigma$), standard error (SE) or 95\% confidence intervals (CI95).
-Standard deviation is calculated from standard error as
-\begin{equation}
-\sigma = SE \sqrt{n}
-\end{equation}
-and from 95\% confidence intervals as
-\begin{equation}
-\sigma = \frac{CI}{1.96} \sqrt{n}
-\end{equation}
-
-\subsection{Box Plots} 
-The mean ($\mu$) and SD are calculated sing the maximum ($b$), upper quartile ($q_3$), median ($m$), lower quartile ($q_1$) and minimum ($a$) as 
-\begin{equation}
-\mu = \frac{(n+3)(a+b) + 2(n-1)(q_1 + m + q_3)}{8n}
-\end{equation}
-following \citet{Bland2015} and
-\begin{equation}
-\sigma = \frac{b-a}{4\Phi^{-1}(\frac{n-0.375}{n+0.25})} + \frac{q_3-q_1}{4\Phi^{-1}(\frac{0.75n-0.125}{n+0.25})}
-\end{equation}
-where $\Phi^{-1}(z)$ is the upper zth percentile of the standard normal distribution, following \citet{Wan2014}.
-
-\subsection{Histograms} 
-For each bar, the user click two point (the top of the bar).
-Using these points, a midpoint (m; mean x coordinates) and a frequency (f; mean y coordinates, rounded to the nearest integer) is calculated for each bar.
-The sample size, mean and SD are calculated as:
-\begin{equation}
-n = \sum_{i=1}^n{f_i}
-\end{equation}
-
-\begin{equation}
-\mu = \frac{\sum_{i=1}^n{m_i f_i}}{n}
-\end{equation}
-
-\begin{equation}
-\sigma = \sqrt{\frac{\sum_{i=1}^n{(m_i f_i - \mu f_i)^2}}{n-1}}
-\end{equation}
-
-\section{Tutorial}
-
 \bibliographystyle{FuncEcol.bst}

### 2018-03-02 R/import_metaDigitise.R
small tweak to import_metaDigitise

@@ -5,6 +5,9 @@
 #' @return Returns a list (summary = FALSE) or data frame (summary = TRUE)
 
 import_menu<-function(dir, summary){
+
+	if(substring(dir, nchar(dir)) != "/") dir <- paste0(dir, "/")
+
 	caldat <- dir_details(dir)
 	filepaths <- caldat$doneCalFiles
 	files <- caldat$calibrations

### 2018-03-02 docs/MEE_submission/manuscript_MEE.tex
MS updated - best word count (main text, legends and refs) 3070

@@
 % remove page numbers
-%\pagenumbering{gobble}
+\pagenumbering{gobble}
 
@@
 
-%% - \Abstract{} almost as usual
+\clearpage
 \section*{Abstract}
@@
 However, exact values of such statistics are commonly hidden in figures.
+
 2.
@@
 Additionally, current software lacks an ability to incorporate important meta-data (e.g., sample sizes, treatment / variable names) about experiments and is not integrated with other software to streamline analysis pipelines.
+
 3.
@@
 Notably, it enables large-scale extraction by automatically loading image files, letting the user stop processing, edit and add to the resulting data fame at any point.
+
 4.
@@
 %% 209 words
+
  Keywords: meta-analysis, comparative analysis, data extraction, \proglang{R}, reproducibility, figures, images, summary statistics
@@
 %%%---------------------------------
+\clearpage
 
@@
 
-In many different contexts, researchers need to make use of data presented in primary literature.
+In many different contexts, researchers make use of data presented in primary literature.
 Most notably, this includes meta-analysis, which is becoming increasingly common in many research fields.
@@
 Meta-analysis relies on descriptive statistics (e.g.
-means, standard deviations, sample sizes, correlation coefficients) extracted from primary literature that have been reported in the text or tables of research papers.
+means, standard deviations (SD), sample sizes, correlation coefficients) extracted from primary literature that have been reported in the text or tables of research papers.
 Descriptive statistics are also, however, frequently presented in figures and so need to be manually extracted using digitising programs.
@@
 
-Although there are several existing tools to perform tasks like this (e.g.
-\proglang{DataThief} \citep{DataThief}, \proglang{GraphClick} \citep{GraphClick}, \proglang{WebPlotDigitizer} \citep{WebPlotDigitizer}), these tools are not appropriate for the needs of meta-analysis for three main reasons.
+Although there are several tools that data extraction from figures (e.g.
+\proglang{DataThief} \citep{DataThief}, \proglang{GraphClick} \citep{GraphClick}, \proglang{WebPlotDigitizer} \citep{WebPlotDigitizer}), these tools do not cater to needs of meta-analysis for three main reasons.
 First, they typically only provide the user with calibrated \textit{x,y} coordinates from imported figures, and do not differentiate between common plot types that are used to present data.
-This means that a large amount of downstream data manipulation is required, that is different across plots types.
-For example, data are frequently presented in mean/error plots (Figure \ref{fig:all_extract}A), for which the user wants a mean and error estimate for each group presented in the figure.
-With existing programs, from \textit{x,y} coordinates users must manually discern between mean and error coordinates and assign points to groups.
-Error then needs to be calculated as the deviation from the mean, and then transformed to a standard deviation, depending on the type of error presented.
-Second, digitising programs do not easily allow the integration of metadata at the time of data extraction, such as experimental group or variable names, and sample sizes.
-This makes the downstream calculations more laborious, as the information has to be added later, in most cases using different software.
-Finally, existing programs do not automatically import new images upon completion to allow the user to systematically work through them.
+Consequently a large amount of downstream data manipulation is required, that is different across plots types.
+For example, data are frequently presented in mean/error plots (Figure \ref{fig:all_extract}A), from which the user wants a mean and SD for each group presented.
+From \textit{x,y} coordinates, users must manually discern between mean and error coordinates and assign points to groups.
+Error then needs to be calculated as the deviation from the mean, and then transformed to SD, according to the type of error presented.
+Second, digitising programs do not allow the integration of metadata at the time of data extraction, such as experimental group or variable names, and sample sizes.
+This makes the downstream calculations laborious, as information has to be added later using different software.
+Finally, existing programs do not import sets of images for the user to systematically work through.
 Instead they require the user to manually import images one by one, and export data into individual files, that need to be imported and edited using different software.
 
-These present major issues because extracting from figures is an incredibly time-consuming process and existing software does not provide an optimized research pipeline to facilitate data extraction and editing.
+Data extraction from figures is therefore an incredibly time-consuming process as existing software does not provide an optimized research pipeline to facilitate data extraction and editing.
 Furthermore, although meta-analysis is an important tool in consolidating the data from multiple studies, many of the processes involved in data extraction are opaque and difficult to reproduce, making extending studies problematic.
@@
 
-%%% This paragraph could be shorter as it is all said in the main text
 Here, we present an interactive \proglang{R} package, \pkg{metaDigitise} (available at https://github.com/daniel1noble/metaDigitise), which is designed for large scale, reproducible data extraction from figures, specifically catering to the the needs of meta-analysts.
-To this end, we provide tools specific to data extraction from common plot types (mean/error plots, box plots, scatter plots and histograms, see Figure \ref{fig:all_extract}).
+To this end, we provide tools to extract data from common plot types (mean/error plots, box plots, scatter plots and histograms, see Figure \ref{fig:all_extract}).
 \pkg{metaDigitise} operates within the \proglang{R} environment making data extraction, analysis and export more streamlined.
@@
 Summary data from multiple figures is returned into a single data frame which can be can easily exported or use in downstream analysis within \proglang{R}.
-Completed digitisations are automatically saved for each figure.
-Users can therefore redraw their digitisations on figures, make corrections and access calibration and proceeded data.
+Completed digitisations are automatically saved for each figure, meaning users can redraw their digitisations on figures, make corrections and access calibration and proceeded data.
 This makes sharing figure digitisation and reproducing the work of others simple and easy, and allows meta-analyses to be updated more efficiently.
 
-%% 647 words
+%% 612 words
 
@@
 \fct{metaDigitise} works on a directory containing images of figures copied from primary literature, in .png, .jpg, .tiff, .pdf format, specified to \fct{metaDigitise} through the \code{dir} argument.
-The user needs to think carefully about their directory structure early on in their project.
+The user should think carefully about their directory structure early on in their project.
 Although different directory structures may be used, we would recommend having all files for one project in a single directory with an informative and unambiguous naming scheme for images to help identify the paper and figure that data come from (e.g.
-paper_figure_trait.png).
+paper\_figure\_trait.png).
 
-When \fct{metaDigitise} is run, it recognizes all the images in a directory and automatically imports them one by one, allowing the user to extract the relevant information about a figure as they go.
-Having all figures in one directory therefore expedites digitising figures by preventing users from having to constantly change directories and / or open new images.
+\fct{metaDigitise} recognizes all the images in a directory and automatically imports them one by one, allowing the user to extract the relevant information about a figure as they go.
+Having all figures in one directory therefore expedites digitisation by preventing users from having to constantly change directories and / or open new images.
 The data from each completed image is automatically saved as a \code{metaDigitise} object in a separate .RDS file to a \code{caldat} directory that is created within the parent directory when first executing \fct{metaDigitise}.
 These files enable re-plotting and editing of images at a later point (see below).
-
-\fct{metaDigitise} identifies images within a directory that have been previously digitised and only import images that have not been digitised in previous calls of the function.
-All figures do not, therefore, need to be extracted at one time and new figures can be added to the directory as the project develops.
-After each image is extracted, the user is asked whether they wish to continue or quit the extraction process.
-Upon rerunning \fct{metaDigitise}, previously digitised figures are ignored during processing, but their data is automatically integrated into the final output.
+When run, \fct{metaDigitise} also identifies the images within a directory that have been previously digitised and only imports new images to process.
+The data of all images is then automatically integrated into the final output.
+This means that all figures do not need to be extracted at one time and new figures can be added to the directory as the project develops.
 
 This directory structure allows the complete digitisation process to be reproduced at a later stage, shared with collaborators and presented as supplementary materials for a publication.
-As long as all the images and the caldat directory are still in the directory, \fct{metaDigitise} will be able to reproduce all figure extractions, regardless of the computer it is run on.
+As long as all the images and the caldat directory are still in one directory, \fct{metaDigitise} will be able to reproduce all figure extractions, regardless of the computer it is run on.
 For an analysis to be updated, new figures can simply be added to the directory and \fct{metaDigitise} run to incorporate the new data.
 
-%% 410 words
+%% 365 words
 %%%---------------------------------
@@
 
-\pkg{metaDigitise} recognises four main types of plot; Mean/error plots, box plots, scatter plots and histograms, shown in Figure \ref{fig:all_extract}.
+\pkg{metaDigitise} recognises four main types of plot; Mean/error plots, box plots, scatter plots and histograms (\ref{fig:all_extract}).
 All plot types can be extracted in a single call of \fct{metaDigitise} and integrated into one output.
@@
 All four plot types are extracted slightly differently (outlined below).
-After completing images, or upon quitting, either summarised or processed data is returned (specified by the user).
-Summarised data consists of a mean, standard deviation and sample size, for each identified group within the plot (should multiple groups exist).
-In the case of scatter plots, the correlation coefficient between the points within each identified group is also returned.
-Alternatively choosing processed data will return a list with slots for each of the four figure types, containing the calibrated points that the user has clicked.
+Upon completing all images, or quitting, either summarised or processed data is returned (specified by the user through the \code{summary} argument).
+Summarised data consists of a mean, SD and sample size, for each identified group within the plot (should multiple groups exist).
+In the case of scatter plots, the correlation coefficient between x and y variables within each identified group is also returned.
+Processed data consists of a list with slots for each of the four figure types, containing the calibrated points that the user has clicked.
 This may be particularly useful in the case of scatter plots.
@@
 For each mean/box, the user is enters group names and sample sizes.
-If the user does not enter a sample size at the time of data extraction (if, for example, the information is not readily available) a standard deviation (SD) is not calculated.
-This can, however, be entered at a later time (see below).
+If the user does not enter a sample size at the time of data extraction (if, for example, the information is not readily available) a SD is not calculated.
+Sample sizes can, however, be entered at a later time (see below).
 For mean/error plots, the user clicks on an error bar and the mean.
@@
 Points are displayed where the user has clicked, with the error in a different colour to the mean (Figure \ref{fig:all_extract}A).
-The user also enters the type of error used in the figure: standard deviation (SD), standard error (SE) or 95\% confidence intervals (CI95).
+The user also enters the type of error used in the figure: SD, standard error (SE) or 95\% confidence intervals (CI95).
 For box plots, the user clicks on the maximum, upper quartile, median, lower quartile and minimum.
 For both plot types, the user can add, edit or remove groups.
-Three functions, \fct{error\_to\_sd}, \fct{rqm\_to\_mean} and \fct{rqm\_to\_sd}, that convert different error types to SD, box plot data to mean and box plot data SD, respectively, are also available in the package (see SM for further details of these conversions).
+Three functions, \fct{error\_to\_sd}, \fct{rqm\_to\_mean} and \fct{rqm\_to\_sd}, that convert different error types to SD, box plot data to mean and box plot data SD, respectively, are also available in the package (see supplements for further details of these conversions).
 
@@
 Data points may overlap with each other making it impossible to know whether points have been missed.
-However, this will result in the sample size of digitised groups conflicting with what is reported in the paper.
+This may result in the sample size of digitised groups conflicting with what is reported in the paper.
 For example, in Figure \ref{fig:all_extract}C only 49 points have been clicked when the sample size is known to be 50.
@@
 \subsection{Histograms}
-The user clicks on the top corners of each bar and alternating colours are used across bars (Figure \ref{fig:all_extract}D).
+The user clicks on the top corners of each bar, which are drawn in alternating colours (Figure \ref{fig:all_extract}D).
 Bars are numbered to allow the the user to edit them.
 As with scatter plots, if the sample size from the extracted data does not match a known sample size, the user can enter an alternate sample size.
-The calculation of mean and SD from this data is shown in the SM.
+The calculation of mean, SD and sample size from this data is shown in the supplements.
 
-%% 777 words
+%% 770 words
 
@@
 \pkg{metaDigitise} is also able to re-import, edit and re-plot previously digitised figures.
-When running \fct{metaDigitise}, the user can choose to `Import existing data', which returns previously digitised data.
-Users can also choose to `Edit existing data' which allows them to re-plot or edit information or digitisations that have previously be done.
-Points added by mistake can be deleted.
-The user can add more groups, edit groups (add or remove points) or delete groups and this is automatically re-incorporated in the data.
+When running \fct{metaDigitise}, the user can choose to `Import existing data', which returns previously digitised data, from sinlge or all figures.
+Alternately, the \fct{import\_metaDigitise} function returns the data of previous digitisations, but without user interaction, allowing easier integration into larger scripts.
+`Edit existing data' allows the user to re-plot or edit information or digitisations that have previously be done.
 
 \subsection{Adding Sample Sizes to Previous Digitisations}
-In many cases important information, such as sample sizes, may not be readily available when digitising figures.
-Such information does not need to be added a the time of digitisation.
+In many cases sample sizes may not be readily available when digitising figures.
+This information does not need to be added a the time of digitisation.
 To expedite finding and adding these sample sizes at a later point, \fct{metaDigitise} has a specific edit option that allows users to enter previously omitted sample sizes.
@@
 \section{Software Validation}
-
-In order to evaluate the consistency of digitisation using \pkg{metaDigitise} between users, we simulated a dataset of two variables with two groups.
-The same simulated datasets were given to 14 different digitisers to compare the inter-observer variability in digitisations.
-We also used simulated data to test the accuracy of digitisations with respect to known values (See SM for more details on how simulations were set up).
-
-Across the plot types we found no evidence for any inter-observer variability in digitisations for the mean (ICC = 0, 95\% CI = 0 to 0.029, \textit{p} = 1), standard deviation (ICC = 0, 95\% CI = 0 to 0.033, \textit{p} = 0.5) or correlation coefficient (ICC = 0.053, 95\% CI = 0 to 0.296, \textit{p} = 0.377).
-There were was also little bias between digitised and true values, on average 1.63\% (mean = 0.02\%, SD = 4.9\%, \textit{r} = -0.03\%) and overall there were only small absolute differences between digitised and true values, deviating, on average 2.18\% (mean = 0.40\%, SD = 5.81\%, \textit{r} = 0.33\%) across all three summary statistics.
-SD estimates from digitisations are clearly more prone to error than means or correlation coefficients.
-If the mean absolute difference is calculated for each plot type, we can see that this effect is driven mainly by extraction from box plots and histograms (\% difference; box plot: 15.805, histogram: 5.210, mean/error: 1.500, scatter plot: 0.433).
+In order to evaluate the consistency of digitisation with \pkg{metaDigitise} between users, we got 14 people to digitise the same set of 14 figures created form a simulated dataset (see supplements).
+We found no evidence for any inter-observer variability in digitisations for the mean (ICC = 0, 95\% CI = 0 to 0.029, \textit{p} = 1), SD (ICC = 0, 95\% CI = 0 to 0.033, \textit{p} = 0.5) or correlation coefficient (ICC = 0.053, 95\% CI = 0 to 0.296, \textit{p} = 0.377).
+There were was little bias between digitised and true values, on average 1.63\% (mean = 0.02\%, SD = 4.9\%, \textit{r} = -0.03\%) and there were small absolute differences between digitised and true values, on average 2.18\% (mean = 0.40\%, SD = 5.81\%, \textit{r} = 0.33\%) across all three summary statistics.
+SD estimates from digitisations are clearly most error prone.
+The mean absolute differences for each plot type clearly show that this effect is driven by extraction from box plots and histograms (\% difference; box plot: 15.805, histogram: 5.210, mean/error: 1.500, scatter plot: 0.433).
 SD estimation from box plot summary statistics is known to be more error prone, especially at small sample sizes \citep{Wan2014}.
 
-\pkg{metaDigitise} was extremely accurate at matching clicked points to their true values essentially being perfectly 
-correlated with the true simulated data for both the \textit{x}-variable (Pearson's correlation; \textit{r} = 0.9999915, \textit{t} = 2137.4, \textit{df} = 78, $p < 0.001$) and \textit{y}-variable (\textit{r} = 0.9999892, \textit{t} = 1897.8, \textit{df} = 78, $p < 0.001$) in scatterplots.
+We also used simulated data to test the accuracy of digitisations with respect to known values (see supplements).
+\pkg{metaDigitise} was extremely accurate at matching clicked points to their true values essentially being perfectly correlated with the true simulated data for both the \textit{x}-variable (Pearson's correlation; \textit{r} = 0.9999915, \textit{t} = 2137.4, \textit{df} = 78, $p < 0.001$) and \textit{y}-variable (\textit{r} = 0.9999892, \textit{t} = 1897.8, \textit{df} = 78, $p < 0.001$) in scatterplots.
 
-%% 571 words
+%% 263 words
 
@@
 
-\section{Limitations and Future Extensions}
+\section{Limitations}
 
-Although \pkg{metaDigitise} is already very flexible, and provides functionality not seen in any other package (see Table S1), it is clear that there are some functions that it does not perform.
+Although \pkg{metaDigitise} is very flexible and provides functionality not seen in any other package, there are some functions that it does not perform (see Table S1).
 Notably \pkg{metaDigitise} lacks automated point detection.
 However, from our experience, manual digitising is more reliable and often equally as fast.
-Given the variation in image quality, calibration for automatic point detection needs to be done for each plot individually.
+Given the variation in image quality, calibration for automatic point detection needs to be done for each figure individually.
 Additionally, auto-detection often misses points which then need to be manually added.
 Based on tests of \pkg{metaDigitise} (see above), figures can be extracted in around 1-2 minutes, including the entry of metadata.
-As a result, we do not believe that current automated point detection techniques provides substantial benefits in terms of time or accuracy.
+As a result, we do not believe that current automated point detection techniques provide substantial benefits in terms of time or accuracy.
 
-\pkg{metaDigitise} also (currently) lacks the ability to zoom in on plots.
+\pkg{metaDigitise} also (currently) lacks the ability to zoom in on figures.
 Zooming may enable users to gain greater accuracy when clicking on points.
-However, from our own experience (and indeed from the results above), if you are using a reasonably sized screen then the accuracy is already high, and so there is not much gain to be had from zooming in on points.
+However, from our own experience (see results above), with a reasonably sized screen accuracy is already high, and so relatively little gain is to be had from zooming in on points.
 
-In contrast to some other packages (Table S1), \pkg{metaDigitise} does not extract lines from figures.
-Line extraction may not particularly useful for most meta-analyses, although we recognise that it may be useful in other fields.
+In contrast to some other packages \pkg{metaDigitise} does not extract lines from figures.
+Line extraction is not particularly useful for most meta-analyses, although we recognise that it may be useful in other fields.
 Should a user like to extract lines with \pkg{metaDigitise}, we would recommend extracting data as a scatter plot, and clicking along the line in question.
@@
 
-%% 271 words
+%% 249 words
 
@@
 
+\clearpage
 \section*{Figures}

### 2018-03-07 docs/MEE_submission/manuscript_MEE.tex
Shinichi's changes

@@
 
+% remove page numbers
+%\pagenumbering{gobble}
+
 %% new custom commands for code formatting
@@
 \let\proglang=\textsf
-
-% remove page numbers
-\pagenumbering{gobble}
 
@@
 Most notably, this includes meta-analysis, which is becoming increasingly common in many research fields.
-Meta-analysis uses effect size estimates and their sampling variance, taken from many studies, to understand whether particular effects are common across studies and to explain variation among these effects \citep{Glass1976,Borenstein2009,Koricheva2013,Naka2017}.
+Meta-analysis uses effect size estimates and their sampling variance, taken from many studies, to understand whether particular effects are common across studies and to explain variation among these effects \citep{Glass1976,Koricheva2013,Naka2017}.
 Meta-analysis relies on descriptive statistics (e.g.
@@
 Furthermore, although meta-analysis is an important tool in consolidating the data from multiple studies, many of the processes involved in data extraction are opaque and difficult to reproduce, making extending studies problematic.
-Having a tool that facilitates reproducibility in meta-analyses will increase transparency and aid in resolving the reproducibility crises seen in many fields \citep{peng_reproducible_2006, peng_reproducible_2011, sandve_ten_2013, Parker2016,Ihle2017}.
+Having a tool that facilitates reproducibility in meta-analyses will increase transparency and aid in resolving the reproducibility crises seen in many fields \citep{peng_reproducible_2006, peng_reproducible_2011, Parker2016}.
 

### 2018-03-07 docs/MEE_submission/supplements.tex
supplements updates

@@
 
-\usepackage{amssymb,rotating,natbib,graphicx,fancyvrb}
+\usepackage{amssymb,rotating,natbib,graphicx,fancyvrb,booktabs}
 
@@
 \usepackage[utf8]{inputenc}
+
+\usepackage[colorlinks=true,linkcolor=blue, citecolor=blue, urlcolor=blue, linktoc=page]{hyperref}
 
@@
 \renewcommand{\thetable}{S\arabic{table}}%
+\renewcommand{\thefigure}{S\arabic{figure}}%
 \renewcommand{\theequation}{S\arabic{equation}}%
@@
 
+\tableofcontents
+
+\clearpage
+
 \section{Tutorial}
@@
 
-We will now demonstrate how \fct{metaDigitise} works using figures generated from the well known iris data set. Users can install the \pkg{metaDigitise} package from GitHub as follows:
+We can demonstrate how \fct{metaDigitise} works using figures generated from the well known iris data set. \pkg{metaDigitise} can installed from GitHub as follows:
 
@@
 
-Assume that the user would like to extract descriptive statistics from studies measuring sepal length or width in iris species for a fictitious project.
+Assume you would like to extract descriptive statistics from studies measuring sepal length or width in iris species for a fictitious project.
 There are a few studies that only present these data in figures.
-As the user reads papers found from a systematic search, they add figures with relevant data to a "FiguresToExtract" folder as follows
+As you read papers found from a systematic search, you add figures with relevant data to a "FiguresToExtract" folder as follows
 
@@
 At first there is one figure in the folder, shown in Figure \ref{fig:Anderson}.
-Running \fct{metaDigitise} brings up a series of prompts for the user using a main menu that provides access to a number of its features ("..." here represents the user's path to the project directory):
+Running \fct{metaDigitise} brings up a series of prompts using a main menu that provides access to a number of its features ("..." here represents the user's path to the project directory):
 
@@
 
-The user simply enters in the numeric value that corresponds to what they would like to do.
-In this case they want to "Process new images".
-The user is then asked whether there are different types of plot(s) in the folder.
-This question is most relevant when there are lots of different figures in the folder because it will then ask the user for the type of figure as they are cycled through.
+You then simply enter in the numeric value that corresponds to what you would like to do.
+In this case we want to "Process new images".
+You are then asked whether there are different types of plot(s) in the folder.
+This question is most relevant when there are lots of different figures in the folder because it will then ask for the type of figure as they are cycled through.
 
@@
 
-\fct{metaDigitise} then asks the user whether the figure needs to be rotated or flipped.
+\fct{metaDigitise} then asks whether the figure needs to be rotated or flipped.
 This can be needed when box plots and mean and error plots are not orientated correctly.
 In some cases, older papers can give slightly off angled images which can be corrected by rotating.
-So, in this prompt the user has three options: \code{f} for ``Flip'', \code{r} for ``rotate'' or \code{c} for ``continue''.
+In this prompt the user has three options: \code{f} for ``Flip'', \code{r} for ``rotate'' or \code{c} for ``continue''.
 
@@
 
-After this, \fct{metaDigitise} will ask the user to specify the plot type. Depending on the figure, the user can specify that it is a figure containing the mean and error (\code{m}), a box plot (\code
-{b}), a scatter plot (\code{s}) or a histogram (\code{h}).
-If the user has specified \code{d} instead of \code{s} in response to the question about whether the plot types are the same or different, this question will pop up for each plot, but will only be asked once if plots are all the same.
+After this, \fct{metaDigitise} will ask you to specify the plot type.
+Depending on the figure, you can specify that it is a figure containing the mean and error (\code{m}), a box plot (\code{b}), a scatter plot (\code{s}) or a histogram (\code{h}).
+If you has specified \code{d} instead of \code{s} in response to the question about whether the plot types are the same or different, this question will pop up for each plot, but will only be asked once if plots are all the same.
 
@@
 
-After selecting the figure type a new set of prompts will come up that will ask the user first what the y and x-axis variables are.
+After selecting the figure type a new set of prompts will come up that will ask what the y and x-axis variables are.
 This is useful as users can keep track of the different variables across figures and papers.
-Here, the user can just add this information in to the \proglang{R} console.
+Here, you can just add this information in to the \proglang{R} console.
 Once complete, details on how to calibrate the x and y-axis appear, so that the relevant statistics / data can be correctly calculated.
-When working with a plot of mean and standard errors, the x-axis is rather useless in terms of calibration so \fct{metaDigitise} just asks the user to calibrate the y-axis.
+When working with a plot of mean and errors, the x-axis is rather useless in terms of calibration so \fct{metaDigitise} just asks you to calibrate the y-axis.
 
@@
 
-The user can just follow the instructions on screen step-by-step (instructions above have been truncated by `...' to simplify), and in the order specified.
-Before moving on, the user is forced to check whether or not the calibration has been set up correctly.
-If \code{n} is chosen because something needs to be fixed then the user can re-calibrate.
+You can just follow the instructions on screen step-by-step (instructions above have been truncated by `...' to simplify), and in the order specified.
+Before moving on, you are forced to check whether or not the calibration has been set up correctly.
+If \code{n} is chosen because something needs to be fixed then you can re-calibrate.
 
@@
 
-Often, plots might contain multiple groups that the meta-analyst wants to extract from.
-\fct{metaDigitise} handles this nicely by prompting the user to enter the group first, followed by digitisation of this groups data.
-After digitising the first group, and having exited, \fct{metaDigitise} will ask the user whether they would like to add another group.
-Users can continually add groups (\code{a}), delete groups (\code{d}), edit groups (\code{e}) or finish a plot and continue to the next one (\code{f} - if another plot exists).
+Often, plots might contain multiple groups that a meta-analyst wants to extract from.
+\fct{metaDigitise} handles this nicely by prompting you to enter the group first, followed by digitisation of this groups data.
+After digitising the first group, \fct{metaDigitise} will ask whether you would like to add another group.
+You can continually add groups (\code{a}), delete groups (\code{d}), edit groups (\code{e}) or finish a plot and continue to the next one (\code{f} - if another plot exists).
 %%% This needs checking - think it is now different
-The number of groups are not really limited and users can just keep adding in groups to accommodate the different numbers that may be presented across figures (although it can get complicated with too many).
+The number of groups are not really limited and you can just keep adding in groups to accommodate the different numbers that may be presented across figures (although it can get complicated with too many).
 
@@
 
-To finish selecting points, the user can exit by clicking on the red button that appears when extracting points. The user is then asked if they want to add or delete points from that group.
+To finish selecting points, you can exit by clicking on the red button that appears when extracting points. You are then asked if you want to add or delete points from that group.
 
@@
 
-Users can access the \code{metaDigitise} object created (001\_Anderson\_1935\_Fig1) at any time using the \fct{metaDigitse} function.
+You can access the \code{metaDigitise} object created (001\_Anderson\_1935\_Fig1) at any time using the \fct{metaDigitse} function.
 In the \proglang{R} console, the summarised data for the digitised figure can be printed on screen or even written to a .csv file:
@@
 However, a meta-analyst will probably realise that the sample sizes here conflict with what is reported in the paper.
-Hence, \pkg{metaDigitise} also provides the user with  options to input the sample sizes directly (see Editing section below), even for scatter plots and histograms where, strictly speaking, this should not be necessary.
-Nonetheless, it is important to recognise the impact that overlapping points can have on summary statistics, particularly its effects on standard deviation (SD) and standard error (SE).
+Hence, \pkg{metaDigitise} also provides you with  options to input the sample sizes directly, even for scatter plots and histograms where, strictly speaking, this should not be necessary.
+Nonetheless, it is important to recognise the impact that overlapping points can have on summary statistics, particularly its effects on sampling variance.
 Here, the mean point estimates are nearly exactly the same as the true values, but the SD's are slightly over-estimated:
@@
 
-Users can add additional figures as new papers with relevant information are found.
+You can add additional figures to the directory as new papers with relevant information are found.
 Each figure should be in its own file with unique naming, even if a single paper has multiple figures for extraction.
@@
 
-The user has already processed one figure (001\_Anderson\_1935\_Fig1.png).
+We have already processed one figure (001\_Anderson\_1935\_Fig1.png).
 We can tell this because the caldat folder has digitised data in it (caldat/001\_Anderson\_1935\_Fig1).
-Now the user has two new figures that have not yet been digitised.
-This example will nicely demonstrate how users can easily pick up from where they left off and how all previous data gets re-integrated.
+Now there are two new figures that have not yet been digitised.
+This example will also nicely demonstrate how you can easily pick up from where you left off and how previous data gets re-integrated.
 It will also demonstrate how different plot types are handled.
@@
 
-The user gets the same set of prompts and simply chooses option one.
-This will permit users to digitise new figures, and will integrate previously completed digitisations along with newly digitised data together at the end of the session, or when the user decides to quit.
+You get the same set of prompts and simply chooses option one.
+This will permit users to digitise new figures, and will integrate previously completed digitisations along with newly digitised data together at the end of the session, or when the you decides to quit.
 This time, 001\_Anderson\_1935\_Fig1.png is ignored and the new plots cycle on screen; first 002\_Doe\_2013\_Fig1.png and then 002\_Doe\_2013\_Fig3.png.
-Since there are a few different figure types, the user answers the first question in the \proglang{R} console as "d":
+Since there are a few different figure types, you answer the first question in the \proglang{R} console as "d":
 
@@
 
-Here, the user specifies the new plot type as \code{m} for 002\_Doe\_2013\_Fig1.png because the user has a plot of the mean and error of sepal length for each of the three species.
+Here, you specify the new plot type as \code{m} for 002\_Doe\_2013\_Fig1.png because this a plot of the mean and error of sepal length for each of the three species.
 The user is then prompted a bit differently from our scatter plot as the x-axis is not needed for calibration:
@@
 
-Again, \fct{metaDigitise} will simply guide the user through digitising each of these figures describing to them exactly what needs to be done.
-At any point if mistakes are made the user can choose relevant options to edit or correct things before ending the figure.
-This process continues for each plot so long as the user would like to continue and after completing a single plot the user is always prompted as follows:
+Again, \fct{metaDigitise} will simply guide you through digitising each of these figures describing to you exactly what needs to be done.
+At any point if mistakes are made you can choose relevant options to edit or correct things before ending the figure.
+This process continues for each plot so long as the you would like to continue and after completing a single plot you are always prompted as follows:
 
@@
 
-This continues until users have completed all non-digitised figures in the folder, at which point \fct{metaDigitise} concatenates the new data with previously digitised data in the object:
+This continues until you have completed all non-digitised figures in the folder, at which point \fct{metaDigitise} concatenates the new data with previously digitised data in the object:
 
@@
 
-If the user chooses "Import existing data", they have the option of either 1) importing data from all digitised images or 2) importing data from a single image that has been digitised.
-If 2, then a list of files are provided to the user that they can select.
-Editing existing data allows users to easily re-plot or edit information or digitisations that have previously be done for any plot.
-This is accomplished by guiding the user through a new set of options:
+If we choose "Import existing data", you have the option of either 1) importing data from all digitised images or 2) importing data from a single image that has been digitised.
+If 2, then a list of files are provided for you to select.
+Editing existing data allows you to easily re-plot or edit information or digitisations that have previously be done for any plot.
+This is accomplished by guiding you through a new set of options:
 
@@
 
-If the user is unsure about the name of the specific figure they need to edit or simply want to just check the digitisations of figures they can choose "Cycle through images", which will bring up each figure, one by one, overlaying the calibrations, group names (if they exist), sample sizes (if they were entered) and the selected points.
+If you is unsure about the name of the specific figure you need to edit or simply want to just check the digitisations of figures they can choose "Cycle through images", which will bring up each figure, one by one, overlaying the calibrations, group names (if they exist), sample sizes (if they were entered) and the selected points.
 The user will then be given the choice to edit individual images.
 Alternatively, choosing option 2, will bring up a list of the completed files in the folder and the specific file can be chosen, at which point it will be replotted.
-Either of these options will cycle through a number of questions asking the user what they would like to edit:
+Either of these options will cycle through a number of questions asking you what they would like to edit:
 
@@
 Variable entered as: 
-\end{CodeOutput}
-\begin{CodeInput}
-R> Sepal length
-\end{CodeInput}
-\begin{CodeOutput}
+Sepal length
 Rename Variables (y/n) 
@@
 
-While \fct{metaDigitise} provides users with the summary statistics by default, for all plot types, in many cases the user may actually be interested in obtaining the processed digitised data from scatter plots (i.e.
+While \fct{metaDigitise} provides you with the summary statistics by default, for all plot types, in many cases the user may actually be interested in obtaining the processed digitised data from scatter plots (i.e.
 calibrated points).
-This is very easy to do my changing the default \code{summary} argument from TRUE to FALSE in \fct{metaDigitise}.
-Instead of providing the user with summary statistics it will return a list containing four slots for each of the figure types (mean error, box plot, histogram and scatter plots).
+This is very easy to do by changing the default \code{summary} argument from TRUE to FALSE in \fct{metaDigitise}.
+Instead of providing you with summary statistics it will return a list containing four slots for each of the figure types (mean error, box plot, histogram and scatter plots).
 An example of a data object returned from digitising figures is as follows:
@@
 
+Additionally there is another function, \fct{import\_metaDigitise}, that imports all digitised data from a directory, as either summarised or processed data. For example,
+
+\begin{CodeChunk}
+\begin{CodeInput}
+>R import_metaDigitise(".../FiguresToExtract", summary=TRUE)
+\end{CodeInput}
+{\scriptsize
+\begin{CodeOutput}
+                  filename    group_id         variable   mean  error error_type n    r    sd   plot_type
+001_Anderson_1935_Fig1.png      setosa  Sepal width (mm)  3.42  0.40  sd        39  0.75  0.40  scatterplot
+001_Anderson_1935_Fig1.png      setosa  Sepal length (mm) 5.00  0.38  sd        39  0.75  0.38  scatterplot
+001_Anderson_1935_Fig1.png  versicolor  Sepal width (mm)  2.77  0.32  sd        44  0.52  0.32  scatterplot
+001_Anderson_1935_Fig1.png  versicolor  Sepal length (mm) 5.95  0.53  sd        44  0.52  0.53  scatterplot
+     002_Doe_2013_Fig1.png      setosa  Sepal length      5.00  0.11  se        50  NA    0.78  mean_error
+     002_Doe_2013_Fig1.png  viriginica  Sepal length      6.59  0.18  se        50  NA    1.26  mean_error
+     002_Doe_2013_Fig1.png  versicolor  Sepal length      5.94  0.14  se        50  NA    1.01  mean_error
+     003_Doe_2013_Fig3.png      catana  Sepal length      4.95  0.36  sd        50  NA    0.36  histogram
+\end{CodeOutput}
+}
+\end{CodeChunk}
+
+This function is useful in scripts the extracted data simply needs to be imported, and where the interactive element of \fct{metaDigitise} becomes problematic.
+
 \subsection{Adding sample sizes to previous Digitisations}
@@
 In many cases important information, such as sample sizes, may not be readily available or clear when digitising figures.
-In these circumstances users will have answered `no' to the question about whether they have sample sizes or not while digitising.
+In these circumstances you would have answered `no' to the question about whether you have sample sizes or not while digitising.
 To expedite finding and adding in these sample sizes to do the necessary calculations (if for example a figure presented 95\% CI's or standard errors), \fct{metaDigitise} has s specific edit option that allows users to enter in previously omitted sample sizes.
-It works by first identifying the missing sample sizes in the digitised output, re-plotting the relevant figure and then prompting the user to enter the sample sizes for the relevant groups in the figure, one by one.
+It works by first identifying the missing sample sizes in the digitised output, re-plotting the relevant figure and then prompting you to enter the sample sizes for the relevant groups in the figure, one by one.
 As an example, assume that we were missing sample sizes for two groups in 002\_Doe\_2013\_Fig1.png:
@@
 Here, we can see that we are missing the sample sizes for setosa and viriginica, and as a result, sd is not calculated because \fct{metaDigitise} needs this information to make the calculation.
-If the user found this information after contacting the authors for clarification then they can add these back in as follows:
+If, for examnple, you found this information after contacting the authors for clarification then you can add these in as follows:
 
@@
 
-\fct{metaDigitise} will replot the figure after this and list, only the groups missing data, for which the user can then update the data.
+\fct{metaDigitise} will replot the figure after this and list, only the groups missing data, for which you can then update the data.
 This is then re-integrated back into the data automatically and the sd calculated.
@@
 
-\subsection{Figure Rotation and Adjustment}
-Figures may have been extracted from old publications, for example from scanned images, and so are not perfectly orientated on the image.
-This will make the calibration of the points in the figure from the image problematic.
-\fct{metaDigitise} allows users to rotate the image.
-By clicking two points on the x-axis, metaDigitse calculates the angle needed to rotate the image so the x-axis is horizontal, and rotates it.
-(Figure \ref{fig:rotate}A,B)
-
-\begin{figure}[!b] 
-\centering
- \includegraphics[width=0.75\textwidth]{fig_rotate.pdf} 
- \caption{Figure rotation. A) and B) show how non-aligned images can be realigned through user defined rotation. C) and D) show how figures can be re-orientated so as to aid data input.}
-\label{fig:rotate}
-\end{figure}
-
-Furthermore, some figures, including mean and error, boxplots or histograms, may be presented with horizontal bars.
-\fct{metaDigitise} assumes that the bars are vertical, but allows the user to flip the image so that the bars are vertical if provided horizontally (Figure \ref{fig:rotate}C,D).
+\clearpage
 
@@
 
-\section{Software Validation: Details of Simulation}
+\clearpage
+
+\section{Software Validation: Details of Simulations}
 
 \subsection{Inter-observer variability in digitisations}
-In order to evaluate the consistency of digitisation using \pkg{metaDigitise} between users, we simulated a dataset of two variables with two groups($n=10$ within groups).
+In order to evaluate the consistency of digitisation using \pkg{metaDigitise} between users, we simulated a dataset of two variables with two groups ($n=10$ within groups).
 Each variable was plotted twice for each plot type (figures were modified slightly to give users a sense that they were digitising new data) generating a total of 14 figures.
@@
 
-Across plot types we found no evidence for any inter-observer variability in digitisations for the mean (ICC = 0, 95\% CI = 0 to 0.029, \textit{p} = 1), standard deviation (ICC = 0, 95\% CI = 0 to 0.033, \textit{p} = 0.5) or correlation coefficient (ICC = 0.053, 95\% CI = 0 to 0.296, \textit{p} = 0.377).
-There were was little bias between digitised and true values, on average 1.63\% (mean = 0.02\%, SD = 4.9\%, \textit{r} = -0.03\%) and overall there were only small absolute differences between digitised and true values, deviating, on average 2.18\% (mean = 0.40\%, SD = 5.81\%, \textit{r} = 0.33\%) across all three summary statistics.
-SD estimates from digitisations are clearly more prone to error than means or correlation coefficients.
-If the mean absolute difference is calculated for each plot type, we can see that this effect is driven mainly by extraction from box plots and histograms (\% difference; box plot: 15.805, histogram: 5.210, mean/error: 1.500, scatter plot: 0.433).
-SD estimation from box plot summary statistics is known to be more error prone, especially at small sample sizes \citep{Wan2014}.
-
 \subsection{Testing the accuracy of digitisations}
@@
 This was done by one digitiser (J.L.P.), as there is no detectable between user variation.
-Data digitised using \pkg{metaDigitise} was essentially perfectly correlated with the true simulated data for both the \textit{x}-variable (Pearson's correlation; \textit{r} = 0.9999915, \textit{t} = 2137.4, \textit{df} = 78, $p < 0.001$) and \textit{y}-variable (\textit{r} = 0.9999892, \textit{t} = 1897.8, \textit{df} = 78, $p < 0.001$).
+
+\clearpage
+
 \section{Comparison with existing software}
-\begin{sidewaystable}[ph!]
-{\centering
+\begin{table}[h!]
+\centering
+\rotatebox{90}{%
+\begin{minipage}{1.35\textwidth}
+\small
 \begin{tabular}{lccccccc}
@@
 \end{tabular}
-}
-{\footnotesize
 \\
@@
 \\$^{11}$ Standalone software.
+\caption{\label{tab:comparison} Comparison of functionality between different digitisation softwares.}
+\end{minipage}
 }
 
-\caption{\label{tab:comparison} Comparison of functionality between different digitisation softwares.}
-
-\end{sidewaystable}
+\end{table}
 

### 2018-03-08 docs/MEE_submission/manuscript_MEE.tex
final manuscript for submission to MEE

@@
 \section*{Abstract}
-1.
-Research synthesis requires data extraction from primary studies with effect sizes for meta-analyses being calculated from summary statistics.
+\begin{enumerate} 
+\item Research synthesis, such as meta-analysis requires the extraction of effect sizes from primary literature.
+Such effect sizes are calculated from summary statistics.
 However, exact values of such statistics are commonly hidden in figures.
 
-2.
-Extracting summary statistics from figures can be a slow process that is not easily reproducible.
+\item Extracting summary statistics from figures can be a slow process that is not easily reproducible.
 Additionally, current software lacks an ability to incorporate important meta-data (e.g., sample sizes, treatment / variable names) about experiments and is not integrated with other software to streamline analysis pipelines.
 
-3.
-Here we present the R package \pkg{metaDigitise} which extracts descriptive statistics such as means, standard deviations and correlations from the four plot types: 1) mean/error plots (e.g.
+\item Here we present the R package \pkg{metaDigitise} which extracts descriptive statistics such as means, standard deviations and correlations from the four plot types: 1) mean/error plots (e.g.
 bar graphs with standard errors), 2) box plots, 3) scatter plots and 4) histograms.
@@
 
-4.
-Digitised data can be easily re-plotted and checked, facilitating reproducible data extraction from plots with little inter-observer bias.
+\item Digitised data can be easily re-plotted and checked, facilitating reproducible data extraction from plots with little inter-observer bias.
 We hope that by making the process of figure extraction more flexible and easy to conduct it will improve the transparency and quality of meta-analyses in the future.
-%% 209 words
+\end{enumerate}
 
- Keywords: meta-analysis, comparative analysis, data extraction, \proglang{R}, reproducibility, figures, images, summary statistics
+\vskip10pt
+ \textbf{Keywords:} meta-analysis, comparative analysis, data extraction, \proglang{R}, reproducibility, figures, images, summary statistics
 
@@
 \pkg{metaDigitise} operates within the \proglang{R} environment making data extraction, analysis and export more streamlined.
-The necessary calculations are carried out on processed data immediately after extraction so that comparable summary statistics can be obtained quickly.
+The necessary calculations are carried out on calibrated data immediately after extraction so that comparable summary statistics can be obtained quickly.
 Summary data from multiple figures is returned into a single data frame which can be can easily exported or use in downstream analysis within \proglang{R}.
@@
 \fct{metaDigitise} then calculates the value of any clicked points in terms of the figure axes.
+This is based on the calibration used in the \pkg{digitize} R package \citep{Poisot2011}.
 For mean/error and box plots, only the y-axis is calibrated (Figure \ref{fig:all_extract}A,B), assuming the x-axis is redundant.
@@
 All four plot types are extracted slightly differently (outlined below).
-Upon completing all images, or quitting, either summarised or processed data is returned (specified by the user through the \code{summary} argument).
+Upon completing all images, or quitting, either summarised or calibrated data is returned (specified by the user through the \code{summary} argument).
 Summarised data consists of a mean, SD and sample size, for each identified group within the plot (should multiple groups exist).
 In the case of scatter plots, the correlation coefficient between x and y variables within each identified group is also returned.
-Processed data consists of a list with slots for each of the four figure types, containing the calibrated points that the user has clicked.
+Calibrated data consists of a list with slots for each of the four figure types, containing the calibrated points that the user has clicked.
 This may be particularly useful in the case of scatter plots.
@@
 When running \fct{metaDigitise}, the user can choose to `Import existing data', which returns previously digitised data, from sinlge or all figures.
-Alternately, the \fct{import\_metaDigitise} function returns the data of previous digitisations, but without user interaction, allowing easier integration into larger scripts.
+Alternately, the \fct{getExtracted} function returns the data of previous digitisations, but without user interaction, allowing easier integration into larger scripts.
 `Edit existing data' allows the user to re-plot or edit information or digitisations that have previously be done.
@@
 Should a user like to extract lines with \pkg{metaDigitise}, we would recommend extracting data as a scatter plot, and clicking along the line in question.
-A model can then be fitted to these points (accessed by choosing to return processed rather than summary data) to estimate the parameters needed.
+A model can then be fitted to these points (accessed by choosing to return calibrated rather than summary data) to estimate the parameters needed.
 

### 2018-03-08 docs/MEE_submission/supplements.tex
final manuscript for submission to MEE

@@
 
-\subsection{Obtaining Processed Data}
+\subsection{Obtaining Calibrated Data}
 
-While \fct{metaDigitise} provides you with the summary statistics by default, for all plot types, in many cases the user may actually be interested in obtaining the processed digitised data from scatter plots (i.e.
-calibrated points).
+While \fct{metaDigitise} provides you with the summary statistics by default, for all plot types, in many cases the user may actually be interested in obtaining the calibrated data from scatter plots (i.e.
+all calibrated points).
 This is very easy to do by changing the default \code{summary} argument from TRUE to FALSE in \fct{metaDigitise}.
@@
 
-Here, the user can easily access the list of processed scatter plot data by simply extracting the scatter plot slot:
+Here, the user can easily access the list of calibrated scatter plot data by simply extracting the scatter plot slot:
 
@@
 
-Additionally there is another function, \fct{import\_metaDigitise}, that imports all digitised data from a directory, as either summarised or processed data. For example,
+Additionally there is another function, \fct{getExtracted}, that imports all digitised data from a directory, as either summarised or calibrated data. For example,
 
@@
 \begin{CodeInput}
->R import_metaDigitise(".../FiguresToExtract", summary=TRUE)
+>R getExtracted(".../FiguresToExtract", summary=TRUE)
 \end{CodeInput}

### 2018-03-08 R/calibrate_coords.R
added getExtracted

@@ -1,7 +1,7 @@
 #' @title print_cal_instructions
 #' @param plot_type plot type
 #' @param ... further arguments passed to or from other methods.
-#' @description Prints instructions for calibration
+#' @description Prints instructions for calibration. Modified from the digitize package
 
 print_cal_instructions <- function(plot_type,...){
   cat( "\nOn the Figure, click IN ORDER: 

### 2018-03-08 R/import_metaDigitise.R

@@ -1,12 +1,30 @@
+#' @title getExtracted
+#' @description Extracts data from a directory that has been previous digitised using metaDigitise
+#' @param dir The directory where figures have already been digitised. There
+#' @param summary Logical indicating whether summarised (default) or calibrated data should be returned. 
+#' @return Returns a data frame (summary = TRUE) or a list with slots for each plot type (summary = FALSE)
+#' @export
+
+getExtracted <- function(dir, summary=TRUE){
+
+	if(substring(dir, nchar(dir)) != "/") dir <- paste0(dir, "/")
+
+	if(length(dir_details(dir)$doneCalFiles) == 0) stop("No digitised files to import!", call. = FALSE)
+
+	import <- import_metaDigitise(dir=dir, summary=summary)
+	
+	return(import)
+}
+
+
+
 #' @title import_metaDigitise
 #' @description Imports metaDigitise calibration files from a directory that is partially or fully digitised already
 #' @param dir The directory where figures have already been digitised
-#' @param summary Logical indicating whether the imported data should be returned in 
+#' @param summary Logical indicating whether the imported data should be returned in summarised or processed form.
 #' @return Returns a list (summary = FALSE) or data frame (summary = TRUE)
 
 import_menu<-function(dir, summary){
-
-	if(substring(dir, nchar(dir)) != "/") dir <- paste0(dir, "/")
 
 	caldat <- dir_details(dir)
 	filepaths <- caldat$doneCalFiles
@@ -43,7 +61,7 @@ import_menu<-function(dir, summary){
 #' @author Daniel Noble - daniel.wa.noble@gmail.com
 #' @export
 
-import_metaDigitise <- function(dir, summary = TRUE ) {
+import_metaDigitise <- function(dir, summary) {
 	
 	# Obtain directory details
 	    details <- dir_details(dir)

### 2018-03-08 tests/testthat/test-import_metaDigitise.R

@@ -35,16 +35,17 @@ mock_object2 <- list(
 class(mock_object2) <- 'metaDigitise'
 
 list <- list(mock_object1=mock_object1, mock_object2=mock_object2)
-tmp_dir <- paste0(tempdir(), "/")
-dir.create(paste0(tmp_dir, "/caldat/"))
+tmp_dir_no_slash <- tempdir()
+tmp_dir <- paste0(tmp_dir_no_slash, "/")
+dir.create(paste0(tmp_dir, "caldat/"))
 
-list.files(paste0(tmp_dir, "/caldat/"))
+list.files(paste0(tmp_dir, "caldat/"))
 list.files(tmp_dir)
 
-saveRDS(mock_object1, file = paste0(tmp_dir, "/caldat/", "mock_object1"))
-saveRDS(mock_object2, file = paste0(tmp_dir, "/caldat/", "mock_object2"))
+saveRDS(mock_object1, file = paste0(tmp_dir, "caldat/", "mock_object1"))
+saveRDS(mock_object2, file = paste0(tmp_dir, "caldat/", "mock_object2"))
 
-doneCalFiles <- paste0(tmp_dir, "/caldat/", list.files(paste0(tmp_dir, "/caldat/")))
+doneCalFiles <- paste0(tmp_dir, "caldat/", list.files(paste0(tmp_dir, "/caldat/")))
 names <- list("mock_object1", "mock_object2")
 	
 
@@ -56,10 +57,34 @@ test_that("Checking load_metaDigitise works as expected..", {
 })
 
 test_that("Checking import_metaDigitise works as expected..", {
-	summaries <- import_metaDigitise(tmp_dir)
+	summaries <- import_metaDigitise(tmp_dir, summary=TRUE)
 	expect_equal(summaries$mean[1], 1.5, info="Object 1 loaded and matches mean after import_metaDigitis..")
 	expect_equal(summaries$mean[2], 4, info="Object 2 loaded and matches mean after import_metaDigitis..")
 })
+
+
+
+
+test_that("Checking getExtracted works as expected..", {
+	
+	expect_equal(getExtracted(tmp_dir, summary=TRUE), 
+		rbind(summary(mock_object1),summary(mock_object2)), 
+		info="Problem with summary=TRUE")
+	
+	expect_equal(getExtracted(tmp_dir_no_slash, summary=TRUE), 
+		rbind(summary(mock_object1),summary(mock_object2)), 
+		info="Problem with directory input")
+	
[…]

### 2018-06-04 docs/MEE_submission/manuscript_MEE.tex
Some small updates to MS abse don reviewers and change to ignore

@@
 \section{Software Validation}
-In order to evaluate the consistency of digitisation with \pkg{metaDigitise} between users, we got 14 people to digitise the same set of 14 figures created form a simulated dataset (see supplements).
+In order to evaluate the consistency of digitisation with \pkg{metaDigitise} between users, we got 14 people to digitise the same set of 14 figures created from a simulated dataset (see supplements).
 We found no evidence for any inter-observer variability in digitisations for the mean (ICC = 0, 95\% CI = 0 to 0.029, \textit{p} = 1), SD (ICC = 0, 95\% CI = 0 to 0.033, \textit{p} = 0.5) or correlation coefficient (ICC = 0.053, 95\% CI = 0 to 0.296, \textit{p} = 0.377).
-There were was little bias between digitised and true values, on average 1.63\% (mean = 0.02\%, SD = 4.9\%, \textit{r} = -0.03\%) and there were small absolute differences between digitised and true values, on average 2.18\% (mean = 0.40\%, SD = 5.81\%, \textit{r} = 0.33\%) across all three summary statistics.
+There was little bias between digitised and true values, on average 1.63\% (mean = 0.02\%, SD = 4.9\%, \textit{r} = -0.03\%) and there were small absolute differences between digitised and true values, on average 2.18\% (mean = 0.40\%, SD = 5.81\%, \textit{r} = 0.33\%) across all three summary statistics.
 SD estimates from digitisations are clearly most error prone.
@@
 We also used simulated data to test the accuracy of digitisations with respect to known values (see supplements).
-\pkg{metaDigitise} was extremely accurate at matching clicked points to their true values essentially being perfectly correlated with the true simulated data for both the \textit{x}-variable (Pearson's correlation; \textit{r} = 0.9999915, \textit{t} = 2137.4, \textit{df} = 78, $p < 0.001$) and \textit{y}-variable (\textit{r} = 0.9999892, \textit{t} = 1897.8, \textit{df} = 78, $p < 0.001$) in scatterplots.
+\pkg{metaDigitise} was extremely accurate at matching clicked points to their true values essentially being perfectly correlated with the true simulated data for both the \textit{x}-variable (Pearson's correlation; \textit{r} > 0.999, \textit{t} = 2137.4, \textit{df} = 78, $p < 0.001$) and \textit{y}-variable (\textit{r} > 0.999, \textit{t} = 1897.8, \textit{df} = 78, $p < 0.001$) in scatterplots.
 

### 2018-06-18 docs/MEE_submission/manuscript_MEE.tex
MS revisions

@@
 \pkg{metaDigitise} is user-friendly and easy to learn as it interactively guides the user through the data extraction process.
-Notably, it enables large-scale extraction by automatically loading image files, letting the user stop processing, edit and add to the resulting data fame at any point.
+Notably, it enables large-scale extraction by automatically loading image files, letting the user stop processing, edit and add to the resulting data-frame at any point.
 
@@
 
+%%Most notably, this includes meta-analysis, which is increasingly common in ecology and evolution.
+Meta-analysis uses effect size estimates and their sampling variance, taken from many studies, to understand whether particular effects are common across studies and to explain variation among these effects \citep{Glass1976,Koricheva2013,Naka2017}.
+
+%% While inferential statistics (e.g., \textit{t}- and \textit{F}-statistics) are often presented along side descriptive statistics, and can be used to derive effect sizes, descriptive statistics are much more appropriate to use because sources of non-independence in experimental designs can be dealt with more easily \citep{Noble2017}.
+
 In many different contexts, researchers make use of data presented in primary literature.
-Most notably, this includes meta-analysis, which is becoming increasingly common in many research fields.
-Meta-analysis uses effect size estimates and their sampling variance, taken from many studies, to understand whether particular effects are common across studies and to explain variation among these effects \citep{Glass1976,Koricheva2013,Naka2017}.
-Meta-analysis relies on descriptive statistics (e.g.
-means, standard deviations (SD), sample sizes, correlation coefficients) extracted from primary literature that have been reported in the text or tables of research papers.
-Descriptive statistics are also, however, frequently presented in figures and so need to be manually extracted using digitising programs.
-While inferential statistics (e.g., \textit{t}- and \textit{F}-statistics) are often presented along side descriptive statistics, and can be used to derive effect sizes, descriptive statistics are much more appropriate to use because sources of non-independence in experimental designs can be dealt with more easily \citep{Noble2017}.
+In the ecology and evolution, this most notably includes comparative and meta-analyses.
+These techniques rely on descriptive statistics (e.g.
+means, standard deviations (SD), sample sizes, correlation coefficients) extracted from primary literature.
+As well as being presented in the text or tables of research papers, descriptive statistics are frequently presented in figures and so need to be manually extracted using digitising programs.
 
-Although there are several tools that data extraction from figures (e.g.
-\proglang{DataThief} \citep{DataThief}, \proglang{GraphClick} \citep{GraphClick}, \proglang{WebPlotDigitizer} \citep{WebPlotDigitizer}), these tools do not cater to needs of meta-analysis for three main reasons.
-First, they typically only provide the user with calibrated \textit{x,y} coordinates from imported figures, and do not differentiate between common plot types that are used to present data.
-Consequently a large amount of downstream data manipulation is required, that is different across plots types.
-For example, data are frequently presented in mean/error plots (Figure \ref{fig:all_extract}A), from which the user wants a mean and SD for each group presented.
-From \textit{x,y} coordinates, users must manually discern between mean and error coordinates and assign points to groups.
-Error then needs to be calculated as the deviation from the mean, and then transformed to SD, according to the type of error presented.
+Although there are several tools that extract data from figures (e.g.
+\proglang{DataThief} \citep{DataThief}, \proglang{GraphClick} \citep{GraphClick}, \proglang{WebPlotDigitizer} \citep{WebPlotDigitizer}), these tools do not cater to needs of meta-analysis for four main reasons (here we focus on meta-analysis, although many  points apply to extraction for comparative analysis).
+First, although meta-analysis is an important tool in consolidating the data from multiple studies, many of the processes involved in data extraction are opaque and difficult to reproduce, making extending or replicating studies problematic.
+Having a tool that facilitates reproducibility in meta-analyses will increase transparency and aid in resolving the reproducibility crises seen in many fields \citep{peng_reproducible_2006, peng_reproducible_2011, Parker2016}.
 Second, digitising programs do not allow the integration of metadata at the time of data extraction, such as experimental group or variable names, and sample sizes.
 This makes the downstream calculations laborious, as information has to be added later using different software.
-Finally, existing programs do not import sets of images for the user to systematically work through.
+Third, existing programs do not import sets of images for the user to systematically work through.
 Instead they require the user to manually import images one by one, and export data into individual files, that need to be imported and edited using different software.
+Finally, digitising programs typically only provide the user with calibrated \textit{x,y} coordinates from imported figures, and do not differentiate between common plot types that are used to present data.
+Consequently a large amount of additional data manipulation is required, that is different across plots types.
+For example, data are frequently presented in plots with means and standard errors or confidence intervals (Figure \ref{fig:all_extract}A), from which the user wants a mean and SD for each group presented.
+From \textit{x,y} coordinates, users must manually discern between mean and error coordinates and assign points to groups.
+The error then needs to be calculated as the deviation from the mean, and then transformed to SD, according to the type of error presented.
 
-Data extraction from figures is therefore an incredibly time-consuming process as existing software does not provide an optimized research pipeline to facilitate data extraction and editing.
-Furthermore, although meta-analysis is an important tool in consolidating the data from multiple studies, many of the processes involved in data extraction are opaque and difficult to reproduce, making extending studies problematic.
-Having a tool that facilitates reproducibility in meta-analyses will increase transparency and aid in resolving the reproducibility crises seen in many fields \citep{peng_reproducible_2006, peng_reproducible_2011, Parker2016}.
-
+Data extraction from figures is therefore an incredibly time-consuming process as existing software does not provide an optimized, reproducible research pipeline to facilitate data extraction and editing.
 Here, we present an interactive \proglang{R} package, \pkg{metaDigitise} (available at https://github.com/daniel1noble/metaDigitise), which is designed for large scale, reproducible data extraction from figures, specifically catering to the the needs of meta-analysts.
@@
 
-\section{Directory Structure and Reproducibility}
-The \pkg{metaDigitise} package was created with the idea that users would have multiple images to extract from and therefore operates in the same way whether the user has one or multiple images.
-There is one main function, \fct{metaDigitise}, which interactively takes the user through the process of extracting data from figures.
+\section{\pkg{metaDigitise} and Reproducibility}
+The \pkg{metaDigitise} package has one main function, \fct{metaDigitise}, which interactively takes the user through the process of extracting data from figures.
 \fct{metaDigitise} works on a directory containing images of figures copied from primary literature, in .png, .jpg, .tiff, .pdf format, specified to \fct{metaDigitise} through the \code{dir} argument.
-The user should think carefully about their directory structure early on in their project.
-Although different directory structures may be used, we would recommend having all files for one project in a single directory with an informative and unambiguous naming scheme for images to help identify the paper and figure that data come from (e.g.
+\fct{metaDigitise} recognizes all the images in the given directory and automatically imports them one by one, allowing the user to extract the relevant information about a figure as they go.
+It is worth the user thinking carefully about their directory structure early on in their project.
+Although different directory structures may be used, we would recommend having all figures for one project in a single directory with an informative and unambiguous naming scheme (e.g.
 paper\_figure\_trait.png).
+This expedites digitisation by preventing users from having to constantly change directories and / or open new images.
 
-\fct{metaDigitise} recognizes all the images in a directory and automatically imports them one by one, allowing the user to extract the relevant information about a figure as they go.
-Having all figures in one directory therefore expedites digitisation by preventing users from having to constantly change directories and / or open new images.
 The data from each completed image is automatically saved as a \code{metaDigitise} object in a separate .RDS file to a \code{caldat} directory that is created within the parent directory when first executing \fct{metaDigitise}.
@@
 Running \fct{metaDigitise} presents the user with three options; `Process new images', `Import existing data' or `Edit existing data'.
-Selecting `Process New Images' starts the digitisation process on images within the directory that have not preciously been digitised; the other functions are discussed below.
+Selecting `Process New Images' starts the digitisation process on images within the directory that have not previously been digitised; the other functions are discussed below.
 
@@
 In many cases sample sizes may not be readily available when digitising figures.
-This information does not need to be added a the time of digitisation.
+This information does not need to be added at the time of digitisation.
 To expedite finding and adding these sample sizes at a later point, \fct{metaDigitise} has a specific edit option that allows users to enter previously omitted sample sizes.
@@
 \section{Software Validation}
-In order to evaluate the consistency of digitisation with \pkg{metaDigitise} between users, we got 14 people to digitise the same set of 14 figures created from a simulated dataset (see supplements).
+In order to evaluate the consistency of digitisation with \pkg{metaDigitise} between users, fourteen people digitized sets of 14 identical images created from a simulated dataset (see supplements).
 We found no evidence for any inter-observer variability in digitisations for the mean (ICC = 0, 95\% CI = 0 to 0.029, \textit{p} = 1), SD (ICC = 0, 95\% CI = 0 to 0.033, \textit{p} = 0.5) or correlation coefficient (ICC = 0.053, 95\% CI = 0 to 0.296, \textit{p} = 0.377).
@@
 In contrast to some other packages \pkg{metaDigitise} does not extract lines from figures.
-Line extraction is not particularly useful for most meta-analyses, although we recognise that it may be useful in other fields.
+Line extraction is not particularly useful for most comparative or meta-analytic work, although we recognise that it may be useful in fields other than these.
 Should a user like to extract lines with \pkg{metaDigitise}, we would recommend extracting data as a scatter plot, and clicking along the line in question.

### 2018-06-18 docs/MEE_submission/supplements.tex
MS revisions

@@
 Multiple image processing & \checkmark   & $\times$   & $\times$   & $\times$   & $\times$       & $\times$       & $\times$    \\
-Reproducable$^9$  & \checkmark   & \checkmark & \checkmark & $\times$   & \checkmark     & $\times$       & $\times$    \\
+Reproducable$^9$  & \checkmark   & \checkmark & \checkmark & $\times$   & \checkmark     & \checkmark       & $\times$    \\
 Automated point detection & $\times$     & \checkmark &     $\times$      & \checkmark & \checkmark     & \checkmark     & $\times$    \\

### 2018-06-18 docs/MEE_submission/tikx/figure.tex
MS revisions

\documentclass{article}
\usepackage[a3paper]{geometry}

\usepackage[active,tightpage]{preview}
\PreviewEnvironment{tikzpicture}

\usepackage{tikz}
\usetikzlibrary{positioning}
\usepackage{graphicx}

\begin{document}

\begin{preview}

\centering
\begin{tikzpicture}

%%%%-------------------------------------------------------------------------------------

\node[] (figure) at (3,10) {\includegraphics[width=.15\textwidth]{figure.png}};
\node[] (figure2) at (7,10) {\includegraphics[width=.15\textwidth]{figure2.png}};

\node[] (figure_D) at (0,5) {\includegraphics[width=.15\textwidth]{figure_digitize.png}};
\draw[->,thick, line width=0.3mm] (figure.south) -- (figure_D.north) ;

\node[] (figure2_D) at (3,5) {\includegraphics[width=.15\textwidth]{figure2_digitize.png}};
\draw[->,thick, line width=0.3mm] (figure2.south) -- (figure2_D.north) ;

\node[] (figure_mD) at (7,5) {\includegraphics[width=.15\textwidth]{figure_metaDigitise.png}};
\draw[->,thick, line width=0.3mm] (figure.south) -- (figure_mD.north);

\node[] (figure2_mD) at (10,5) {\includegraphics[width=.15\textwidth]{figure2_metaDigitise.png}};
\draw[->,thick, line width=0.3mm] (figure2.south) -- (figure2_mD.north);

\node[] (label_D) at (0,8) {digitize};
\node[] (label_D) at (10,8) {metaDigitise};

\setlength{\tabcolsep}{0.25em}
\node[inner sep=0pt, text width=2.5cm, anchor=north] (output_D) at (0,2) {\tiny   
\begin{tabular}{c c}
x & y\\
2.329895 & 9.605761\\
2.114273 & 7.212674\\
1.583472 & 5.246919\\
4.747202 & 3.914303\\
6.648952 & 5.837144\\
3.518762 & 7.556821\\
4.027756 & 7.544506\\
7.247960 & 4.248326\\
9.070899 & 6.302404\\
9.867233 & 2.323294\\
\end{tabular}
};

\node[inner sep=0pt, text width=2.5cm, anchor=north] (output2_D) at (3,2) {\tiny   
\begin{tabular}{c c}
x & y\\
1.00381 & 3.366921\\
1.00332 & 1.344158\\
1.00672 & 0.664347\\
1.00313 & -0.098183\\
0.99579 & -0.741522\\
2.00424 & 3.031860\\
2.00210 & 1.240030\\
2.01120 & 0.112484\\
1.99990 & -0.532548\\
2.00292 & -1.421841\\
2.99886 & 1.659290\\
2.99899 & 0.452495\\
3.01687 & -0.358099\\
2.99800 & -0.717574\\
2.99575 & -1.749286\\
\end{tabular}
};

\node[inner sep=0pt, anchor=north] (output_mD) at (8.5,2) {\tiny   
\begin{tabular}{c c c c c c c c}
filename & group\_id & variable & mean & sd & n & r & plot\_type\\
figure.png & black & seed size & 3.51610 & 2.12806 & 5 & -0.40887 & scatterplot\\
figure.png & black & seed number & 6.38350 & 2.12550 & 5 & -0.40887 & scatterplot\\
figure.png & red & seed size & 6.77229 & 2.87290 & 5 & -0.79541 & scatterplot\\
figure.png & red & seed number & 5.60961 & 2.24468 & 5 & -0.79541 & scatterplot\\
figure2.png & 1 & seed size & 0.81727 & 1.03320 & 40 & NA & boxplot\\
figure2.png & 2 & seed size & 0.41903 & 1.19537 & 40 & NA & boxplot\\
figure2.png & 3 & seed size & -0.14119 & 0.84790 & 40 & NA & boxplot\\
% filename & group\_id & variable & mean & sd & n & plot\_type\\
% figure2.png & 1 & seed size & 0.81727 & 1.03320 & 40 & boxplot\\
% figure2.png & 2 & seed size & 0.41903 & 1.19537 & 40 & boxplot\\
% figure2.png & 3 & seed size & -0.14119 & 0.84789 & 40 & boxplot\\
\end{tabular}
};

\draw[->,thick, line width=0.3mm] (figure_D) -- (output_D);
\draw[->,thick, line width=0.3mm] (figure2_D) -- (output2_D);
\draw[->,thick, line width=0.3mm] (figure_mD.south) -- (output_mD);
\draw[->,thick, line width=0.3mm] (figure2_mD.south) -- (output_mD);

\end{tikzpicture}

%\end{frame}
\end{preview}
\end{document}

### 2018-06-18 docs/MEE_submission/tikx/figure_gen.R

@@ -1,0 +1,66 @@
+
+
+png("/Users/jpick/Dropbox/0_postdoc/10_metaDigitise/docs/MEE_submission/tikx/figure.png")
+set.seed(25)
+plot(runif(10,1,10),runif(10,1,10), pch=c(17,19),col=c(1,2),cex=2, xlab="seed size", ylab="seed number", xlim=c(1,10), ylim=c(1,10), cex.axis=1.5)
+dev.off()
+
+png("/Users/jpick/Dropbox/0_postdoc/10_metaDigitise/docs/MEE_submission/tikx/figure2.png")
+set.seed(25)
+boxplot(c(rnorm(60,1,1),rnorm(60,0,1))~gl(3,40))
+dev.off()
+
+library(digitize)
+
+digitize("~/Dropbox/0_postdoc/10_metaDigitise/docs/MEE_submission/tikx/figure.png")
+
+
+digitize("~/Dropbox/0_postdoc/10_metaDigitise/docs/MEE_submission/tikx/figure2.png")
+
+
+
+#install.packages("devtools")
+devtools::install_github("daniel1noble/metaDigitise")
+library(metaDigitise)
+
+metaDigitise("~/Dropbox/0_postdoc/10_metaDigitise/docs/MEE_submission/tikx")
+
+
+  x        y
+2.329895 9.605761
+2.114273 7.212674
+1.583472 5.246919
+4.747202 3.914303
+6.648952 5.837144
+3.518762 7.556821
+4.027756 7.544506
+7.247960 4.248326
+9.070899 6.302404
+9.867233 2.323294
+
+    x           y
+1.0038093  3.36692124
+1.0033249  1.34415790
+1.0067158  0.66434661
+1.0031267 -0.09818274
+0.9957944 -0.74152193
+2.0042386  3.03186020
+2.0021028  1.24003046
+2.0111966  0.11248387
+1.9999009 -0.53254776
+2.0029175 -1.42184095
+2.9988550  1.65929044
+2.9989871  0.45249529
+3.0168665 -0.35809939
+2.9979963 -0.71757389
+2.9957503 -1.74928600
+
+
+filename	 group_id	 variable       mean	  sd		n	 	r		plot_type
+figure.png   	black   seed size  3.5161025 2.1280607  	5	-0.4088671 scatterplot
+figure.png    	black 	seed number  6.3834989 2.1254968	5	-0.4088671 scatterplot
[…]

### 2018-06-19 docs/MEE_submission/supplements.tex

--- 2018-03-08 docs/MEE_submission/supplements.tex
@@
 Scatterplots              & \checkmark   & \checkmark & \checkmark & \checkmark & \checkmark     & \checkmark$^7$ & \checkmark \\
-Mean/error plots      & \checkmark   & \checkmark & \checkmark & $\times$   & $\times$       & \checkmark$^7$ & $\times$    \\
-Boxplots                  & \checkmark   & $\times$   & $\times$   & $\times$   & $\times$       & $\times$       & $\times$    \\
-Histograms                & \checkmark   & $\times$   & $\times$   & $\times$   & \checkmark$^7$ & $\times$       & $\times$    \\
-Graph rotation$^8$        & \checkmark   & \checkmark & \checkmark & \checkmark & \checkmark     & $\times$       & $\times$    \\
-Groups                    & \checkmark   & \checkmark & $\times$   & \checkmark & \checkmark     & $\times$       & $\times$    \\
-Entry of metadata         & \checkmark   & $\times$   & $\times$   & $\times$   & $\times$       & $\times$       & $\times$    \\
-Summarising data          & \checkmark   & $\times$   & $\times$   & $\times$   & $\times$       & $\times$       & $\times$    \\
-Multiple image processing & \checkmark   & $\times$   & $\times$   & $\times$   & $\times$       & $\times$       & $\times$    \\
-Reproducable$^9$  & \checkmark   & \checkmark & \checkmark & $\times$   & \checkmark     & $\times$       & $\times$    \\
-Automated point detection & $\times$     & \checkmark &     $\times$      & \checkmark & \checkmark     & \checkmark     & $\times$    \\
-Line extraction           & $\times$     & \checkmark & \checkmark & \checkmark & \checkmark     & $\times$       & $\times$    \\
-Zoom                      & $\times$     & \checkmark & \checkmark & \checkmark & \checkmark     & $\times$       & $\times$    \\
-Log axis                  & $\times$   & \checkmark & \checkmark & \checkmark & \checkmark     & $\times$       & $\times$    \\
-Dates                     & $\times$   & $\times$   & \checkmark & $\times$   & \checkmark     & $\times$       & $\times$    \\
-Asymmetric error bars     & $\times$     & $\times$   & \checkmark & $\times$   & $\times$       & $\times$       & $\times$    \\
-Freeware                  & \checkmark$^{10}$ & \checkmark$^{11}$ & \checkmark$^{11}$ & $\times$$^{11}$ & \checkmark$^{11}$ & \checkmark$^{10}$ & \checkmark$^{10}$\\
+Mean/error plots      & \checkmark   & \checkmark & \checkmark & $\color{red}\times$   & $\color{red}\times$       & \checkmark$^7$ & $\color{red}\times$    \\
+Boxplots                  & \checkmark   & $\color{red}\times$   & $\color{red}\times$   & $\color{red}\times$   & $\color{red}\times$       & $\color{red}\times$       & $\color{red}\times$    \\
+Histograms                & \checkmark   & $\color{red}\times$   & $\color{red}\times$   & $\color{red}\times$   & \checkmark$^7$ & $\color{red}\times$       & $\color{red}\times$    \\
+Graph rotation$^8$        & \checkmark   & \checkmark & \checkmark & \checkmark & \checkmark     & $\color{red}\times$       & $\color{red}\times$    \\
+Groups                    & \checkmark   & \checkmark & $\color{red}\times$   & \checkmark & \checkmark     & $\color{red}\times$       & $\color{red}\times$    \\
+Entry of metadata         & \checkmark   & $\color{red}\times$   & $\color{red}\times$   & $\color{red}\times$   & $\color{red}\times$       & $\color{red}\times$       & $\color{red}\times$    \\
+Summarising data          & \checkmark   & $\color{red}\times$   & $\color{red}\times$   & $\color{red}\times$   & $\color{red}\times$       & $\color{red}\times$       & $\color{red}\times$    \\
+Multiple image processing & \checkmark   & $\color{red}\times$   & $\color{red}\times$   & $\color{red}\times$   & $\color{red}\times$       & $\color{red}\times$       & $\color{red}\times$    \\
+Reproducable$^9$  & \checkmark   & \checkmark & \checkmark & $\color{red}\times$   & \checkmark     & \checkmark       & $\color{red}\times$    \\
+Automated point detection & $\color{red}\times$     & \checkmark &     $\color{red}\times$      & \checkmark & \checkmark     & \checkmark     & $\color{red}\times$    \\
+Line extraction           & $\color{red}\times$     & \checkmark & \checkmark & \checkmark & \checkmark     & $\color{red}\times$       & $\color{red}\times$    \\
+Zoom                      & $\color{red}\times$     & \checkmark & \checkmark & \checkmark & \checkmark     & $\color{red}\times$       & $\color{red}\times$    \\
+Log axis                  & \checkmark   & \checkmark & \checkmark & \checkmark & \checkmark     & $\color{red}\times$       & $\color{red}\times$    \\
+Dates                     & $\color{red}\times$   & $\color{red}\times$   & \checkmark & $\color{red}\times$   & \checkmark     & $\color{red}\times$       & $\color{red}\times$    \\
+Asymmetric error bars     & $\color{red}\times$     & $\color{red}\times$   & \checkmark & $\color{red}\times$   & $\color{red}\times$       & $\color{red}\times$       & $\color{red}\times$    \\
+Freeware                  & \checkmark$^{10}$ & \checkmark$^{11}$ & \checkmark$^{11}$ & $\color{red}\times$$^{11}$ & \checkmark$^{11}$ & \checkmark$^{10}$ & \checkmark$^{10}$\\
 \hline

### 2018-06-20 R/H_extract.R
Cleaned some code, resolved come conflicts and added example to getExtracted

@@ -1,9 +1,10 @@
 #' @title histogram_extract
+#' @description Extraction of data from histograms
 #' @param edit logical; whether in edit mode 
 #' @param raw_data raw data
 #' @param calpoints The calibration points
 #' @param ... arguments to pass to internal_redraw
-#' @description Extraction of data from histograms
+
 histogram_extract <- function(edit=FALSE, raw_data = data.frame(), calpoints, ...){
 
 	if(edit){ 

### 2018-06-20 R/edit_metaDigitise.R

@@ -79,9 +79,9 @@ N_needed <- function(filepaths){
 
 
 #' @title enter_N
-#' @description ...
+#' @description Enter sample sizes for a group
 #' @param raw_data raw_data
-#' @param ... ...
+#' @param ... Pass additional arguments
 #' @author Joel Pick
 
 enter_N <- function(raw_data,...){
@@ -93,16 +93,11 @@ enter_N <- function(raw_data,...){
 }
 
 
-
-
-
-
 #' @title edit_metaDigitise
 #' @description Function for editing previous data extraction through `metaDigitise`
 #' @param object an R object of class ‘metaDigitise’
 #' @return Data.frame
 #' @author Joel Pick
-#' @export
 
 edit_metaDigitise <- function(object){
 	graphics::plot(object)

### 2018-06-20 R/import_metaDigitise.R

@@ -3,6 +3,37 @@
 #' @param dir The directory where figures have already been digitised. There
 #' @param summary Logical indicating whether summarised (default) or calibrated data should be returned. 
 #' @return Returns a data frame (summary = TRUE) or a list with slots for each plot type (summary = FALSE)
+#' @example
+#' \dontrun{
+#' # Make some mock metaDigitise object
+#' 	mock_metaDig <- list(
+#' 			image_file = "./image.png",
+#' 			flip=FALSE,
+#' 			rotate=0,
+#' 			plot_type="mean_error",
+#' 			variable="y",
+#' 			calpoints = data.frame(x=c(0,0),y=c(0,100)), 
+#' 			point_vals = c(1,2), 
+#' 			entered_N=TRUE,
+#' 			raw_data = data.frame(id=rep("control",2), x=c(60,60), y=c(75,50), n=rep(20,2)),
+#' 			knownN = NULL,
+#' 			error_type="sd",
+#' 			processed_data=data.frame(id=as.factor("control"),mean=1.5, error=0.25, n=20, 	variable="y", stringsAsFactors = FALSE)	
+#' 		)
+#' class(mock_metaDig) <- 'metaDigitise'
+#' 
+#' # write image file to tmpdir()
+#' 		dir <- tempdir()
+#' 
+#' # Setup directory as it would be if digitised images existed		
+#' 		setup_calibration_dir(dir)
+#' 
+#' # Save the digitised data
+#' 		saveRDS(mock_metaDig, file = paste0(dir, "/caldat/", "image"))
+#' 
+#' #metaDigitise figures
+#' data <- getExtracted(dir)
+#' }
 #' @export
 
 getExtracted <- function(dir, summary=TRUE){

### 2018-06-21 docs/MEE_submission/manuscript_MEE.tex
MS updates

@@
 
-\usepackage{amssymb,rotating,natbib,graphicx,fancyvrb}
-
+\usepackage{amssymb,rotating,natbib,graphicx,fancyvrb,booktabs,xcolor} 
 \usepackage[parfill]{parskip} 
@@
 Although there are several tools that extract data from figures (e.g.
-\proglang{DataThief} \citep{DataThief}, \proglang{GraphClick} \citep{GraphClick}, \proglang{WebPlotDigitizer} \citep{WebPlotDigitizer}), these tools do not cater to needs of meta-analysis for four main reasons (here we focus on meta-analysis, although many  points apply to extraction for comparative analysis).
+\proglang{DataThief} \citep{DataThief}, \proglang{GraphClick} \citep{GraphClick}, \proglang{WebPlotDigitizer} \citep{WebPlotDigitizer}, see Table \ref{tab:comparison}), these tools do not cater to needs of meta-analysis for four main reasons (here we focus on meta-analysis, although many  points apply to extraction for comparative analysis).
 First, although meta-analysis is an important tool in consolidating the data from multiple studies, many of the processes involved in data extraction are opaque and difficult to reproduce, making extending or replicating studies problematic.
@@
 This makes calibration of the points in the figure problematic.
-\fct{metaDigitise} allows users to rotate the image (Figure \ref{fig:rotate}A,B).
+\fct{metaDigitise} allows users to rotate the image (Figure S2A,B).
 Furthermore, mean/error plots, box plots and histograms, may be presented with horizontal bars.
-\fct{metaDigitise} assumes that bars are vertical, but allows the user to flip the image to make the bars are vertical (Figure \ref{fig:rotate}C,D).
+\fct{metaDigitise} assumes that bars are vertical, but allows the user to flip the image to make the bars are vertical (Figure S2C,D).
 
@@
 
-\begin{figure}[!b] 
+\clearpage
+
+\section*{Tables}
+
+\begin{table}[h!]
 \centering
- \includegraphics[width=0.75\textwidth]{fig_rotate.pdf} 
- \caption{Figure rotation. A) and B) show how non-aligned images can be realigned through user defined rotation. C) and D) show how figures can be re-orientated so as to aid data input.}
-\label{fig:rotate}
-\end{figure}
+\rotatebox{90}{%
+\begin{minipage}{1.35\textwidth}
+\small
+\begin{tabular}{lccccccc}
+\hline
+Function                  & metaDigitise & GraphClick$^1$ & DataThief$^2$  & DigitizeIt$^3$ & WebPlotDigitizer$^4$ & metagear$^5$ & digitize$^6$\\
+\hline
+Scatterplots              & \checkmark   & \checkmark & \checkmark & \checkmark & \checkmark     & \checkmark$^7$ & \checkmark \\
+Mean/error plots      & \checkmark   & \checkmark & \checkmark & $\color{red}\times$   & $\color{red}\times$       & \checkmark$^7$ & $\color{red}\times$    \\
+Boxplots                  & \checkmark   & $\color{red}\times$   & $\color{red}\times$   & $\color{red}\times$   & $\color{red}\times$       & $\color{red}\times$       & $\color{red}\times$    \\
+Histograms                & \checkmark   & $\color{red}\times$   & $\color{red}\times$   & $\color{red}\times$   & \checkmark$^7$ & $\color{red}\times$       & $\color{red}\times$    \\
+Graph rotation$^8$        & \checkmark   & \checkmark & \checkmark & \checkmark & \checkmark     & $\color{red}\times$       & $\color{red}\times$    \\
+Grouped Data            & \checkmark   & \checkmark & $\color{red}\times$   & \checkmark & \checkmark     & $\color{red}\times$       & $\color{red}\times$    \\
+Entry of metadata         & \checkmark   & $\color{red}\times$   & $\color{red}\times$   & $\color{red}\times$   & $\color{red}\times$       & $\color{red}\times$       & $\color{red}\times$    \\
+Summarising data          & \checkmark   & $\color{red}\times$   & $\color{red}\times$   & $\color{red}\times$   & $\color{red}\times$       & $\color{red}\times$       & $\color{red}\times$    \\
+Multiple image processing & \checkmark   & $\color{red}\times$   & $\color{red}\times$   & $\color{red}\times$   & $\color{red}\times$       & $\color{red}\times$       & $\color{red}\times$    \\
+Reproducable$^9$  & \checkmark   & \checkmark & \checkmark & $\color{red}\times$   & \checkmark     & \checkmark       & $\color{red}\times$    \\
+Automated point detection & $\color{red}\times$     & \checkmark &     $\color{red}\times$      & \checkmark & \checkmark     & \checkmark     & $\color{red}\times$    \\
+Line extraction           & $\color{red}\times$     & \checkmark & \checkmark & \checkmark & \checkmark     & $\color{red}\times$       & $\color{red}\times$    \\
+Zoom                      & $\color{red}\times$     & \checkmark & \checkmark & \checkmark & \checkmark     & $\color{red}\times$       & $\color{red}\times$    \\
+Log axis                  & \checkmark   & \checkmark & \checkmark & \checkmark & \checkmark     & $\color{red}\times$       & $\color{red}\times$    \\
+Dates                     & $\color{red}\times$   & $\color{red}\times$   & \checkmark & $\color{red}\times$   & \checkmark     & $\color{red}\times$       & $\color{red}\times$    \\
+Asymmetric error bars     & $\color{red}\times$     & $\color{red}\times$   & \checkmark & $\color{red}\times$   & $\color{red}\times$       & $\color{red}\times$       & $\color{red}\times$    \\
+Freeware                  & \checkmark$^{10}$ & \checkmark$^{11}$ & \checkmark$^{11}$ & $\color{red}\times$$^{11}$ & \checkmark$^{11}$ & \checkmark$^{10}$ & \checkmark$^{10}$\\
+\hline
+
+%http://datathief.org/DatathiefManual.pdf
+%http://www.digitizeit.de/#Features
+%https://automeris.io/WebPlotDigitizer/userManual.pdf
+
+\end{tabular}
+\\
+$^1$ \citet{GraphClick}
+$^2$ \citet{DataThief}
+$^3$ \citet{DigitizeIt}
+$^4$ \citet{WebPlotDigitizer}
+$^5$ \citet{Lajeunesse2016}
+$^6$ \citet{Poisot2011}
+\\$^7$ Only automated, no manual extraction.
+\\$^8$ Or handles rotated graphs. 
+\\$^9$ Allows saving, re-plotting and editing of data extraction.
+\\$^{10}$ \proglang{R} package.
+\\$^{11}$ Standalone software.
+\caption{\label{tab:comparison} Comparison of functionality between different digitisation softwares.}
+\end{minipage}
+}
+
+\end{table}
 

### 2018-06-21 docs/MEE_submission/supplements.tex
MS updates

--- 2018-03-08 docs/MEE_submission/supplements.tex
@@
 \usepackage{amssymb,rotating,natbib,graphicx,fancyvrb,booktabs} 
-
 \usepackage[parfill]{parskip} 
@@
 \fct{metaDigitise} then asks whether the figure needs to be rotated or flipped.
-This can be needed when box plots and mean and error plots are not orientated correctly.
+This can be needed when box plots and mean and error plots are not orientated correctly (as shown in Figure \ref{fig:rotate}).
 In some cases, older papers can give slightly off angled images which can be corrected by rotating.
 In this prompt the user has three options: \code{f} for ``Flip'', \code{r} for ``rotate'' or \code{c} for ``continue''.
+
+\begin{figure}[!b] 
+\centering
+ \includegraphics[width=0.75\textwidth]{fig_rotate.pdf} 
+ \caption{Figure rotation. A) and B) show how non-aligned images can be realigned through user defined rotation. C) and D) show how figures can be re-orientated so as to aid data input.}
+\label{fig:rotate}
+\end{figure}
 
@@
 
-\section{Comparison with existing software}
-\begin{table}[h!]
-\centering
-\rotatebox{90}{%
-\begin{minipage}{1.35\textwidth}
-\small
-\begin{tabular}{lccccccc}
-\hline
-Function                  & metaDigitise & GraphClick$^1$ & DataThief$^2$  & DigitizeIt$^3$ & WebPlotDigitizer$^4$ & metagear$^5$ & digitize$^6$\\
-\hline
-Scatterplots              & \checkmark   & \checkmark & \checkmark & \checkmark & \checkmark     & \checkmark$^7$ & \checkmark \\
-Mean/error plots      & \checkmark   & \checkmark & \checkmark & $\times$   & $\times$       & \checkmark$^7$ & $\times$    \\
-Boxplots                  & \checkmark   & $\times$   & $\times$   & $\times$   & $\times$       & $\times$       & $\times$    \\
-Histograms                & \checkmark   & $\times$   & $\times$   & $\times$   & \checkmark$^7$ & $\times$       & $\times$    \\
-Graph rotation$^8$        & \checkmark   & \checkmark & \checkmark & \checkmark & \checkmark     & $\times$       & $\times$    \\
-Groups                    & \checkmark   & \checkmark & $\times$   & \checkmark & \checkmark     & $\times$       & $\times$    \\
-Entry of metadata         & \checkmark   & $\times$   & $\times$   & $\times$   & $\times$       & $\times$       & $\times$    \\
-Summarising data          & \checkmark   & $\times$   & $\times$   & $\times$   & $\times$       & $\times$       & $\times$    \\
-Multiple image processing & \checkmark   & $\times$   & $\times$   & $\times$   & $\times$       & $\times$       & $\times$    \\
-Reproducable$^9$  & \checkmark   & \checkmark & \checkmark & $\times$   & \checkmark     & $\times$       & $\times$    \\
-Automated point detection & $\times$     & \checkmark &     $\times$      & \checkmark & \checkmark     & \checkmark     & $\times$    \\
-Line extraction           & $\times$     & \checkmark & \checkmark & \checkmark & \checkmark     & $\times$       & $\times$    \\
-Zoom                      & $\times$     & \checkmark & \checkmark & \checkmark & \checkmark     & $\times$       & $\times$    \\
-Log axis                  & $\times$   & \checkmark & \checkmark & \checkmark & \checkmark     & $\times$       & $\times$    \\
-Dates                     & $\times$   & $\times$   & \checkmark & $\times$   & \checkmark     & $\times$       & $\times$    \\
-Asymmetric error bars     & $\times$     & $\times$   & \checkmark & $\times$   & $\times$       & $\times$       & $\times$    \\
-Freeware                  & \checkmark$^{10}$ & \checkmark$^{11}$ & \checkmark$^{11}$ & $\times$$^{11}$ & \checkmark$^{11}$ & \checkmark$^{10}$ & \checkmark$^{10}$\\
-\hline
-
-%http://datathief.org/DatathiefManual.pdf
-%http://www.digitizeit.de/#Features
-%https://automeris.io/WebPlotDigitizer/userManual.pdf
-
-\end{tabular}
-\\
-$^1$ \citet{GraphClick}
-$^2$ \citet{DataThief}
-$^3$ \citet{DigitizeIt}
-$^4$ \citet{WebPlotDigitizer}
-$^5$ \citet{Lajeunesse2016}
-$^6$ \citet{Poisot2011}
-\\$^7$ Only automated, no manual extraction.
-\\$^8$ Or handles rotated graphs. 
-\\$^9$ Allows saving, re-plotting and editing of data extraction.
-\\$^{10}$ \proglang{R} package.
-\\$^{11}$ Standalone software.
-\caption{\label{tab:comparison} Comparison of functionality between different digitisation softwares.}
-\end{minipage}
-}
-
-\end{table}
-
-\clearpage
-
 \bibliographystyle{FuncEcol.bst}

### 2018-06-22 docs/MEE_submission/manuscript_MEE.tex
Edit MS and added some stuff

@@
 
-\textbf{Reproducible, flexible and high throughput data extraction from primary literature: The \pkg{metaDigitise} \proglang{R} package}
+\textbf{Reproducible, flexible and high-throughput data extraction from primary literature: The \pkg{metaDigitise} \proglang{R} package}
 
@@
 
-In many different contexts, researchers make use of data presented in primary literature.
-In the ecology and evolution, this most notably includes comparative and meta-analyses.
+In the fields of ecology and evolution, researchers make use of data presented in primary literature for comparative- and meta-analyses.
 These techniques rely on descriptive statistics (e.g.
@@
 Although there are several tools that extract data from figures (e.g.
-\proglang{DataThief} \citep{DataThief}, \proglang{GraphClick} \citep{GraphClick}, \proglang{WebPlotDigitizer} \citep{WebPlotDigitizer}, see Table \ref{tab:comparison}), these tools do not cater to needs of meta-analysis for four main reasons (here we focus on meta-analysis, although many  points apply to extraction for comparative analysis).
+\proglang{DataThief} \citep{DataThief}, \proglang{GraphClick} \citep{GraphClick}, \proglang{WebPlotDigitizer} \citep{WebPlotDigitizer}, see Table \ref{tab:comparison}), these tools do not cater to needs of meta-analysts for four main reasons (here we focus on meta-analysis, although many  points apply to extraction for comparative analysis).
 First, although meta-analysis is an important tool in consolidating the data from multiple studies, many of the processes involved in data extraction are opaque and difficult to reproduce, making extending or replicating studies problematic.
@@
 Third, existing programs do not import sets of images for the user to systematically work through.
-Instead they require the user to manually import images one by one, and export data into individual files, that need to be imported and edited using different software.
+Instead they require the user to manually import images and export the resulting digitised data into individual files one-by-one.
+These data often subsequently need to be imported and edited using different software.
 Finally, digitising programs typically only provide the user with calibrated \textit{x,y} coordinates from imported figures, and do not differentiate between common plot types that are used to present data.
-Consequently a large amount of additional data manipulation is required, that is different across plots types.
+Consequently, a large amount of additional data manipulation is required, that is different across plots types.
 For example, data are frequently presented in plots with means and standard errors or confidence intervals (Figure \ref{fig:all_extract}A), from which the user wants a mean and SD for each group presented.
@@
 The necessary calculations are carried out on calibrated data immediately after extraction so that comparable summary statistics can be obtained quickly.
-Summary data from multiple figures is returned into a single data frame which can be can easily exported or use in downstream analysis within \proglang{R}.
-Completed digitisations are automatically saved for each figure, meaning users can redraw their digitisations on figures, make corrections and access calibration and proceeded data.
+Summary data from multiple figures is returned into a single data frame which can be can easily exported or used in downstream analysis within \proglang{R}.
+Completed digitisations are automatically saved for each figure, meaning users can redraw their digitisations (along with metadata) on figures, make corrections and access calibration and processed (i.e., summarised) data.
 This makes sharing figure digitisation and reproducing the work of others simple and easy, and allows meta-analyses to be updated more efficiently.
@@
 \fct{metaDigitise} works on a directory containing images of figures copied from primary literature, in .png, .jpg, .tiff, .pdf format, specified to \fct{metaDigitise} through the \code{dir} argument.
-\fct{metaDigitise} recognizes all the images in the given directory and automatically imports them one by one, allowing the user to extract the relevant information about a figure as they go.
-It is worth the user thinking carefully about their directory structure early on in their project.
-Although different directory structures may be used, we would recommend having all figures for one project in a single directory with an informative and unambiguous naming scheme (e.g.
+\fct{metaDigitise} recognizes all the images in the given directory and automatically imports them one-by-one, allowing the user to extract the relevant information about a figure as they go.
+Figures can be organised in different ways for a project, but we would recommend having all figures for one project in a single directory with an informative and unambiguous naming scheme (e.g.
 paper\_figure\_trait.png).
@@
 
-The data from each completed image is automatically saved as a \code{metaDigitise} object in a separate .RDS file to a \code{caldat} directory that is created within the parent directory when first executing \fct{metaDigitise}.
+The data from each completed image is automatically saved as a \code{metaDigitise} object in a separate .RDS file to a \code{caldat} folder that is created within the parent directory when first executing \fct{metaDigitise}.
 These files enable re-plotting and editing of images at a later point (see below).
@@
 
-This directory structure allows the complete digitisation process to be reproduced at a later stage, shared with collaborators and presented as supplementary materials for a publication.
-As long as all the images and the caldat directory are still in one directory, \fct{metaDigitise} will be able to reproduce all figure extractions, regardless of the computer it is run on.
+The complete digitisation process can then be reproduced at a later stage, shared with collaborators and presented as supplementary materials for a publication, regardless of the computer it is run on.
 For an analysis to be updated, new figures can simply be added to the directory and \fct{metaDigitise} run to incorporate the new data.
@@
 \section{Image Processing}
-Running \fct{metaDigitise} presents the user with three options; `Process new images', `Import existing data' or `Edit existing data'.
-Selecting `Process New Images' starts the digitisation process on images within the directory that have not previously been digitised; the other functions are discussed below.
-
+Running \fct{metaDigitise} presents the user with three options; `Process new images', `Import existing data' or `Edit existing data', which can be used during and after digitisation to execute a range of functions (see Figure \ref{fig:all_extract} – 'Editing' and 'Importing' are discussed in the next section).
+Selecting `Process New Images' starts the digitisation process on images within the directory that have not previously been digitised.
 For all plot types, \fct{metaDigitise} requires the user to calibrate the axes in the figure, by clicking on two known points on the axis in question, and entering the value of those points (Figure \ref{fig:all_extract}).
@@
 This is based on the calibration used in the \pkg{digitize} R package \citep{Poisot2011}.
-For mean/error and box plots, only the y-axis is calibrated (Figure \ref{fig:all_extract}A,B), assuming the x-axis is redundant.
-For scatter plots and histograms both axes are calibrated (Figure \ref{fig:all_extract}C,D).
+For mean/error and box plots, only the y-axis is calibrated (Figure \ref{fig:all_extract}), assuming the x-axis is redundant.
+For scatter plots and histograms both axes are calibrated (Figure \ref{fig:all_extract}).
 
@@
 \fct{metaDigitise} handles mean/error and box plots in a very similar way.
-For each mean/box, the user is enters group names and sample sizes.
+For each mean/box, the user enters group name(s) and sample size(s).
 If the user does not enter a sample size at the time of data extraction (if, for example, the information is not readily available) a SD is not calculated.
-Sample sizes can, however, be entered at a later time (see below).
-For mean/error plots, the user clicks on an error bar and the mean.
+Sample sizes can, however, be entered at a later time (see next section).
+For mean/error plots, the user clicks on an error bar followed by the mean.
 Error bars above or below the mean can be clicked, as sometimes one is clearer than the other.
@@
 For box plots, the user clicks on the maximum, upper quartile, median, lower quartile and minimum.
-For both plot types, the user can add, edit or remove groups.
+For both plot types, the user can add, edit or remove groups while digitising for when finished.
 Three functions, \fct{error\_to\_sd}, \fct{rqm\_to\_mean} and \fct{rqm\_to\_sd}, that convert different error types to SD, box plot data to mean and box plot data SD, respectively, are also available in the package (see supplements for further details of these conversions).
+
+\subsection{Histograms}
+The user clicks on the top corners of each bar, which are drawn in alternating colours (Figure \ref{fig:all_extract}C).
+Bars are numbered to allow the the user to edit them.
+As with scatter plots, if the sample size from the extracted data does not match a known sample size, the user can enter an alternate sample size.
+The formulas for calculation of mean, SD and sample size are provided in the supplement.
 
@@
 Users can extract points from multiple groups from scatter plots.
-Different groups are plotted in different colours and shapes to enable them to be distinguished, with a legend at the bottom of the figure (Figure \ref{fig:all_extract}C).
+Different groups are plotted in different colours and shapes to enable them to be distinguished, with a legend at the bottom of the figure (Figure \ref{fig:all_extract}D).
 Mean, SD and sample size are calculated from the clicked points, for each group.
@@
 This may result in the sample size of digitised groups conflicting with what is reported in the paper.
-For example, in Figure \ref{fig:all_extract}C only 49 points have been clicked when the sample size is known to be 50.
-Hence, \pkg{metaDigitise} also provides the user with the option to input known sample sizes directly.
+However, users also have the option to input known sample sizes directly, if required.
 Nonetheless, it is important to recognise the impact that overlapping points can have on summary statistics, and in particular on sampling variance.
-
-\subsection{Histograms}
-The user clicks on the top corners of each bar, which are drawn in alternating colours (Figure \ref{fig:all_extract}D).
-Bars are numbered to allow the the user to edit them.
-As with scatter plots, if the sample size from the extracted data does not match a known sample size, the user can enter an alternate sample size.
-The calculation of mean, SD and sample size from this data is shown in the supplements.
 
@@
 \pkg{metaDigitise} is also able to re-import, edit and re-plot previously digitised figures.
-When running \fct{metaDigitise}, the user can choose to `Import existing data', which returns previously digitised data, from sinlge or all figures.
+When running \fct{metaDigitise}, the user can choose to 'Import existing data', which returns previously digitised data, from a single figure or all figures.
 Alternately, the \fct{getExtracted} function returns the data of previous digitisations, but without user interaction, allowing easier integration into larger scripts.
-`Edit existing data' allows the user to re-plot or edit information or digitisations that have previously be done.
+'Edit existing data' allows the user to re-plot or edit information for digitisations that have previously be done.
 
@@
 To expedite finding and adding these sample sizes at a later point, \fct{metaDigitise} has a specific edit option that allows users to enter previously omitted sample sizes.
-This first identifies missing sample sizes in the digitised output, re-plots the relevant figures and prompts the user to enter the sample sizes for the relevant groups in the figure, one by one.
+This first identifies missing sample sizes in the digitised output, re-plots the relevant figures and prompts the user to enter the sample sizes for the relevant groups in the figure.
 

### 2018-07-09 docs/MEE_submission/manuscript_MEE.tex
MS changes
MS updates

@@
 \begin{enumerate} 
-\item Research synthesis, such as meta-analysis requires the extraction of effect sizes from primary literature.
-Such effect sizes are calculated from summary statistics.
-However, exact values of such statistics are commonly hidden in figures.
+\item Research synthesis, such as comparative and meta-analyses, requires the extraction of effect sizes from primary literature, which are commonly calculated from descriptive statistics.
+However, the exact values of such statistics are commonly hidden in figures.
 
-\item Extracting summary statistics from figures can be a slow process that is not easily reproducible.
+\item Extracting descriptive statistics from figures can be a slow process that is not easily reproducible.
 Additionally, current software lacks an ability to incorporate important meta-data (e.g., sample sizes, treatment / variable names) about experiments and is not integrated with other software to streamline analysis pipelines.
 
-\item Here we present the R package \pkg{metaDigitise} which extracts descriptive statistics such as means, standard deviations and correlations from the four plot types: 1) mean/error plots (e.g.
+\item Here we present the R package \pkg{metaDigitise} which extracts descriptive statistics such as means, standard deviations and correlations from four plot types: 1) mean/error plots (e.g.
 bar graphs with standard errors), 2) box plots, 3) scatter plots and 4) histograms.
@@
 \vskip10pt
- \textbf{Keywords:} meta-analysis, comparative analysis, data extraction, \proglang{R}, reproducibility, figures, images, summary statistics
+ \textbf{Keywords:} meta-analysis, comparative analysis, data extraction, \proglang{R}, reproducibility, figures, images, descriptive statistics
 
@@
 
-%%Most notably, this includes meta-analysis, which is increasingly common in ecology and evolution.
-Meta-analysis uses effect size estimates and their sampling variance, taken from many studies, to understand whether particular effects are common across studies and to explain variation among these effects \citep{Glass1976,Koricheva2013,Naka2017}.
-
-%% While inferential statistics (e.g., \textit{t}- and \textit{F}-statistics) are often presented along side descriptive statistics, and can be used to derive effect sizes, descriptive statistics are much more appropriate to use because sources of non-independence in experimental designs can be dealt with more easily \citep{Noble2017}.
-
-In the fields of ecology and evolution, researchers make use of data presented in primary literature for comparative- and meta-analyses.
+In the fields of ecology and evolution, researchers make use of data presented in primary literature for comparative and meta-analyses.
 These techniques rely on descriptive statistics (e.g.
@@
 \pkg{metaDigitise} operates within the \proglang{R} environment making data extraction, analysis and export more streamlined.
-The necessary calculations are carried out on calibrated data immediately after extraction so that comparable summary statistics can be obtained quickly.
+The necessary calculations are carried out on calibrated data immediately after extraction so that comparable descriptive statistics can be obtained quickly.
 Summary data from multiple figures is returned into a single data frame which can be can easily exported or used in downstream analysis within \proglang{R}.
@@
 
-\pkg{metaDigitise} recognises four main types of plot; Mean/error plots, box plots, scatter plots and histograms (\ref{fig:all_extract}).
+%%% \pkg{metaDigitise} also allows back calculation of data presented on log axes.
+
+\pkg{metaDigitise} recognises four main types of plot; Mean/error plots, box plots, scatter plots and histograms (Figure \ref{fig:all_extract}).
 All plot types can be extracted in a single call of \fct{metaDigitise} and integrated into one output.
@@
 
-\subsection{Histograms}
-The user clicks on the top corners of each bar, which are drawn in alternating colours (Figure \ref{fig:all_extract}C).
-Bars are numbered to allow the the user to edit them.
-As with scatter plots, if the sample size from the extracted data does not match a known sample size, the user can enter an alternate sample size.
-The formulas for calculation of mean, SD and sample size are provided in the supplement.
-
 \subsection{Scatter plots}
@@
 However, users also have the option to input known sample sizes directly, if required.
-Nonetheless, it is important to recognise the impact that overlapping points can have on summary statistics, and in particular on sampling variance.
+Nonetheless, it is important to recognise the impact that overlapping points can have on descriptive statistics, and in particular on sampling variance.
+
+\subsection{Histograms}
+The user clicks on the top corners of each bar, which are drawn in alternating colours (Figure \ref{fig:all_extract}C).
+Bars are numbered to allow the the user to edit them.
+As with scatter plots, if the sample size from the extracted data does not match a known sample size, the user can enter an alternate sample size.
+The formulas for calculation of mean, SD and sample size are provided in the supplement.
 
@@
 When running \fct{metaDigitise}, the user can choose to 'Import existing data', which returns previously digitised data, from a single figure or all figures.
-Alternately, the \fct{getExtracted} function returns the data of previous digitisations, but without user interaction, allowing easier integration into larger scripts.
+Alternately, the \fct{getExtracted} function returns the data from previous digitisations, but without user interaction, allowing easier integration into larger scripts.
 'Edit existing data' allows the user to re-plot or edit information for digitisations that have previously be done.
+Re-plotting digitisations with all metadata is an important reproducibility feature, as it allows users to see exactly what information has been extracted, as well as making it easy to spot and data extraction errors.
 
@@
 We found no evidence for any inter-observer variability in digitisations for the mean (ICC = 0, 95\% CI = 0 to 0.029, \textit{p} = 1), SD (ICC = 0, 95\% CI = 0 to 0.033, \textit{p} = 0.5) or correlation coefficient (ICC = 0.053, 95\% CI = 0 to 0.296, \textit{p} = 0.377).
-There was little bias between digitised and true values, on average 1.63\% (mean = 0.02\%, SD = 4.9\%, \textit{r} = -0.03\%) and there were small absolute differences between digitised and true values, on average 2.18\% (mean = 0.40\%, SD = 5.81\%, \textit{r} = 0.33\%) across all three summary statistics.
+There was little bias between digitised and true values, on average 1.63\% (mean = 0.02\%, SD = 4.9\%, \textit{r} = -0.03\%) and there were small absolute differences between digitised and true values, on average 2.18\% (mean = 0.40\%, SD = 5.81\%, \textit{r} = 0.33\%) across all three descriptive statistics.
 SD estimates from digitisations are clearly most error prone.
 The mean absolute differences for each plot type clearly show that this effect is driven by extraction from box plots and histograms (\% difference; box plot: 15.805, histogram: 5.210, mean/error: 1.500, scatter plot: 0.433).
-SD estimation from box plot summary statistics is known to be more error prone, especially at small sample sizes \citep{Wan2014}.
+SD estimation from box plot descriptive statistics is known to be more error prone, especially at small sample sizes \citep{Wan2014}.
 
 We also used simulated data to test the accuracy of digitisations with respect to known values (see supplements).
-\pkg{metaDigitise} was extremely accurate at matching clicked points to their true values essentially being perfectly correlated with the true simulated data for both the \textit{x}-variable (Pearson's correlation; \textit{r} > 0.999, \textit{t} = 2137.4, \textit{df} = 78, $p < 0.001$) and \textit{y}-variable (\textit{r} > 0.999, \textit{t} = 1897.8, \textit{df} = 78, $p < 0.001$) in scatterplots.
+\pkg{metaDigitise} was extremely accurate at matching clicked points to their true values essentially being perfectly correlated with the true simulated data for both the \textit{x}-variable (Pearson's correlation; $r > 0.999$, \textit{t} = 2137.4, \textit{df} = 78, $p < 0.001$) and \textit{y}-variable ($r > 0.999$, \textit{t} = 1897.8, \textit{df} = 78, $p < 0.001$) in scatterplots.
 
@@
 
-Although \pkg{metaDigitise} is very flexible and provides functionality not seen in any other package, there are some functions that it does not perform (see Table S1).
+Although \pkg{metaDigitise} is very flexible and provides functionality not seen in any other package, there are some functions that it does not perform (see Table \ref{tab:comparison}).
 Notably \pkg{metaDigitise} lacks automated point detection.
@@
 \centering 
- \includegraphics[width=0.75\textwidth]{fig_all_extract.pdf} 
- \caption{Four plot types that \pkg{metaDigitise} is designed to extract data from: A) mean/error plot, B) box plot, C) scatter plot and D) histogram.
-Data is taken from the iris dataset in R.
-A and B are plotted with the whole dataset, C and D are just the data for the species \textit{setosa}.
-Digitisation of the images is shown.
-All figures are clearly labelled at the top to remind users of the filename and plot type.
+\rotatebox{90}{%
+
+ \includegraphics[width=1.3\textwidth]{Adobe_fig1_V3.png} 
+ }
+ \caption{Functionality of \pkg{metaDigitise}.
+Using the iris dataset in R, digitisation of different plot types, A) mean/error plot, B) box plot, C) histogram and D) scatter plot, is shown in \pkg{metaDigitise} (left) compared with other common softwares (right).
+A) and B) are plotted with the whole dataset, C) is just the data for the species \textit{setosa} and D) a subset from all three species.
+Notable functions of metaDigitise are listed in the center.
+Other software also perform points 3 and 4 (see Table \ref{tab:comparison}), although these functions are more developed in \pkg{metaDigitise}.
+As shown on the left hand side of the figure, \pkg{metaDigitise} clearly displays the stages of the digitisation to aid the transparency of the process, and returns concatenated summary data for all images.
+%  All figures are clearly labelled at the top to remind users of the filename and plot type.
 This reduces errors throughout the digitisation process.
@@
 In A) and B), user entered group names and sample sizes are displayed beside the relevant points.
-In C) the names and sample sizes for each group are shown below the figure.}
+In C) the names and sample sizes for each group are shown below the figure.
+}
 \label{fig:all_extract}
@@
 Histograms                & \checkmark   & $\color{red}\times$   & $\color{red}\times$   & $\color{red}\times$   & \checkmark$^7$ & $\color{red}\times$       & $\color{red}\times$    \\
-Graph rotation$^8$        & \checkmark   & \checkmark & \checkmark & \checkmark & \checkmark     & $\color{red}\times$       & $\color{red}\times$    \\
+Entry of metadata         & \checkmark   & $\color{red}\times$   & $\color{red}\times$   & $\color{red}\times$   & $\color{red}\times$       & $\color{red}\times$       & $\color{red}\times$    \\
 Grouped Data            & \checkmark   & \checkmark & $\color{red}\times$   & \checkmark & \checkmark     & $\color{red}\times$       & $\color{red}\times$    \\
-Entry of metadata         & \checkmark   & $\color{red}\times$   & $\color{red}\times$   & $\color{red}\times$   & $\color{red}\times$       & $\color{red}\times$       & $\color{red}\times$    \\
+Reproducable$^8$  & \checkmark   & \checkmark & \checkmark & $\color{red}\times$   & \checkmark     & \checkmark       & $\color{red}\times$    \\
 Summarising data          & \checkmark   & $\color{red}\times$   & $\color{red}\times$   & $\color{red}\times$   & $\color{red}\times$       & $\color{red}\times$       & $\color{red}\times$    \\
 Multiple image processing & \checkmark   & $\color{red}\times$   & $\color{red}\times$   & $\color{red}\times$   & $\color{red}\times$       & $\color{red}\times$       & $\color{red}\times$    \\
-Reproducable$^9$  & \checkmark   & \checkmark & \checkmark & $\color{red}\times$   & \checkmark     & \checkmark       & $\color{red}\times$    \\
 Automated point detection & $\color{red}\times$     & \checkmark &     $\color{red}\times$      & \checkmark & \checkmark     & \checkmark     & $\color{red}\times$    \\
@@
 Zoom                      & $\color{red}\times$     & \checkmark & \checkmark & \checkmark & \checkmark     & $\color{red}\times$       & $\color{red}\times$    \\
+Graph rotation$^9$        & \checkmark   & \checkmark & \checkmark & \checkmark & \checkmark     & $\color{red}\times$       & $\color{red}\times$    \\
 Log axis                  & \checkmark   & \checkmark & \checkmark & \checkmark & \checkmark     & $\color{red}\times$       & $\color{red}\times$    \\
@@
 \\$^7$ Only automated, no manual extraction.
-\\$^8$ Or handles rotated graphs. 
-\\$^9$ Allows saving, re-plotting and editing of data extraction.
+\\$^8$ Allows saving, re-plotting and editing of data extraction.
+\\$^9$ Or handles rotated graphs. 
 \\$^{10}$ \proglang{R} package.

### 2018-07-13 docs/MEE_submission/manuscript_MEE.tex
final MS updates

@@
 \section{\pkg{metaDigitise} and Reproducibility}
-The \pkg{metaDigitise} package has one main function, \fct{metaDigitise}, which interactively takes the user through the process of extracting data from figures.
+The \pkg{metaDigitise} package has one main function, \fct{metaDigitise}, which interactively takes the user through the process of extracting data from figures (see Supplementary Material S1 for a full tutorial).
 \fct{metaDigitise} works on a directory containing images of figures copied from primary literature, in .png, .jpg, .tiff, .pdf format, specified to \fct{metaDigitise} through the \code{dir} argument.
@@
 \fct{metaDigitise} assumes that bars are vertical, but allows the user to flip the image to make the bars are vertical (Figure S2C,D).
-
-%%% \pkg{metaDigitise} also allows back calculation of data presented on log axes.
+\pkg{metaDigitise} also allows back calculation of data presented on log axes.
 
@@
 \centering 
-\rotatebox{90}{%
+% \rotatebox{90}{%
 
- \includegraphics[width=1.3\textwidth]{Adobe_fig1_V3.png} 
- }
- \caption{Functionality of \pkg{metaDigitise}.
+%  \includegraphics[width=1.3\textwidth]{Adobe_fig1_V3.png} 
+%  }
+ \caption{\doublespacing
+Functionality of \pkg{metaDigitise}.
 Using the iris dataset in R, digitisation of different plot types, A) mean/error plot, B) box plot, C) histogram and D) scatter plot, is shown in \pkg{metaDigitise} (left) compared with other common softwares (right).

### 2018-07-13 docs/MEE_submission/supplements.tex
final MS updates

@@
 In order to evaluate the consistency of digitisation using \pkg{metaDigitise} between users, we simulated a dataset of two variables with two groups ($n=10$ within groups).
-Each variable was plotted twice for each plot type (figures were modified slightly to give users a sense that they were digitising new data) generating a total of 14 figures.
+Each variable was plotted twice for each plot type (figures were modified slightly to give users a sense that they were digitising new data) generating a total of 14 figures (example figures are available at http://doi.org/10.5281/zenodo.1311681).
 Fourteen independent digitisers (including the authors) were provided with a directory with all 14 figures in a randomised order.
@@
 Sampling variance for ICC estimates was generated based on 1000 parametric bootstraps of the model and the significance was tested using likelihood ratio tests, using \pkg{rptR} \citep{stoffel2017}.
+Data and code are available at http://doi.org/10.5281/zenodo.1311681.
 

### 2018-07-13 R/metaDigitise.R
bug fix for get_notDone_file_details()

@@ -269,24 +269,27 @@ setup_calibration_dir <- function(dir){
 
 get_notDone_file_details <- function(dir){
 	
-	      details <- dir_details(dir)
+	details <- dir_details(dir)
 
 	# Check whether there are new files still needing to be done.
-
-	if(length(details$calibrations) == length(details$name)) {
-		stop("\nCongratulations! Looks like you have finished digitising all figures in this directory.\n", call. = FALSE)
+	if(!any(!details$name %in% details$calibrations)) {
+		stop("\r Congratulations! Looks like you have finished digitising all figures in this directory.\n", call. = FALSE)
 	}
 
 	# Find what files are already done. Remove these from our list
 	if (length(details$calibrations) >= 1){
-		done_figures <- match(details$calibrations, details$name)
-	
-	# Remove the files that are already done.
-		details$images <- details$images[-done_figures]
-		  details$name <- details$name[-done_figures]
-		 details$paths <- details$paths[-done_figures]
+		done_figures <- details$name %in% details$calibrations
+
+		# Remove the files that are already done.
+		details$images <- details$images[!done_figures]
+		details$name <- details$name[!done_figures]
+		details$paths <- details$paths[!done_figures]
 	}
 
+	cal_no_fig <- details$calibrations[!details$calibrations %in% details$name  ]
+	if(length(cal_no_fig)>0){
+	    message("Some calibration files do not have matching figures - a figure has been either deleted or renamed.")
+	}
 	return(details)
 }
 

### 2018-07-13 R/metaDigitise.R
second big fix for get_notDone_file_details()

@@ -272,6 +272,11 @@ get_notDone_file_details <- function(dir){
 	details <- dir_details(dir)
 
 	# Check whether there are new files still needing to be done.
+	cal_no_fig <- details$calibrations[!details$calibrations %in% details$name  ]
+	if(length(cal_no_fig)>0){
+	    message("Some calibration files do not have a corresponding figure - a figure has been either deleted or renamed.\n")
+	}
+
 	if(!any(!details$name %in% details$calibrations)) {
 		stop("\r Congratulations! Looks like you have finished digitising all figures in this directory.\n", call. = FALSE)
 	}
@@ -286,10 +291,7 @@ get_notDone_file_details <- function(dir){
 		details$paths <- details$paths[!done_figures]
 	}
 
-	cal_no_fig <- details$calibrations[!details$calibrations %in% details$name  ]
-	if(length(cal_no_fig)>0){
-	    message("Some calibration files do not have matching figures - a figure has been either deleted or renamed.")
-	}
+
 	return(details)
 }
 

### 2018-07-13 R/calibrate_coords.R
remove log transformation for mean_error and histogram, until correct transformations have been implemented

@@ -108,7 +108,12 @@ user_calibrate <- function(object){
       object$calpoints <- do.call(cal_coords, object)
       object$point_vals <- do.call(getVals, object)
       do.call(internal_redraw, c(object, calibration=TRUE, points=FALSE))
-      object$log_axes <- logAxes()
+      if(object$plot_type %in% c("boxplot","scatterplot")){
+        object$log_axes <- logAxes()
+      }else{
+        object$log_axes <- c(axes="n",transformed=NULL,base=NULL)
+      }
+      
     }
     cal_Q <- readline("\nRe-calibrate? (y/n) \n")
   }

### 2018-07-14 R/calibrate_coords.R
testing bug fix for user_calibrate()

@@ -112,8 +112,7 @@ user_calibrate <- function(object){
         object$log_axes <- logAxes()
       }else{
         object$log_axes <- c(axes="n",transformed=NULL,base=NULL)
-      }
-      
+      }      
     }
     cal_Q <- readline("\nRe-calibrate? (y/n) \n")
   }

### 2018-07-14 tests/testthat/test-calibrate_coords.R

@@ -56,11 +56,10 @@ testthat::test_that("Checking user_calibrate..", {
 		`metaDigitise::logAxes` = function(...) c(axes="n"),
 		readline = function(...) "n",
 		testthat::expect_equal(
-			user_calibrate(object=list()),
+			user_calibrate(object=list(plot_type="scatterplot")),
 			list(calpoints=data.frame(x=c(0,0,0,100),y=c(0,100,0,0)), point_vals=c(y1=1,y2=2,x1=3,x2=4), log_axes=c(axes="n")), 
 			info = "user_calibrate failed"
 		)
 	)
-
 })
 

### 2018-09-28 docs/MEE_submission/manuscript_MEE.tex
updated MS

@@
 
-In the fields of ecology and evolution, researchers make use of data presented in primary literature for comparative and meta-analyses.
-These techniques rely on descriptive statistics (e.g.
+In many different contexts, researchers make use of data presented in primary literature.
+In the fields of ecology and evolution (E\&E), this data is most commonly used for comparative and meta-analyses.
+Meta-analysis in particular is a rapidly expanding methodology in E\&E, not only in terms of the number of meta-analyses (in plant ecology alone the yearly number of published meta-analyses doubled from 2006 to 2012 (20-40); \cite{Koricheva2014}), but also in terms of their size (a recent meta-analysis, for example, included 6440 effect sizes from 175 publications; \cite{Noble2018}).
+The integrative nature of these studies makes them extremely important to the progression of E\&E \citep{Gurevitch2018}.
+These techniques rely heavily on descriptive statistics (e.g.
 means, standard deviations (SD), sample sizes, correlation coefficients) extracted from primary literature.
@@
 
-Although there are several tools that extract data from figures (e.g.
-\proglang{DataThief} \citep{DataThief}, \proglang{GraphClick} \citep{GraphClick}, \proglang{WebPlotDigitizer} \citep{WebPlotDigitizer}, see Table \ref{tab:comparison}), these tools do not cater to needs of meta-analysts for four main reasons (here we focus on meta-analysis, although many  points apply to extraction for comparative analysis).
+Although there are several tools that extract data from figures, including both standalone programs and \proglang{R} packages (reviewed in Table \ref{tab:comparison}), these tools do not cater to the general needs of meta-analysts for four main reasons (here we focus on meta-analysis, although many  points apply to extraction for comparative analysis).
+%
  First, although meta-analysis is an important tool in consolidating the data from multiple studies, many of the processes involved in data extraction are opaque and difficult to reproduce, making extending or replicating studies problematic.
 Having a tool that facilitates reproducibility in meta-analyses will increase transparency and aid in resolving the reproducibility crises seen in many fields \citep{peng_reproducible_2006, peng_reproducible_2011, Parker2016}.
+%
  Second, digitising programs do not allow the integration of metadata at the time of data extraction, such as experimental group or variable names, and sample sizes.
-This makes the downstream calculations laborious, as information has to be added later using different software.
+This makes the downstream calculations laborious, as information has to be added later, typically using different software.
+% Given the ubiquity of the R platform in E\&E, and that the most popular meta-analysis software in E\&E (e.g., metafor and MCMCglmm) are available as R packages, it is highly likely to be used for some (if not all) stages of the research synthesis process.
+It would therefore be very beneficial to have comprehensive, and robust, digitisation capabilities in R that provides users with lots of flexibility, and makes the process of figure extraction more transparent and easier to reproduce.
+%
  Third, existing programs do not import sets of images for the user to systematically work through.
@@
 These data often subsequently need to be imported and edited using different software.
+%
  Finally, digitising programs typically only provide the user with calibrated \textit{x,y} coordinates from imported figures, and do not differentiate between common plot types that are used to present data.
 Consequently, a large amount of additional data manipulation is required, that is different across plots types.
-For example, data are frequently presented in plots with means and standard errors or confidence intervals (Figure \ref{fig:all_extract}A), from which the user wants a mean and SD for each group presented.
+For example, in E\&E data are commonly presented in plots with means and standard errors or confidence intervals (Figure \ref{fig:all_extract}A), from which the user wants a mean and SD for each group presented.
 From \textit{x,y} coordinates, users must manually discern between mean and error coordinates and assign points to groups.
 The error then needs to be calculated as the deviation from the mean, and then transformed to SD, according to the type of error presented.
+Histograms and box plots are also frequently used in E\&E to presented data, and whilst their downsream calculations are even more laborious, there are few (if any; see Table \ref{tab:comparison}) tools to extract data from these plot types.
 
-Data extraction from figures is therefore an incredibly time-consuming process as existing software does not provide an optimized, reproducible research pipeline to facilitate data extraction and editing.
-Here, we present an interactive \proglang{R} package, \pkg{metaDigitise} (available at https://github.com/daniel1noble/metaDigitise), which is designed for large scale, reproducible data extraction from figures, specifically catering to the the needs of meta-analysts.
-To this end, we provide tools to extract data from common plot types (mean/error plots, box plots, scatter plots and histograms, see Figure \ref{fig:all_extract}).
+Data extraction from figures is therefore a time-consuming process as existing software does not provide an optimized, reproducible research pipeline to facilitate data extraction and editing.
+Given the ubiquity of the R platform in E\&E, and that it hosts the most popular meta-analysis software in E\&E (e.g., metafor \citep{Viechtbauer2010} and MCMCglmm \citep{Hadfield2010b}), it is highly likely to be used for some (if not all) stages of the research synthesis process.
+It is therefore important to have comprehensive, robust and flexible digitisation capabilities in R to make the process of figure extraction more streamline,  transparent and easier to reproduce.
+Here, we present an interactive \proglang{R} package, \pkg{metaDigitise} (available on CRAN), which is designed for large scale, reproducible data extraction from figures, specifically catering to the the needs of meta-analysts.
+To this end, we provide tools to extract data from common plot types in E\&E (mean/error plots, box plots, scatter plots and histograms, see Figure \ref{fig:all_extract}).
 \pkg{metaDigitise} operates within the \proglang{R} environment making data extraction, analysis and export more streamlined.
@@
 The \pkg{metaDigitise} package has one main function, \fct{metaDigitise}, which interactively takes the user through the process of extracting data from figures (see Supplementary Material S1 for a full tutorial).
+Running \fct{metaDigitise} presents the user with three options; `Process new images', `Import existing data' or `Edit existing data', which can be used during and after digitisation to execute a range of functions (see Figure \ref{fig:all_extract} – `Processing images' is discussed in Section 3, and `Editing' and `Importing' in Section 4).
 \fct{metaDigitise} works on a directory containing images of figures copied from primary literature, in .png, .jpg, .tiff, .pdf format, specified to \fct{metaDigitise} through the \code{dir} argument.
@@
 
-The complete digitisation process can then be reproduced at a later stage, shared with collaborators and presented as supplementary materials for a publication, regardless of the computer it is run on.
-For an analysis to be updated, new figures can simply be added to the directory and \fct{metaDigitise} run to incorporate the new data.
+The complete digitisation process can be reproduced at a later stage, shared with collaborators and presented as supplementary materials for a publication, regardless of the computer it is run on.
+To update an analysis, new figures can simply be added to the directory and \fct{metaDigitise} run to incorporate the new data.
 
@@
 \section{Image Processing}
-Running \fct{metaDigitise} presents the user with three options; `Process new images', `Import existing data' or `Edit existing data', which can be used during and after digitisation to execute a range of functions (see Figure \ref{fig:all_extract} – 'Editing' and 'Importing' are discussed in the next section).
-Selecting `Process New Images' starts the digitisation process on images within the directory that have not previously been digitised.
+ Selecting `Process New Images', after running \fct{metaDigitise}, starts the digitisation process on images within the directory that have not previously been digitised.
 For all plot types, \fct{metaDigitise} requires the user to calibrate the axes in the figure, by clicking on two known points on the axis in question, and entering the value of those points (Figure \ref{fig:all_extract}).
@@
 
-As figures may have been copied from older, scanned publications, they may not be perfectly orientated.
-This makes calibration of the points in the figure problematic.
+Calibration of points in figures from older, scanned publications can be problematic, as the figures may not be perfectly orientated.
 \fct{metaDigitise} allows users to rotate the image (Figure S2A,B).
@@
 \pkg{metaDigitise} is also able to re-import, edit and re-plot previously digitised figures.
-When running \fct{metaDigitise}, the user can choose to 'Import existing data', which returns previously digitised data, from a single figure or all figures.
+When running \fct{metaDigitise}, the user can choose to `Import existing data' , which returns previously digitised data, from a single figure or all figures.
 Alternately, the \fct{getExtracted} function returns the data from previous digitisations, but without user interaction, allowing easier integration into larger scripts.
-'Edit existing data' allows the user to re-plot or edit information for digitisations that have previously be done.
+`Edit existing data' allows the user to re-plot or edit information for digitisations that have previously be done.
 Re-plotting digitisations with all metadata is an important reproducibility feature, as it allows users to see exactly what information has been extracted, as well as making it easy to spot and data extraction errors.
@@
 In order to evaluate the consistency of digitisation with \pkg{metaDigitise} between users, fourteen people digitized sets of 14 identical images created from a simulated dataset (see supplements).
-We found no evidence for any inter-observer variability in digitisations for the mean (ICC = 0, 95\% CI = 0 to 0.029, \textit{p} = 1), SD (ICC = 0, 95\% CI = 0 to 0.033, \textit{p} = 0.5) or correlation coefficient (ICC = 0.053, 95\% CI = 0 to 0.296, \textit{p} = 0.377).
+We found no evidence for any inter-observer variability in digitisations for the mean (ICC = 0, 95\% CI = 0 to 0.029, \textit{p} $>$ 0.999), SD (ICC = 0, 95\% CI = 0 to 0.033, \textit{p} $>$ 0.999) or correlation coefficient (ICC = 0.053, 95\% CI = 0 to 0.296, \textit{p} = 0.377).
 There was little bias between digitised and true values, on average 1.63\% (mean = 0.02\%, SD = 4.9\%, \textit{r} = -0.03\%) and there were small absolute differences between digitised and true values, on average 2.18\% (mean = 0.40\%, SD = 5.81\%, \textit{r} = 0.33\%) across all three descriptive statistics.
 SD estimates from digitisations are clearly most error prone.
-The mean absolute differences for each plot type clearly show that this effect is driven by extraction from box plots and histograms (\% difference; box plot: 15.805, histogram: 5.210, mean/error: 1.500, scatter plot: 0.433).
+The mean absolute differences for each plot type clearly show that this effect is driven by extraction from box plots and histograms (\% difference; box plot: 15.81, histogram: 5.21, mean/error: 1.50, scatter plot: 0.43).
 SD estimation from box plot descriptive statistics is known to be more error prone, especially at small sample sizes \citep{Wan2014}.
@@
 We also used simulated data to test the accuracy of digitisations with respect to known values (see supplements).
-\pkg{metaDigitise} was extremely accurate at matching clicked points to their true values essentially being perfectly correlated with the true simulated data for both the \textit{x}-variable (Pearson's correlation; $r > 0.999$, \textit{t} = 2137.4, \textit{df} = 78, $p < 0.001$) and \textit{y}-variable ($r > 0.999$, \textit{t} = 1897.8, \textit{df} = 78, $p < 0.001$) in scatterplots.
+\pkg{metaDigitise} was very accurate at matching clicked points to their true values essentially being perfectly correlated with the true simulated data for both the \textit{x}-variable (Pearson's correlation; $r > 0.999$, \textit{t} = 2137.4, \textit{df} = 78, $p < 0.001$) and \textit{y}-variable ($r > 0.999$, \textit{t} = 1897.8, \textit{df} = 78, $p < 0.001$) in scatterplots.
 
@@
 In contrast to some other packages \pkg{metaDigitise} does not extract lines from figures.
-Line extraction is not particularly useful for most comparative or meta-analytic work, although we recognise that it may be useful in fields other than these.
+Although line extraction is not generally necessary in meta-analysis and comparative analysis, outside of these fields researchers may need to extract parameters of a line from a figure.
 Should a user like to extract lines with \pkg{metaDigitise}, we would recommend extracting data as a scatter plot, and clicking along the line in question.
@@
 \centering 
-% \rotatebox{90}{%
-
-%  \includegraphics[width=1.3\textwidth]{Adobe_fig1_V3.png} 
-%  }
- \caption{\doublespacing
+\rotatebox{90}{%
+  \includegraphics[width=1.3\textwidth]{Adobe_fig1_V3.png} 
+ }
+ \caption{%\doublespacing
 Functionality of \pkg{metaDigitise}.
@@
 As shown on the left hand side of the figure, \pkg{metaDigitise} clearly displays the stages of the digitisation to aid the transparency of the process, and returns concatenated summary data for all images.
-%  All figures are clearly labelled at the top to remind users of the filename and plot type.
-This reduces errors throughout the digitisation process.
-Names of the variables and calibration (in blue) are plotted alongside the digitised points.
-In A) and B), user entered group names and sample sizes are displayed beside the relevant points.
-In C) the names and sample sizes for each group are shown below the figure.
 }

### 2018-09-29 docs/MEE_submission/manuscript_MEE.tex
Updated the MS with some small changes

@@
 In many different contexts, researchers make use of data presented in primary literature.
-In the fields of ecology and evolution (E\&E), this data is most commonly used for comparative and meta-analyses.
-Meta-analysis in particular is a rapidly expanding methodology in E\&E, not only in terms of the number of meta-analyses (in plant ecology alone the yearly number of published meta-analyses doubled from 2006 to 2012 (20-40); \cite{Koricheva2014}), but also in terms of their size (a recent meta-analysis, for example, included 6440 effect sizes from 175 publications; \cite{Noble2018}).
-The integrative nature of these studies makes them extremely important to the progression of E\&E \citep{Gurevitch2018}.
+In the fields of ecology and evolution (E\&E), these data are most commonly used for comparative and meta-analyses.
+The use of meta-analysis in E\&E in particular, is rapidly growing, not only in terms of the number of meta-analyses (in plant ecology alone the yearly number of published meta-analyses doubled from 2006 to 2012 (20-40); \cite{Koricheva2014}), but also in terms of their size (a recent meta-analysis, for example, included 6440 effect sizes from 175 publications; \cite{Noble2018}).
+Meta-analyses are extremely important in providing a means of quantitatively synthesizing experimental and/or observational studies to evaluate empirical support for fundamental theory in E\&E \citep{Gurevitch2018}.
 These techniques rely heavily on descriptive statistics (e.g.
 means, standard deviations (SD), sample sizes, correlation coefficients) extracted from primary literature.
-As well as being presented in the text or tables of research papers, descriptive statistics are frequently presented in figures and so need to be manually extracted using digitising programs.
+As well as being presented in the text or tables of research papers, descriptive statistics are frequently presented in figures.
+For example, in some cases up to 42\% of the papers used have one or more figures for data extraction \cite{Noble2018}.
+These data often need to be manually extracted using digitising programs.
 
@@
 The error then needs to be calculated as the deviation from the mean, and then transformed to SD, according to the type of error presented.
-Histograms and box plots are also frequently used in E\&E to presented data, and whilst their downsream calculations are even more laborious, there are few (if any; see Table \ref{tab:comparison}) tools to extract data from these plot types.
+Histograms and box plots are also frequently used in E\&E to presented data, and whilst their downstream calculations are even more laborious, there are few (if any; see Table \ref{tab:comparison}) tools to extract data from these plot types.
 
@@
 In contrast to some other packages \pkg{metaDigitise} does not extract lines from figures.
-Although line extraction is not generally necessary in meta-analysis and comparative analysis, outside of these fields researchers may need to extract parameters of a line from a figure.
+Although line extraction is not generally necessary in comparative and meta-analysis, outside of these fields researchers may need to extract parameters of a line from a figure.
 Should a user like to extract lines with \pkg{metaDigitise}, we would recommend extracting data as a scatter plot, and clicking along the line in question.

### 2018-10-01 docs/MEE_submission/manuscript_MEE.tex
tiny MS changes

@@
 As well as being presented in the text or tables of research papers, descriptive statistics are frequently presented in figures.
-For example, in some cases up to 42\% of the papers used have one or more figures for data extraction \cite{Noble2018}.
+For example, in some cases up to 42\% of the papers used have one or more figures for data extraction \citep{Noble2018}.
 These data often need to be manually extracted using digitising programs.
@@
 Although there are several tools that extract data from figures, including both standalone programs and \proglang{R} packages (reviewed in Table \ref{tab:comparison}), these tools do not cater to the general needs of meta-analysts for four main reasons (here we focus on meta-analysis, although many points apply to extraction for comparative analysis).
-%
 First, although meta-analysis is an important tool in consolidating the data from multiple studies, many of the processes involved in data extraction are opaque and difficult to reproduce, making extending or replicating studies problematic.
 Having a tool that facilitates reproducibility in meta-analyses will increase transparency and aid in resolving the reproducibility crises seen in many fields \citep{peng_reproducible_2006, peng_reproducible_2011, Parker2016}.
-%
 Second, digitising programs do not allow the integration of metadata at the time of data extraction, such as experimental group or variable names, and sample sizes.
 This makes the downstream calculations laborious, as information has to be added later, typically using different software.
-% Given the ubiquity of the R platform in E\&E, and that the most popular meta-analysis software in E\&E (e.g., metafor and MCMCglmm) are available as R packages, it is highly likely to be used for some (if not all) stages of the research synthesis process.
-It would therefore be very beneficial to have comprehensive, and robust, digitisation capabilities in R that provides users with lots of flexibility, and makes the process of figure extraction more transparent and easier to reproduce.
-%
 Third, existing programs do not import sets of images for the user to systematically work through.
@@
 These data often subsequently need to be imported and edited using different software.
-%
 Finally, digitising programs typically only provide the user with calibrated \textit{x,y} coordinates from imported figures, and do not differentiate between common plot types that are used to present data.

### 2018-10-02 docs/MEE_submission/manuscript_MEE.tex
MS tweaks
another MS tweak

@@
 \section{Software Validation}
-In order to evaluate the consistency of digitisation with \pkg{metaDigitise} between users, fourteen people digitized sets of 14 identical images created from a simulated dataset (see supplements).
+To evaluate the consistency of digitisation with \pkg{metaDigitise} between users, fourteen people digitized sets of 14 identical images created from a simulated dataset (see supplements).
 We found no evidence for any inter-observer variability in digitisations for the mean (ICC = 0, 95\% CI = 0 to 0.029, \textit{p} $>$ 0.999), SD (ICC = 0, 95\% CI = 0 to 0.033, \textit{p} $>$ 0.999) or correlation coefficient (ICC = 0.053, 95\% CI = 0 to 0.296, \textit{p} = 0.377).
@@
 As a result, we do not believe that current automated point detection techniques provide substantial benefits in terms of time or accuracy.
+Indeed, in a recent project developing automated point extraction techniques, only 15/136 (11\%) of studies screened contained figures suitable for the presented method, and in only 12/27 (44\%) of the resulting figures was the data correctly extracted \citep{Hartgerink2017}.
 

### 2018-12-11 docs/MEE_submission/manuscript_MEE.tex

@@
 \newcommand{\class}[1]{`\code{#1}'}
-%\newcommand{\fct}[1]{#1}
 \newcommand{\fct}[1]{\texttt{#1()}}
@@
 
-\DefineVerbatimEnvironment{Code}{Verbatim}{}
-\DefineVerbatimEnvironment{CodeInput}{Verbatim}{fontshape=sl}
-\DefineVerbatimEnvironment{CodeOutput}{Verbatim}{}
-\newenvironment{CodeChunk}{}{}
-
 \raggedright
@@
 
-Joel L. Pick$^{1,*}$, Shinichi Nakagawa$^1$, Daniel W.A. Noble$^1$
+Joel L. Pick$^{1,2,*}$, Shinichi Nakagawa$^1$, Daniel W.A. Noble$^1$
 
 $^1$
-  Ecology and Evolution Research Centre, School of Biological, Earth and Environmental Sciences,  University of New South Wales, Kensington, NSW 2052, Sydney, AUSTRALIA
+  Ecology and Evolution Research Centre, School of Biological, Earth and Environmental Sciences, University of New South Wales, Kensington, NSW 2052, Sydney, Australia
+
+$^2$
+  Current Address: Institute of Evolutionary Biology, School of Biological Sciences, University of Edinburgh, Edinburgh EH9 3JT, United Kingdom
 
  $^*$Corresponding Author: joel.l.pick@gmail.com\\
+
+\vskip10pt
+ \textbf{Running Head:} Data extraction from figures with metaDigitise
 
@@
 In the fields of ecology and evolution (E\&E), these data are most commonly used for comparative and meta-analyses.
-The use of meta-analysis in E\&E in particular, is rapidly growing, not only in terms of the number of meta-analyses (in plant ecology alone the yearly number of published meta-analyses doubled from 2006 to 2012 (20-40); \cite{Koricheva2014}), but also in terms of their size (a recent meta-analysis, for example, included 6440 effect sizes from 175 publications; \cite{Noble2018}).
+The use of meta-analysis in E\&E in particular, is rapidly growing, not only in terms of the number of meta-analyses (in plant ecology alone the yearly number of published meta-analyses doubled from 2006 to 2012 (20-40) \citep{Koricheva2014}), but also in terms of their size (a recent meta-analysis, for example, included 6440 effect sizes from 175 publications \citep{Noble2018}).
 Meta-analyses are extremely important in providing a means of quantitatively synthesizing experimental and/or observational studies to evaluate empirical support for fundamental theory in E\&E \citep{Gurevitch2018}.
@@
 As well as being presented in the text or tables of research papers, descriptive statistics are frequently presented in figures.
-For example, in some cases up to 42\% of the papers used have one or more figures for data extraction \citep{Noble2018}.
-These data often need to be manually extracted using digitising programs.
+For example, 42\% of the papers used in a recent meta-analysis presented some or all of the required data in figures \citep{Noble2018}.
+These data need to be manually extracted using digitising programs.
 
@@
 Data extraction from figures is therefore a time-consuming process as existing software does not provide an optimized, reproducible research pipeline to facilitate data extraction and editing.
-Given the ubiquity of the R platform in E\&E, and that it hosts the most popular meta-analysis software in E\&E (e.g., metafor \citep{Viechtbauer2010} and MCMCglmm \citep{Hadfield2010b}), it is highly likely to be used for some (if not all) stages of the research synthesis process.
-It is therefore important to have comprehensive, robust and flexible digitisation capabilities in R to make the process of figure extraction more streamline, transparent and easier to reproduce.
+Given the ubiquity of the \proglang{R} platform in E\&E, and that it hosts the most popular meta-analysis software in E\&E (e.g., metafor \citep{Viechtbauer2010} and MCMCglmm \citep{Hadfield2010b}), it is highly likely to be used for some (if not all) stages of the research synthesis process.
+It is therefore important to have comprehensive, robust and flexible digitisation capabilities in \proglang{R} to make the process of figure extraction more streamline, transparent and easier to reproduce.
 Here, we present an interactive \proglang{R} package, \pkg{metaDigitise} (available on CRAN), which is designed for large scale, reproducible data extraction from figures, specifically catering to the the needs of meta-analysts.
@@
 \section*{Acknowledgments}
-We thank the I-DEEL group and colleagues at UNSW for for testing, providing feedback and digitising including: Rose O'Dea, Fonti Kar, Malgorzata Lagisz, Julia Riley, Diego Barneche, Erin Macartney, Ivan Beltran, Gihan Samarasinghe, Dax Kellie, Jonathan Noble, Yian Noble and Alison Pick.
+We thank the I-DEEL group and colleagues at UNSW for for testing, providing feedback and digitising including: Rose O'Dea, Fonti Kar, Malgorzata Lagisz, Julia Riley, Diego Barneche, Erin Macartney, Ivan Beltran, Gihan Samarasinghe, Dax Kellie, Jonathan Noble, Yian Noble, Elena Noble and Alison Pick.
 J.L.P.
@@
 \rotatebox{90}{%
-  \includegraphics[width=1.3\textwidth]{Adobe_fig1_V3.png} 
+  \includegraphics[width=1.3\textwidth]{Adobe_fig1_final.pdf} 
  }
